### 2024-01-17 frontiers_submission/frontiers_submission.Rmd
initial commit

---
title: "A meta-analysis of the impacts of best management practices on nonpoint source pollutants."
runtitle: "Meta-analysis of best management practices"
documentclass: FrontiersinHarvard # for articles in journals using the Harvard Referencing Style (Author-Date), for Frontiers Reference Styles by Journal: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal
# documentclass: FrontiersinVancouver # for articles in journals using the Vancouver Reference Style (Numbered), for Frontiers Reference Styles by Journal: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal
# documentclass: frontiersinFPHY_FAMS # Vancouver Reference Style (Numbered) for articles in the journals "Frontiers in Physics" and "Frontiers in Applied Mathematics and Statistics" 
classoption: utf8
author:
  - name: Michael Schramm
    affiliation: '1'
    etal: Schramm # First author's last name. 
    email: michael.schramm@ag.tamu.edu # Indicates corresponding Author
    institution: Texas Water Resources Institute, Texas A&M AgriLife Research
    street: 1001 Holleman Dr E.
    city: College Station
    state: Texas # only USA, Australia, Canada
    zip: 77840
    country: USA
  - name: Duncan Kikoyo
    affiliation: '1'
  - name: Janelle Wright
    affiliation: '1'
  - name: Shubham Jain
    affiliation: '1'
affiliation:
  - id: '1'
    department: Texas Water Resources Institute
    institution: Texas A&M AgriLife Research
    city: College Station
    state: Texas # only USA, Australia, Canada
    country: USA

# Abstract can also be provided at the 
abstract: |
  Abstract length and content varies depending on article type. Refer to 
  <http://www.frontiersin.org/about/AuthorGuidelines> for abstract requirement
  and length according to article type.
keywords:
  - best management practice
  - water quality
  - nonpoint source pollution
  - fecal indicator bacteria
  - nutrients
output: rticles::frontiers_article
# bib yaml is commented to manually insert in correct section
# csl: frontiers.csl
# bibliography: bibliography.bib
# Match bibliostyle with the documentclass
# biblio-style: Frontiers-Harvard # Many Frontiers journals use the Harvard referencing system (Author-date), to find the style and resources for the journal you are submitting to: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal. For Humanities and Social Sciences articles please include page numbers in the in-text citations 
# biblio-style: Frontiers-Vancouver # Many Frontiers journals use the numbered referencing system, to find the style and resources for the journal you are submitting to: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal
---

```{r setup, include=FALSE}
## load libraries
library(flextable)
library(ggtext)
library(kableExtra)
library(officer)
library(ragg)
library(targets)
library(tidyverse)

## this sets our default code chunk options
knitr::opts_chunk$set(dev = "ragg_png",
                      echo = FALSE,
                      ## figure chunk options
                      fig.cap = TRUE,
                      fig.width = 6,
                      dpi = 300)

## make fonts available
font_hoist <- function(family, silent = FALSE) {
  font_specs <- systemfonts::system_fonts() |> 
    filter(family == {{family}}) |> 
    mutate(family = paste(family, style)) |> 
    select(plain = path, name = family)

  purrr::pwalk(as.list(font_specs), systemfonts::register_font)

  if (!silent)  message(paste0("Hoisted ", nrow(font_specs), " variants:\n",
                               paste(font_specs$name, collapse = "\n")))
}

font_hoist("Manrope")

## grab output from targets to use in manuscript
store <- "../_targets"
bac_missing_sd <- tar_read(bac_missing_sd,
                           store = store)
tn_missing_sd <- tar_read(tn_missing_sd,
                           store = store)
tp_missing_sd <- tar_read(tp_missing_sd,
                           store = store)
tss_missing_sd <- tar_read(tss_missing_sd,
                           store = store)

eggers_bac <- tar_read(bac_sse,
                       store = store)
## set flextable defaults
set_flextable_defaults(fonts_ignore=TRUE)

```

# Introduction

In the United States (U.S.), major progress towards water quality goals have been achieved through the Clean Water Act of 1972.
This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge [@benhamLessonsLearnedTMDL2008; @nationalresearchcouncilAssessingTMDLApproach2001; @schrammTotalMaximumDaily2022].
Increased pollutant loads and pollutant concentrations in runoff are a particular challenge when dealing with nonpoint source pollution resulting from land use change.
The impacts of land use change on hydrology and water quality are well established [@allanLandscapesRiverscapesInfluence2004; @carpenterNonpointPollutionSurface1998; @bernhardtUnderstandingManagingMinimizing2008; @careyEvaluatingNutrientImpacts2013; @freemanImpactsUrbanizationDevelopment2019].
Nonpoint source driven fecal indicator bacteria (FIB), nitrogen, phosphorus, and suspended sediment remain major causes of water quality impairments in U.S. rivers and streams despite decades of work.
In 2017, the Environmental Protection Agency (EPA) estimated 41% or more of the nation's rivers and streams rated poorly for biological condition due to excess nitrogen or phosphorus [@epaNationalWaterQuality2017]. 
FIB remains the leading cause of water body impairment on the Clean Water Act 303(d) list in the United States [@epaNationalWaterQuality2017].

Best management practices (BMPs) have been the primary suite of tools for addressing nonpoint source pollution. 
BMPs are structural or non-structural controls used to mitigate the effects of increased runoff volume, pollutant loads, or pollutant concentrations emanating from diffuse nonpoint sources. 
BMPs control the delivery of pollutants through a few possible mechanisms. 
Structural BMPs reduce and retard total volume of runoff, thus reducing both the volume of water and pollutant load. 
Structural BMPs may also provide a mechanism for physical, chemical, or biological removal of pollutant constituents suspended or dissolved in runoff. 
Non-structural BMPs (such as nutrient management or livestock management) are utilized to reduce the generation of pollutant runoff by avoiding pollutant generation during critical periods.

Practitioners rely extensively on mechanistic models to plan and evaluate BMP scenarios and resulting water quality.
@linternBestManagementPractices2020 found 43% of reviewed BMP effectiveness studies relied completely on modeled outputs, with modeled outputs almost always predicting water quality improvements following BMP implementation. 
However, field studies are much more likely to demonstrate mixed results including net releases of pollutants under certain conditions  [@linternBestManagementPractices2020; @liuReviewEffectivenessBest2017].
The disconnect between modeled outcomes and field studies might be attributed to (1) overly simplified or incorrect estimates of parameters that represent management practices [@ullrichApplicationSoilWater2009; @fuReviewCatchmentscaleWater2019; @linternBestManagementPractices2020], (2) failure to incorporate uncertainty in estimates [@tasdighiBayesianTotalUncertainty2018; @fuReviewCatchmentscaleWater2019; @linternBestManagementPractices2020], and (3) assumption of static performance over time [@mealsLagTimeWater2010; @liuReviewEffectivenessBest2017; @fuReviewCatchmentscaleWater2019].

A source of uncertainty comes from the substantial variability of performance metrics reported in empirical BMP studies [@linternBestManagementPractices2020].
Previous empirical reviews generally describe high variability and uncertainty in nitrogen and phosphorus removal and consistent reduction in total suspended sediment concentrations across BMP types [@linternBestManagementPractices2020;@liuReviewEffectivenessBest2017; @kochNitrogenRemovalStormwater2014; @claryBMPPerformanceAnalysis2011; @barrettComparisonBMPPerformance2008; @grudzinskiDoesRiparianFencing2020]. 
The review literature on the effects of BMPs on FIBs are sparse but generally find extremely high variance in performance across BMPs [@claryBMPPerformanceAnalysis2011; @grudzinskiDoesRiparianFencing2020].
Accurately characterizing BMP treated runoff quality is challenging due to the high variability in BMP performance.
However, influent water quality, seasonality, and BMP age provide some explanatation to highliy heterogenous BMP efficiency data [@barrettPerformanceComparisonStructural2005; @liuReviewEffectivenessBest2017].  
A recent meta-analysis demonstrated the further linkages between local climatic conditions and BMP performance on nitrogen and phosphorus concentrations [@horvathEffectsRegionalClimate2023].

# Methods

We conducted a systematic review of recent (2000-2022) literature to compile field studies documenting the effectiveness of best management practices on fecal indicator bacteria and nutrients concentrations. 
The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines [@collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018].
In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
The inclusion criteria filtered out (1) non-field studies, (2) modelling results, (3) studies that did not evaluate specific BMPs, (4) studies conducted outside of the U.S. or published in a language other than English.
We ran search queries in Texas A&M Library Catalog, Web of Science, and Google Scholar.
Although results from Google Scholar are not always replicable, we utilized the service to maximize search results for studies not published in academic journals and presumably increase the chance of identifying studies with negative effects.
Fecal indicator bacteria study seraches included the following query: “fecal indicator bacteria” OR “E. coli” OR “Escherichia coli” OR “enterococci” OR “enterococcus” AND “best management practices” OR “BMPs” AND “effectiveness” OR “performance”. Nutrient BMP studies utilized a similar query: “nutrient” OR “nitrogen” OR “phosphorus” OR “sediment” OR “TSS” AND “best management practices” OR “BMPs” AND “effectiveness” OR “performance”.

Results from each database were first filtered to remove duplicates.
After removal of duplicates, each member of the research team (*n* = 4) was assigned a subset of studies to evaluate if they should be included.
**include supplementary tables with inclusion criteria and data variables**
Each study was reviewed by two team members and differences in opinion were collectively discussed and agreed upon before progressing.
The remaining studies were split among team members for data extraction, again with at least two team members reviewing each study. 
If data was provided in figures, the data was extracted with the WebPlotDigitizer tool [@rohatgiWebPlotDigitizer2022].

```{r criteria}

tibble(
  "Attribute" = c("Study type", "Outcomes", "Geographical context", "Timeframe"),
  "Inclusion Criteria" = c("Journal articles, book chapters, conference papers, unpublished research reports, thesis and dissertations, organizational and agency white papers.",
                           "Field studies with measured effects on fecal indicator bacteria or nutrient concentrations.",
                           "Studies conducted within the United States.",
                           "Studies published from 2000 through 2022."),
  "Excluision Criteria" = c("Synopsis or review studies, peports with reductions based on modeled or other estimated reductions (e.g. TMDLs, watershed plans, or modelling studies.",
                            "Studies not explicitly linking reductions to a specific BMP or insufficient information to quantify reductions.",
                            "Studies outside of the United States.",
                            "Studies published prior to 2000 or after 2022.")
) |> 
  kableExtra::kbl(booktabs = TRUE,
                  caption = "Criteria applied for including or excluding studies within the review database.") |> 
  kableExtra::kable_styling(full_width = FALSE) |>
  kableExtra::column_spec(1, width = "5em") |> 
  kableExtra::column_spec(2, width = "15em") |> 
  kableExtra::column_spec(3, width = "15em")

```

```{r datvars}

tibble(Variable = c(
  "Publication Year", "Parameter", "Runoff source", "Source type",
  "BMP", "BMP Classification", "BMP Category", "BMP Subcategory",
  "Study scale", "Location", "State", "Study area", "Longitude", "Latitude",
  "Study years", "N control", "N experiment", "X control", "X experiment", 
  "SE control", "SE experiment", "Minimum control", "Minimum experiment", 
  "Maximum control", "Maximum experiment", "SD control", "SD experiment",
  "Units", "Percent reduction"),
  Description = c("Year the study was published",
                  "The specific pollutant measured.",
                  "Dominant source of runoff (rop fields, livestock pasture, commerical, residential",
                  "Major categorization of runoff source: agricultural or urban",
                  "BMP evaluated",
                  "BMP description based on NRCS conservation practice standards and EPA BMP fact sheets",
                  "BMP categorization based on structural or management",
                  "BMP subcategorization based on pollutant removal processes",
                  'Spatial scale of the study area ("lot/field", "community", "watershed")',
                  "Location name used in the study description",
                  "State where the study was conducted",
                  "Drainage area in hectares",
                  "Approximated or reported latitude coordinate",
                  "Approximated or reported longitude coordinate",
                  "Year or years when data were collected",
                  "Number of control measurements",
                  "Number of experiemental measurements",
                  "Mean concentration for control measurements",
                  "Mean concentration for experiemental measurements",
                  "Standard error of control measurements",
                  "Standard error of experimental measurements",
                  "Minimum control measurement",
                  "Minimum experiment measurement",
                  "Maximum control measurement",
                  "Maximum experiment measurement",
                  "Standard deviation of control measreuments",
                  "Standard deviation of experimental measurements",
                  "Units reported by the study",
                  "BMP efficiency for studies that only reported efficiency")
) |> 
  kableExtra::kbl(booktabs = TRUE,
                  caption = "Study and effect variables extracted for review.") |> 
  kableExtra::kable_styling(full_width = FALSE) |>
  kableExtra::column_spec(1, width = "10em") |> 
  kableExtra::column_spec(2, width = "25em")

```

## Effect size calculation {-}

Prior BMP reviews routinely report the effect of BMPs using efficiency values [see @agouridisLivestockGrazingManagement2005; @krogerReviewBestManagement2012; @liuEnhancingRainfallrunoffModel2015; @simpsonDevelopingBestManagement2009].
Efficiency values are calculated as:

$$
\text{Efficiency}=\frac{x_{control}-x_{experiment}}{x_{control}}\times 100,
$$

where *x~control~* is the pre-treatment or control pollutant concentration and *x~experiment~* is the pollutant concentration measured after the BMP intervention. 
There are several statistical shortcomings (distributional asymetry, skewness, and non-additive properties) when using efficiency to estimate overall effect sizes [@nuzzoPercentDifferencesAnother2018; @coleStatisticsNotesWhat2017].
In comparision, the log ratio of means (*ROM*) provide preferable statistical properties for regression analysis [@osenbergEffectSizeEcological1997; @hedgesMetaanalysisResponseRatios1999]. 
*ROM* quantifies the difference in means between the control and experimental group [@hedgesMetaanalysisResponseRatios1999]:

$$
ROM_i = ln\left(\frac{x_{i,control}}{x_{i,experiment}}\right) = ln(x_{i, control})-ln(x_{i, experiment}),
$$
where *x~i,control~* and *x~i,experiment~* are the mean pollutant concentrations for experiment *i*. 
The statistical properties of *lnRR* (normal distribution around zero and additive properties) are preferable to the more commonly reported efficiency metric [@osenbergEffectSizeEcological1997; @hedgesMetaanalysisResponseRatios1999]. 
Use of *ROM* required the exclusion of studies that only provided measures of BMP efficiency and not the underlying data used to derive the metric.

## Statistical models

We used the "rma.mv" function in the *metafor* R package to fit multilevel random effects linear models with *ROM* as the effect variable [@viechtbauerConductingMetaanalysesMetafor2010; @rcoreteamLanguageEnvironmentStatistical2023].
Our models specified a nested random effects term accounting for heterogeneity between effect sizes from the same study and for heterogeneity between studies.
A key feature of meta-analysis is the weighting of effects using sampling variance of individual effect sizes.
`r paste0(round(bac_missing_sd$missing / bac_missing_sd$total * 100, 0), "%")` of fecal indicator bacteria, `r paste0(round(tn_missing_sd$missing / tn_missing_sd$total * 100, 0), "%")` of TN, `r paste0(round(tp_missing_sd$missing / tp_missing_sd$total * 100, 0), "%")` of TP, and `r paste0(round(tss_missing_sd$missing / tss_missing_sd$total * 100, 0), "%")` of TSS effect sizes were missing standard deviations required to estimate sampling variance. 
Removal of studies due to missing variance information can introduce substantial bias [@kambachConsequencesMultipleImputation2020], so we imputed missing standard deviations using random forest based multivariate imputation with chained equations using the *mice* R package [@buurenMiceMultivariateImputation2011].
Sampling variance was estimated with the variance estimator described in Doncaster et al. to reduce bias in small sample studies:

$$
v(ROM) = \frac{\sum_{i=1}^{K}{(CV^2_{control,i})/K}}{n_{control}} + \frac{\sum_{i=1}^{K}{(CV^2_{experiment,i})/K}}{n_{experiment}},
$$

where *v* represents the sampling variance, $CV_{control,i}$ and $CV^2_{experiment,i}$ are the coefficients of variation from the *i*th study for studies 1, 2, ..., *K*.

Our initial models included log transformed pre-treatment pollutant concentration, study scale (plot/field, community, watershed), study length (*n* years), BMP subcategory (), and pollutant concentration:BMP subcategory interactions were included as fixed effect terms.
We used an information-theoretic approach to select the most parsimonious model from the subset of candidate models based on corrected Akaike information criterion (AIC~c~) estimated with maximum likelihood [@cinarUsingInformationTheoretic2021].
The final model was selected from candidate models, which included all combination and subsets of the full model, by eliminating candidate models until the $\Delta$*AIC~c~* reached $\leq$ 2 [@burnhamAICModelSelection2011].
We estimated regression coefficients of the selected model using restricted maximum likelihood.
Relative heterogeneity between and within studies were calculated using the *I^2^* metric described in @nakagawaMethodologicalIssuesAdvances2012.
Marginal *R^2^* was used to describe the amount of variance explained by fixed effects [@nakagawaGeneralSimpleMethod2013].

```{r}
bac_z <- eggers_bac$zval[[3]] |> prettyNum(digits = 3, nsmall = 2)
bac_p <- eggers_bac$pval[[3]] |> format.pval(eps = .01, digits = 2, nsmall = 1)
```

We tested for evidence of publication bias, in the form of small study effect, by using the extension of Egger's regression applied to the multilevel model framework that included adjusted sampling error as a moderator [@nakagawaQuantitativeEvidenceSynthesis2023]. 
We did not identify evidence of publication bias in the surveyed studies for fecal indicator bacteria (*Z* = `r bac_z`, *p* = `r bac_p`), total nitrogen, total phosphorus, or total suspended solids; therefore, adjustments for publication bias were not included in the final models.
We conducted a sensitivity analysis of the robustness of overall effect sizes to individual studies using leave-one-out analysis [@nakagawaQuantitativeEvidenceSynthesis2023].
Outlier studies identified in the sensitivity analysis are highlighted in Table X and were removed prior to fitting the final models.

<!-- Comments should be HTML-style comments -->

# Results

## Summary of BMP literature

## Regression results

Some evidence of a relationship between input concentrations and fecal indicator bacteria reductions.
No evidence strong evidence that the type of BMP changed the overall relationship.

```{r bacsummary}

tar_read(bac_table, store = store) |> 
  pack_rows(group_label = "BMP Subcategories",
            start_row = 2,
            end_row = 5)

```

```{r nitsummary}

tar_read(tn_table, store = store)

```

```{r phossummary}

tar_read(tp_table, store = store)

```

## Subsection 2

Frontiers requires figures to be submitted individually, in the same order as
they are referred to in the manuscript. Figures will then be automatically
embedded at the bottom of the submitted manuscript. Kindly ensure that each
table and figure is mentioned in the text and in numerical order. Permission
must be obtained for use of copyrighted material from other sources (including
the web). Please note that it is compulsory to follow figure instructions.
Figures which are not according to the guidelines will cause substantial delay
during the production process.

# Discussion

Linking BMP performance to watershed scale improvements remains a challenge (Tomer and Locke, Melland et al, Meals). 
The succesful implementation and maintence of BMPs at watershed scales remains an area for more investigation (Liu et al., Lintern) 
Long-term trans-disciplinary monitoring projects not just on BMP performance but maintance, management, and social apsects may elucidate unknown factors influencing the ability to scale BMP reductions to watershed scale improvements (Lintern et al.)

Runoff volume was not incorporated into our models, largely attributed to lack of available data. However, BMP size is typically scaled to accommodate runoff volumes of a particular return period volume (1-yr or 2-yr flows for example) (Hoss et al.).

# Data availability statement {-}

The data and R code used in this study are deposited in Zenodo: TBD.

# Disclosure/Conflict-of-Interest Statement {-}

The authors declare that the research was conducted in the absence of any
commercial or financial relationships that could be construed as a potential
conflict of interest.

# Author Contributions {-}

MS: Conceptualization, Formal analysis, Funding acquisition, Methodology, Supervision, Writing—original draft, Writing—review & editing.
DK: Conceptualization, Formal analysis, Data collection, Writing—original draft, Writing—review & editing.
JW: Data collection, Writing—review & editing.
SJ: Data collection; Writing—review & editing.

# Acknowledgments {-}

Funding: This project was supported by a state nonpoint source grant from the Texas State Soil and Water Conservation Board.

# Supplemental Data {-}

Supplementary Material should be uploaded separately on submission, if there are
Supplementary Figures, please include the caption in the same file as the
figure. LaTeX Supplementary Material templates can be found in the Frontiers
LaTeX folder

\bibliographystyle{Frontiers-Harvard}
\bibliography{bibliography.bib}

# Figure captions {-}

```{r studymap}
#| fig.cap: "Distribution of studies identified in the systematic review by state and parameter."
#| fig.pos: "H"
#| out.extra: ""
#| out.width: "100%"

knitr::include_graphics("../figures/study_map.pdf")

```

```{r bmpsummary}
#| fig.cap: "Summary of (A) study scale, (B) dominant runoff source, and (C) BMPs identified in the systematic review."
#| #| fig.pos: "H"
#| out.extra: ""
#| out.width: "100%"

knitr::include_graphics("../figures/bmp_summary.pdf")

```

```{r overalleffect}
#| fig.cap: "Estimated effect sizes and intervals from the null multilevel random effects model. Individual points represent studies, with size scaled by sampling variance. The point estimate with uncertaintity bars indicate the estimated overall effect, 95% confidence intervals, and 95% prediction intervals. Here, \\textit{k} indicates the number of overall effects with the number of unique studies in parenthesis."
#| fig.pos: "H"
#| out.extra: ""
#| out.width: "100%"

knitr::include_graphics("../figures/overall_effect.pdf")
```

<!-- #  -->
<!-- # ```{r marginalmeans} -->
<!-- # #| fig.cap: "Marginal effects of influent pollutant concentrations on the log ratio of means (\\textit{ROM}). Individual dots represent each effect size identified in the literature with the size scaled by sampling variance. The solid lines are the marginal mean effect, dashed lines are the 95% confidence interval around the mean, and the dotted lines are 95% prediction intervals. Here, *k* represents the number of effects with the number of studies in parenthesis." -->
<!-- # #| fig.pos: "H" -->
<!-- # #| out.extra: "" -->
<!-- # #| out.width: "100%" -->
<!-- #  -->
<!-- # knitr::include_graphics("../figures/marginal_means.pdf") -->
<!-- #  -->
<!-- #  -->
<!-- #``` -->

### 2024-01-17 frontiers_submission/frontiers_submission.tex
initial commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.   
%                                              %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.4 Generated 2022/06/14 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
%%%  NB logo1.jpg is required in the path in order to correctly compile front page header %%%

\documentclass[utf8]{FrontiersinHarvard}

%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}

\linenumbers

% BELOW TAKEN FROM rticles plos template
%
% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Below is from frontiers

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\def\Authors{
  Michael Schramm\,\textsuperscript{1*},
  Duncan Kikoyo\,\textsuperscript{1},
  Janelle Wright\,\textsuperscript{1},
  Shubham Jain\,\textsuperscript{1}}

\def\Address{

  \textsuperscript{1} Texas Water Resources Institute, Texas A\&M
AgriLife Research,  College Station,  Texas,  USA
  }

  \def\corrAuthor{Michael Schramm}\def\corrAddress{Texas Water Resources
Institute, Texas A\&M AgriLife Research\\1001 Holleman Dr E.\\College
Station, Texas, 77840 USA}\def\corrEmail{\href{mailto:michael.schramm@ag.tamu.edu}{\nolinkurl{michael.schramm@ag.tamu.edu}}}
  \def\firstAuthorLast{Schramm {et~al.}}

\begin{document}

\onecolumn
\firstpage{1}

\title[Meta-analysis of best management practices]{A meta-analysis of
the impacts of best management practices on nonpoint source pollutants.}
\author[\firstAuthorLast]{\Authors}
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}
Abstract length and content varies depending on article type. Refer to
\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract
requirement and length according to article type.

\tiny
 \keyFont{ \section{Keywords:} best management practice, water
quality, nonpoint source pollution, fecal indicator
bacteria, nutrients} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

In the United States (U.S.), major progress towards water quality goals
have been achieved through the Clean Water Act of 1972. This progress
has been largely attributed to investments and reductions in point
source discharges while reductions in nonpoint source pollutants remain
a substantial challenge
\citep{benhamLessonsLearnedTMDL2008, nationalresearchcouncilAssessingTMDLApproach2001, schrammTotalMaximumDaily2022}.
Increased pollutant loads and pollutant concentrations in runoff are a
particular challenge when dealing with nonpoint source pollution
resulting from land use change. The impacts of land use change on
hydrology and water quality are well established
\citep{allanLandscapesRiverscapesInfluence2004, carpenterNonpointPollutionSurface1998, bernhardtUnderstandingManagingMinimizing2008, careyEvaluatingNutrientImpacts2013, freemanImpactsUrbanizationDevelopment2019}.
Nonpoint source driven fecal indicator bacteria (FIB), nitrogen,
phosphorus, and suspended sediment remain major causes of water quality
impairments in U.S. rivers and streams despite decades of work. In 2017,
the Environmental Protection Agency (EPA) estimated 41\% or more of the
nation's rivers and streams rated poorly for biological condition due to
excess nitrogen or phosphorus \citep{epaNationalWaterQuality2017}. FIB
remains the leading cause of water body impairment on the Clean Water
Act 303(d) list in the United States
\citep{epaNationalWaterQuality2017}.

Best management practices (BMPs) have been the primary suite of tools
for addressing nonpoint source pollution. BMPs are structural or
non-structural controls used to mitigate the effects of increased runoff
volume, pollutant loads, or pollutant concentrations emanating from
diffuse nonpoint sources. BMPs control the delivery of pollutants
through a few possible mechanisms. Structural BMPs reduce and retard
total volume of runoff, thus reducing both the volume of water and
pollutant load. Structural BMPs may also provide a mechanism for
physical, chemical, or biological removal of pollutant constituents
suspended or dissolved in runoff. Non-structural BMPs (such as nutrient
management or livestock management) are utilized to reduce the
generation of pollutant runoff by avoiding pollutant generation during
critical periods.

Practitioners rely extensively on mechanistic models to plan and
evaluate BMP scenarios and resulting water quality.
\citet{linternBestManagementPractices2020} found 43\% of reviewed BMP
effectiveness studies relied completely on modeled outputs, with modeled
outputs almost always predicting water quality improvements following
BMP implementation. However, field studies are much more likely to
demonstrate mixed results including net releases of pollutants under
certain conditions
\citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017}.
The disconnect between modeled outcomes and field studies might be
attributed to (1) overly simplified or incorrect estimates of parameters
that represent management practices
\citep{ullrichApplicationSoilWater2009, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020},
(2) failure to incorporate uncertainty in estimates
\citep{tasdighiBayesianTotalUncertainty2018, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020},
and (3) assumption of static performance over time
\citep{mealsLagTimeWater2010, liuReviewEffectivenessBest2017, fuReviewCatchmentscaleWater2019}.

A source of uncertainty comes from the substantial variability of
performance metrics reported in empirical BMP studies
\citep{linternBestManagementPractices2020}. Previous empirical reviews
generally describe high variability and uncertainty in nitrogen and
phosphorus removal and consistent reduction in total suspended sediment
concentrations across BMP types
\citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017, kochNitrogenRemovalStormwater2014, claryBMPPerformanceAnalysis2011, barrettComparisonBMPPerformance2008, grudzinskiDoesRiparianFencing2020}.
The review literature on the effects of BMPs on FIBs are sparse but
generally find extremely high variance in performance across BMPs
\citep{claryBMPPerformanceAnalysis2011, grudzinskiDoesRiparianFencing2020}.
Accurately characterizing BMP treated runoff quality is challenging due
to the high variability in BMP performance. However, influent water
quality, seasonality, and BMP age provide some explanatation to highliy
heterogenous BMP efficiency data
\citep{barrettPerformanceComparisonStructural2005, liuReviewEffectivenessBest2017}.\\
A recent meta-analysis demonstrated the further linkages between local
climatic conditions and BMP performance on nitrogen and phosphorus
concentrations \citep{horvathEffectsRegionalClimate2023}.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

We conducted a systematic review of recent (2000-2022) literature to
compile field studies documenting the effectiveness of best management
practices on fecal indicator bacteria and nutrients concentrations. The
systematic review followed guidance provided in the Collaboration for
Environmental Evidence systematic review guidelines
\citep{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}.
In order to maximize the number of studies included in the review, we
included both peer-reviewed studies and unpublished white papers to
reduce potential bias against negative results. The inclusion criteria
filtered out (1) non-field studies, (2) modelling results, (3) studies
that did not evaluate specific BMPs, (4) studies conducted outside of
the U.S. or published in a language other than English. We ran search
queries in Texas A\&M Library Catalog, Web of Science, and Google
Scholar. Although results from Google Scholar are not always replicable,
we utilized the service to maximize search results for studies not
published in academic journals and presumably increase the chance of
identifying studies with negative effects. Fecal indicator bacteria
study seraches included the following query: ``fecal indicator
bacteria'' OR ``E. coli'' OR ``Escherichia coli'' OR ``enterococci'' OR
``enterococcus'' AND ``best management practices'' OR ``BMPs'' AND
``effectiveness'' OR ``performance''. Nutrient BMP studies utilized a
similar query: ``nutrient'' OR ``nitrogen'' OR ``phosphorus'' OR
``sediment'' OR ``TSS'' AND ``best management practices'' OR ``BMPs''
AND ``effectiveness'' OR ``performance''.

Results from each database were first filtered to remove duplicates.
After removal of duplicates, each member of the research team (\emph{n}
= 4) was assigned a subset of studies to evaluate if they should be
included. \textbf{include supplementary tables with inclusion criteria
and data variables} Each study was reviewed by two team members and
differences in opinion were collectively discussed and agreed upon
before progressing. The remaining studies were split among team members
for data extraction, again with at least two team members reviewing each
study. If data was provided in figures, the data was extracted with the
WebPlotDigitizer tool \citep{rohatgiWebPlotDigitizer2022}.

\begin{table}

\caption{\label{tab:criteria}Criteria applied for including or excluding studies within the review database.}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5em}>{\raggedright\arraybackslash}p{15em}>{\raggedright\arraybackslash}p{15em}}
\toprule
Attribute & Inclusion Criteria & Excluision Criteria\\
\midrule
Study type & Journal articles, book chapters, conference papers, unpublished research reports, thesis and dissertations, organizational and agency white papers. & Synopsis or review studies, peports with reductions based on modeled or other estimated reductions (e.g. TMDLs, watershed plans, or modelling studies.\\
Outcomes & Field studies with measured effects on fecal indicator bacteria or nutrient concentrations. & Studies not explicitly linking reductions to a specific BMP or insufficient information to quantify reductions.\\
Geographical context & Studies conducted within the United States. & Studies outside of the United States.\\
Timeframe & Studies published from 2000 through 2022. & Studies published prior to 2000 or after 2022.\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:datvars}Study and effect variables extracted for review.}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{25em}}
\toprule
Variable & Description\\
\midrule
Publication Year & Year the study was published\\
Parameter & The specific pollutant measured.\\
Runoff source & Dominant source of runoff (rop fields, livestock pasture, commerical, residential\\
Source type & Major categorization of runoff source: agricultural or urban\\
BMP & BMP evaluated\\
\addlinespace
BMP Classification & BMP description based on NRCS conservation practice standards and EPA BMP fact sheets\\
BMP Category & BMP categorization based on structural or management\\
BMP Subcategory & BMP subcategorization based on pollutant removal processes\\
Study scale & Spatial scale of the study area ("lot/field", "community", "watershed")\\
Location & Location name used in the study description\\
\addlinespace
State & State where the study was conducted\\
Study area & Drainage area in hectares\\
Longitude & Approximated or reported latitude coordinate\\
Latitude & Approximated or reported longitude coordinate\\
Study years & Year or years when data were collected\\
\addlinespace
N control & Number of control measurements\\
N experiment & Number of experiemental measurements\\
X control & Mean concentration for control measurements\\
X experiment & Mean concentration for experiemental measurements\\
SE control & Standard error of control measurements\\
\addlinespace
SE experiment & Standard error of experimental measurements\\
Minimum control & Minimum control measurement\\
Minimum experiment & Minimum experiment measurement\\
Maximum control & Maximum control measurement\\
Maximum experiment & Maximum experiment measurement\\
\addlinespace
SD control & Standard deviation of control measreuments\\
SD experiment & Standard deviation of experimental measurements\\
Units & Units reported by the study\\
Percent reduction & BMP efficiency for studies that only reported efficiency\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{effect-size-calculation}{%
\subsection*{Effect size calculation}\label{effect-size-calculation}}
\addcontentsline{toc}{subsection}{Effect size calculation}

Prior BMP reviews routinely report the effect of BMPs using efficiency
values
\citep[see][]{agouridisLivestockGrazingManagement2005, krogerReviewBestManagement2012, liuEnhancingRainfallrunoffModel2015, simpsonDevelopingBestManagement2009}.
Efficiency values are calculated as:

\[
\text{Efficiency}=\frac{x_{control}-x_{experiment}}{x_{control}}\times 100,
\]

where \emph{x\textsubscript{control}} is the pre-treatment or control
pollutant concentration and \emph{x\textsubscript{experiment}} is the
pollutant concentration measured after the BMP intervention. There are
several statistical shortcomings (distributional asymetry, skewness, and
non-additive properties) when using efficiency to estimate overall
effect sizes
\citep{nuzzoPercentDifferencesAnother2018, coleStatisticsNotesWhat2017}.
In comparision, the log ratio of means (\emph{ROM}) provide preferable
statistical properties for regression analysis
\citep{osenbergEffectSizeEcological1997, hedgesMetaanalysisResponseRatios1999}.
\emph{ROM} quantifies the difference in means between the control and
experimental group \citep{hedgesMetaanalysisResponseRatios1999}:

\[
ROM_i = ln\left(\frac{x_{i,control}}{x_{i,experiment}}\right) = ln(x_{i, control})-ln(x_{i, experiment}),
\] where \emph{x\textsubscript{i,control}} and
\emph{x\textsubscript{i,experiment}} are the mean pollutant
concentrations for experiment \emph{i}. The statistical properties of
\emph{lnRR} (normal distribution around zero and additive properties)
are preferable to the more commonly reported efficiency metric
\citep{osenbergEffectSizeEcological1997, hedgesMetaanalysisResponseRatios1999}.
Use of \emph{ROM} required the exclusion of studies that only provided
measures of BMP efficiency and not the underlying data used to derive
the metric.

\hypertarget{statistical-models}{%
\subsection{Statistical models}\label{statistical-models}}

We used the ``rma.mv'' function in the \emph{metafor} R package to fit
multilevel random effects linear models with \emph{ROM} as the effect
variable
\citep{viechtbauerConductingMetaanalysesMetafor2010, rcoreteamLanguageEnvironmentStatistical2023}.
Our models specified a nested random effects term accounting for
heterogeneity between effect sizes from the same study and for
heterogeneity between studies. A key feature of meta-analysis is the
weighting of effects using sampling variance of individual effect sizes.
45\% of fecal indicator bacteria, 68\% of TN, 68\% of TP, and 70\% of
TSS effect sizes were missing standard deviations required to estimate
sampling variance. Removal of studies due to missing variance
information can introduce substantial bias
\citep{kambachConsequencesMultipleImputation2020}, so we imputed missing
standard deviations using random forest based multivariate imputation
with chained equations using the \emph{mice} R package
\citep{buurenMiceMultivariateImputation2011}. Sampling variance was
estimated with the variance estimator described in Doncaster et al.~to
reduce bias in small sample studies:

\[
v(ROM) = \frac{\sum_{i=1}^{K}{(CV^2_{control,i})/K}}{n_{control}} + \frac{\sum_{i=1}^{K}{(CV^2_{experiment,i})/K}}{n_{experiment}},
\]

where \emph{v} represents the sampling variance, \(CV_{control,i}\) and
\(CV^2_{experiment,i}\) are the coefficients of variation from the
\emph{i}th study for studies 1, 2, \ldots, \emph{K}.

Our initial models included log transformed pre-treatment pollutant
concentration, study scale (plot/field, community, watershed), study
length (\emph{n} years), BMP subcategory (), and pollutant
concentration:BMP subcategory interactions were included as fixed effect
terms. We used an information-theoretic approach to select the most
parsimonious model from the subset of candidate models based on
corrected Akaike information criterion (AIC\textsubscript{c}) estimated
with maximum likelihood \citep{cinarUsingInformationTheoretic2021}. The
final model was selected from candidate models, which included all
combination and subsets of the full model, by eliminating candidate
models until the \(\Delta\)\emph{AIC\textsubscript{c}} reached \(\leq\)
2 \citep{burnhamAICModelSelection2011}. We estimated regression
coefficients of the selected model using restricted maximum likelihood.
Relative heterogeneity between and within studies were calculated using
the \emph{I\textsuperscript{2}} metric described in
\citet{nakagawaMethodologicalIssuesAdvances2012}. Marginal
\emph{R\textsuperscript{2}} was used to describe the amount of variance
explained by fixed effects \citep{nakagawaGeneralSimpleMethod2013}.

We tested for evidence of publication bias, in the form of small study
effect, by using the extension of Egger's regression applied to the
multilevel model framework that included adjusted sampling error as a
moderator \citep{nakagawaQuantitativeEvidenceSynthesis2023}. We did not
identify evidence of publication bias in the surveyed studies for fecal
indicator bacteria (\emph{Z} = -1.17, \emph{p} = 0.25), total nitrogen,
total phosphorus, or total suspended solids; therefore, adjustments for
publication bias were not included in the final models. We conducted a
sensitivity analysis of the robustness of overall effect sizes to
individual studies using leave-one-out analysis
\citep{nakagawaQuantitativeEvidenceSynthesis2023}. Outlier studies
identified in the sensitivity analysis are highlighted in Table X and
were removed prior to fitting the final models.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{summary-of-bmp-literature}{%
\subsection{Summary of BMP literature}\label{summary-of-bmp-literature}}

\hypertarget{regression-results}{%
\subsection{Regression results}\label{regression-results}}

Some evidence of a relationship between input concentrations and fecal
indicator bacteria reductions. No evidence strong evidence that the type
of BMP changed the overall relationship.

\begin{table}

\caption{Summary table of moderator effects for performance of BMPs on fecal indicator bacteria concentrations.}
\centering
\begin{tabular}[t]{lrlrrrl}
\toprule
Moderator & \makecell[r]{Estimate\\(\textit{lnRR})} & 95\% CI & SE & \textit{T}-statistic & df & \textit{p}-value\\
\midrule
Intercept & 0.21 & {}[-2.07,2.48] & 1.06 & 0.19 & 14 & 0.85\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{BMP Subcategories}}\\
\hspace{1em}Filtration & -0.39 & {}[-6.42,5.63] & 2.81 & -0.14 & 14 & 0.89\\
\hspace{1em}Infiltration & -1.24 & {}[-3.50,1.01] & 1.13 & -1.10 & 64 & 0.28\\
\hspace{1em}Livestock & -1.18 & {}[-3.25,0.89] & 0.96 & -1.23 & 14 & 0.24\\
\hspace{1em}Treatment & -1.56 & {}[-2.63,-0.49] & 0.53 & -2.92 & 64 & \textless 0.01\\
log(Concentration) & 0.20 & {}[0.084,0.32] & 0.059 & 3.42 & 65 & \textless 0.01\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}$I^{2}_{\text{total}}$=21.48, $I^{2}_{\text{study}}$=0, $I^{2}_{\text{effect}}$=21.48; $R^{2}_{\text{marginal}}$=0.46}\\
\end{tabular}
\end{table}

\begin{table}

\caption{Summary table of moderator effects for performance of BMPs on nitrogen removal.}
\centering
\begin{tabular}[t]{lrlrrrl}
\toprule
Moderator & \makecell[r]{Estimate\\(\textit{lnRR})} & 95\% CI & SE & \textit{T}-statistic & df & \textit{p}-value\\
\midrule
Intercept & 0.95 & {}[0.38,1.52] & 0.28 & 3.41 & 28 & \textless 0.01\\
log(Concentration) & -0.034 & {}[-0.14,0.071] & 0.053 & -0.64 & 124 & 0.525\\
Adjusted sampling error & -1.54 & {}[-3.30,0.22] & 0.89 & -1.74 & 124 & 0.085\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}$I^{2}_{\text{total}}$=96.57, $I^{2}_{\text{study}}$=52.45, $I^{2}_{\text{effect}}$=44.12; $R^{2}_{\text{marginal}}$=0.05}\\
\end{tabular}
\end{table}

\begin{table}

\caption{Summary table of moderator effects for performance of BMPs on phosphorus removal.}
\centering
\begin{tabular}[t]{lrlrrrl}
\toprule
Moderator & \makecell[r]{Estimate\\(\textit{lnRR})} & 95\% CI & SE & \textit{T}-statistic & df & \textit{p}-value\\
\midrule
Intercept & 0.40 & {}[-0.14,0.93] & 0.26 & 1.52 & 28 & 0.14\\
log(Concentration) & 0.27 & {}[0.15,0.40] & 0.063 & 4.30 & 97 & \textless 0.01\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}$I^{2}_{\text{total}}$=15.54, $I^{2}_{\text{study}}$=0, $I^{2}_{\text{effect}}$=15.54; $R^{2}_{\text{marginal}}$=0.32}\\
\end{tabular}
\end{table}

\hypertarget{subsection-2}{%
\subsection{Subsection 2}\label{subsection-2}}

Frontiers requires figures to be submitted individually, in the same
order as they are referred to in the manuscript. Figures will then be
automatically embedded at the bottom of the submitted manuscript. Kindly
ensure that each table and figure is mentioned in the text and in
numerical order. Permission must be obtained for use of copyrighted
material from other sources (including the web). Please note that it is
compulsory to follow figure instructions. Figures which are not
according to the guidelines will cause substantial delay during the
production process.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Linking BMP performance to watershed scale improvements remains a
challenge (Tomer and Locke, Melland et al, Meals). The succesful
implementation and maintence of BMPs at watershed scales remains an area
for more investigation (Liu et al., Lintern) Long-term
trans-disciplinary monitoring projects not just on BMP performance but
maintance, management, and social apsects may elucidate unknown factors
influencing the ability to scale BMP reductions to watershed scale
improvements (Lintern et al.)

Runoff volume was not incorporated into our models, largely attributed
to lack of available data. However, BMP size is typically scaled to
accommodate runoff volumes of a particular return period volume (1-yr or
2-yr flows for example) (Hoss et al.).

\hypertarget{data-availability-statement}{%
\section*{Data availability
statement}\label{data-availability-statement}}
\addcontentsline{toc}{section}{Data availability statement}

The data and R code used in this study are deposited in Zenodo: TBD.

\hypertarget{disclosureconflict-of-interest-statement}{%
\section*{Disclosure/Conflict-of-Interest
Statement}\label{disclosureconflict-of-interest-statement}}
\addcontentsline{toc}{section}{Disclosure/Conflict-of-Interest
Statement}

The authors declare that the research was conducted in the absence of
any commercial or financial relationships that could be construed as a
potential conflict of interest.

\hypertarget{author-contributions}{%
\section*{Author Contributions}\label{author-contributions}}
\addcontentsline{toc}{section}{Author Contributions}

MS: Conceptualization, Formal analysis, Funding acquisition,
Methodology, Supervision, Writing---original draft, Writing---review \&
editing. DK: Conceptualization, Formal analysis, Data collection,
Writing---original draft, Writing---review \& editing. JW: Data
collection, Writing---review \& editing. SJ: Data collection;
Writing---review \& editing.

\hypertarget{acknowledgments}{%
\section*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{section}{Acknowledgments}

Funding: This project was supported by a state nonpoint source grant
from the Texas State Soil and Water Conservation Board.

\hypertarget{supplemental-data}{%
\section*{Supplemental Data}\label{supplemental-data}}
\addcontentsline{toc}{section}{Supplemental Data}

Supplementary Material should be uploaded separately on submission, if
there are Supplementary Figures, please include the caption in the same
file as the figure. LaTeX Supplementary Material templates can be found
in the Frontiers LaTeX folder

\bibliographystyle{Frontiers-Harvard}
\bibliography{bibliography.bib}

\hypertarget{figure-captions}{%
\section*{Figure captions}\label{figure-captions}}
\addcontentsline{toc}{section}{Figure captions}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../figures/study_map} \caption{Distribution of studies identified in the systematic review by state and parameter.}\label{fig:studymap}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth,]{../figures/bmp_summary} \caption{Summary of (A) study scale, (B) dominant runoff source, and (C) BMPs identified in the systematic review.}\label{fig:bmpsummary}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../figures/overall_effect} \caption{Estimated effect sizes and intervals from the null multilevel random effects model. Individual points represent studies, with size scaled by sampling variance. The point estimate with uncertaintity bars indicate the estimated overall effect, 95\% confidence intervals, and 95\% prediction intervals. Here, \textit{k} indicates the number of overall effects with the number of unique studies in parenthesis.}\label{fig:overalleffect}
\end{figure}

\end{document}

### 2024-01-17 manuscript/manuscript.Rmd
initial commit

---
title: "A meta-analysis of best management practice effects on nonpoint source pollutants."
author:
  - Michael Schramm ^1^, Duncan Kikoyo ^2^, Janelle Wright ^3^, Shubham Jain ^4^
  - ^1^ Research Specialist IV, Texas Water Resources Institute, Texas A\&M AgriLife Research
  - ^2^ Research Specialist III, Texas Water Resources Institute, Texas A\&M AgriLife Research
  - ^3^ Graduate Research Assistant, Texas Water Resources Institute, Texas A\&M AgriLife Research
  - ^4^ Graduate Research Assistant, Texas Water Resources Institute, Texas A\&M AgriLife Research
date: "April 1, 2021"
output: 
  officedown::rdocx_document:
    reference_docx: format.docx
    plots:
      style: Normal
      align: center
      fig.lp: 'fig:'
      topcaption: false
      caption:
        style: Image Caption
        pre: 'Figure '
        sep: '. '
        tnd: 0
        tns: '-'
        fp_text: !expr officer::fp_text_lite(bold = FALSE)
    tables:
      style: Table
      layout: autofit
      width: 1.0
      topcaption: true
      tab.lp: 'tab:'
      caption:
        style: Table Caption
        pre: 'Table '
        sep: '. '
        tnd: 0
        tns: '-'
        fp_text: !expr officer::fp_text_lite(bold = FALSE)
bibliography: bibliography.bib
csl: chicago-author-date.csl
abstract: |
  Lorem ipsum

keywords: |
  best management practice; water quality
---

```{r setup, include=FALSE}
## load libraries
library(officer)
library(officedown)
library(ggtext)
library(tidyverse)
library(flextable)
library(ragg)
library(targets)

## this sets our default code chunk options
knitr::opts_chunk$set(dev = "ragg_png",
                      echo = FALSE,
                      ## figure chunk options
                      fig.cap = TRUE,
                      fig.width = 6,
                      dpi = 300)

## make fonts available
font_hoist <- function(family, silent = FALSE) {
  font_specs <- systemfonts::system_fonts() |> 
    filter(family == {{family}}) |> 
    mutate(family = paste(family, style)) |> 
    select(plain = path, name = family)

  purrr::pwalk(as.list(font_specs), systemfonts::register_font)

  if (!silent)  message(paste0("Hoisted ", nrow(font_specs), " variants:\n",
                               paste(font_specs$name, collapse = "\n")))
}

font_hoist("Manrope")

## grab output from targets to use in manuscript
store <- "../_targets"
bac_missing_sd <- tar_read(bac_missing_sd,
                           store = store)

```

```{r}
## leave in, ends word document section
## after table of abbreviations
block_section(prop_section(type = "nextPage"))
```

# Introduction {-}

The impact of land use change on hydrology and water quality is well established.
Best management practices (BMPs) are control measures used to mitigate increased runoff volume or pollutant concentrations resulting from diffuse nonpoint sources.
A suite of empirical approaches and mechanistic watershed models have been developed for practitioners to develop planning scenarios and forecast the hydrologic and water quality impacts of BMP implementation from site to basin scales.
Resource managers depend on these models and global databases, such as the International Stormwater Database, to understand the performance of BMPs and plan their design and implementation within impaired watersheds.
Despite progress in the development and implementation of BMPs, nonpoint source driven impairments remain a substantial challenge in the United States.
Fecal indicator bacteria (FIB) and nutrients, largely driven by nonpoint sources, remain the major cause of water quality impairments in the nation's rivers and streams despite decades of work.
In 2017, the Environmental Protection Agency (EPA) estimated 41% or more of the nation's rivers and streams rated poorly for biological condition due to excess nitrogen or phosphorus [@epaNationalWaterQuality2017].

# Methods {-}

## Review and data extraction {-}

We conducted a systematic review of recent (2000-2022) literature to compile field studies documenting the effectiveness of best management practices on fecal indicator bacteria and nutrients concentrations. 
The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines [@collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018].
In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
The inclusion criteria filtered out (1) non-field studies, (2) modelling results, (3) studies that did not evaluate specific BMPs, (4) studies conducted outside of the U.S. or published in a language other than English.
We ran search queries in Texas A&M Library Catalog, Web of Science, and Google Scholar.
Although results from Google Scholar are not always replicable, we utilized the service to maximize search results for studies not published in academic journals.

Results from each database were first filtered to remove duplicates.
After removal of duplicates, each member of the research team (*n* = 4) was assigned a subset of studies to evaluate if they should be included.
**include supplementary tables with inclusion criteria and data variables**
Each study was reviewed by two team members and differences in opinion were collectively discussed and agreed upon before progressing.
The remaining studies were split among team members for data extraction, again with at least two team members reviewing each study. 
If data was provided in figures, the data was extracted with the WebPlotDigitizer tool [@rohatgiWebPlotDigitizer2022].

## Meta-analysis {-}

Robust meta-analysis produces estimates of overall effect size of interventions by weighting individual studies using samples sizes or uncertainty estimates in the calculation of overall effect size [@hedgesMetaanalysisResponseRatios1999].
Prior BMP reviews routinely report measures of central tendency (mean or median) of BMP efficiency values [see @agouridisLivestockGrazingManagement2005; @krogerReviewBestManagement2012; @liuEnhancingRainfallrunoffModel2015; @simpsonDevelopingBestManagement2009].
Efficiency values are calculated as:

$$
\text{Efficiency}=\frac{x_{control}-x_{experiment}}{x_{control}}\times 100
$$

where *x~control~* is the pre-treatment or control pollutant concentration and *x~experiment~* is the pollutant concentration measured after the BMP intervention. 
The resulting efficiency or percent change values are asymmetric, skewed, and are not additive [@nuzzoPercentDifferencesAnother2018; @coleStatisticsNotesWhat2017].
Log response ratios (*R*), commonly used in ecological meta-analysis, quantifies the difference in means between the control and experimental group [@hedgesMetaanalysisResponseRatios1999]:

$$
R_i = ln\left(\frac{x_{i,control}}{x_{i,experiment}}\right) = ln(x_{i, control})-ln(x_{i, experiment})
$$
where *x~i,control~* and *x~i,experiment~* are the mean pollutant concentrations for experiment *i*. 
The statistical properties of the log response ratio (normal distribution around zero and additive properties) are preferable to the more commonly reported efficiency metric [@osenbergEffectSizeEcological1997; @hedgesMetaanalysisResponseRatios1999]. 
This required the exclusion of studies that only provided measures of BMP efficiency and not the underlying data used to derive the metric.

The application of overall BMP efficiency values have been shown to result in inaccurate characterization of treated runoff quality due primarily to the fact that efficiency is a largely a function of influent pollutant concentration [@barrettPerformanceComparisonStructural2005]. 
In general, higher influent concentrations yield higher efficiencies while relatively clean water yield lower efficiencies.
We explored relationships between *R* and influent concentration using meta-analytic multi-level random effects models using the "metafor" package [@viechtbauerConductingMetaanalysesMetafor2010] in R [@rcoreteamLanguageEnvironmentStatistical2023].

A key feature of meta-analysis is the weighting of effects based on sampling variance.
`r paste0(round(bac_missing_sd$missing / bac_missing_sd$total * 100, 0), "%")` of effect sizes for fecal indicator bacteria were missing standard deviation required to estimate sampling variance. 
Sampling variance for missing cases was imputed via the weighted average coefficient of variation within groups to avoid bias associated with excluding incomplete cases [@nakagawaRobustReadilyImplementable2023; @kambachConsequencesMultipleImputation2020].
Log transformed pre-treatment pollutant concentration, BMP category, and interactions were included as fixed effects in the regression model.
We also specified a between study random effects term and within study random effects term to account for non-independence of multiple effect sizes report within individual studies.
The design of many BMP experiments (pre-post or upstream-downstream studies) are prone to statistical non-independence that inflates Type I error.
To account for statistical non-independence among sampling errors we applied robust variance estimation (RVE) methods to obtain adjusted standard errors and confidence intervals [@hedgesRobustVarianceEstimation2010].
Relative heterogeneities between and within studies were calculated using the *I^2^* metric described in @nakagawaMethodologicalIssuesAdvances2012.
Marginal *R^2^* was used to describe the amount of variance explained by the independent variables [@nakagawaGeneralSimpleMethod2013].
We tested for evidence of publication bias, in the form of small study effect, by using the extension of Egger's regression applied to the multilevel model framework that included adjusted sampling error as a moderator [@nakagawaQuantitativeEvidenceSynthesis2023]. 
We did not find evidence of publication bias in the surveyed studies for fecal indicator bacteria (include z stat and p value), total nitrogen, total phosphorus, or total suspended solids. 
Adjustments for publication bias were not included in the final models.
We conducted a sensitivity analysis of the overall effect sizes to individual studies using leave-one-out analysis [@nakagawaQuantitativeEvidenceSynthesis2023].
Outlier studies identified in the sensitivity analysis are highlighted in Table X and were removed prior to fitting the final models.

# Results {-}

```{r marginalmeans, fig.width=8, fig.height=8, fig.cap="Marginal means....", out.width="81%", dpi=300}
tar_read(marginal_means,
         store = store)
```

# Discussion {-}

## Study limitations {-}

**probably goes in discussion**
Our analysis was constrained by the lack of reported sampling variation and potential confounding variables such as watershed conditions or climatic variables that can serve as independent variables in a meta-analysis context.
@eagleMetaAnalysisConstrained2017 reported that five independent teams of researchers identified similar constraints when conducting a meta-analysis of nutrient management BMP studies.
The lack of consistent study design and reporting standards presents a significant challenge for pooling results from different studies.
Reporting requirements that facilitate meta-analysis are much more common in epidemiology and human health studies where meta-analysis is routinely conducted. 
The adoption of meta-analysis in fields such as ecological science is more recent. 
As meta-analysis becomes increasingly utilized in fields such as hydrology and watershed science, we hope that field coalesces around a standard requirement of data required for publication.**probably goes in discussion**

# Conclusion {-}

# Bibliography {-}

<div id="refs"></div>

```{r}
## leave in, ends word document section
## after table of abbreviations
block_section(prop_section(type = "nextPage"))
```

### 2024-01-17 R/01_read_data.R

@@ -1,0 +1,51 @@
+## Read data
+
+
+read_data <- function(path, sheet) {
+  
+  df <- read_xlsx(path, sheet, na = "N/A") |>
+    mutate(Longitude = as.numeric(Longitude),
+           Latitude = as.numeric(Latitude),
+           Pre = as.numeric(Pre),
+           Post = as.numeric(Post),
+           log_Pre = log(Pre),
+           log_Post = log(Post),
+           StudyID = as.factor(StudyID))
+  
+  ##calculate study length based on number of years in "Data_Year"
+  tlength <- lapply(strsplit(df$Data_Year,
+                             '\\s*;\\s*'),
+                    function(x) unlist(lapply(strsplit(x,'-'), function(y){ z <- as.integer(y); if (length(z)==1L) z else z[1L]:z[2L]; })))
+  
+  df <- df |> 
+    mutate(study_length = tlength) |> 
+    mutate(study_length = map_dbl(study_length,
+                                  \(x) length(x)))
+  
+  
+  return(df)
+  
+}
+
+load_df <- function(dataset) {
+  if(dataset == "bacteria") {
+    path <- "data/NPS-BMP-DB-Part1-v1.05.xlsx"
+    sheet <- "FIB_Database"
+    df <- read_data(path, sheet) |> 
+      filter(Unit != "% Exceeded standard (2000/100ml)",
+             Unit != "% Exceeded standard (200/100ml)") |> 
+      filter(!(Parameter %in% c("Streptococci")),
+             BMP != "No BMP")
+  } else {
+    path <- "data/NPS-BMP-DB-Part2-v1.05.xlsx"
+    sheet <- "Database"
+    df <- read_data(path, sheet) |> 
+      mutate(Area = case_when(
+        Area == "NA" ~ NA,
+        Area != "NA" ~ Area
+      )) |>
+      mutate(Area = as.numeric(Area)) |> 
+      filter(Unit == "mg/L")
+  }
+  return(df)
+}

### 2024-01-17 R/02_fit_models.R

@@ -1,0 +1,613 @@
+
+
+#' fit random effect meta regression to bacteria data
+#'
+#' @param data dataframe with `StudyID`, `Study`, `ValueID`, `Source_Type`,
+#'   `BMP_SubCat`, `n_Pre`, `n_Post`, `Pre`, `Post`, `sd_Pre`, `sd_Post`,
+#'   `Unit`. Or the output from `load_df()`
+#'
+#' @return list with model output of class `rma.mv` and dataframe with formatted
+#'   data used to fit the model.
+fit_bac_model <- function(data) {
+  df <- data |> 
+    select(StudyID, Study, ValueID, Parameter, Source_Type, BMP_SubCat, Scale, study_length, n_Pre, n_Post, Pre, Post, sd_Pre, sd_Post, Unit) |> 
+    filter(Unit %in% c("CFU/100ml", "MPN/100ml", "Geometric mean (CFU/100ml)", "Geomean (CFU/100 mL)")) |> 
+    filter(!is.na(Scale)) |> 
+    filter(!is.na(Pre), !is.na(Post)) |> 
+    filter(BMP_SubCat != "Drainage") |> # only one drainage observations, can't model the drainage x lnPre interaction.
+    mutate(StudyID = fct_drop(StudyID),
+           Study = fct_drop(Study),
+           lnPre = log(Pre),
+           BMP_SubCat = fct_drop(BMP_SubCat),
+           missing_sd = is.na(sd_Pre),
+           Scale = forcats::fct_relevel(Scale, "Lot", "Community", "Watershed"))
+  
+  ##impute SD
+  predMatrix <- make.predictorMatrix(df)
+  predMatrix[,"StudyID"] <- 0
+  predMatrix[,"Study"] <- 0
+  predMatrix[,"ValueID"] <- 0
+  predMatrix[,"Parameter"] <- 0
+  impMethod <- make.method(df)
+  impMethod[c("sd_Pre", "sd_Post")] <- "rf"
+  
+  imp <- mice(df, print=FALSE, m=20, predictorMatrix=predMatrix, method=impMethod, seed=1234)
+  df <- complete(imp)
+
+  
+  ## Prep data for rma
+  df <- escalc(measure = "ROM",
+               vtype = "AV",
+               m1i = Pre,
+               m2i = Post,
+               n1i = n_Pre,
+               n2i = n_Post,
+               sd1i = sd_Pre,
+               sd2i = sd_Post,
+               slab = Study,
+               data = df)
+  
+
[…]

### 2024-01-17 R/03_plot_models.R

@@ -1,0 +1,777 @@
+## load custom fonts
+## https://yjunechoe.github.io/posts/2021-06-24-setting-up-and-debugging-custom-fonts/
+font_hoist <- function(family, silent = FALSE) {
+  font_specs <- systemfonts::system_fonts() |> 
+    dplyr::filter(family == {{family}}) |> 
+    dplyr::mutate(family = paste(family, style)) |> 
+    dplyr::select(plain = path, name = family)
+  
+  purrr::pwalk(as.list(font_specs), systemfonts::register_font)
+  
+  if (!silent)  message(paste0("Hoisted ", nrow(font_specs), " variants:\n",
+                               paste(font_specs$name, collapse = "\n")))
+}
+
+
+
+
+
+#' Plots the estimated marginal means from the model with bubble_plot
+#'
+#' @param model rma model object
+#' @param xlab character object with the x-axis label
+#' @param ... additional arguments for bubble plot
+#'
+#' @return ggobject
+
+plot_marginal_effect_lnRR <- function(model,
+                                      xlab,
+                                      ylab,
+                                      mod = "lnPre",
+                                      by = "BMP_SubCat",
+                                      at = NULL,
+                                      subset = NULL,
+                                      sec_y_breaks = waiver(),
+                                      sec_y_labels = waiver(),
+                                      ...) {
+  
+  ## use mod_results from orchaRd to generate model grid 
+  mod_table <- mod_results(model, 
+                           mod = mod, 
+                           group = "ValueID", 
+                           weights = "prop", 
+                           by = by,
+                           at = at,
+                           subset = subset)
+  
+  ## back transform lnPre for plotting purposes
+  ##would like to back transform the moderator
+  if(mod == "lnPre"){
+    mod_table <- as.list(mod_table) |> 
+      purrr::modify_at("mod_table",
+                       ~mutate(.,
+                               moderator = exp(moderator))) |> 
+      purrr::modify_at("data",
+                       ~mutate(.,
+                               moderator = exp(moderator)))    
+  }
+
+  
+  ## remove NA values from "data"
+  if(!is.null(at)){
+    mod_table <- mod_table |> 
+      purrr::modify_at("data",
[…]

### 2024-01-17 R/04_tables.R

@@ -1,0 +1,70 @@
+## create a flextable output
+tidy_metafor <- function(model, 
+                         conf.level = 0.95, 
+                         terms = NULL,
+                         i2 = NULL,
+                         r2 = NULL,
+                         caption = NULL) {
+  
+  s_data <- summary(model)
+  
+  term_tab <- broom::tidy(model,
+                          conf.int = TRUE,
+                          conf.level = conf.level) |> 
+    mutate(estimate = prettyNum(estimate, digits = 2, nsmall = 2),
+           std.error = prettyNum(std.error, digits = 2, nsmall = 2),
+           statistic = prettyNum(statistic, digits = 2, nsmall = 2),
+           df = s_data$ddf,
+           `95\\% CI` = paste0("[",
+                             prettyNum(conf.low, digits = 2, nsmall = 2), 
+                             ",", 
+                             prettyNum(conf.high, digits = 2, nsmall = 2),
+                             "]"),
+           `p-value`= format.pval(p.value, eps = .01, digits = 2, nsmall = 1)) |>
+    mutate(`p-value` = gsub('<', '\\\\textless ', `p-value`)) |> 
+    select(Term = term,
+           `Estimate\\nLog response ratio` = estimate,
+           `95\\% CI`,
+           SE = std.error,
+           statistic,
+           df,
+           `\\textit{p}-value` = `p-value`)
+  
+  if(!is.null(terms)) {
+    term_tab <- term_tab |> 
+      mutate(Term = terms)
+
+  }
+  
+  kableExtra::kbl(term_tab,
+                  format = "latex",
+                  escape = FALSE,
+                  booktabs = TRUE,
+                  caption = caption,
+                  col.names = linebreak(c("Moderator", 
+                                          "Estimate\n(\\textit{lnRR})",
+                                          "95\\% CI",
+                                          "SE",
+                                          "\\textit{T}-statistic",
+                                          "df",
+                                          "\\textit{p}-value"),
+                                        align = c("l", "r", "l", "r", "r", "r")),
+                  align = c("l", "r", "l", "r", "r", "r")
+                  )|>
+    # add_indent(c(5:7, 9:10, 12:15)) |> 
+    footnote(general = paste0("$I^{2}_{\\\\text{total}}$=",round(i2[[1]], 2), 
[…]

### 2024-01-17 R/05_literature_statistics.R

@@ -1,0 +1,271 @@
+plot_map <- function(bac_df,
+                     nut_df) {
+  bac_df <- bac_df |>  
+    group_by(StudyID, State) |> 
+    summarise(n = 1) |> 
+    ungroup() |> 
+    group_by(State) |> 
+    summarise(Bacteria = sum(n))
+  
+  nut_df <- nut_df |> 
+    mutate(Parameter = case_when(
+      Parameter %in% c("TN", "DON", "NOX", "NO3", "NO2", "TDN") ~ "Nitrogen",
+      Parameter %in% c("TP", "PO4", "DOP", "DP", "TDP") ~ "Phosphorus",
+      Parameter == "TSS" ~ "Suspended Solids",
+      .default = NA
+    )) |> 
+    group_by(StudyID, State, Parameter) |>
+    summarise(n = 1) |> 
+    filter(!is.na(Parameter)) |> 
+    ungroup() |> 
+    group_by(State, Parameter) |> 
+    summarise(n = sum(n)) |> 
+    pivot_wider(names_from = Parameter,
+                values_from = n)
+  
+  
+  us_file <- tigris::states(cb = TRUE, progress_bar = FALSE)
+  
+  
+  ## shold make the following left join a bac field, then nut field, then pivot longer 
+  
+  us_file |>  
+    left_join(bac_df, by = c("NAME" = "State")) |> 
+    left_join(nut_df, by = c("NAME" = "State")) |> 
+    pivot_longer(cols = Bacteria:Phosphorus,
+                 names_to = "Parameter",
+                 values_to = "n") |> 
+    mutate(n_scale = case_when(
+      is.na(n) ~ 0,
+      .default = n
+    )) |> 
+    filter(!(NAME %in% c("American Samoa",
+                         "Commonwealth of the Northern Mariana Islands",
+                         "Guam",
+                         "United States Virgin Islands"))) |>
+    shift_geometry() |> 
+    ggplot() +
+    geom_sf(aes(fill = n)) +
+    geom_text_repel(aes(label = n, geometry = geometry), size = 3, 
+                    stat = "sf_coordinates", bg.color = "white", bg.r = 0.15,
+                    family = "Manrope SemiBold", fontface = "plain",
+                    force = 0) +
+    scale_fill_distiller(palette = "Greens", direction = 1, na.value = "grey80",
+                         trans = "log10") +
+    facet_wrap(~Parameter) +
+    labs(x = "", y = "") +
+    theme_mps_noto(base_family = "Manrope Regular") +
[…]

### 2024-01-17 _targets.R

@@ -1,0 +1,230 @@
+# _targets.R
+library(targets)
+
+tar_option_set(packages = c("broom",
+                            "dplyr",
+                            "flextable",
+                            "forcats",
+                            "ggplot2",
+                            "ggtext",
+                            "ggrepel",
+                            "kableExtra",
+                            "latex2exp",
+                            "metafor",
+                            "metagear",
+                            "mice",
+                            "modelr",
+                            "mpsTemplates",
+                            "MuMIn",
+                            "officer",
+                            "orchaRd",
+                            "patchwork",
+                            "purrr",
+                            "RColorBrewer",
+                            "readxl",
+                            "scales",
+                            "sf",
+                            "tibble",
+                            "tidyr",
+                            "tigris",
+                            "units"
+))
+
+options(knitr.kable.NA = '',
+        na.action = "na.fail")
+
+
+## source functions
+source("R/01_read_data.R")
+source("R/02_fit_models.R")
+source("R/03_plot_models.R")
+source("R/04_tables.R")
+source("R/05_literature_statistics.R")
+
+## prep fancy figure font
+font_hoist("Manrope")
+
+## for runnning dredge() in MuMIn on rma models
+eval(metafor:::.MuMIn)
+
+
+
+
+list(
+  tar_target(bacteria_df, load_df("bacteria")),
+  tar_target(nutrient_df, load_df("nutrients")),
+  # Bacteria data analysis -----------------------------------------------------
+  ## fit multilevel random effects model ---------------------------------------
+  tar_target(bac_model, fit_bac_model(bacteria_df)),
+  
+  ## Diagnostics and assessment ------------------------------------------------
+  ## small study effect/publication bias using extension of egger's regression
+  tar_target(bac_sse, fit_sse_model(bac_model)),
+
+  ## leave one out sensitivity analysis identifies outliers or influential studies having disproportionate effects on the model estimates
+  tar_target(bac_sens, run_sensitvity(bac_model)),
+
+  ## missing SDs
+  tar_target(bac_missing_sd, assess_missing(bac_model)),
+
+  # i2
+  # i2_total represents the relative variance due to difference between and within studies
+  # i2_studyid represent the relative variance variation attributed to difference between studies
+  # i2_valueid represents the relative variation due to difference within studies.
+  tar_target(bac_i2, i2_ml(bac_model$sel_mod, method = "ratio")),
+
+  ## marginal r2 quantifies the percentage of variation explained by the moderators.
[…]

### 2024-01-31 frontiers_submission/frontiers_submission.Rmd
finish regressions

@@
 ---
-title: "A meta-analysis of the impacts of best management practices on nonpoint source pollutants."
+title: "A meta-analysis of the impacts of best management practices on nonpoint source pollutant concentration."
 runtitle: "Meta-analysis of best management practices"
@@
   - nutrients
-output: rticles::frontiers_article
+output: 
+  bookdown::pdf_book:
+    base_format: rticles::frontiers_article
+    pandoc_args:  !expr rmdfiltr::add_wordcount_filter()
 # bib yaml is commented to manually insert in correct section
@@
 ## load libraries
-library(flextable)
 library(ggtext)
@@
                            store = store)
+din_missing_sd <- tar_read(tin_missing_sd,
+                           store = store)
 tp_missing_sd <- tar_read(tp_missing_sd,
+                           store = store)
+po4_missing_sd <- tar_read(po4_missing_sd,
                            store = store)
@@
 
-eggers_bac <- tar_read(bac_sse,
-                       store = store)
-## set flextable defaults
-set_flextable_defaults(fonts_ignore=TRUE)
+missing_sd <- list(bac_missing_sd,
+                tn_missing_sd,
+                din_missing_sd,
+                tp_missing_sd,
+                po4_missing_sd,
+                tss_missing_sd) 
+missing <- missing_sd |>
+  map_dbl("missing") |> 
+  sum()
+total <- missing_sd |> 
+  map_dbl("total") |> 
+  sum()
+
+percent_missing <- round(missing/total * 100,0)
+
+fib_r2 <- tar_read(bac_r2, store = store)
+fib_i2 <- tar_read(bac_i2, store = store)
+fib_i2_null <- tar_read(bac_i2_null, store = store)
+
+tn_i2_null <- tar_read(tn_i2_null, store = store)
+din_i2_null <- tar_read(tin_i2_null, store = store)
+
+tp_i2_null <- tar_read(tp_i2_null, store = store)
+po4_i2_null <- tar_read(po4_i2_null, store = store)
+
+tss_i2_null <- tar_read(tss_i2_null, store = store)
 
@@
 
-In the United States (U.S.), major progress towards water quality goals have been achieved through the Clean Water Act of 1972.
+In the United States (U.S.), major improvement in water quality has been achieved under the Clean Water Act of 1972.
 This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge [@benhamLessonsLearnedTMDL2008; @nationalresearchcouncilAssessingTMDLApproach2001; @schrammTotalMaximumDaily2022].
-Increased pollutant loads and pollutant concentrations in runoff are a particular challenge when dealing with nonpoint source pollution resulting from land use change.
+Increased pollutant loads and concentrations in runoff resulting from land use changes are a particular challenge.
 The impacts of land use change on hydrology and water quality are well established [@allanLandscapesRiverscapesInfluence2004; @carpenterNonpointPollutionSurface1998; @bernhardtUnderstandingManagingMinimizing2008; @careyEvaluatingNutrientImpacts2013; @freemanImpactsUrbanizationDevelopment2019].
@@
 @linternBestManagementPractices2020 found 43% of reviewed BMP effectiveness studies relied completely on modeled outputs, with modeled outputs almost always predicting water quality improvements following BMP implementation.
-However, field studies are much more likely to demonstrate mixed results including net releases of pollutants under certain conditions  [@linternBestManagementPractices2020; @liuReviewEffectivenessBest2017].
-The disconnect between modeled outcomes and field studies might be attributed to (1) overly simplified or incorrect estimates of parameters that represent management practices [@ullrichApplicationSoilWater2009; @fuReviewCatchmentscaleWater2019; @linternBestManagementPractices2020], (2) failure to incorporate uncertainty in estimates [@tasdighiBayesianTotalUncertainty2018; @fuReviewCatchmentscaleWater2019; @linternBestManagementPractices2020], and (3) assumption of static performance over time [@mealsLagTimeWater2010; @liuReviewEffectivenessBest2017; @fuReviewCatchmentscaleWater2019].
+However, field studies are much more likely to demonstrate mixed results including net releases (leaching) of pollutants under certain conditions  [@linternBestManagementPractices2020; @liuReviewEffectivenessBest2017].
+The disconnect between modeled outcomes and field studies might be attributed to (1) overly simplified or incorrect estimates of model parameters that represent management practices [@ullrichApplicationSoilWater2009; @fuReviewCatchmentscaleWater2019; @linternBestManagementPractices2020], (2) the failure to incorporate the appropriate types of uncertainty into estimates [@tasdighiBayesianTotalUncertainty2018; @fuReviewCatchmentscaleWater2019; @linternBestManagementPractices2020], and (3) the assumption of static performance over time [@mealsLagTimeWater2010; @liuReviewEffectivenessBest2017; @fuReviewCatchmentscaleWater2019].
 
-A source of uncertainty comes from the substantial variability of performance metrics reported in empirical BMP studies [@linternBestManagementPractices2020].
-Previous empirical reviews generally describe high variability and uncertainty in nitrogen and phosphorus removal and consistent reduction in total suspended sediment concentrations across BMP types [@linternBestManagementPractices2020;@liuReviewEffectivenessBest2017; @kochNitrogenRemovalStormwater2014; @claryBMPPerformanceAnalysis2011; @barrettComparisonBMPPerformance2008; @grudzinskiDoesRiparianFencing2020].
+An underlying source of uncertainty comes from the substantial variability of performance metrics reported in empirical BMP studies [@linternBestManagementPractices2020].
+There have been varied attempts at synthesizing estimates of BMP efficiency to provide resource managers with knowledge for improved decision-making [@agouridisLivestockGrazingManagement2005; @barrettComparisonBMPPerformance2008; @claryBMPPerformanceAnalysis2011; @grudzinskiDoesRiparianFencing2020; @horvathEffectsRegionalClimate2023; @kochNitrogenRemovalStormwater2014; @krogerReviewBestManagement2012; @liuReviewEffectivenessBest2017; @simpsonDevelopingBestManagement2009].
+These reviews generally describe high variability and uncertainty in nitrogen and phosphorus removal and consistent reduction in total suspended sediment concentrations across BMP types [@linternBestManagementPractices2020;@liuReviewEffectivenessBest2017; @kochNitrogenRemovalStormwater2014; @claryBMPPerformanceAnalysis2011; @barrettComparisonBMPPerformance2008; @grudzinskiDoesRiparianFencing2020].
 The review literature on the effects of BMPs on FIBs are sparse but generally find extremely high variance in performance across BMPs [@claryBMPPerformanceAnalysis2011; @grudzinskiDoesRiparianFencing2020].
+
+## Sources of variance in BMP performance
+
+While it is assumed that site specific conditions are responsible for some of the heterogeneity in observed BMP performance, it is not clear how much of that variance is due to any one specific factor.
+Influent concentration is likely to have some effect on certain types of structural BMPs.
+@barrettPerformanceComparisonStructural2005 demonstrated that percent pollutant reduction is often a function of influent quality.
+Specifically, for certain types of BMPs percent removal is low at low influent concentrations, and increases with increasing influent concentrations. 
+However, for some types of BMPs and pollutant parameters, effluent concentration is unrelated to influent concentration.
+Second, local climatic conditions can be expected to influence BMP performance.
+BMPs in dry climates have been shown to be more likely to leach phosphorus than those in wetter climates [@horvathEffectsRegionalClimate2023].
+However, elucidating possible confounders such as climate and soil condition has been constrained by the lack of reported local condition data included in most BMP studies [@horvathEffectsRegionalClimate2023; @kochNitrogenRemovalStormwater2014; @eagleMetaAnalysisConstrained2017].
+A third variable that has been explored is the infuence of BMP age and maintence on performance.
+
 Accurately characterizing BMP treated runoff quality is challenging due to the high variability in BMP performance.
-However, influent water quality, seasonality, and BMP age provide some explanatation to highliy heterogenous BMP efficiency data [@barrettPerformanceComparisonStructural2005; @liuReviewEffectivenessBest2017].
-A recent meta-analysis demonstrated the further linkages between local climatic conditions and BMP performance on nitrogen and phosphorus concentrations [@horvathEffectsRegionalClimate2023].
+However, influent water quality, seasonality, and BMP age provide some explanation to highly heterogeneous BMP efficiency data [@barrettPerformanceComparisonStructural2005; @liuReviewEffectivenessBest2017].
+A recent meta-analysis demonstrated the strong linkages between local climatic conditions and phosphorus removal in BMPs [@horvathEffectsRegionalClimate2023].
+
+## Effect size calculations
+
+Results from BMP studies are often reported as BMP efficiency (or percent reduction):
+
+$$
+\text{BMP}_{\text{eff}}=\frac{x_{control}-x_{experiment}}{x_{control}}\times 100,
+$$
+
+where *x~control~* is the pre-treatment or control pollutant concentration and *x~experiment~* is the pollutant concentration measured after the BMP intervention. 
+Several BMP data synthesis efforts have applied statistical summaries or regressions using BMP efficiency as the response variable of interest [@agouridisLivestockGrazingManagement2005; @claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014; @krogerReviewBestManagement2012; @liuReviewEffectivenessBest2017; @simpsonDevelopingBestManagement2009].
+There are several statistical shortcomings (distributional asymmetry, skewness, and non-additive properties) when using efficiency to estimate overall effect sizes across multiple studies that are cause for concern for metrics estimated using this approach [@nuzzoPercentDifferencesAnother2018; @coleStatisticsNotesWhat2017].
+@barrettPerformanceComparisonStructural2005 demonstrated the use of effluent concentration directly as a response variable improved the ability to describe BMP performance. 
+More recently, researchers have applied effect size calculations more commonly used in ecological meta-analysis.
+@horvathEffectsRegionalClimate2023 used the standardized mean difference between influent and effluent, calculated as the difference in the means divided by the pooled standard deviation of the two groups (**hedges**).
+@grudzinskiDoesRiparianFencing2020 applied the log ratio of means (*ROM*) to summarize performance of livestock BMPs.
+*ROM* quantifies the difference in means between the control and experimental group [@hedgesMetaanalysisResponseRatios1999]:
+
+$$
+ROM_i = ln\left(\frac{x_{i,control}}{x_{i,experiment}}\right) = ln(x_{i, control})-ln(x_{i, experiment}),
+$$
+
+where *x~i,control~* and *x~i,experiment~* are the mean pollutant concentrations for experiment *i*. 
+The statistical properties of *ROM* (normal distribution around zero and additive properties) are preferable to using BMP efficiency [@osenbergEffectSizeEcological1997; @hedgesMetaanalysisResponseRatios1999].
+*ROM* \textgreater 0 indicates higher percent reductions and *ROM* \textless indicates pollutant leaching.
+One advantage of *ROM* is that the statistical results calculated using *ROM* are easily transformed to BMP efficiency for interpretation:
+
+$$
+\text{BMP}_{\text{eff}} = \left( 1 - \frac{1}{e^{ROM}}\right) \times 100
+$$
+
+**still needs work**
+Building on previous work, the objective of this project is to (1) improve understanding of the general effectiveness of BMPs on pollutant concentration, and (2) identify important contributors to variability in BMP performance.
+To accomplish this, we conducted a systematic review of relevant published literature and applied a meta-analytic approaches to develop weighted results across studies and identify contributors to variance in BMP performance.
 
@@
 We conducted a systematic review of recent (2000-2022) literature to compile field studies documenting the effectiveness of best management practices on fecal indicator bacteria and nutrients concentrations.
+Prior meta-analysis have utilized data reported in the International Stormwater Database, which consists of self-reported and quality checked BMP data [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014; @horvathEffectsRegionalClimate2023].
+The International Stormwater Database only recently added agricultural BMPs and has relatively sparse FIB data [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014].
+Since we had interest in both FIB performance and agricultural BMPs we chose to utilize a systematic review.
+We also assume that most published reports undergo some method of internal or external peer review prior to publication, which is not necessarily the case with all of the data published in the International Stormwater Database.
+
 The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines [@collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018].
@@
 Although results from Google Scholar are not always replicable, we utilized the service to maximize search results for studies not published in academic journals and presumably increase the chance of identifying studies with negative effects.
-Fecal indicator bacteria study seraches included the following query: “fecal indicator bacteria” OR “E.
-coli” OR “Escherichia coli” OR “enterococci” OR “enterococcus” AND “best management practices” OR “BMPs” AND “effectiveness” OR “performance”.
+Fecal indicator bacteria study seraches included the following query: “fecal indicator bacteria” OR “*E.
+coli*” OR “*Escherichia coli*” OR “enterococci” OR “enterococcus” AND “best management practices” OR “BMPs” AND “effectiveness” OR “performance”.
 Nutrient BMP studies utilized a similar query: “nutrient” OR “nitrogen” OR “phosphorus” OR “sediment” OR “TSS” AND “best management practices” OR “BMPs” AND “effectiveness” OR “performance”.
@@
 Results from each database were first filtered to remove duplicates.
-After removal of duplicates, each member of the research team (*n* = 4) was assigned a subset of studies to evaluate if they should be included.
-**include supplementary tables with inclusion criteria and data variables**
+After removal of duplicates, each member of the research team (*n* = 4) was assigned a subset of studies to evaluate if they should be included (Table \@ref(tab:criteria)).
 Each study was reviewed by two team members and differences in opinion were collectively discussed and agreed upon before progressing.
-The remaining studies were split among team members for data extraction, again with at least two team members reviewing each study. 
+The remaining studies were split among team members for data extraction (Table \@ref(tab:datvars)), again with at least two team members reviewing each study. 
 If data was provided in figures, the data was extracted with the WebPlotDigitizer tool [@rohatgiWebPlotDigitizer2022].
@@
                   "The specific pollutant measured.",
-                  "Dominant source of runoff (rop fields, livestock pasture, commerical, residential",
+                  "Dominant source of runoff (crop fields, livestock pasture, commerical, residential",
                   "Major categorization of runoff source: agricultural or urban",
@@
 
-## Effect size calculation {-}
-
-Prior BMP reviews routinely report the effect of BMPs using efficiency values [see @agouridisLivestockGrazingManagement2005; @krogerReviewBestManagement2012; @liuEnhancingRainfallrunoffModel2015; @simpsonDevelopingBestManagement2009].
-Efficiency values are calculated as:
-
-$$
-\text{Efficiency}=\frac{x_{control}-x_{experiment}}{x_{control}}\times 100,
-$$
-
-where *x~control~* is the pre-treatment or control pollutant concentration and *x~experiment~* is the pollutant concentration measured after the BMP intervention. 
-There are several statistical shortcomings (distributional asymetry, skewness, and non-additive properties) when using efficiency to estimate overall effect sizes [@nuzzoPercentDifferencesAnother2018; @coleStatisticsNotesWhat2017].
-In comparision, the log ratio of means (*ROM*) provide preferable statistical properties for regression analysis [@osenbergEffectSizeEcological1997; @hedgesMetaanalysisResponseRatios1999]. 
-*ROM* quantifies the difference in means between the control and experimental group [@hedgesMetaanalysisResponseRatios1999]:
-
-$$
-ROM_i = ln\left(\frac{x_{i,control}}{x_{i,experiment}}\right) = ln(x_{i, control})-ln(x_{i, experiment}),
-$$
-where *x~i,control~* and *x~i,experiment~* are the mean pollutant concentrations for experiment *i*. 
-The statistical properties of *lnRR* (normal distribution around zero and additive properties) are preferable to the more commonly reported efficiency metric [@osenbergEffectSizeEcological1997; @hedgesMetaanalysisResponseRatios1999].
-Use of *ROM* required the exclusion of studies that only provided measures of BMP efficiency and not the underlying data used to derive the metric.
-
 ## Statistical models
 
-We used the "rma.mv" function in the *metafor* R package to fit multilevel random effects linear models with *ROM* as the effect variable [@viechtbauerConductingMetaanalysesMetafor2010; @rcoreteamLanguageEnvironmentStatistical2023].
+We used the "rma.mv" function in the *metafor* R package to fit multilevel random effects regression models with *ROM* as the effect variable [@viechtbauerConductingMetaanalysesMetafor2010; @rcoreteamLanguageEnvironmentStatistical2023].
 Our models specified a nested random effects term accounting for heterogeneity between effect sizes from the same study and for heterogeneity between studies.
+*ROM* was used as the effect size which required the exclusion of studies that only provided measures of BMP efficiency and not the underlying data used to derive the metric.
 A key feature of meta-analysis is the weighting of effects using sampling variance of individual effect sizes.
-`r paste0(round(bac_missing_sd$missing / bac_missing_sd$total * 100, 0), "%")` of fecal indicator bacteria, `r paste0(round(tn_missing_sd$missing / tn_missing_sd$total * 100, 0), "%")` of TN, `r paste0(round(tp_missing_sd$missing / tp_missing_sd$total * 100, 0), "%")` of TP, and `r paste0(round(tss_missing_sd$missing / tss_missing_sd$total * 100, 0), "%")` of TSS effect sizes were missing standard deviations required to estimate sampling variance.
-Removal of studies due to missing variance information can introduce substantial bias [@kambachConsequencesMultipleImputation2020], so we imputed missing standard deviations using random forest based multivariate imputation with chained equations using the *mice* R package [@buurenMiceMultivariateImputation2011].
-Sampling variance was estimated with the variance estimator described in Doncaster et al. to reduce bias in small sample studies:
+`r percent_missing |> xfun::numbers_to_words(cap = TRUE)` percent of `r total` effect sizes were missing standard deviations required to estimate sampling variance. 
+Removal of studies due to missing variance information can introduce substantial bias [@kambachConsequencesMultipleImputation2020]. 
+Missing standard deviations were imputed using the pooled ratio of the mean effect size to coefficient of variation (CV) [@brackenStatisticalMethodsAnalysis1992].
+Sampling variance was estimated utilizing the average squared CV across all studies divided by sample size for each effect [@nakagawaRobustReadilyImplementable2023; @doncasterCorrectionBiasMeta2018]:
 
@@
 
-Our initial models included log transformed pre-treatment pollutant concentration, study scale (plot/field, community, watershed), study length (*n* years), BMP subcategory (), and pollutant concentration:BMP subcategory interactions were included as fixed effect terms.
+Our initial models included log transformed influent concentration, BMP subcategory (drainage modification, crop field management, livestock management, filtration, treatment, detention, or infiltration), aridity index (mean-centered), influent concentration\texttimes BMP subcategory interactions, and aridity index\texttimes BMP subcategory interactions were included as fixed effect terms.
+Aridity index was the only moderator not obtained directly in the systematic review (Table \@ref(tab:datvars)).
+We mapped study location coordinates to aridity index values published in the "Global Aridity Index and Potential Evapotranspiration Database - Version 3" (Global-AI_PET_v3) which provides gridded 30 arc-second annual average precipitation and potential evapotranspiration estimates [@zomerVersionGlobalAridity2022].
+The aridity index is calculated as the ratio of mean annual precipitation to mean annual evapotraspiration with values between 0 and 0.5 considered hyper to semi-arid, and values above 0.65 as humid.
+
 We used an information-theoretic approach to select the most parsimonious model from the subset of candidate models based on corrected Akaike information criterion (AIC~c~) estimated with maximum likelihood [@cinarUsingInformationTheoretic2021].
-The final model was selected from candidate models, which included all combination and subsets of the full model, by eliminating candidate models until the $\Delta$*AIC~c~* reached $\leq$ 2 [@burnhamAICModelSelection2011].
-We estimated regression coefficients of the selected model using restricted maximum likelihood.
+Candidate models used for variable selection were fit with maximum likelihood (ML).
+The final model was selected from candidate models, which included all combination and subsets of the full model, by selecting the model with the lowest AIC~c~ score [@burnhamAICModelSelection2011; @cinarUsingInformationTheoretic2021].
+Regression coefficients of the selected model were estimated using restricted maximum likelihood (REML).
 Relative heterogeneity between and within studies were calculated using the *I^2^* metric described in @nakagawaMethodologicalIssuesAdvances2012.
@@
 
-```{r}
-bac_z <- eggers_bac$zval[[3]] |> prettyNum(digits = 3, nsmall = 2)
-bac_p <- eggers_bac$pval[[3]] |> format.pval(eps = .01, digits = 2, nsmall = 1)
-```
-
 We tested for evidence of publication bias, in the form of small study effect, by using the extension of Egger's regression applied to the multilevel model framework that included adjusted sampling error as a moderator [@nakagawaQuantitativeEvidenceSynthesis2023].
-We did not identify evidence of publication bias in the surveyed studies for fecal indicator bacteria (*Z* = `r bac_z`, *p* = `r bac_p`), total nitrogen, total phosphorus, or total suspended solids; therefore, adjustments for publication bias were not included in the final models.
+We did not identify evidence of publication bias in the surveyed studies (FIB: ROM = 1.19, 95% CI [-3.23, 5.61]; TN: ROM = 0.31, 95% CI [-1.15, 1.78]; DIN: ROM = -2.4, 95% CI [-6.14, 1.59]; TP: ROM = -1.97, 95% CI [-5.18, 1.23]; PO~4~: -0.05, 95% CI [-2.82, 2.72]; TSS: ROM: 0.31, 95% CI [-1.15, 1.78]; Figure S1-S6); therefore, adjustments for publication bias were not included in the final models.
 We conducted a sensitivity analysis of the robustness of overall effect sizes to individual studies using leave-one-out analysis [@nakagawaQuantitativeEvidenceSynthesis2023].
-Outlier studies identified in the sensitivity analysis are highlighted in Table X and were removed prior to fitting the final models.
+This approach repeatedly fits the selected model leaving out an individual value each time. 
+The overall effect and 95% CI from each refit model is compared to the overall effect and 95% CI of the model fit to the full dataset.
+We did not identify evidence of outliers or overly influential studies for any of our models (Figure S7-S12).
 
@@
 
-## Regression results
+<!-- ** add study length figure.** -->
 
-Some evidence of a relationship between input concentrations and fecal indicator bacteria reductions.
-No evidence strong evidence that the type of BMP changed the overall relationship.
+There were a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effects for PO~4~, and 33 studies with 125 effect sizes for TSS.
+The majority of studies were identified as smaller scaled lot or field studies (Figure \@ref(fig:bmpsummary) A).
+FIB studies had a roughly equal proportion large watershed/catchment studies and studies conducted at the commuity/farm scale or smaller.
+We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \@ref(fig:bmpsummary) B).
+We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \@ref(fig:bmpsummary) C).
+Our review was resticted to studies published after 1999. 
+The number of studies published for each parameter were roughly uniformly distributed over time (Figure \@ref(fig:bmptemporal) A) and are not indicative of increases or decreases in the number of published studies.
+Study length was strongly skewed for all parameters (Figure \@ref(fig:bmptemporal) B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO~4), 2.5 (TN), 2.5 (TP), and 2 (TSS). 
+There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \@ref(fig:studymap)).
+This might suggest that spatially correlated predictors such as climate and soil variables would be difficult to model due to the poor spatial representation of the collected data.
 
-```{r bacsummary}
+## Regression models
 
-tar_read(bac_table, store = store) |> 
-  pack_rows(group_label = "BMP Subcategories",
-            start_row = 2,
-            end_row = 5)
+```{r modelaic}
+
+add_cand_models <- function(res) {
+  df <- data.frame(res) |> 
+    arrange(df) |> 
+    mutate(model = c(
+      "$\\sim$ Int",
+      "$\\sim$ Int+Influent",
+      "$\\sim$ Int+AI",
+      "$\\sim$ Int+AI+Influent",
+      "$\\sim$ Int+BMP",
+      "$\\sim$ Int+BMP+Influent",
+      "$\\sim$ Int+AI+BMP",
+      "$\\sim$ Int+AI+BMP+Influent",
+      "$\\sim$ Int+BMP\\texttimes Influent",
+      "$\\sim$ Int+AI\\texttimes BMP",
+      "$\\sim$ Int+Influent+AI\\texttimes BMP",
+      "$\\sim$ Int+AI+BMP\\texttimes Influent",
+      "$\\sim$ Int+AI\\texttimes BMP+BMP\\texttimes Influent"
+    )) |> 
+    mutate(AICc = round(AICc, 1))
+
+  df$AICc <- cell_spec(df$AICc, bold = ifelse(df$AICc == min(df$AICc),
+                                              TRUE, FALSE))
+
+  df
+}
+
+bac_res <- tar_read(bac_model, store = store)$res |> 
+  add_cand_models() |> 
+  select(model, "FIB" = AICc)
+
+tn_res <- tar_read(tn_model, store = store)$res |> 
+  add_cand_models() |> 
+  select(model, "TN" = AICc)
+
+din_res <- tar_read(tin_model, store = store)$res |> 
+  add_cand_models() |> 
+  select(model, "DIN" = AICc)
+
+tp_res <- tar_read(tp_model, store = store)$res |> 
+  add_cand_models() |> 
+  select(model, "TP" = AICc)
+
+po4_res <- tar_read(po4_model, store = store)$res |> 
+  add_cand_models() |> 
+  select(model, "PO\\textsubscript4" = AICc)
+
+tss_res <- tar_read(tss_model, store = store)$res |> 
+  add_cand_models() |> 
+  select(model, "TSS" = AICc)
+
+modelaic <- left_join(bac_res, tn_res, by = c("model" = "model")) |> 
+  left_join(din_res, by = c("model" = "model")) |> 
+  left_join(tp_res, by = c("model" = "model")) |> 
+  left_join(po4_res, by = c("model" = "model")) |> 
+  left_join(tss_res, by = c("model" = "model")) |> 
+  rename("Candidate Models" = model)
+
+kableExtra::kbl(modelaic,
+                escape = FALSE, booktabs = TRUE,
+                caption = "Model selection AICc") |> 
+  add_header_above(c(" ", "AICc" = 6)) |> 
+  footnote(general = "Int = intercept; AI = ariditiy index; BMP = BMP subcategory", general_title = "")
+```
+
+### Fecal Indicator Bacteria
+
+```{r}
+fib_i2_null <- paste0("\\textit{I\\textsuperscript{2}\\textsubscript{total}} = ",
+                      round(fib_i2_null[[1]],2),
+                      ", \\textit{I\\textsuperscript{2}\\textsubscript{study}} = ",
+                      round(fib_i2_null[[2]],2),
+                      ", \\textit{I\\textsuperscript{2}\\textsubscript{effect}} = ",
+                      round(fib_i2_null[[3]],2))
+
+fib_r2 <- paste0("\\textit{R\\textsuperscript{2}\\textsubscript{marginal}} =",
+                 fib_r2[[1]] |> round(2))
+```
+
+There were only 19 studies and 63 FIB effect sizes available to model after removal of studies and effects that only reported BMP~eff~.
+The overall mean effect (estimated with the intercept only multilevel random effects model) showed significant mean reductions in FIB (ROM = 0.85, 95% CI [0.36, 1.34]; BMP~eff~ = 57.4%, 95% CI [30.4%, 73.9%]; Figure \@ref(fig:overalleffect)) resulting from BMPs.
+Total heterogeneity was moderate with a relatively large amount of heterogeneity observed due to differences within studies (`r fib_i2_null`).
+
+AICc scores included log transformed influent concentration and the aridity index\texttimes BMP subcategory interaction as moderators for the FIB model (Table \@ref(tab:modelaic)).
+Moderator terms and interactions explained a high proportion of effect size variance (`r fib_r2`) in the FIB model.
+Increased influent concentrations ($\beta$ = 0.25, 95% CI [0.14, 0.37]) resulted in significantly larger ROM effect for FIB (Figure \@ref(fig:fibeffect).
+Compared to the baseline aridity index\texttimes detention BMP subcategory interaction, infiltration ($\beta$ = -29.90, 95% CI [-50.34, -9.47]), livestock management ($\beta$ = -30.37, 95% CI [-50.93, -9.81]), and treatment ($\beta$ = -30.33, 95% CI [-49.62, -11.03]) interactions had significantly smaller slopes.
+However, the data had uneven coverage of BMP subcategories across the aridity index. 
+Effects for detention BMPs were clustered in humid climates (aridity index > 0.65) and the resulting estimate for the baseline interaction ($\beta$ = 32.63, 95% CI [12.57, 52.69]) may not be reliable.
+
+### Nitrogen
+
+```{r}
+tn_i2_null <- paste0("\\textit{I\\textsuperscript{2}\\textsubscript{total}} = ",
+                      round(tn_i2_null[[1]],2),
+                      ", \\textit{I\\textsuperscript{2}\\textsubscript{study}} = ",
+                      round(tn_i2_null[[2]],2),
+                      ", \\textit{I\\textsuperscript{2}\\textsubscript{effect}} = ",
+                      round(tn_i2_null[[3]],2))
+
+din_i2_null <- paste0("\\textit{I\\textsuperscript{2}\\textsubscript{total}} = ",
+                      round(din_i2_null[[1]],2),
+                      ", \\textit{I\\textsuperscript{2}\\textsubscript{study}} = ",
+                      round(din_i2_null[[2]],2),
+                      ", \\textit{I\\textsuperscript{2}\\textsubscript{effect}} = ",
+                      round(din_i2_null[[3]],2))
 
@@
 
-```{r nitsummary}
+We identified 13 eligible TN studies and 14 DIN studies and 31 and 44 effect sizes respectively that could be included in the regression model.
+Overall effects showed that BMPs resulted in significant mean reductions in TN (ROM = 0.42, 95% CI [0.21, 0.62]; BMP~eff~ = 34.0%, 95% CI [18.7%, 46.4%]; Figure \@ref(fig:overalleffect)) but not in DIN (ROM = 0.64, 95% CI [-0.08, 1.35]; BMP~eff~ = 47.1%, 95% CI [-8.1%, 74.1%]; Figure \@ref(fig:overalleffect)).
+Heterogeneity was high for TN with a large proportion of heterogeneity attributed to within study effect (`r tn_i2_null`).
+The DIN model had even higher heterogeneity with a larger proportion attributed between studies (`r din_i2_null`).
+AICc scores indicated that none of the moderators resulted in substantial improvement over the intercept only model (Table \@ref(tab:modelaic)).
 
-tar_read(tn_table, store = store)
+### Phosphorus
+
+```{r}
+tp_i2_null <- paste0("\\textit{I\\textsuperscript{2}\\textsubscript{total}} = ",
+                      round(tp_i2_null[[1]],2),
+                      ", \\textit{I\\textsuperscript{2}\\textsubscript{study}} = ",
+                      round(tp_i2_null[[2]],2),
+                      ", \\textit{I\\textsuperscript{2}\\textsubscript{effect}} = ",
+                      round(tp_i2_null[[3]],2))
+
+po4_i2_null <- paste0("\\textit{I\\textsuperscript{2}\\textsubscript{total}} = ",
+                      round(po4_i2_null[[1]],2),
+                      ", \\textit{I\\textsuperscript{2}\\textsubscript{study}} = ",
+                      round(po4_i2_null[[2]],2),
+                      ", \\textit{I\\textsuperscript{2}\\textsubscript{effect}} = ",
+                      round(po4_i2_null[[3]],2))
 
@@
 
-```{r phossummary}
+We found 17 TP studies with 37 effect sizes and 9 PO~4~ studies with 21 effect sizes for inclusion in regression models.
+There was a significant overall reduction found for TP (ROM = 0.40, 95% CI [0.03, 0.76]; BMP~eff~ = 32.7%, 95% CI [3.4%, 53.2%]) but no evidence of negative or positive effect for PO~4~ (ROM = -0.18, 95% CI [-0.56, 0.19]; BMP~eff~ = -20.1%, 95% CI [-75.3%, 17.7%]).
+For both the TP and PO~4~ models, heterogenity was high, with moderate to high within study variance and low to moderate between study variance (TP: `r tp_i2_null`; PO~4~: `r po4_i2_null`).
+The best model for both parameters only included influent as a moderator (Table \@ref(tab:modelaic)).
+Influent concentration ($\beta$ = 0.23, 95% CI [-0.035, 0.49]) was not significant at the 95% confidence level for the TP model (Figure \@ref(fig:phoseffect); Table S4).
+Influent concentration ($\beta$ = 0.27, 95% CI [0.085, 0.44]) was significant for the PO~4~ model (Figure \@ref(fig:phoseffect); Table S5).
 
-tar_read(tp_table, store = store)
+### Sediment
+
+```{r}
+tss_i2_null <- paste0("\\textit{I\\textsuperscript{2}\\textsubscript{total}} = ",
+                      round(tss_i2_null[[1]],2),
+                      ", \\textit{I\\textsuperscript{2}\\textsubscript{study}} = ",
+                      round(tss_i2_null[[2]],2),
+                      ", \\textit{I\\textsuperscript{2}\\textsubscript{effect}} = ",
+                      round(tss_i2_null[[3]],2))
 
@@
 
-## Subsection 2
-
-Frontiers requires figures to be submitted individually, in the same order as
-they are referred to in the manuscript. Figures will then be automatically
-embedded at the bottom of the submitted manuscript. Kindly ensure that each
-table and figure is mentioned in the text and in numerical order. Permission
-must be obtained for use of copyrighted material from other sources (including
-the web). Please note that it is compulsory to follow figure instructions.
-Figures which are not according to the guidelines will cause substantial delay
-during the production process.
+There were 12 eligible TSS studies with 26 effect sizes for regression modelling.
+We found a significant and large reduction in TSS concentrations across studies (ROM = 1.65, 95% CI [0.96, 2.34]; BMP~eff~ = 80.9%, 95% CI [61.9%, 90.4%]).
+Heterogeneity was high for TSS with a large proportion of heterogeneity attributed to within study effect (`r tss_i2_null`).
+Similar to nitrogen, we did not find strong evidence linking any of the tested moderators to BMP performance (Table \@ref(tab:modelaic)).
 
@@
 
-Linking BMP performance to watershed scale improvements remains a challenge (Tomer and Locke, Melland et al, Meals). 
+Our systematic review revealed strong spatial disparities in published BMP studies (Figure \@ref(fig:studymap).
+
+Meta-analysis indicated that BMPs showed significant overall reductions in FIB, TN, TP, and TSS concentrations when using ROM as an effect size. 
+We did not find strong evidence of overall leaching or reductions of DIN or PO~4~ in BMP studies.
+The results are in agreement to some extent with previous reviews that found effective (but highly variable) removal efficiencies for nitrogen, phosphorus, and sediment [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014; @liuReviewEffectivenessBest2017].
+
+compare overall effects to koch, barrett, clary
+
+Linking BMP performance to watershed scale improvement remains a challenge (Tomer and Locke, Melland et al, Meals). 
 The succesful implementation and maintence of BMPs at watershed scales remains an area for more investigation (Liu et al., Lintern) 
@@
 
+```{r bmpsummary}
+#| fig.cap: "Summary of (A) study scale, (B) dominant runoff source, and (C) BMPs identified in the systematic review."
+#| #| fig.pos: "p"
+#| out.extra: ""
+#| out.width: "100%"
+
+knitr::include_graphics("../figures/bmp_summary.pdf")
+
+```
+
+```{r bmptemporal}
+#| fig.cap: "Number of studies identified in the systematic review summarized by (A) publication date and (B) study length."
+#| #| fig.pos: "p"
+#| out.extra: ""
+#| out.width: "100%"
+
+knitr::include_graphics("../figures/bmp_temporal_summary.pdf")
+
+```
+
 ```{r studymap}
 #| fig.cap: "Distribution of studies identified in the systematic review by state and parameter."
-#| fig.pos: "H"
+#| fig.pos: "p"
 #| out.extra: ""
@@
 
-```{r bmpsummary}
-#| fig.cap: "Summary of (A) study scale, (B) dominant runoff source, and (C) BMPs identified in the systematic review."
-#| #| fig.pos: "H"
-#| out.extra: ""
-#| out.width: "100%"
-
-knitr::include_graphics("../figures/bmp_summary.pdf")
-
-```
-
 ```{r overalleffect}
-#| fig.cap: "Estimated effect sizes and intervals from the null multilevel random effects model.
+#| fig.cap: "Estimated effect sizes and intervals from the intercept only multilevel random effects model.
 Individual points represent studies, with size scaled by sampling variance.
@@
 Here, \\textit{k} indicates the number of overall effects with the number of unique studies in parenthesis."
-#| fig.pos: "H"
+#| fig.pos: "p"
 #| out.extra: ""
@@
 
+```{r fibeffect}
+#| fig.cap: "Predicted marignal effect of influent FIB and aridity index (conditioned on BMP subcategory).
+Solid lines are the predicted mean effect, dashed lines are the 95% confidence intervals, and the dotted lines are the 95% prediction intervals.
+Individual dots represent each effect size identified in the literature with the size scaled by sampling variance."
+#| fig.pos: "p"
+#| out.extra: ""
+#| out.width: "100%"
+
+knitr::include_graphics("../figures/bac_predicted.pdf")
+```
+
+```{r phoseffect}
+#| fig.cap: "Predicted marignal effect of influent pollutant concentration on TP (Panel A) and PO\\textsubscript{4} (Panel B) reductions.
+Solid lines are the predicted mean effect, dashed lines are the 95% confidence intervals, and the dotted lines are the 95% prediction intervals.
+Individual dots represent each effect size identified in the literature with the size scaled by sampling variance."
+#| fig.pos: "p"
+#| out.extra: ""
+#| out.width: "100%"
+
+knitr::include_graphics("../figures/phos_predicted.pdf")
+```
 <!-- #  -->

### 2024-01-31 frontiers_submission/frontiers_submission.tex
finish regressions

@@
 
+% From pandoc table feature
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+
 \usepackage{booktabs}
@@
 
-  \textsuperscript{1} Texas Water Resources Institute, Texas A\&M
-AgriLife Research,  College Station,  Texas,  USA
+  \textsuperscript{1} Texas Water Resources Institute, Texas A\&M AgriLife Research,  College Station,  Texas,  USA
   }
 
-  \def\corrAuthor{Michael Schramm}\def\corrAddress{Texas Water Resources
-Institute, Texas A\&M AgriLife Research\\1001 Holleman Dr E.\\College
-Station, Texas, 77840 USA}\def\corrEmail{\href{mailto:michael.schramm@ag.tamu.edu}{\nolinkurl{michael.schramm@ag.tamu.edu}}}
+  \def\corrAuthor{Michael Schramm}\def\corrAddress{Texas Water Resources Institute, Texas A\&M AgriLife Research\\1001 Holleman Dr E.\\College Station, Texas, 77840 USA}\def\corrEmail{\href{mailto:michael.schramm@ag.tamu.edu}{\nolinkurl{michael.schramm@ag.tamu.edu}}}
   \def\firstAuthorLast{Schramm {et~al.}}
@@
 
-\title[Meta-analysis of best management practices]{A meta-analysis of
-the impacts of best management practices on nonpoint source pollutants.}
+\title[Meta-analysis of best management practices]{A meta-analysis of the impacts of best management practices on nonpoint source pollutants.}
 \author[\firstAuthorLast]{\Authors}
@@
 Abstract length and content varies depending on article type. Refer to
-\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract
-requirement and length according to article type.
+\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract requirement
+and length according to article type.
 
 \tiny
- \keyFont{ \section{Keywords:} best management practice, water
-quality, nonpoint source pollution, fecal indicator
-bacteria, nutrients} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
+ \keyFont{ \section{Keywords:} best management practice, water quality, nonpoint source pollution, fecal indicator bacteria, nutrients} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
 \end{abstract}
@@
 
-In the United States (U.S.), major progress towards water quality goals
-have been achieved through the Clean Water Act of 1972. This progress
-has been largely attributed to investments and reductions in point
-source discharges while reductions in nonpoint source pollutants remain
-a substantial challenge
-\citep{benhamLessonsLearnedTMDL2008, nationalresearchcouncilAssessingTMDLApproach2001, schrammTotalMaximumDaily2022}.
-Increased pollutant loads and pollutant concentrations in runoff are a
-particular challenge when dealing with nonpoint source pollution
-resulting from land use change. The impacts of land use change on
-hydrology and water quality are well established
-\citep{allanLandscapesRiverscapesInfluence2004, carpenterNonpointPollutionSurface1998, bernhardtUnderstandingManagingMinimizing2008, careyEvaluatingNutrientImpacts2013, freemanImpactsUrbanizationDevelopment2019}.
-Nonpoint source driven fecal indicator bacteria (FIB), nitrogen,
-phosphorus, and suspended sediment remain major causes of water quality
-impairments in U.S. rivers and streams despite decades of work. In 2017,
-the Environmental Protection Agency (EPA) estimated 41\% or more of the
-nation's rivers and streams rated poorly for biological condition due to
-excess nitrogen or phosphorus \citep{epaNationalWaterQuality2017}. FIB
-remains the leading cause of water body impairment on the Clean Water
-Act 303(d) list in the United States
-\citep{epaNationalWaterQuality2017}.
+In the United States (U.S.), major progress towards water quality goals have been achieved through the Clean Water Act of 1972.
+This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge \citep{benhamLessonsLearnedTMDL2008, nationalresearchcouncilAssessingTMDLApproach2001, schrammTotalMaximumDaily2022}.
+Increased pollutant loads and pollutant concentrations in runoff are a particular challenge when dealing with nonpoint source pollution resulting from land use change.
+The impacts of land use change on hydrology and water quality are well established \citep{allanLandscapesRiverscapesInfluence2004, carpenterNonpointPollutionSurface1998, bernhardtUnderstandingManagingMinimizing2008, careyEvaluatingNutrientImpacts2013, freemanImpactsUrbanizationDevelopment2019}.
+Nonpoint source driven fecal indicator bacteria (FIB), nitrogen, phosphorus, and suspended sediment remain major causes of water quality impairments in U.S. rivers and streams despite decades of work.
+In 2017, the Environmental Protection Agency (EPA) estimated 41\% or more of the nation's rivers and streams rated poorly for biological condition due to excess nitrogen or phosphorus \citep{epaNationalWaterQuality2017}.
+FIB remains the leading cause of water body impairment on the Clean Water Act 303(d) list in the United States \citep{epaNationalWaterQuality2017}.
 
-Best management practices (BMPs) have been the primary suite of tools
-for addressing nonpoint source pollution. BMPs are structural or
-non-structural controls used to mitigate the effects of increased runoff
-volume, pollutant loads, or pollutant concentrations emanating from
-diffuse nonpoint sources. BMPs control the delivery of pollutants
-through a few possible mechanisms. Structural BMPs reduce and retard
-total volume of runoff, thus reducing both the volume of water and
-pollutant load. Structural BMPs may also provide a mechanism for
-physical, chemical, or biological removal of pollutant constituents
-suspended or dissolved in runoff. Non-structural BMPs (such as nutrient
-management or livestock management) are utilized to reduce the
-generation of pollutant runoff by avoiding pollutant generation during
-critical periods.
+Best management practices (BMPs) have been the primary suite of tools for addressing nonpoint source pollution.
+BMPs are structural or non-structural controls used to mitigate the effects of increased runoff volume, pollutant loads, or pollutant concentrations emanating from diffuse nonpoint sources.
+BMPs control the delivery of pollutants through a few possible mechanisms.
+Structural BMPs reduce and retard total volume of runoff, thus reducing both the volume of water and pollutant load.
+Structural BMPs may also provide a mechanism for physical, chemical, or biological removal of pollutant constituents suspended or dissolved in runoff.
+Non-structural BMPs (such as nutrient management or livestock management) are utilized to reduce the generation of pollutant runoff by avoiding pollutant generation during critical periods.
 
-Practitioners rely extensively on mechanistic models to plan and
-evaluate BMP scenarios and resulting water quality.
-\citet{linternBestManagementPractices2020} found 43\% of reviewed BMP
-effectiveness studies relied completely on modeled outputs, with modeled
-outputs almost always predicting water quality improvements following
-BMP implementation. However, field studies are much more likely to
-demonstrate mixed results including net releases of pollutants under
-certain conditions
-\citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017}.
-The disconnect between modeled outcomes and field studies might be
-attributed to (1) overly simplified or incorrect estimates of parameters
-that represent management practices
-\citep{ullrichApplicationSoilWater2009, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020},
-(2) failure to incorporate uncertainty in estimates
-\citep{tasdighiBayesianTotalUncertainty2018, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020},
-and (3) assumption of static performance over time
-\citep{mealsLagTimeWater2010, liuReviewEffectivenessBest2017, fuReviewCatchmentscaleWater2019}.
+Practitioners rely extensively on mechanistic models to plan and evaluate BMP scenarios and resulting water quality.
+\citet{linternBestManagementPractices2020} found 43\% of reviewed BMP effectiveness studies relied completely on modeled outputs, with modeled outputs almost always predicting water quality improvements following BMP implementation.
+However, field studies are much more likely to demonstrate mixed results including net releases of pollutants under certain conditions \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017}.
+The disconnect between modeled outcomes and field studies might be attributed to (1) overly simplified or incorrect estimates of parameters that represent management practices \citep{ullrichApplicationSoilWater2009, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020}, (2) failure to incorporate uncertainty in estimates \citep{tasdighiBayesianTotalUncertainty2018, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020}, and (3) assumption of static performance over time \citep{mealsLagTimeWater2010, liuReviewEffectivenessBest2017, fuReviewCatchmentscaleWater2019}.
 
-A source of uncertainty comes from the substantial variability of
-performance metrics reported in empirical BMP studies
-\citep{linternBestManagementPractices2020}. Previous empirical reviews
-generally describe high variability and uncertainty in nitrogen and
-phosphorus removal and consistent reduction in total suspended sediment
-concentrations across BMP types
-\citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017, kochNitrogenRemovalStormwater2014, claryBMPPerformanceAnalysis2011, barrettComparisonBMPPerformance2008, grudzinskiDoesRiparianFencing2020}.
-The review literature on the effects of BMPs on FIBs are sparse but
-generally find extremely high variance in performance across BMPs
-\citep{claryBMPPerformanceAnalysis2011, grudzinskiDoesRiparianFencing2020}.
-Accurately characterizing BMP treated runoff quality is challenging due
-to the high variability in BMP performance. However, influent water
-quality, seasonality, and BMP age provide some explanatation to highliy
-heterogenous BMP efficiency data
-\citep{barrettPerformanceComparisonStructural2005, liuReviewEffectivenessBest2017}.\\
-A recent meta-analysis demonstrated the further linkages between local
-climatic conditions and BMP performance on nitrogen and phosphorus
-concentrations \citep{horvathEffectsRegionalClimate2023}.
+One source of uncertainty comes from the substantial variability of performance metrics reported in empirical BMP studies \citep{linternBestManagementPractices2020}.
+There have been varied attempts at synthesizing estimates of BMP efficiency to provide resource managers with knowledge for improved decision-making \citep{agouridisLivestockGrazingManagement2005, barrettComparisonBMPPerformance2008, claryBMPPerformanceAnalysis2011, grudzinskiDoesRiparianFencing2020, horvathEffectsRegionalClimate2023, kochNitrogenRemovalStormwater2014, krogerReviewBestManagement2012, liuReviewEffectivenessBest2017, simpsonDevelopingBestManagement2009}.
+These reviews generally describe high variability and uncertainty in nitrogen and phosphorus removal and consistent reduction in total suspended sediment concentrations across BMP types \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017, kochNitrogenRemovalStormwater2014, claryBMPPerformanceAnalysis2011, barrettComparisonBMPPerformance2008, grudzinskiDoesRiparianFencing2020}.
+The review literature on the effects of BMPs on FIBs are sparse but generally find extremely high variance in performance across BMPs \citep{claryBMPPerformanceAnalysis2011, grudzinskiDoesRiparianFencing2020}.
+Accurately characterizing BMP treated runoff quality is challenging due to the high variability in BMP performance.
+However, influent water quality, seasonality, and BMP age provide some explanation to highly heterogeneous BMP efficiency data \citep{barrettPerformanceComparisonStructural2005, liuReviewEffectivenessBest2017}.\\
+A recent meta-analysis demonstrated the strong linkages between local climatic conditions and phosphorus removal in BMPs \citep{horvathEffectsRegionalClimate2023}.
+
+\hypertarget{effect-size-calculations}{%
+\subsection{Effect size calculations}\label{effect-size-calculations}}
+
+Results from BMP studies are often reported as BMP efficiency (or percent reduction):
+
+\[
+\text{BMP}_{\text{eff}}=\frac{x_{control}-x_{experiment}}{x_{control}}\times 100,
+\]
+
+where \emph{x\textsubscript{control}} is the pre-treatment or control pollutant concentration and \emph{x\textsubscript{experiment}} is the pollutant concentration measured after the BMP intervention.
+Several BMP data synthesis efforts have applied statistical summaries or regressions using BMP efficiency as the response variable of interest \citep{agouridisLivestockGrazingManagement2005, claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, krogerReviewBestManagement2012, liuReviewEffectivenessBest2017, simpsonDevelopingBestManagement2009}.
+There are several statistical shortcomings (distributional asymmetry, skewness, and non-additive properties) when using efficiency to estimate overall effect sizes across multiple studies that are cause for concern for metrics estimated using this approach \citep{nuzzoPercentDifferencesAnother2018, coleStatisticsNotesWhat2017}.
+\citet{barrettPerformanceComparisonStructural2005} demonstrated the use of effluent concentration directly as a response variable improved the ability to describe BMP performance.
+More recently, researchers have applied effect size calculations more commonly used in ecological meta-analysis.
+\citet{horvathEffectsRegionalClimate2023} used the standardized mean difference between influent and effluent, calculated as the difference in the means divided by the pooled standard deviation of the two groups (\textbf{hedges}).
+\citet{grudzinskiDoesRiparianFencing2020} applied the log ratio of means (\emph{ROM}) to summarize performance of livestock BMPs.
+\emph{ROM} quantifies the difference in means between the control and experimental group \citep{hedgesMetaanalysisResponseRatios1999}:
+
+\[
+ROM_i = ln\left(\frac{x_{i,control}}{x_{i,experiment}}\right) = ln(x_{i, control})-ln(x_{i, experiment}),
+\]
+
+where \emph{x\textsubscript{i,control}} and \emph{x\textsubscript{i,experiment}} are the mean pollutant concentrations for experiment \emph{i}.
+The statistical properties of \emph{ROM} (normal distribution around zero and additive properties) are preferable to using BMP efficiency \citep{osenbergEffectSizeEcological1997, hedgesMetaanalysisResponseRatios1999}.
+\emph{ROM} \textgreater 0 indicates higher percent reductions and \emph{ROM} \textless indicates pollutant leaching.
+One advantage of \emph{ROM} is that the statistical results calculated using \emph{ROM} are easily transformed to BMP efficiency for interpretation:
+
+\[
+\text{BMP}_{\text{eff}} = \left( 1 - \frac{1}{e^{ROM}}\right) \times 100
+\]
+
+The goals of this project were to (1) conduct a systematic review of BMP field studies; (2) develop estimates of FIB, nitrogen, phosphorus, and TSS removal using a meta-analytic approach; and (3) explore the effect of multiple moderators using multilevel random effects meta-analytic regression models.
 
@@
 
-We conducted a systematic review of recent (2000-2022) literature to
-compile field studies documenting the effectiveness of best management
-practices on fecal indicator bacteria and nutrients concentrations. The
-systematic review followed guidance provided in the Collaboration for
-Environmental Evidence systematic review guidelines
-\citep{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}.
-In order to maximize the number of studies included in the review, we
-included both peer-reviewed studies and unpublished white papers to
-reduce potential bias against negative results. The inclusion criteria
-filtered out (1) non-field studies, (2) modelling results, (3) studies
-that did not evaluate specific BMPs, (4) studies conducted outside of
-the U.S. or published in a language other than English. We ran search
-queries in Texas A\&M Library Catalog, Web of Science, and Google
-Scholar. Although results from Google Scholar are not always replicable,
-we utilized the service to maximize search results for studies not
-published in academic journals and presumably increase the chance of
-identifying studies with negative effects. Fecal indicator bacteria
-study seraches included the following query: ``fecal indicator
-bacteria'' OR ``E. coli'' OR ``Escherichia coli'' OR ``enterococci'' OR
-``enterococcus'' AND ``best management practices'' OR ``BMPs'' AND
-``effectiveness'' OR ``performance''. Nutrient BMP studies utilized a
-similar query: ``nutrient'' OR ``nitrogen'' OR ``phosphorus'' OR
-``sediment'' OR ``TSS'' AND ``best management practices'' OR ``BMPs''
-AND ``effectiveness'' OR ``performance''.
+We conducted a systematic review of recent (2000-2022) literature to compile field studies documenting the effectiveness of best management practices on fecal indicator bacteria and nutrients concentrations.
+Prior meta-analysis have utilized data reported in the International Stormwater Database, which consists of self-reported and quality checked BMP data \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, horvathEffectsRegionalClimate2023}.
+The International Stormwater Database only recently added agricultural BMPs and has relatively sparse FIB data \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014}.
+Since we had interest in both FIB performance and agricultural BMPs we chose to utilize a systematic review.
+We also assume that most published reports undergo some method of internal or external peer review prior to publication, which is not necessarily the case with all of the data published in the International Stormwater Database.
+
+The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines \citep{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}.
+In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
+The inclusion criteria filtered out (1) non-field studies, (2) modelling results, (3) studies that did not evaluate specific BMPs, (4) studies conducted outside of the U.S.
+or published in a language other than English.
+We ran search queries in Texas A\&M Library Catalog, Web of Science, and Google Scholar.
+Although results from Google Scholar are not always replicable, we utilized the service to maximize search results for studies not published in academic journals and presumably increase the chance of identifying studies with negative effects.
+Fecal indicator bacteria study seraches included the following query: ``fecal indicator bacteria'' OR ``\emph{E.
+coli}'' OR ``\emph{Escherichia coli}'' OR ``enterococci'' OR ``enterococcus'' AND ``best management practices'' OR ``BMPs'' AND ``effectiveness'' OR ``performance''.
+Nutrient BMP studies utilized a similar query: ``nutrient'' OR ``nitrogen'' OR ``phosphorus'' OR ``sediment'' OR ``TSS'' AND ``best management practices'' OR ``BMPs'' AND ``effectiveness'' OR ``performance''.
 
 Results from each database were first filtered to remove duplicates.
-After removal of duplicates, each member of the research team (\emph{n}
-= 4) was assigned a subset of studies to evaluate if they should be
-included. \textbf{include supplementary tables with inclusion criteria
-and data variables} Each study was reviewed by two team members and
-differences in opinion were collectively discussed and agreed upon
-before progressing. The remaining studies were split among team members
-for data extraction, again with at least two team members reviewing each
-study. If data was provided in figures, the data was extracted with the
-WebPlotDigitizer tool \citep{rohatgiWebPlotDigitizer2022}.
+After removal of duplicates, each member of the research team (\emph{n} = 4) was assigned a subset of studies to evaluate if they should be included (Table \ref{tab:criteria}).
+Each study was reviewed by two team members and differences in opinion were collectively discussed and agreed upon before progressing.
+The remaining studies were split among team members for data extraction (Table \ref{tab:datvars}), again with at least two team members reviewing each study.
+If data was provided in figures, the data was extracted with the WebPlotDigitizer tool \citep{rohatgiWebPlotDigitizer2022}.
 
@@
 Parameter & The specific pollutant measured.\\
-Runoff source & Dominant source of runoff (rop fields, livestock pasture, commerical, residential\\
+Runoff source & Dominant source of runoff (crop fields, livestock pasture, commerical, residential\\
 Source type & Major categorization of runoff source: agricultural or urban\\
@@
 
-\hypertarget{effect-size-calculation}{%
-\subsection*{Effect size calculation}\label{effect-size-calculation}}
-\addcontentsline{toc}{subsection}{Effect size calculation}
-
-Prior BMP reviews routinely report the effect of BMPs using efficiency
-values
-\citep[see][]{agouridisLivestockGrazingManagement2005, krogerReviewBestManagement2012, liuEnhancingRainfallrunoffModel2015, simpsonDevelopingBestManagement2009}.
-Efficiency values are calculated as:
-
-\[
-\text{Efficiency}=\frac{x_{control}-x_{experiment}}{x_{control}}\times 100,
-\]
-
-where \emph{x\textsubscript{control}} is the pre-treatment or control
-pollutant concentration and \emph{x\textsubscript{experiment}} is the
-pollutant concentration measured after the BMP intervention. There are
-several statistical shortcomings (distributional asymetry, skewness, and
-non-additive properties) when using efficiency to estimate overall
-effect sizes
-\citep{nuzzoPercentDifferencesAnother2018, coleStatisticsNotesWhat2017}.
-In comparision, the log ratio of means (\emph{ROM}) provide preferable
-statistical properties for regression analysis
-\citep{osenbergEffectSizeEcological1997, hedgesMetaanalysisResponseRatios1999}.
-\emph{ROM} quantifies the difference in means between the control and
-experimental group \citep{hedgesMetaanalysisResponseRatios1999}:
-
-\[
-ROM_i = ln\left(\frac{x_{i,control}}{x_{i,experiment}}\right) = ln(x_{i, control})-ln(x_{i, experiment}),
-\] where \emph{x\textsubscript{i,control}} and
-\emph{x\textsubscript{i,experiment}} are the mean pollutant
-concentrations for experiment \emph{i}. The statistical properties of
-\emph{lnRR} (normal distribution around zero and additive properties)
-are preferable to the more commonly reported efficiency metric
-\citep{osenbergEffectSizeEcological1997, hedgesMetaanalysisResponseRatios1999}.
-Use of \emph{ROM} required the exclusion of studies that only provided
-measures of BMP efficiency and not the underlying data used to derive
-the metric.
-
 \hypertarget{statistical-models}{%
@@
 
-We used the ``rma.mv'' function in the \emph{metafor} R package to fit
-multilevel random effects linear models with \emph{ROM} as the effect
-variable
-\citep{viechtbauerConductingMetaanalysesMetafor2010, rcoreteamLanguageEnvironmentStatistical2023}.
-Our models specified a nested random effects term accounting for
-heterogeneity between effect sizes from the same study and for
-heterogeneity between studies. A key feature of meta-analysis is the
-weighting of effects using sampling variance of individual effect sizes.
-45\% of fecal indicator bacteria, 68\% of TN, 68\% of TP, and 70\% of
-TSS effect sizes were missing standard deviations required to estimate
-sampling variance. Removal of studies due to missing variance
-information can introduce substantial bias
-\citep{kambachConsequencesMultipleImputation2020}, so we imputed missing
-standard deviations using random forest based multivariate imputation
-with chained equations using the \emph{mice} R package
-\citep{buurenMiceMultivariateImputation2011}. Sampling variance was
-estimated with the variance estimator described in Doncaster et al.~to
-reduce bias in small sample studies:
+We used the ``rma.mv'' function in the \emph{metafor} R package to fit multilevel random effects regression models with \emph{ROM} as the effect variable \citep{viechtbauerConductingMetaanalysesMetafor2010, rcoreteamLanguageEnvironmentStatistical2023}.
+Our models specified a nested random effects term accounting for heterogeneity between effect sizes from the same study and for heterogeneity between studies.
+\emph{ROM} was used as the effect size which required the exclusion of studies that only provided measures of BMP efficiency and not the underlying data used to derive the metric.
+A key feature of meta-analysis is the weighting of effects using sampling variance of individual effect sizes.
+Fifty-nine percent of 222 effect sizes were missing standard deviations required to estimate sampling variance.
+Removal of studies due to missing variance information can introduce substantial bias \citep{kambachConsequencesMultipleImputation2020}.
+Missing standard deviations were imputed using the pooled ratio of the mean effect size to coefficient of variation (CV) \citep{brackenStatisticalMethodsAnalysis1992}.
+Sampling variance was estimated utilizing the average squared CV across all studies divided by sample size for each effect \citep{nakagawaRobustReadilyImplementable2023, doncasterCorrectionBiasMeta2018}:
 
@@
 
-where \emph{v} represents the sampling variance, \(CV_{control,i}\) and
-\(CV^2_{experiment,i}\) are the coefficients of variation from the
-\emph{i}th study for studies 1, 2, \ldots, \emph{K}.
+where \emph{v} represents the sampling variance, \(CV_{control,i}\) and \(CV^2_{experiment,i}\) are the coefficients of variation from the \emph{i}th study for studies 1, 2, \ldots, \emph{K}.
 
-Our initial models included log transformed pre-treatment pollutant
-concentration, study scale (plot/field, community, watershed), study
-length (\emph{n} years), BMP subcategory (), and pollutant
-concentration:BMP subcategory interactions were included as fixed effect
-terms. We used an information-theoretic approach to select the most
-parsimonious model from the subset of candidate models based on
-corrected Akaike information criterion (AIC\textsubscript{c}) estimated
-with maximum likelihood \citep{cinarUsingInformationTheoretic2021}. The
-final model was selected from candidate models, which included all
-combination and subsets of the full model, by eliminating candidate
-models until the \(\Delta\)\emph{AIC\textsubscript{c}} reached \(\leq\)
-2 \citep{burnhamAICModelSelection2011}. We estimated regression
-coefficients of the selected model using restricted maximum likelihood.
-Relative heterogeneity between and within studies were calculated using
-the \emph{I\textsuperscript{2}} metric described in
-\citet{nakagawaMethodologicalIssuesAdvances2012}. Marginal
-\emph{R\textsuperscript{2}} was used to describe the amount of variance
-explained by fixed effects \citep{nakagawaGeneralSimpleMethod2013}.
+Our initial models included log transformed influent concentration, BMP subcategory (drainage modification, crop field management, livestock management, filtration, treatment, detention, or infiltration), aridity index (mean-centered), influent concentration\texttimes BMP subcategory interactions, and aridity index\texttimes BMP subcategory interactions were included as fixed effect terms.
+Aridity index was the only moderator not obtained directly in the systematic review (Table \ref{tab:datvars}).
+We mapped study location coordinates to aridity index values published in the ``Global Aridity Index and Potential Evapotranspiration Database - Version 3'' (Global-AI\_PET\_v3) which provides gridded 30 arc-second annual average precipitation and potential evapotranspiration estimates \citep{zomerVersionGlobalAridity2022}.
+The aridity index is calculated as the ratio of mean annual precipitation to mean annual evapotraspiration with values between 0 and 0.5 considered hyper to semi-arid, and values above 0.65 as humid.
 
-We tested for evidence of publication bias, in the form of small study
-effect, by using the extension of Egger's regression applied to the
-multilevel model framework that included adjusted sampling error as a
-moderator \citep{nakagawaQuantitativeEvidenceSynthesis2023}. We did not
-identify evidence of publication bias in the surveyed studies for fecal
-indicator bacteria (\emph{Z} = -1.17, \emph{p} = 0.25), total nitrogen,
-total phosphorus, or total suspended solids; therefore, adjustments for
-publication bias were not included in the final models. We conducted a
-sensitivity analysis of the robustness of overall effect sizes to
-individual studies using leave-one-out analysis
-\citep{nakagawaQuantitativeEvidenceSynthesis2023}. Outlier studies
-identified in the sensitivity analysis are highlighted in Table X and
-were removed prior to fitting the final models.
+We used an information-theoretic approach to select the most parsimonious model from the subset of candidate models based on corrected Akaike information criterion (AIC\textsubscript{c}) estimated with maximum likelihood \citep{cinarUsingInformationTheoretic2021}.
+Candidate models used for variable selection were fit with maximum likelihood (ML).
+The final model was selected from candidate models, which included all combination and subsets of the full model, by selecting the model with the lowest AIC\textsubscript{c} score \citep{burnhamAICModelSelection2011, cinarUsingInformationTheoretic2021}.
+Regression coefficients of the selected model were estimated using restricted maximum likelihood (REML).
+Relative heterogeneity between and within studies were calculated using the \emph{I\textsuperscript{2}} metric described in \citet{nakagawaMethodologicalIssuesAdvances2012}.
+Marginal \emph{R\textsuperscript{2}} was used to describe the amount of variance explained by fixed effects \citep{nakagawaGeneralSimpleMethod2013}.
+
+We tested for evidence of publication bias, in the form of small study effect, by using the extension of Egger's regression applied to the multilevel model framework that included adjusted sampling error as a moderator \citep{nakagawaQuantitativeEvidenceSynthesis2023}.
+We did not identify evidence of publication bias in the surveyed studies (FIB: ROM = 1.19, 95\% CI {[}-3.23, 5.61{]}; TN: ROM = 0.31, 95\% CI {[}-1.15, 1.78{]}; DIN: ROM = -2.4, 95\% CI {[}-6.14, 1.59{]}; TP: ROM = -1.97, 95\% CI {[}-5.18, 1.23{]}; PO\textsubscript{4}: -0.05, 95\% CI {[}-2.82, 2.72{]}; TSS: ROM: 0.31, 95\% CI {[}-1.15, 1.78{]}; Figure S1-S6); therefore, adjustments for publication bias were not included in the final models.
+We conducted a sensitivity analysis of the robustness of overall effect sizes to individual studies using leave-one-out analysis \citep{nakagawaQuantitativeEvidenceSynthesis2023}.
+This approach repeatedly fits the selected model leaving out an individual value each time.
+The overall effect and 95\% CI from each refit model is compared to the overall effect and 95\% CI of the model fit to the full dataset.
+We did not identify evidence of outliers or overly influential studies for any of our models (Figure S7-S12).
 
@@
 
-\hypertarget{regression-results}{%
-\subsection{Regression results}\label{regression-results}}
+There were a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effects for PO\textsubscript{4}, and 33 studies with 125 effect sizes for TSS.
+The majority of studies were identified as smaller scaled lot or field studies (Figure \ref{fig:bmpsummary} A).
+FIB studies had a roughly equal proportion large watershed/catchment studies and studies conducted at the commuity/farm scale or smaller.
+We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \ref{fig:bmpsummary} B).
+We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \ref{fig:bmpsummary} C).
+Our review was resticted to studies published after 1999.
+The number of studies published for each parameter were roughly uniformly distributed over time (Figure \ref{fig:bmptemporal} A) and are not indicative of increases or decreases in the number of published studies.
+Study length was strongly skewed for all parameters (Figure \ref{fig:bmptemporal} B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO\textasciitilde4), 2.5 (TN), 2.5 (TP), and 2 (TSS).
+There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \ref{fig:studymap}).
+This might suggest that spatially correlated predictors such as climate and soil variables would be difficult to model due to the poor spatial representation of the collected data.
 
-Some evidence of a relationship between input concentrations and fecal
-indicator bacteria reductions. No evidence strong evidence that the type
-of BMP changed the overall relationship.
+\hypertarget{regression-models}{%
+\subsection{Regression models}\label{regression-models}}
 
@@
 
-\caption{Summary table of moderator effects for performance of BMPs on fecal indicator bacteria concentrations.}
+\caption{\label{tab:modelaic}Model selection AICc}
 \centering
-\begin{tabular}[t]{lrlrrrl}
+\begin{tabular}[t]{lllllll}
 \toprule
-Moderator & \makecell[r]{Estimate\\(\textit{lnRR})} & 95\% CI & SE & \textit{T}-statistic & df & \textit{p}-value\\
+\multicolumn{1}{c}{ } & \multicolumn{6}{c}{AICc} \\
+\cmidrule(l{3pt}r{3pt}){2-7}
+Candidate Models & FIB & TN & DIN & TP & PO\textsubscript4 & TSS\\
 \midrule
-Intercept & 0.21 & {}[-2.07,2.48] & 1.06 & 0.19 & 14 & 0.85\\
-\addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{BMP Subcategories}}\\
-\hspace{1em}Filtration & -0.39 & {}[-6.42,5.63] & 2.81 & -0.14 & 14 & 0.89\\
-\hspace{1em}Infiltration & -1.24 & {}[-3.50,1.01] & 1.13 & -1.10 & 64 & 0.28\\
-\hspace{1em}Livestock & -1.18 & {}[-3.25,0.89] & 0.96 & -1.23 & 14 & 0.24\\
-\hspace{1em}Treatment & -1.56 & {}[-2.63,-0.49] & 0.53 & -2.92 & 64 & \textless 0.01\\
-log(Concentration) & 0.20 & {}[0.084,0.32] & 0.059 & 3.42 & 65 & \textless 0.01\\
+$\sim$ Int & 208.9 & \textbf{37.7} & \textbf{112.6} & 96.7 & 42.1 & \textbf{103.9}\\
+$\sim$ Int+Influent & 197.4 & 39.8 & 114.6 & \textbf{96.3} & \textbf{36.8} & 106.4\\
+$\sim$ Int+AI & 209 & 40.3 & 114.8 & 98.2 & 45.2 & 106.5\\
+$\sim$ Int+AI+Influent & 195 & 42.5 & 116.9 & 98.9 & 40.2 & 109.4\\
+$\sim$ Int+BMP & 209.1 & 47.3 & 121 & 101.8 & 51.8 & 115.3\\
+\addlinespace
+$\sim$ Int+BMP+Influent & 196.7 & 48.9 & 124 & 102.3 & 45.6 & 120.4\\
+$\sim$ Int+AI+BMP & 210.1 & 50.8 & 124.1 & 104.8 & 56.4 & 120.7\\
+$\sim$ Int+AI+BMP+Influent & 195.6 & 52.9 & 127.3 & 105.6 & 50.8 & 126.6\\
+$\sim$ Int+BMP\texttimes Influent & 201.3 & 67.5 & 133.7 & 121.5 & 62.9 & 174.1\\
+$\sim$ Int+AI\texttimes BMP & 208.1 & 75.1 & 137.6 & 121 & 67.7 & 175.5\\
+\addlinespace
+$\sim$ Int+Influent+AI\texttimes BMP & \textbf{193.7} & 77.2 & 141.4 & 121.6 & 70.7 & 191.4\\
+$\sim$ Int+AI+BMP\texttimes Influent & 202 & 83.2 & 145.9 & 127.3 & 72.9 & 191.9\\
+$\sim$ Int+AI\texttimes BMP+BMP\texttimes Influent & 201.9 & 116.8 & 156.2 & 142.7 & 94.8 & 188.5\\
 \bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}$I^{2}_{\text{total}}$=21.48, $I^{2}_{\text{study}}$=0, $I^{2}_{\text{effect}}$=21.48; $R^{2}_{\text{marginal}}$=0.46}\\
+\multicolumn{7}{l}{\rule{0pt}{1em}Int = intercept; AI = ariditiy index; BMP = BMP subcategory}\\
 \end{tabular}
@@
 
-\begin{table}
+\hypertarget{fecal-indicator-bacteria}{%
+\subsubsection{Fecal Indicator Bacteria}\label{fecal-indicator-bacteria}}
 
-\caption{Summary table of moderator effects for performance of BMPs on nitrogen removal.}
-\centering
-\begin{tabular}[t]{lrlrrrl}
-\toprule
-Moderator & \makecell[r]{Estimate\\(\textit{lnRR})} & 95\% CI & SE & \textit{T}-statistic & df & \textit{p}-value\\
-\midrule
-Intercept & 0.95 & {}[0.38,1.52] & 0.28 & 3.41 & 28 & \textless 0.01\\
-log(Concentration) & -0.034 & {}[-0.14,0.071] & 0.053 & -0.64 & 124 & 0.525\\
-Adjusted sampling error & -1.54 & {}[-3.30,0.22] & 0.89 & -1.74 & 124 & 0.085\\
-\bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}$I^{2}_{\text{total}}$=96.57, $I^{2}_{\text{study}}$=52.45, $I^{2}_{\text{effect}}$=44.12; $R^{2}_{\text{marginal}}$=0.05}\\
-\end{tabular}
-\end{table}
+There were only 19 studies and 63 FIB effect sizes available to model after removal of studies and effects that only reported BMP\textsubscript{eff}.
+The overall mean effect (estimated with the intercept only multilevel random effects model) showed significant mean reductions in FIB (ROM = 0.85, 95\% CI {[}0.36, 1.34{]}; BMP\textsubscript{eff} = 57.4\%, 95\% CI {[}30.4\%, 73.9\%{]}; Figure \ref{fig:overalleffect}) resulting from BMPs.
+Total heterogeneity was moderate with a relatively large amount of heterogeneity observed due to differences within studies (\textit{I\textsuperscript{2}\textsubscript{total}} = 53.54, \textit{I\textsuperscript{2}\textsubscript{study}} = 10.03, \textit{I\textsuperscript{2}\textsubscript{effect}} = 43.51).
 
-\begin{table}
+AICc scores included log transformed influent concentration and the aridity index\texttimes BMP subcategory interaction as moderators for the FIB model (Table \ref{tab:modelaic}).
+Moderator terms and interactions explained a high proportion of effect size variance (\textit{R\textsuperscript{2}\textsubscript{marginal}} =0.89) in the FIB model.
+Increased influent concentrations (\(\beta\) = 0.25, 95\% CI {[}0.14, 0.37{]}) resulted in significantly larger ROM effect for FIB (Figure \ref{fig:fibeffect}.
+Compared to the baseline aridity index\texttimes detention BMP subcategory interaction, infiltration (\(\beta\) = -29.90, 95\% CI {[}-50.34, -9.47{]}), livestock management (\(\beta\) = -30.37, 95\% CI {[}-50.93, -9.81{]}), and treatment (\(\beta\) = -30.33, 95\% CI {[}-49.62, -11.03{]}) interactions had significantly smaller slopes.
+However, the data had uneven coverage of BMP subcategories across the aridity index.
+Effects for detention BMPs were clustered in humid climates (aridity index \textgreater{} 0.65) and the resulting estimate for the baseline interaction (\(\beta\) = 32.63, 95\% CI {[}12.57, 52.69{]}) may not be reliable.
 
-\caption{Summary table of moderator effects for performance of BMPs on phosphorus removal.}
-\centering
-\begin{tabular}[t]{lrlrrrl}
-\toprule
-Moderator & \makecell[r]{Estimate\\(\textit{lnRR})} & 95\% CI & SE & \textit{T}-statistic & df & \textit{p}-value\\
-\midrule
-Intercept & 0.40 & {}[-0.14,0.93] & 0.26 & 1.52 & 28 & 0.14\\
-log(Concentration) & 0.27 & {}[0.15,0.40] & 0.063 & 4.30 & 97 & \textless 0.01\\
-\bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}$I^{2}_{\text{total}}$=15.54, $I^{2}_{\text{study}}$=0, $I^{2}_{\text{effect}}$=15.54; $R^{2}_{\text{marginal}}$=0.32}\\
-\end{tabular}
-\end{table}
+\hypertarget{nitrogen}{%
+\subsubsection{Nitrogen}\label{nitrogen}}
 
-\hypertarget{subsection-2}{%
-\subsection{Subsection 2}\label{subsection-2}}
+We identified 13 eligible TN studies and 14 DIN studies and 31 and 44 effect sizes respectively that could be included in the regression model.
+Overall effects showed that BMPs resulted in significant mean reductions in TN (ROM = 0.42, 95\% CI {[}0.21, 0.62{]}; BMP\textsubscript{eff} = 34.0\%, 95\% CI {[}18.7\%, 46.4\%{]}; Figure \ref{fig:overalleffect}) but not in DIN (ROM = 0.64, 95\% CI {[}-0.08, 1.35{]}; BMP\textsubscript{eff} = 47.1\%, 95\% CI {[}-8.1\%, 74.1\%{]}; Figure \ref{fig:overalleffect}).
+Heterogeneity was high for TN with a large proportion of heterogeneity attributed to within study effect (\textit{I\textsuperscript{2}\textsubscript{total}} = 77.12, \textit{I\textsuperscript{2}\textsubscript{study}} = 23.2, \textit{I\textsuperscript{2}\textsubscript{effect}} = 53.92).
+The DIN model had even higher heterogeneity with a larger proportion attributed between studies (\textit{I\textsuperscript{2}\textsubscript{total}} = 99.51, \textit{I\textsuperscript{2}\textsubscript{study}} = 83.53, \textit{I\textsuperscript{2}\textsubscript{effect}} = 15.97).
+AICc scores indicated that none of the moderators resulted in substantial improvement over the intercept only model (Table \ref{tab:modelaic}).
 
-Frontiers requires figures to be submitted individually, in the same
-order as they are referred to in the manuscript. Figures will then be
-automatically embedded at the bottom of the submitted manuscript. Kindly
-ensure that each table and figure is mentioned in the text and in
-numerical order. Permission must be obtained for use of copyrighted
-material from other sources (including the web). Please note that it is
-compulsory to follow figure instructions. Figures which are not
-according to the guidelines will cause substantial delay during the
-production process.
+\hypertarget{phosphorus}{%
+\subsubsection{Phosphorus}\label{phosphorus}}
+
+We found 17 TP studies with 37 effect sizes and 9 PO\textsubscript{4} studies with 21 effect sizes for inclusion in regression models.
+There was a significant overall reduction found for TP (ROM = 0.40, 95\% CI {[}0.03, 0.76{]}; BMP\textsubscript{eff} = 32.7\%, 95\% CI {[}3.4\%, 53.2\%{]}) but no evidence of negative or positive effect for PO\textsubscript{4} (ROM = -0.18, 95\% CI {[}-0.56, 0.19{]}; BMP\textsubscript{eff} = -20.1\%, 95\% CI {[}-75.3\%, 17.7\%{]}).
+For both the TP and PO\textsubscript{4} models, heterogenity was high, with moderate to high within study variance and low to moderate between study variance (TP: \textit{I\textsuperscript{2}\textsubscript{total}} = 96.13, \textit{I\textsuperscript{2}\textsubscript{study}} = 32.15, \textit{I\textsuperscript{2}\textsubscript{effect}} = 63.99; PO\textsubscript{4}: \textit{I\textsuperscript{2}\textsubscript{total}} = 97.28, \textit{I\textsuperscript{2}\textsubscript{study}} = 33.36, \textit{I\textsuperscript{2}\textsubscript{effect}} = 63.92).
+The best model for both parameters only included influent as a moderator (Table \ref{tab:modelaic}).
+Influent concentration (\(\beta\) = 0.23, 95\% CI {[}-0.035, 0.49{]}) was not significant at the 95\% confidence level for the TP model (Figure \ref{fig:phoseffect}; Table S4).
+Influent concentration (\(\beta\) = 0.27, 95\% CI {[}0.085, 0.44{]}) was significant for the PO\textsubscript{4} model (Figure \ref{fig:phoseffect}; Table S5).
+
+\hypertarget{sediment}{%
+\subsubsection{Sediment}\label{sediment}}
+
+There were 12 eligible TSS studies with 26 effect sizes for regression modelling.
+We found a significant and large reduction in TSS concentrations across studies (ROM = 1.65, 95\% CI {[}0.96, 2.34{]}; BMP\textsubscript{eff} = 80.9\%, 95\% CI {[}61.9\%, 90.4\%{]}).
+Heterogeneity was high for TSS with a large proportion of heterogeneity attributed to within study effect (\textit{I\textsuperscript{2}\textsubscript{total}} = 99.57, \textit{I\textsuperscript{2}\textsubscript{study}} = 0, \textit{I\textsuperscript{2}\textsubscript{effect}} = 99.57).
+Similar to nitrogen, we did not find strong evidence linking any of the tested moderators to BMP performance (Table \ref{tab:modelaic}).
 
@@
 
-Linking BMP performance to watershed scale improvements remains a
-challenge (Tomer and Locke, Melland et al, Meals). The succesful
-implementation and maintence of BMPs at watershed scales remains an area
-for more investigation (Liu et al., Lintern) Long-term
-trans-disciplinary monitoring projects not just on BMP performance but
-maintance, management, and social apsects may elucidate unknown factors
-influencing the ability to scale BMP reductions to watershed scale
-improvements (Lintern et al.)
+Our systematic review revealed strong spatial disparities in published BMP studies (Figure \ref{fig:studymap}.
 
-Runoff volume was not incorporated into our models, largely attributed
-to lack of available data. However, BMP size is typically scaled to
-accommodate runoff volumes of a particular return period volume (1-yr or
-2-yr flows for example) (Hoss et al.).
+Meta-analysis indicated that BMPs showed significant overall reductions in FIB, TN, TP, and TSS concentrations when using ROM as an effect size.
+We did not find strong evidence of overall leaching or reductions of DIN or PO\textsubscript{4} in BMP studies.
+The results are in agreement to some extent with previous reviews that found effective (but highly variable) removal efficiencies for nitrogen, phosphorus, and sediment \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, liuReviewEffectivenessBest2017}.
+
+compare overall effects to koch, barrett, clary
+
+Linking BMP performance to watershed scale improvement remains a challenge (Tomer and Locke, Melland et al, Meals).
+The succesful implementation and maintence of BMPs at watershed scales remains an area for more investigation (Liu et al., Lintern)
+Long-term trans-disciplinary monitoring projects not just on BMP performance but maintance, management, and social apsects may elucidate unknown factors influencing the ability to scale BMP reductions to watershed scale improvements (Lintern et al.)
+
+Runoff volume was not incorporated into our models, largely attributed to lack of available data.
+However, BMP size is typically scaled to accommodate runoff volumes of a particular return period volume (1-yr or 2-yr flows for example) (Hoss et al.).
 
 \hypertarget{data-availability-statement}{%
-\section*{Data availability
-statement}\label{data-availability-statement}}
+\section*{Data availability statement}\label{data-availability-statement}}
 \addcontentsline{toc}{section}{Data availability statement}
@@
 \hypertarget{disclosureconflict-of-interest-statement}{%
-\section*{Disclosure/Conflict-of-Interest
-Statement}\label{disclosureconflict-of-interest-statement}}
-\addcontentsline{toc}{section}{Disclosure/Conflict-of-Interest
-Statement}
+\section*{Disclosure/Conflict-of-Interest Statement}\label{disclosureconflict-of-interest-statement}}
+\addcontentsline{toc}{section}{Disclosure/Conflict-of-Interest Statement}
 
-The authors declare that the research was conducted in the absence of
-any commercial or financial relationships that could be construed as a
-potential conflict of interest.
+The authors declare that the research was conducted in the absence of any
+commercial or financial relationships that could be construed as a potential
+conflict of interest.
 
@@
 
-MS: Conceptualization, Formal analysis, Funding acquisition,
-Methodology, Supervision, Writing---original draft, Writing---review \&
-editing. DK: Conceptualization, Formal analysis, Data collection,
-Writing---original draft, Writing---review \& editing. JW: Data
-collection, Writing---review \& editing. SJ: Data collection;
-Writing---review \& editing.
+MS: Conceptualization, Formal analysis, Funding acquisition, Methodology, Supervision, Writing---original draft, Writing---review \& editing.
+DK: Conceptualization, Formal analysis, Data collection, Writing---original draft, Writing---review \& editing.
+JW: Data collection, Writing---review \& editing.
+SJ: Data collection; Writing---review \& editing.
 
@@
 
-Funding: This project was supported by a state nonpoint source grant
-from the Texas State Soil and Water Conservation Board.
+Funding: This project was supported by a state nonpoint source grant from the Texas State Soil and Water Conservation Board.
 
@@
 
-Supplementary Material should be uploaded separately on submission, if
-there are Supplementary Figures, please include the caption in the same
-file as the figure. LaTeX Supplementary Material templates can be found
-in the Frontiers LaTeX folder
+Supplementary Material should be uploaded separately on submission, if there are
+Supplementary Figures, please include the caption in the same file as the
+figure. LaTeX Supplementary Material templates can be found in the Frontiers
+LaTeX folder
 
@@
 
-\begin{figure}[H]
-\includegraphics[width=1\linewidth,]{../figures/study_map} \caption{Distribution of studies identified in the systematic review by state and parameter.}\label{fig:studymap}
-\end{figure}
-
 \begin{figure}
@@
 
-\begin{figure}[H]
-\includegraphics[width=1\linewidth,]{../figures/overall_effect} \caption{Estimated effect sizes and intervals from the null multilevel random effects model.
+\begin{figure}
+\includegraphics[width=1\linewidth,]{../figures/bmp_temporal_summary} \caption{Number of studies identified in the systematic review summarized by (A) publication date and (B) study length.}\label{fig:bmptemporal}
+\end{figure}
+
+\begin{figure}[p]
+\includegraphics[width=1\linewidth,]{../figures/study_map} \caption{Distribution of studies identified in the systematic review by state and parameter.}\label{fig:studymap}
+\end{figure}
+
+\begin{figure}[p]
+\includegraphics[width=1\linewidth,]{../figures/overall_effect} \caption{Estimated effect sizes and intervals from the intercept only multilevel random effects model.
 Individual points represent studies, with size scaled by sampling variance.
@@
 
+\begin{figure}[p]
+\includegraphics[width=1\linewidth,]{../figures/bac_predicted} \caption{Predicted marignal effect of influent FIB and aridity index (conditioned on BMP subcategory).
+Solid lines are the predicted mean effect, dashed lines are the 95\% confidence intervals, and the dotted lines are the 95\% prediction intervals.
+Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.}\label{fig:fibeffect}
+\end{figure}
+
+\begin{figure}[p]
+\includegraphics[width=1\linewidth,]{../figures/phos_predicted} \caption{Predicted marignal effect of influent pollutant concentration on TP (Panel A) and PO\textsubscript{4} (Panel B) reductions.
+Solid lines are the predicted mean effect, dashed lines are the 95\% confidence intervals, and the dotted lines are the 95\% prediction intervals.
+Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.}\label{fig:phoseffect}
+\end{figure}
+
 \end{document}

### 2024-01-31 R/01_read_data.R

@@ -49,3 +49,32 @@ load_df <- function(dataset) {
   }
   return(df)
 }
+
+
+# download Global Aridity data https://figshare.com/ndownloader/files/34377245
+download_aridity <- function(url) {
+  dir <- "data"
+  temp_path <- tempfile(fileext = ".zip")
+  if (file.exists(temp_path))  'file alredy exists' else httr::GET(url = url, httr::write_disk(temp_path))
+  unzip(zipfile = temp_path, exdir = dir)
+}
+
+
+
+add_aridity <- function(df) {
+  ai <- rast("data/Global-AI_ET0_v3_annual/ai_v3_yr.tif")
+  ai <- ai * 0.0001 
+  
+  df <- df |> 
+    filter(!is.na(Longitude), !is.na(Latitude))
+  ai_vals <- terra::extract(ai, vect(df, geom = c("Longitude", "Latitude"), crs ="+proj=longlat +datum=WGS84"))
+  df$ai <- ai_vals$awi_pm_sr_yr
+  
+  ai_centered <- scale(df$ai, scale = FALSE)
+  df <- df |> 
+    mutate(ai = ai_centered)
+  
+  return(list(model_df = df,
+              mean = attr(ai_centered, "scaled:center")))
+  
+}

### 2024-01-31 R/02_fit_models.R

@@ -10,11 +10,13 @@
 #'   data used to fit the model.
 fit_bac_model <- function(data) {
   df <- data |> 
-    select(StudyID, Study, ValueID, Parameter, Source_Type, BMP_SubCat, Scale, study_length, n_Pre, n_Post, Pre, Post, sd_Pre, sd_Post, Unit) |> 
+    select(StudyID, Study, ValueID, Parameter, Source_Type, BMP_SubCat, Scale, study_length, n_Pre, n_Post, Pre, Post, sd_Pre, sd_Post, Unit, ai) |> 
     filter(Unit %in% c("CFU/100ml", "MPN/100ml", "Geometric mean (CFU/100ml)", "Geomean (CFU/100 mL)")) |> 
     filter(!is.na(Scale)) |> 
-    filter(!is.na(Pre), !is.na(Post)) |> 
-    filter(BMP_SubCat != "Drainage") |> # only one drainage observations, can't model the drainage x lnPre interaction.
+    filter(!is.na(Pre), !is.na(Post)) |>
+    ## removing categories that result in rank deficiencies
+    filter(BMP_SubCat != "Drainage") |> 
+    filter(BMP_SubCat != "Filtration") |> 
     mutate(StudyID = fct_drop(StudyID),
            Study = fct_drop(Study),
            lnPre = log(Pre),
@@ -23,16 +25,10 @@ fit_bac_model <- function(data) {
            Scale = forcats::fct_relevel(Scale, "Lot", "Community", "Watershed"))
   
   ##impute SD
-  predMatrix <- make.predictorMatrix(df)
-  predMatrix[,"StudyID"] <- 0
-  predMatrix[,"Study"] <- 0
-  predMatrix[,"ValueID"] <- 0
-  predMatrix[,"Parameter"] <- 0
-  impMethod <- make.method(df)
-  impMethod[c("sd_Pre", "sd_Post")] <- "rf"
-  
-  imp <- mice(df, print=FALSE, m=20, predictorMatrix=predMatrix, method=impMethod, seed=1234)
-  df <- complete(imp)
+  df <- metagear::impute_SD(data.frame(df),
+                            c("sd_Pre", "sd_Post"),
+                            c("Pre", "Post"),
+                            method = "Bracken1992")
 
   
   ## Prep data for rma
@@ -51,6 +47,7 @@ fit_bac_model <- function(data) {
   # impute the variance covariance matrix
   VCV <- vcalc(vi = vi,
                cluster = StudyID,
[…]

### 2024-01-31 R/03_plot_models.R

@@ -33,6 +33,8 @@ plot_marginal_effect_lnRR <- function(model,
                                       subset = NULL,
                                       sec_y_breaks = waiver(),
                                       sec_y_labels = waiver(),
+                                      ylim = NULL,
+                                      x_adj = NULL,
                                       ...) {
   
   ## use mod_results from orchaRd to generate model grid 
@@ -56,6 +58,16 @@ plot_marginal_effect_lnRR <- function(model,
                                moderator = exp(moderator)))    
   }
 
+  ## back transform centered ai data
+  if(mod == "ai") {
+    mod_table <- as.list(mod_table) |> 
+      purrr::modify_at("mod_table",
+                       ~mutate(.,
+                               moderator = moderator + x_adj)) |> 
+      purrr::modify_at("data",
+                       ~mutate(.,
+                               moderator = moderator + x_adj))    
+  }
   
   ## remove NA values from "data"
   if(!is.null(at)){
@@ -144,43 +156,22 @@ plot_marginal_effect_lnRR <- function(model,
                                     color = "black",
                                     face = "plain",
                                     size = rel(1), hjust = 0))
-  # p1 <- mod_table |>
-  #   ## use modified bubble_plot
-  #   bubble_plot(group = "StudyID", mod = mod, g = TRUE, condition.nrow = 1, ...) +
-  #   geom_hline(yintercept = 0, alpha = 0.25, linetype = "solid", color = "gray75") +
-  #   scale_y_continuous(sec.axis = sec_axis(~(1-(1/exp(.))) * 100,
-  #                                          name = "Efficiency (%)",
-  #                                          breaks = sec_y_breaks,
-  #                                          labels = sec_y_labels)) +
-    # labs(x = xlab,
-    #      y = ylab) +
-    # theme_mps_noto(base_family = "Manrope Regular") +
-    # theme(axis.title = element_text(family = "Manrope SemiBold", face = "plain", size = rel(1)),
-    #       axis.title.x = element_text(hjust = 0),
-    #       axis.title.y = element_text(hjust = 1),
-    #       axis.title.y.right = element_text(hjust = 0, angle = 270),
-    #       axis.text = element_text(family = "Manrope Light", face = "plain", size = rel(0.8)),
-    #       axis.text.y.left = element_text(hjust = 1),
[…]

### 2024-01-31 R/04_tables.R

@@ -42,7 +42,7 @@ tidy_metafor <- function(model,
                   booktabs = TRUE,
                   caption = caption,
                   col.names = linebreak(c("Moderator", 
-                                          "Estimate\n(\\textit{lnRR})",
+                                          "Estimate",
                                           "95\\% CI",
                                           "SE",
                                           "\\textit{T}-statistic",

### 2024-01-31 R/05_literature_statistics.R

@@ -1,3 +1,61 @@
+
+study_count_summary <- function(bac_df, nut_df) {
+  
+  ## summarize study
+  study_bac <- bac_df |> 
+    group_by(StudyID) |> 
+    summarize(n = 1) |> 
+    ungroup() |> 
+    summarise(Study_n = sum(n)) |> 
+    mutate(Parameter = "FIB")
+  ##summarize effect
+  effect_bac <- bac_df |> 
+    group_by(ValueID) |> 
+    summarize(n = 1) |> 
+    ungroup() |> 
+    summarise(Effect_n = sum(n)) |> 
+    mutate(Parameter = "FIB")
+  
+  
+  study_df <- nut_df |> 
+    mutate(Parameter = case_when(
+      Parameter %in% c("TN") ~ "TN",
+      Parameter %in% c("NO3", "NOX", "NH3") ~ "DIN",
+      Parameter %in% c("TP") ~ "TP",
+      Parameter %in% c("PO4") ~ "PO4",
+      Parameter == "TSS" ~ "TSS",
+      .default = NA
+    )) |> 
+    filter(!is.na(Parameter)) |> 
+    group_by(StudyID, Parameter) |> 
+    summarize(n = 1) |> 
+    ungroup() |> 
+    group_by(Parameter) |> 
+    summarise(Study_n = sum(n)) |> 
+    bind_rows(study_bac)
+  
+  effect_df <- nut_df |> 
+    mutate(Parameter = case_when(
+      Parameter %in% c("TN") ~ "TN",
+      Parameter %in% c("NO3", "NOX", "NH3") ~ "DIN",
+      Parameter %in% c("TP") ~ "TP",
+      Parameter %in% c("PO4") ~ "PO4",
+      Parameter == "TSS" ~ "TSS",
+      .default = NA
+    )) |> 
+    filter(!is.na(Parameter)) |> 
+    group_by(ValueID, Parameter) |> 
+    summarize(n = 1) |> 
+    ungroup() |> 
+    group_by(Parameter) |> 
+    summarise(Effect_n = sum(n)) |> 
+    bind_rows(effect_bac)
+  
+  df <- left_join(study_df, effect_df) 
+  df
+}
+
+
 plot_map <- function(bac_df,
                      nut_df) {
   bac_df <- bac_df |>  
@@ -5,13 +63,15 @@ plot_map <- function(bac_df,
     summarise(n = 1) |> 
     ungroup() |> 
[…]

### 2024-01-31 _targets.R

@@ -3,20 +3,21 @@ library(targets)
 
 tar_option_set(packages = c("broom",
                             "dplyr",
-                            "flextable",
                             "forcats",
                             "ggplot2",
+                            "ggforce",
                             "ggtext",
                             "ggrepel",
+                            "grid",
+                            "gridtext",
+                            "httr",
                             "kableExtra",
                             "latex2exp",
                             "metafor",
                             "metagear",
-                            "mice",
                             "modelr",
                             "mpsTemplates",
                             "MuMIn",
-                            "officer",
                             "orchaRd",
                             "patchwork",
                             "purrr",
@@ -24,6 +25,7 @@ tar_option_set(packages = c("broom",
                             "readxl",
                             "scales",
                             "sf",
+                            "terra",
                             "tibble",
                             "tidyr",
                             "tigris",
@@ -51,11 +53,37 @@ eval(metafor:::.MuMIn)
 
 
 list(
+  # Data -----------------------------------------------------------------------
+  
   tar_target(bacteria_df, load_df("bacteria")),
   tar_target(nutrient_df, load_df("nutrients")),
+  #https://stackoverflow.com/questions/70281450/dealing-with-zip-files-in-a-targets-workflow
+  #https://figshare.com/ndownloader/files/34377245
+
+  tar_target(global_ai, download_aridity("https://api.figshare.com/v2/file/download/34377245"),
+             format = "file"),
+  tar_target(ai_file_names, global_ai), # not a format = "file" target
+  tar_target(
+    files, {
+      global_ai # Re-hash all the files separately if any file changes.
+      ai_file_names
+    },
+    pattern = map(ai_file_names),
+    format = "file"
+  ),
+  
+  ## add aridity index to data -------------------------------------------------
+  tar_target(bac_model_df, add_aridity(bacteria_df)),
+  tar_target(nut_model_df, add_aridity(nutrient_df)),
+  
+  # Summary stats
+  tar_target(study_count, study_count_summary(bacteria_df,
+                                              nutrient_df)),
+  
+  
   # Bacteria data analysis -----------------------------------------------------
   ## fit multilevel random effects model ---------------------------------------
-  tar_target(bac_model, fit_bac_model(bacteria_df)),
+  tar_target(bac_model, fit_bac_model(bac_model_df$model_df)),
   
   ## Diagnostics and assessment ------------------------------------------------
[…]

### 2024-01-31 supplementary_materials/supplementary_materials.Rmd
start supplementary materials

---
title: "Supplementary Materials"
author:
  - name: Michael Schramm
    affiliation: '1'
    etal: Schramm # First author's last name. 
    email: michael.schramm@ag.tamu.edu # Indicates corresponding Author
    institution: Texas Water Resources Institute, Texas A&M AgriLife Research
    street: 1001 Holleman Dr E.
    city: College Station
    state: Texas # only USA, Australia, Canada
    zip: 77840
    country: USA
  - name: Duncan Kikoyo
    affiliation: '1'
  - name: Janelle Wright
    affiliation: '1'
  - name: Shubham Jain
    affiliation: '1'
affiliation:
  - id: '1'
    department: Texas Water Resources Institute
    institution: Texas A&M AgriLife Research
    city: College Station
    state: Texas # only USA, Australia, Canada
    country: USA
output: 
  bookdown::pdf_book:
    template: template.tex
# bib yaml is commented to manually insert in correct section
# csl: frontiers.csl
# bibliography: bibliography.bib
# Match bibliostyle with the documentclass
# biblio-style: Frontiers-Harvard # Many Frontiers journals use the Harvard referencing system (Author-date), to find the style and resources for the journal you are submitting to: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal. For Humanities and Social Sciences articles please include page numbers in the in-text citations 
# biblio-style: Frontiers-Vancouver # Many Frontiers journals use the numbered referencing system, to find the style and resources for the journal you are submitting to: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal
---

```{r setup, include=FALSE}
## load libraries
library(ggtext)
library(kableExtra)
library(officer)
library(ragg)
library(targets)
library(tidyverse)

## this sets our default code chunk options
knitr::opts_chunk$set(dev = "ragg_png",
                      echo = FALSE,
                      ## figure chunk options
                      fig.cap = TRUE,
                      fig.width = 6,
                      dpi = 300)

## make fonts available
font_hoist <- function(family, silent = FALSE) {
  font_specs <- systemfonts::system_fonts() |> 
    filter(family == {{family}}) |> 
    mutate(family = paste(family, style)) |> 
    select(plain = path, name = family)

  purrr::pwalk(as.list(font_specs), systemfonts::register_font)

  if (!silent)  message(paste0("Hoisted ", nrow(font_specs), " variants:\n",
                               paste(font_specs$name, collapse = "\n")))
}

font_hoist("Manrope")

## grab output from targets to use in manuscript
store <- "../_targets"

```

```{r bacsummary}

tar_read(bac_table, store = store) |> 
  pack_rows(group_label = "BMP Subcategories",
            start_row = 3,
            end_row = 5) |> 
  pack_rows(group_label = "Arididty:BMP Subcategory Interaction",
            start_row = 7,
            end_row = 9)

```

```{r tnsummary}

tar_read(tn_table, store = store)

```

```{r dinsummary}

tar_read(tin_table, store = store)

```

```{r tpsummary}

tar_read(tp_table, store = store)

```

```{r po4summary}

tar_read(po4_table, store = store)

```

```{r tsssummary}

tar_read(tss_table, store = store)

```

```{r bacpub}
#| fig.cap: "Plot of adjusted sampling error and predicted effect size for FIB multilevel random effects model with moderators. Slope and 95% confidence interval of the adjusted sampling error term is annoted."
#| fig.height: 3
#| fig.width: 6
tar_read(bac_sse, store = store)$plot
```

```{r tnpub}
#| fig.cap: "Plot of adjusted sampling error and predicted effect size for TN multilevel random effects model with moderators. Slope and 95% confidence interval of the adjusted sampling error term is annoted."
#| fig.height: 3
#| fig.width: 6
tar_read(tn_sse, store = store)$plot
```

```{r dinpub}
#| fig.cap: "Plot of adjusted sampling error and predicted effect size for DIN multilevel random effects model with moderators. Slope and 95% confidence interval of the adjusted sampling error term is annoted."
#| fig.height: 3
#| fig.width: 6
tar_read(tin_sse, store = store)$plot
```

```{r tppub}
#| fig.cap: "Plot of adjusted sampling error and predicted effect size for TP multilevel random effects model with moderators. Slope and 95% confidence interval of the adjusted sampling error term is annoted."
#| fig.height: 3
#| fig.width: 6
tar_read(tp_sse, store = store)$plot
```

```{r po4pub}
#| fig.cap: "Plot of adjusted sampling error and predicted effect size for PO\\textsubscript{4} multilevel random effects model with moderators. Slope and 95% confidence interval of the adjusted sampling error term is annoted."
#| fig.height: 3
#| fig.width: 6
tar_read(po4_sse, store = store)$plot
```

```{r tsspub}
#| fig.cap: "Plot of adjusted sampling error and predicted effect size for TSS multilevel random effects model with moderators. Slope and 95% confidence interval of the adjusted sampling error term is annoted."
#| fig.height: 3
#| fig.width: 6
tar_read(tn_sse, store = store)$plot
```

```{r bacsens}
#| fig.cap: "Plot of intercept estimates from sensitivity analysis for the FIB regression model. Individual points are the intercept estimates with 95% confince intervals for the regression model fit leaving out the study value indicated on the y-axis. The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values. Any study values outside of the full model confidence intervals would be considered and outlier."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(bac_sens, store = store)$p1
```

```{r tnsens}
#| fig.cap: "Plot of intercept estimates from sensitivity analysis for the TN regression model. Individual points are the intercept estimates with 95% confince intervals for the regression model fit leaving out the study value indicated on the y-axis. The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values. Any study values outside of the full model confidence intervals would be considered and outlier."
#| fig.height: 5.5
#| fig.width: 8
#| out.width: "100%"
tar_read(tn_sens, store = store)$p1
```

```{r dinsens}
#| fig.cap: "Plot of intercept estimates from sensitivity analysis for the DIN regression model. Individual points are the intercept estimates with 95% confince intervals for the regression model fit leaving out the study value indicated on the y-axis. The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values. Any study values outside of the full model confidence intervals would be considered and outlier."
#| fig.height: 7
#| fig.width: 8
#| out.width: "100%"
tar_read(tin_sens, store = store)$p1
```

```{r tpsens}
#| fig.cap: "Plot of intercept estimates from sensitivity analysis for the TP regression model. Individual points are the intercept estimates with 95% confince intervals for the regression model fit leaving out the study value indicated on the y-axis. The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values. Any study values outside of the full model confidence intervals would be considered and outlier."
#| fig.height: 5.5
#| fig.width: 8
#| out.width: "100%"
tar_read(tp_sens, store = store)$p1
```

```{r po4sens}
#| fig.cap: "Plot of intercept estimates from sensitivity analysis for the PO\\textsubscript{4} regression model. Individual points are the intercept estimates with 95% confince intervals for the regression model fit leaving out the study value indicated on the y-axis. The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values. Any study values outside of the full model confidence intervals would be considered and outlier."
#| fig.height: 4
#| fig.width: 8
#| out.width: "100%"
tar_read(po4_sens, store = store)$p1
```

```{r tsssens}
#| fig.cap: "Plot of intercept estimates from sensitivity analysis for the TSS regression model. Individual points are the intercept estimates with 95% confince intervals for the regression model fit leaving out the study value indicated on the y-axis. The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values. Any study values outside of the full model confidence intervals would be considered and outlier."
#| fig.height: 4.5
#| fig.width: 8
#| out.width: "100%"
tar_read(tss_sens, store = store)$p1
```

```{r bacforest}
#| fig.cap: "Forest plot of FIB effect size estimates and 95% confidence intervals. Size of points are scaled to inverse of the sampling variance."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(bac_forest_plot, store = store)
```

```{r tnforest}
#| fig.cap: "Forest plot of TN effect size estimates and 95% confidence intervals. Size of points are scaled to inverse of the sampling variance."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(tn_forest_plot, store = store)
```

```{r dinforest}
#| fig.cap: "Forest plot of DIN effect size estimates and 95% confidence intervals. Size of points are scaled to inverse of the sampling variance."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(tin_forest_plot, store = store)
```

```{r tpforest}
#| fig.cap: "Forest plot of TP effect size estimates and 95% confidence intervals. Size of points are scaled to inverse of the sampling variance."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(tp_forest_plot, store = store)
```

```{r po4forest}
#| fig.cap: "Forest plot of PO\\textsubscript{4} effect size estimates and 95% confidence intervals. Size of points are scaled to inverse of the sampling variance."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(bac_forest_plot, store = store)
```

```{r tssforest}
#| fig.cap: "Forest plot of TSS effect size estimates and 95% confidence intervals. Size of points are scaled to inverse of the sampling variance."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(tss_forest_plot, store = store)
```

### 2024-02-06 frontiers_submission/frontiers_submission.Rmd
manuscript revisions

@@
 
+tp_r2 <- tar_read(tp_r2, store = store)
+po4_r2 <- tar_read(po4_r2, store = store)
 tp_i2_null <- tar_read(tp_i2_null, store = store)
@@
 
-## Sources of variance in BMP performance
-
 While it is assumed that site specific conditions are responsible for some of the heterogeneity in observed BMP performance, it is not clear how much of that variance is due to any one specific factor.
@@
 However, elucidating possible confounders such as climate and soil condition has been constrained by the lack of reported local condition data included in most BMP studies [@horvathEffectsRegionalClimate2023; @kochNitrogenRemovalStormwater2014; @eagleMetaAnalysisConstrained2017].
-A third variable that has been explored is the infuence of BMP age and maintence on performance.
-
-Accurately characterizing BMP treated runoff quality is challenging due to the high variability in BMP performance.
-However, influent water quality, seasonality, and BMP age provide some explanation to highly heterogeneous BMP efficiency data [@barrettPerformanceComparisonStructural2005; @liuReviewEffectivenessBest2017].
-A recent meta-analysis demonstrated the strong linkages between local climatic conditions and phosphorus removal in BMPs [@horvathEffectsRegionalClimate2023].
-
-## Effect size calculations
+The age and upkeep of BMPs is a third factor in BMP performance. 
+On one hand, the observed effects from BMPs are a function of various physical and biological processes that vary in the time required to produce desired reductions, especially as the spatial scale of the deployed project increases [@mealsLagTimeWater2010].
+These "lag times" between implementation and effect, which can be multiple years, have been shown to vary between parameter and BMP type [@mealsLagTimeWater2010].
+On the other hand the ability of BMPs to function effectively may also change over time.
+There has not been overwhelming published evidence to demonstrate the change or lack of change in BMP performance over time [@liuReviewEffectivenessBest2017]. 
+Many of the papers and data available for assessing BMP performance are short term monitoring project, typically around 1 year or less in length [@liuReviewEffectivenessBest2017; @kochNitrogenRemovalStormwater2014], suggesting our ability to assess the long-term performance of BMPs is limited.
 
@@
 More recently, researchers have applied effect size calculations more commonly used in ecological meta-analysis.
-@horvathEffectsRegionalClimate2023 used the standardized mean difference between influent and effluent, calculated as the difference in the means divided by the pooled standard deviation of the two groups (**hedges**).
+@horvathEffectsRegionalClimate2023 used the standardized mean difference between influent and effluent, calculated as the difference in the means divided by the pooled standard deviation of the two groups [@hedgesStatisticalMethodsMetaanalysis1985].
 @grudzinskiDoesRiparianFencing2020 applied the log ratio of means (*ROM*) to summarize performance of livestock BMPs.
@@
 
-**still needs work**
-Building on previous work, the objective of this project is to (1) improve understanding of the general effectiveness of BMPs on pollutant concentration, and (2) identify important contributors to variability in BMP performance.
-To accomplish this, we conducted a systematic review of relevant published literature and applied a meta-analytic approaches to develop weighted results across studies and identify contributors to variance in BMP performance.
+Building on previous work, the objectives of this paper are to (1) assess the general performance of BMPs in the published literature, and (2) identify relationships between BMP performance and potential effect size moderators.
+To accomplish this, we conducted a systematic review of relevant published literature and applied a meta-analytic approaches to develop weighted results across studies and identify variables that explain heterogeneity in BMP performance.
+Based on the existing literature we hypothesized that influent pollutant concentration, BMP type, and climate condition are influential in BMP performance and could be used to predict effluent concentration or percent reductions.
 
@@
 
-The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines [@collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018].
+The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines [@collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018; Figures S1-S2].
 In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
@@
 If data was provided in figures, the data was extracted with the WebPlotDigitizer tool [@rohatgiWebPlotDigitizer2022]. 
+Searches, review, and data extractions were conducted separately for FIB and nutrient related parameters.
 
@@
 We used the "rma.mv" function in the *metafor* R package to fit multilevel random effects regression models with *ROM* as the effect variable [@viechtbauerConductingMetaanalysesMetafor2010; @rcoreteamLanguageEnvironmentStatistical2023].
+We fit separate models for FIB, total nitrogen (TN), dissolved inorganic nitrogen (DIN), total phosphorus (TP), orthophosphate (PO~4~), and total suspended sediment (TSS).
 Our models specified a nested random effects term accounting for heterogeneity between effect sizes from the same study and for heterogeneity between studies.
@@
 We tested for evidence of publication bias, in the form of small study effect, by using the extension of Egger's regression applied to the multilevel model framework that included adjusted sampling error as a moderator [@nakagawaQuantitativeEvidenceSynthesis2023].
-We did not identify evidence of publication bias in the surveyed studies (FIB: ROM = 1.19, 95% CI [-3.23, 5.61]; TN: ROM = 0.31, 95% CI [-1.15, 1.78]; DIN: ROM = -2.4, 95% CI [-6.14, 1.59]; TP: ROM = -1.97, 95% CI [-5.18, 1.23]; PO~4~: -0.05, 95% CI [-2.82, 2.72]; TSS: ROM: 0.31, 95% CI [-1.15, 1.78]; Figure S1-S6); therefore, adjustments for publication bias were not included in the final models.
+We did not identify evidence of publication bias in the surveyed studies (FIB: ROM = 1.19, 95% CI [-3.23, 5.61]; TN: ROM = 0.31, 95% CI [-1.15, 1.78]; DIN: ROM = -2.4, 95% CI [-6.14, 1.59]; TP: ROM = -1.97, 95% CI [-5.18, 1.23]; PO~4~: -0.05, 95% CI [-2.82, 2.72]; TSS: ROM: 0.31, 95% CI [-1.15, 1.78]; Figure S3-S8); therefore, adjustments for publication bias were not included in the final models.
 We conducted a sensitivity analysis of the robustness of overall effect sizes to individual studies using leave-one-out analysis [@nakagawaQuantitativeEvidenceSynthesis2023].
@@
 The overall effect and 95% CI from each refit model is compared to the overall effect and 95% CI of the model fit to the full dataset.
-We did not identify evidence of outliers or overly influential studies for any of our models (Figure S7-S12).
+We did not identify evidence of outliers or overly influential studies for any of our models (Figure S9-S14).
 
@@
 
-There were a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effects for PO~4~, and 33 studies with 125 effect sizes for TSS.
+Our systematic review identified a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effects for PO~4~, and 33 studies with 125 effect sizes for TSS.
 The majority of studies were identified as smaller scaled lot or field studies (Figure \@ref(fig:bmpsummary) A).
@@
 There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \@ref(fig:studymap)).
-This might suggest that spatially correlated predictors such as climate and soil variables would be difficult to model due to the poor spatial representation of the collected data.
 
@@
                 escape = FALSE, booktabs = TRUE,
-                caption = "Model selection AICc") |> 
+                caption = "Model selection AICc.") |> 
   add_header_above(c(" ", "AICc" = 6)) |> 
@@
 
+tp_r2 <- paste0("\\textit{R\\textsuperscript{2}\\textsubscript{marginal}} =",
+                 tp_r2[[1]] |> round(2))
+
 po4_i2_null <- paste0("\\textit{I\\textsuperscript{2}\\textsubscript{total}} = ",
@@
 
+po4_r2 <- paste0("\\textit{R\\textsuperscript{2}\\textsubscript{marginal}} =",
+                 po4_r2[[1]] |> round(2))
+
 ```
@@
 The best model for both parameters only included influent as a moderator (Table \@ref(tab:modelaic)).
+Moderators explained a relatively small amount of variance for both models (TP: `r tp_r2`, PO~4~: `r po4_r2`).
 Influent concentration ($\beta$ = 0.23, 95% CI [-0.035, 0.49]) was not significant at the 95% confidence level for the TP model (Figure \@ref(fig:phoseffect); Table S4).
@@
 
+**Not complete**
+
 Our systematic review revealed strong spatial disparities in published BMP studies (Figure \@ref(fig:studymap).
+Similar spatial disparities have been identified and discussed in @kochNitrogenRemovalStormwater2014, @grudzinskiDoesRiparianFencing2020 and can be problematic for extrapolating results to other regions of interest.
+Inconsistent spatial coverage presents a challenge for disentangling confounding spatially correlated predictors such as climate and soil due to poor representation within the dataset.
+@horvathEffectsRegionalClimate2023 found overlapping BMP type and climate groups within their dataset that reduce the ability to distinguish effects due to either BMP type or climate.
+Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \@ref(fig:fibeffect).
+Not only were there spatially disparities, but we observed that the relative distribution of aridity index values does not resemble the distribution of aridity values across the U.S.
+(Figure \@ref(fig:aiindex).
+Our review indicates that BMP studies are over represented in the generally humid regions of the country and underrepresented the more arid regions.
+Study scale, runoff sources and BMP types appeared well distributed, with the caveat that we are not aware of the actual distribution of these values in BMPs deployed across the country.
 
-Meta-analysis indicated that BMPs showed significant overall reductions in FIB, TN, TP, and TSS concentrations when using ROM as an effect size. 
-We did not find strong evidence of overall leaching or reductions of DIN or PO~4~ in BMP studies.
-The results are in agreement to some extent with previous reviews that found effective (but highly variable) removal efficiencies for nitrogen, phosphorus, and sediment [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014; @liuReviewEffectivenessBest2017].
+We did not see an obvious trend in the number of published studies over time.
+However, there was a clearly skewed distribution in study length for all of the reviewed parameters.
+The prevalence of short-term studies has been observed in similar domains such as stream/river restoration [@bernhardtSynthesizingRiverRestoration2005].
+Given the nature of funding resources, this is not a surprising result but does have implications for developing a full understanding of BMP performance.
+First, there is strong evidence that certain BMPs and larger scale projects require extended time to establish and demonstrate positive benefit [@mealsLagTimeWater2010; @grudzinskiDoesRiparianFencing2020].
+@mealsLagTimeWater2010 documented lag times in the improvement of receiving water ranging from less than 1 year to upwards of 30 years, in particular sediment associated nutrients were assumed to have some of the longest effects.
+Second, BMP maintenance is an important components of BMP performance and success [@kochNitrogenRemovalStormwater2014; @heidariGreenStormwaterInfrastructure2023].
+Relatively little work has been published investigating how the performance of BMPs change over time, but there is scattered evidence that BMP performance may change as a function of BMP type and pollutant type [@liuReviewEffectivenessBest2017].
+While securing long term support for BMP monitoring and maintenance is a substantial hurdle [@heidariGreenStormwaterInfrastructure2023], unmaintained BMPs may see reduced performance [@kochNitrogenRemovalStormwater2014; @liuReviewEffectivenessBest2017].
 
-compare overall effects to koch, barrett, clary
+Study design prevented us from properly assessing BMP effectiveness as a function of age.
+Conducting a meta-analysis of BMP effectivness over time is hampered both by the lack of long-term studies and lack of standardized reporting mechanisms.
+Some studies simply describe the change in pollutant concentrations or loads at the beginning and end of the study as a percent change [@haileHydraulicPerformancePollutant2016] which presents statistical problems, espcially when sampling variance is not reported.
+Changes in performance can also be described using a linear regression using date (transformed as a numeric variable) as an independent variable and log-transformed water quality as the dependent variable [@mitschCreatingWetlandsPrimary2012; @mitschValidationEcosystemServices2014; @pausAssessmentHydraulicToxic2014].
+Slopes are a valid effect size for use in meta-analysis but the set of covairates used between studies should be the same since the coeffecient of interest is adjusted to account for other terms in the the regression model [@beckerSynthesisRegressionSlopes2007].
+It would be reasonable to assume that regressions equations vary between studies to adjust results for seasonality, flow rates, and other variables.
+Future efforts for assessing the performance of BMPs over time would benefit not only from more studies, but a more standardized method for providing comparable results.
 
-Linking BMP performance to watershed scale improvement remains a challenge (Tomer and Locke, Melland et al, Meals). 
-The succesful implementation and maintence of BMPs at watershed scales remains an area for more investigation (Liu et al., Lintern) 
-Long-term trans-disciplinary monitoring projects not just on BMP performance but maintance, management, and social apsects may elucidate unknown factors influencing the ability to scale BMP reductions to watershed scale improvements (Lintern et al.)
+Meta-analysis indicated that BMPs resulted in significant overall reductions in FIB, TN, TP, and TSS concentrations. 
+We did not find strong evidence of leaching or reductions of DIN or PO~4~ across BMP studies.
+The results are in general agreement with previous reviews that found effective (but highly variable) removal efficiencies for nitrogen, phosphorus, and sediment [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014; @liuReviewEffectivenessBest2017].
+The FIB results are useful in particular because the general consensus has been that the effectiveness of BMPs at reducing FIBs and pathogens has been sparsely reviewed and understudied [@hagerPerformanceLowimpactDevelopment2019].
+The FIB reductions generally agreed with our hypothesis that BMP type, influent concentration, and aridity moderate the effectiveness.
+The lower predicted performance in more arid regions comes with the caveat that data coverage in arid regions was quite poor, in particular for detention type BMPs.
+Despite this, the results are promising considering the major limitations of using FIB as a water quality criteria.
+For example, FIB can originate from non-human source, naturalize in soils, and have considerably different fate and transport mechanisms than the human pathogens of underlying concern.
 
-Runoff volume was not incorporated into our models, largely attributed to lack of available data.
-However, BMP size is typically scaled to accommodate runoff volumes of a particular return period volume (1-yr or 2-yr flows for example) (Hoss et al.).
+The effect of higher FIB removal rates at high concentrations is expected as filtration and other removal processes associated with BMPs become saturated(?)...
+something about sediment associated fib...
+The impact of aridity might be due to differential fate and transport processes in arid versus humid environments. 
+On one hand, we assume that arid conditions might be less hospitable to FIBs due to increased UV exposure and osmotic stress. 
+Conversly, these conditions are also less hospitable to the protoza, bacteriophages, and micro-zooplankton that can play a strong role in predating on and controlling FIB concentrations within BMP media  [@zhangCaptureDestructionEscherichia2010; @burtchettMicrozooplanktonGrazingMeans2017; @deanMetaanalysisAddressingImplications2022].
+Site specific conditions (such as retained soil moisture, turbidity, vegetation, and other factors) play an important role in these fate and transport mechanisms. 
+While our models capture some of the variance due to these differences as *between study* effects, including these as moderators in a meta-regression model would be valuable but these details are under reported in BMP studies.
+
+Although we anticipated increases in nitrogen removal rates with increases in influent concentration, we did not find evidence to support this.
+Increased flow rates, which can reduce residence time and increase BMP flushing, lowers nitrogen retention [@wollheimRetentionUrbanizingHeadwater2005; @craigStreamRestorationStrategies2008].
+High nitrogen influent concentrations might be associated with higher flows and decresed BMP retention times in the included studies.
+However, we did not collect associated flow data or discern between flow-weighted and mean concentration data within this study.
+Many of the reviewed studies appear to fail to include associated flow volume information.
+
+We also did not find evidence that BMP type or aridity moderated nitrogen removal.
+This is result is largely consistent with findings in reviews by @kochNitrogenRemovalStormwater2014, @hagerPerformanceLowimpactDevelopment2019 and @horvathEffectsRegionalClimate2023.
+There are a large number of abiotic and biotic processes that control nitrogen retention and removal in BMPs and these processes are moderated by both site specific climate and design factors [@lefevreReviewDissolvedPollutants2015; @valencaNitrateRemovalUncertainty2021].
+It is likely that these site specific factors (retained soil moisture, submerged anoxic zones, vegetation, media composition) are not captured by our broad categorization of BMP types and aridity index values.
+For example, @valencaNitrateRemovalUncertainty2021, using data from the International Stormwater Database, showed that the relative importance of climate and design variables for moderating nitrogen removal varied by BMP type.
+
+*compare nutrient effects to koch, barrett, clary*
+
+*Linking BMP performance to watershed scale improvement remains a challenge (Tomer and Locke, Melland et al, Meals).*
+*The succesful implementation and maintence of BMPs at watershed scales remains an area for more investigation (Liu et al., Lintern).*
+*Long-term trans-disciplinary monitoring projects not just on BMP performance but maintance, management, and social apsects may elucidate unknown factors influencing the ability to scale BMP reductions to watershed scale improvements (Lintern et al.)*
+
+*Runoff volume was not incorporated into our models, largely attributed to lack of available data.
+However, BMP size is typically scaled to accommodate runoff volumes of a particular return period volume (1-yr or 2-yr flows for example) (Hoss et al.).*
+*data exploration did not suggest this was the case, but it is reasonable to suggest the the underlying relationship between ROM and any of the moderators are a nonlinear function.
+Semi-parametric approaches might provide some additional  insight*
 
@@
 ```
+
+```{r aiindex}
+#| fig.cap: "Comparison of the relative distributions (density) of aridity index values across the U.S. against aridity index values for studies in the systematic review."
+#| fig.pos: "p"
+#| out.extra: ""
+#| out.width: "50%"
+
+knitr::include_graphics("../figures/aridity_density.pdf")
+```
+
 <!-- #  -->

### 2024-02-06 frontiers_submission/frontiers_submission.tex
manuscript revisions

@@
 
-\title[Meta-analysis of best management practices]{A meta-analysis of the impacts of best management practices on nonpoint source pollutants.}
+\title[Meta-analysis of best management practices]{A meta-analysis of the impacts of best management practices on nonpoint source pollutant concentration.}
 \author[\firstAuthorLast]{\Authors}
@@
 
-In the United States (U.S.), major progress towards water quality goals have been achieved through the Clean Water Act of 1972.
+In the United States (U.S.), major improvement in water quality has been achieved under the Clean Water Act of 1972.
 This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge \citep{benhamLessonsLearnedTMDL2008, nationalresearchcouncilAssessingTMDLApproach2001, schrammTotalMaximumDaily2022}.
-Increased pollutant loads and pollutant concentrations in runoff are a particular challenge when dealing with nonpoint source pollution resulting from land use change.
+Increased pollutant loads and concentrations in runoff resulting from land use changes are a particular challenge.
 The impacts of land use change on hydrology and water quality are well established \citep{allanLandscapesRiverscapesInfluence2004, carpenterNonpointPollutionSurface1998, bernhardtUnderstandingManagingMinimizing2008, careyEvaluatingNutrientImpacts2013, freemanImpactsUrbanizationDevelopment2019}.
@@
 \citet{linternBestManagementPractices2020} found 43\% of reviewed BMP effectiveness studies relied completely on modeled outputs, with modeled outputs almost always predicting water quality improvements following BMP implementation.
-However, field studies are much more likely to demonstrate mixed results including net releases of pollutants under certain conditions \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017}.
-The disconnect between modeled outcomes and field studies might be attributed to (1) overly simplified or incorrect estimates of parameters that represent management practices \citep{ullrichApplicationSoilWater2009, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020}, (2) failure to incorporate uncertainty in estimates \citep{tasdighiBayesianTotalUncertainty2018, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020}, and (3) assumption of static performance over time \citep{mealsLagTimeWater2010, liuReviewEffectivenessBest2017, fuReviewCatchmentscaleWater2019}.
+However, field studies are much more likely to demonstrate mixed results including net releases (leaching) of pollutants under certain conditions \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017}.
+The disconnect between modeled outcomes and field studies might be attributed to (1) overly simplified or incorrect estimates of model parameters that represent management practices \citep{ullrichApplicationSoilWater2009, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020}, (2) the failure to incorporate the appropriate types of uncertainty into estimates \citep{tasdighiBayesianTotalUncertainty2018, fuReviewCatchmentscaleWater2019, linternBestManagementPractices2020}, and (3) the assumption of static performance over time \citep{mealsLagTimeWater2010, liuReviewEffectivenessBest2017, fuReviewCatchmentscaleWater2019}.
 
-One source of uncertainty comes from the substantial variability of performance metrics reported in empirical BMP studies \citep{linternBestManagementPractices2020}.
+An underlying source of uncertainty comes from the substantial variability of performance metrics reported in empirical BMP studies \citep{linternBestManagementPractices2020}.
 There have been varied attempts at synthesizing estimates of BMP efficiency to provide resource managers with knowledge for improved decision-making \citep{agouridisLivestockGrazingManagement2005, barrettComparisonBMPPerformance2008, claryBMPPerformanceAnalysis2011, grudzinskiDoesRiparianFencing2020, horvathEffectsRegionalClimate2023, kochNitrogenRemovalStormwater2014, krogerReviewBestManagement2012, liuReviewEffectivenessBest2017, simpsonDevelopingBestManagement2009}.
@@
 The review literature on the effects of BMPs on FIBs are sparse but generally find extremely high variance in performance across BMPs \citep{claryBMPPerformanceAnalysis2011, grudzinskiDoesRiparianFencing2020}.
-Accurately characterizing BMP treated runoff quality is challenging due to the high variability in BMP performance.
-However, influent water quality, seasonality, and BMP age provide some explanation to highly heterogeneous BMP efficiency data \citep{barrettPerformanceComparisonStructural2005, liuReviewEffectivenessBest2017}.\\
-A recent meta-analysis demonstrated the strong linkages between local climatic conditions and phosphorus removal in BMPs \citep{horvathEffectsRegionalClimate2023}.
 
-\hypertarget{effect-size-calculations}{%
-\subsection{Effect size calculations}\label{effect-size-calculations}}
+While it is assumed that site specific conditions are responsible for some of the heterogeneity in observed BMP performance, it is not clear how much of that variance is due to any one specific factor.
+Influent concentration is likely to have some effect on certain types of structural BMPs.
+\citet{barrettPerformanceComparisonStructural2005} demonstrated that percent pollutant reduction is often a function of influent quality.
+Specifically, for certain types of BMPs percent removal is low at low influent concentrations, and increases with increasing influent concentrations.
+However, for some types of BMPs and pollutant parameters, effluent concentration is unrelated to influent concentration.
+Second, local climatic conditions can be expected to influence BMP performance.
+BMPs in dry climates have been shown to be more likely to leach phosphorus than those in wetter climates \citep{horvathEffectsRegionalClimate2023}.
+However, elucidating possible confounders such as climate and soil condition has been constrained by the lack of reported local condition data included in most BMP studies \citep{horvathEffectsRegionalClimate2023, kochNitrogenRemovalStormwater2014, eagleMetaAnalysisConstrained2017}.
+The age and upkeep of BMPs is a third factor in BMP performance.
+On one hand, the observed effects from BMPs are a function of various physical and biological processes that vary in the time required to produce desired reductions, especially as the spatial scale of the deployed project increases \citep{mealsLagTimeWater2010}.
+These ``lag times'' between implementation and effect, which can be multiple years, have been shown to vary between parameter and BMP type \citep{mealsLagTimeWater2010}.
+On the other hand the ability of BMPs to function effectively may also change over time.
+There has not been overwhelming published evidence to demonstrate the change or lack of change in BMP performance over time \citep{liuReviewEffectivenessBest2017}.
+Many of the papers and data available for assessing BMP performance are short term monitoring project, typically around 1 year or less in length \citep{liuReviewEffectivenessBest2017, kochNitrogenRemovalStormwater2014}, suggesting our ability to assess the long-term performance of BMPs is limited.
 
@@
 More recently, researchers have applied effect size calculations more commonly used in ecological meta-analysis.
-\citet{horvathEffectsRegionalClimate2023} used the standardized mean difference between influent and effluent, calculated as the difference in the means divided by the pooled standard deviation of the two groups (\textbf{hedges}).
+\citet{horvathEffectsRegionalClimate2023} used the standardized mean difference between influent and effluent, calculated as the difference in the means divided by the pooled standard deviation of the two groups \citep{hedgesStatisticalMethodsMetaanalysis1985}.
 \citet{grudzinskiDoesRiparianFencing2020} applied the log ratio of means (\emph{ROM}) to summarize performance of livestock BMPs.
@@
 
-The goals of this project were to (1) conduct a systematic review of BMP field studies; (2) develop estimates of FIB, nitrogen, phosphorus, and TSS removal using a meta-analytic approach; and (3) explore the effect of multiple moderators using multilevel random effects meta-analytic regression models.
+Building on previous work, the objectives of this paper are to (1) assess the general performance of BMPs in the published literature, and (2) identify relationships between BMP performance and potential effect size moderators.
+To accomplish this, we conducted a systematic review of relevant published literature and applied a meta-analytic approaches to develop weighted results across studies and identify variables that explain heterogeneity in BMP performance.
+Based on the existing literature we hypothesized that influent pollutant concentration, BMP type, and climate condition are influential in BMP performance and could be used to predict effluent concentration or percent reductions.
 
@@
 
-The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines \citep{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}.
+The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines {[}\citet{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}; Figures S1-S2{]}.
 In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
@@
 If data was provided in figures, the data was extracted with the WebPlotDigitizer tool \citep{rohatgiWebPlotDigitizer2022}.
+Searches, review, and data extractions were conducted separately for FIB and nutrient related parameters.
 
@@
 We used the ``rma.mv'' function in the \emph{metafor} R package to fit multilevel random effects regression models with \emph{ROM} as the effect variable \citep{viechtbauerConductingMetaanalysesMetafor2010, rcoreteamLanguageEnvironmentStatistical2023}.
+We fit separate models for FIB, total nitrogen (TN), dissolved inorganic nitrogen (DIN), total phosphorus (TP), orthophosphate (PO\textsubscript{4}), and total suspended sediment (TSS).
 Our models specified a nested random effects term accounting for heterogeneity between effect sizes from the same study and for heterogeneity between studies.
@@
 We tested for evidence of publication bias, in the form of small study effect, by using the extension of Egger's regression applied to the multilevel model framework that included adjusted sampling error as a moderator \citep{nakagawaQuantitativeEvidenceSynthesis2023}.
-We did not identify evidence of publication bias in the surveyed studies (FIB: ROM = 1.19, 95\% CI {[}-3.23, 5.61{]}; TN: ROM = 0.31, 95\% CI {[}-1.15, 1.78{]}; DIN: ROM = -2.4, 95\% CI {[}-6.14, 1.59{]}; TP: ROM = -1.97, 95\% CI {[}-5.18, 1.23{]}; PO\textsubscript{4}: -0.05, 95\% CI {[}-2.82, 2.72{]}; TSS: ROM: 0.31, 95\% CI {[}-1.15, 1.78{]}; Figure S1-S6); therefore, adjustments for publication bias were not included in the final models.
+We did not identify evidence of publication bias in the surveyed studies (FIB: ROM = 1.19, 95\% CI {[}-3.23, 5.61{]}; TN: ROM = 0.31, 95\% CI {[}-1.15, 1.78{]}; DIN: ROM = -2.4, 95\% CI {[}-6.14, 1.59{]}; TP: ROM = -1.97, 95\% CI {[}-5.18, 1.23{]}; PO\textsubscript{4}: -0.05, 95\% CI {[}-2.82, 2.72{]}; TSS: ROM: 0.31, 95\% CI {[}-1.15, 1.78{]}; Figure S3-S8); therefore, adjustments for publication bias were not included in the final models.
 We conducted a sensitivity analysis of the robustness of overall effect sizes to individual studies using leave-one-out analysis \citep{nakagawaQuantitativeEvidenceSynthesis2023}.
@@
 The overall effect and 95\% CI from each refit model is compared to the overall effect and 95\% CI of the model fit to the full dataset.
-We did not identify evidence of outliers or overly influential studies for any of our models (Figure S7-S12).
+We did not identify evidence of outliers or overly influential studies for any of our models (Figure S9-S14).
 
@@
 
-There were a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effects for PO\textsubscript{4}, and 33 studies with 125 effect sizes for TSS.
+Our systematic review identified a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effects for PO\textsubscript{4}, and 33 studies with 125 effect sizes for TSS.
 The majority of studies were identified as smaller scaled lot or field studies (Figure \ref{fig:bmpsummary} A).
@@
 There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \ref{fig:studymap}).
-This might suggest that spatially correlated predictors such as climate and soil variables would be difficult to model due to the poor spatial representation of the collected data.
 
@@
 
-\caption{\label{tab:modelaic}Model selection AICc}
+\caption{\label{tab:modelaic}Model selection AICc.}
 \centering
@@
 The best model for both parameters only included influent as a moderator (Table \ref{tab:modelaic}).
+Moderators explained a relatively small amount of variance for both models (TP: \textit{R\textsuperscript{2}\textsubscript{marginal}} =0.12, PO\textsubscript{4}: \textit{R\textsuperscript{2}\textsubscript{marginal}} =0.35).
 Influent concentration (\(\beta\) = 0.23, 95\% CI {[}-0.035, 0.49{]}) was not significant at the 95\% confidence level for the TP model (Figure \ref{fig:phoseffect}; Table S4).
@@
 
+\textbf{Not complete}
+
 Our systematic review revealed strong spatial disparities in published BMP studies (Figure \ref{fig:studymap}.
+Similar spatial disparities have been identified and discussed in \citet{kochNitrogenRemovalStormwater2014}, \citet{grudzinskiDoesRiparianFencing2020} and can be problematic for extrapolating results to other regions of interest.
+Inconsistent spatial coverage presents a challenge for disentangling confounding spatially correlated predictors such as climate and soil due to poor representation within the dataset.
+\citet{horvathEffectsRegionalClimate2023} found overlapping BMP type and climate groups within their dataset that reduce the ability to distinguish effects due to either BMP type or climate.
+Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \ref{fig:fibeffect}.
+Not only were there spatially disparities, but we observed that the relative distribution of aridity index values does not resemble the distribution of aridity values across the U.S.
+(Figure \ref{fig:aiindex}.
+Our review indicates that BMP studies are over represented in the generally humid regions of the country and underrepresented the more arid regions.
+Study scale, runoff sources and BMP types appeared well distributed, with the caveat that we are not aware of the actual distribution of these values in BMPs deployed across the country.
 
-Meta-analysis indicated that BMPs showed significant overall reductions in FIB, TN, TP, and TSS concentrations when using ROM as an effect size.
-We did not find strong evidence of overall leaching or reductions of DIN or PO\textsubscript{4} in BMP studies.
-The results are in agreement to some extent with previous reviews that found effective (but highly variable) removal efficiencies for nitrogen, phosphorus, and sediment \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, liuReviewEffectivenessBest2017}.
+We did not see an obvious trend in the number of published studies over time.
+However, there was a clearly skewed distribution in study length for all of the reviewed parameters.
+The prevalence of short-term studies has been observed in similar domains such as stream/river restoration \citep{bernhardtSynthesizingRiverRestoration2005}.
+Given the nature of funding resources, this is not a surprising result but does have implications for developing a full understanding of BMP performance.
+First, there is strong evidence that certain BMPs and larger scale projects require extended time to establish and demonstrate positive benefit \citep{mealsLagTimeWater2010, grudzinskiDoesRiparianFencing2020}.
+\citet{mealsLagTimeWater2010} documented lag times in the improvement of receiving water ranging from less than 1 year to upwards of 30 years, in particular sediment associated nutrients were assumed to have some of the longest effects.
+Second, BMP maintenance is an important components of BMP performance and success \citep{kochNitrogenRemovalStormwater2014, heidariGreenStormwaterInfrastructure2023}.
+Relatively little work has been published investigating how the performance of BMPs change over time, but there is scattered evidence that BMP performance may change as a function of BMP type and pollutant type \citep{liuReviewEffectivenessBest2017}.
+While securing long term support for BMP monitoring and maintenance is a substantial hurdle \citep{heidariGreenStormwaterInfrastructure2023}, unmaintained BMPs may see reduced performance \citep{kochNitrogenRemovalStormwater2014, liuReviewEffectivenessBest2017}.
 
-compare overall effects to koch, barrett, clary
+Study design prevented us from properly assessing BMP effectiveness as a function of age.
+Conducting a meta-analysis of BMP effectivness over time is hampered both by the lack of long-term studies and lack of standardized reporting mechanisms.
+Some studies simply describe the change in pollutant concentrations or loads at the beginning and end of the study as a percent change \citep{haileHydraulicPerformancePollutant2016} which presents statistical problems, espcially when sampling variance is not reported.
+Changes in performance can also be described using a linear regression using date (transformed as a numeric variable) as an independent variable and log-transformed water quality as the dependent variable \citep{mitschCreatingWetlandsPrimary2012, mitschValidationEcosystemServices2014, pausAssessmentHydraulicToxic2014}.
+Slopes are a valid effect size for use in meta-analysis but the set of covairates used between studies should be the same since the coeffecient of interest is adjusted to account for other terms in the the regression model \citep{beckerSynthesisRegressionSlopes2007}.
+It would be reasonable to assume that regressions equations vary between studies to adjust results for seasonality, flow rates, and other variables.
+Future efforts for assessing the performance of BMPs over time would benefit not only from more studies, but a more standardized method for providing comparable results.
 
-Linking BMP performance to watershed scale improvement remains a challenge (Tomer and Locke, Melland et al, Meals).
-The succesful implementation and maintence of BMPs at watershed scales remains an area for more investigation (Liu et al., Lintern)
-Long-term trans-disciplinary monitoring projects not just on BMP performance but maintance, management, and social apsects may elucidate unknown factors influencing the ability to scale BMP reductions to watershed scale improvements (Lintern et al.)
+Meta-analysis indicated that BMPs resulted in significant overall reductions in FIB, TN, TP, and TSS concentrations.
+We did not find strong evidence of leaching or reductions of DIN or PO\textsubscript{4} across BMP studies.
+The results are in general agreement with previous reviews that found effective (but highly variable) removal efficiencies for nitrogen, phosphorus, and sediment \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, liuReviewEffectivenessBest2017}.
+The FIB results are useful in particular because the general consensus has been that the effectiveness of BMPs at reducing FIBs and pathogens has been sparsely reviewed and understudied \citep{hagerPerformanceLowimpactDevelopment2019}.
+The FIB reductions generally agreed with our hypothesis that BMP type, influent concentration, and aridity moderate the effectiveness.
+The lower predicted performance in more arid regions comes with the caveat that data coverage in arid regions was quite poor, in particular for detention type BMPs.
+Despite this, the results are promising considering the major limitations of using FIB as a water quality criteria.
+For example, FIB can originate from non-human source, naturalize in soils, and have considerably different fate and transport mechanisms than the human pathogens of underlying concern.
 
-Runoff volume was not incorporated into our models, largely attributed to lack of available data.
-However, BMP size is typically scaled to accommodate runoff volumes of a particular return period volume (1-yr or 2-yr flows for example) (Hoss et al.).
+The effect of higher FIB removal rates at high concentrations is expected as filtration and other removal processes associated with BMPs become saturated(?)\ldots{} something about sediment associated fib\ldots{}
+The impact of aridity might be due to differential fate and transport processes in arid versus humid environments.
+On one hand, we assume that arid conditions might be less hospitable to FIBs due to increased UV exposure and osmotic stress.
+Conversly, these conditions are also less hospitable to the protoza, bacteriophages, and micro-zooplankton that can play a strong role in predating on and controlling FIB concentrations within BMP media \citep{zhangCaptureDestructionEscherichia2010, burtchettMicrozooplanktonGrazingMeans2017, deanMetaanalysisAddressingImplications2022}.
+Site specific conditions (such as retained soil moisture, turbidity, vegetation, and other factors) play an important role in these fate and transport mechanisms.
+While our models capture some of the variance due to these differences as \emph{between study} effects, including these as moderators in a meta-regression model would be valuable but these details are under reported in BMP studies.
+
+Although we anticipated increases in nitrogen removal rates with increases in influent concentration, we did not find evidence to support this.
+Increased flow rates, which can reduce residence time and increase BMP flushing, lowers nitrogen retention \citep{wollheimRetentionUrbanizingHeadwater2005, craigStreamRestorationStrategies2008}.
+High nitrogen influent concentrations might be associated with higher flows and decresed BMP retention times in the included studies.
+However, we did not collect associated flow data or discern between flow-weighted and mean concentration data within this study.
+Many of the reviewed studies appear to fail to include associated flow volume information.
+
+We also did not find evidence that BMP type or aridity moderated nitrogen removal.
+This is result is largely consistent with findings in reviews by \citet{kochNitrogenRemovalStormwater2014}, \citet{hagerPerformanceLowimpactDevelopment2019} and \citet{horvathEffectsRegionalClimate2023}.
+There are a large number of abiotic and biotic processes that control nitrogen retention and removal in BMPs and these processes are moderated by both site specific climate and design factors \citep{lefevreReviewDissolvedPollutants2015, valencaNitrateRemovalUncertainty2021}.
+It is likely that these site specific factors (retained soil moisture, submerged anoxic zones, vegetation, media composition) are not captured by our broad categorization of BMP types and aridity index values.
+For example, \citet{valencaNitrateRemovalUncertainty2021}, using data from the International Stormwater Database, showed that the relative importance of climate and design variables for moderating nitrogen removal varied by BMP type.
+
+\emph{compare nutrient effects to koch, barrett, clary}
+
+\emph{Linking BMP performance to watershed scale improvement remains a challenge (Tomer and Locke, Melland et al, Meals).}
+\emph{The succesful implementation and maintence of BMPs at watershed scales remains an area for more investigation (Liu et al., Lintern).}
+\emph{Long-term trans-disciplinary monitoring projects not just on BMP performance but maintance, management, and social apsects may elucidate unknown factors influencing the ability to scale BMP reductions to watershed scale improvements (Lintern et al.)}
+
+\emph{Runoff volume was not incorporated into our models, largely attributed to lack of available data.
+However, BMP size is typically scaled to accommodate runoff volumes of a particular return period volume (1-yr or 2-yr flows for example) (Hoss et al.).}
+\emph{data exploration did not suggest this was the case, but it is reasonable to suggest the the underlying relationship between ROM and any of the moderators are a nonlinear function.
+Semi-parametric approaches might provide some additional insight}
 
@@
 
+\begin{figure}[p]
+\includegraphics[width=0.5\linewidth,]{../figures/aridity_density} \caption{Comparison of the relative distributions (density) of aridity index values across the U.S.
+against aridity index values for studies in the systematic review.}\label{fig:aiindex}
+\end{figure}
+
 \end{document}

### 2024-02-06 supplementary_materials/supplementary_materials.Rmd
manuscript revisions

@@
 store <- "../_targets"
+
+```
+
+```{r roses1}
+#| fig.cap: "RepOrting standards for Systematic Evidence Syntheses (ROSES) flow diagram for review process of fecal indicator bacteria (FIB) best management practice (BMP) studies."
+#| fig.pos: "H"
+#| out.extra: ""
+#| out.width: "80%"
+
+knitr::include_graphics("../figures/roses_bac.pdf")
+
+```
+
+```{r roses2}
+#| fig.cap: "RepOrting standards for Systematic Evidence Syntheses (ROSES) flow diagram for review process of nutrient and sediment best management practice (BMP) studies."
+#| fig.pos: "H"
+#| out.extra: ""
+#| out.width: "80%"
+
+knitr::include_graphics("../figures/roses_nut.pdf")
 

### 2024-02-06 R/05_literature_statistics.R

@@ -511,3 +511,69 @@ plot_bmp_temporal <- function(bac_df, nut_df) {
   return(list(p1 = p1,
               stdy_len_stats = stdy_len_stats))
 }
+
+
+plot_ai_summary <- function(bac_df, nut_df) {
+
+  ai <- rast("data/Global-AI_ET0_v3_annual/ai_v3_yr.tif")
+  ai <- ai * 0.0001 
+  
+  
+  us_file <- tigris::states(cb = TRUE, progress_bar = FALSE)
+  
+  us_file <- us_file |> 
+    filter(!(NAME %in% c("Alaska",
+                         "American Samoa",
+                         "Guam",
+                         "Commonwealth of the Northern Mariana Islands",
+                         "United States Virgin Islands")))
+  
+  us_file_trans <- us_file |> 
+    vect() |> 
+    project(ai)
+  
+  
+  cropped_ai <- crop(ai, us_file_trans) |> 
+    mask(us_file_trans)
+  
+  plot(cropped_ai)
+  
+  cropped_ai_tibble <- as_tibble(cropped_ai, na.rm = TRUE)
+  
+  bac_ai <- tibble(awi_pm_sr_yr = bac_df$model_df$ai[,1] + bac_df$mean,
+                   label = "BMP AI Values")
+  
+  nut_ai <- tibble(awi_pm_sr_yr = nut_df$model_df$ai[,1] + bac_df$mean,
+                   label = "BMP AI Values")
+  
+  
+  
+  cropped_ai_tibble |> 
+    mutate(label = "U.S. AI Values") |> 
+    bind_rows(bac_ai) |> 
+    bind_rows(nut_ai) |> 
+    ggplot() +
+    geom_density(aes(awi_pm_sr_yr,
+                     color = label,
+                     fill = label),
+                 alpha = 0.5,
+                 trim = TRUE,
+                 bw = "bcv") +
+    scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
+    scale_x_continuous(expand = expansion(mult = c(0,0.05))) +
+    scale_color_brewer("", palette = "Dark2") +
+    scale_fill_brewer("", palette = "Dark2") +
+    coord_cartesian(xlim = c(0,4)) +
+    labs(x = "Aridity index", y = "Density") +
+    theme_mps_noto(base_family = "Manrope Regular") +
+    theme(axis.title = element_text(family = "Manrope SemiBold", face = "plain", size = rel(1)),
[…]

### 2024-02-06 _targets.R

@@ -321,17 +321,17 @@ list(
                                              height = 3.5,
                                              units = "in"),
              format = "file"),
-  # tar_target(marginal_means, patch_marginal_means(plots = list(bac_marginal_means,
-  #                                                              tn_marginal_means,
-  #                                                              tp_marginal_means,
-  #                                                              tss_marginal_means))),
-  # tar_target(marginal_means_file, ggsave("figures/marginal_means.pdf",
-  #                                        plot = marginal_means,
-  #                                        device = cairo_pdf,
-  #                                        width = 8.5,
-  #                                        height = 8.5,
-  #                                        units = "in"),
-  #            format = "file"),
+  
+  tar_target(aridity_density, plot_ai_summary(bac_model_df,
+                                              nut_model_df)),
+  
+  tar_target(aridity_file, ggsave("figures/aridity_density.pdf",
+                                  plot = aridity_density,
+                                  device = cairo_pdf,
+                                  width = 4.5,
+                                  height = 3.5,
+                                  units = "in"),
+             format = "file"),
 
   NULL
   )

### 2024-02-07 frontiers_submission/frontiers_submission.Rmd
wrap up discussion

@@
 abstract: |
-  Abstract length and content varies depending on article type. Refer to 
-  <http://www.frontiersin.org/about/AuthorGuidelines> for abstract requirement
-  and length according to article type.
+  We conducted a systematic review of published nonpoint source water quality best management practice (BMP) studies and used a meta-analysis approach to describe variance in pollutant removal performance.
+Despite high variance, we found the BMPs on average were effective at reducing fecal indicator bacteria (FIB), total nitrogen (TN), total phosphorus (TP), and total suspended sediment (TSS) concentrations.
+We explored how much BMP pollutant removal process, influent pollutant concentration, and aridity effected BMP performance.
+We found that influent concentration and interaction effect between the BMP pollutant removal process and aridity explained a substantial amount of variance in BMP performance in FIB removal.
+Influent concentration explained a small amount of variability in BMP removal of TP and orthophosphate (PO~4~).
+We did not find evidence that any of our chosen variables moderated BMP performance in nitrogen or TSS removal.
+Through our systematic review, we found inadequate spatial representation of BMP studies to capture the underlying variability in climate, soil, and other conditions that could impact BMP performance.
 keywords:
@@
 In the United States (U.S.), major improvement in water quality has been achieved under the Clean Water Act of 1972.
-This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge [@benhamLessonsLearnedTMDL2008; @nationalresearchcouncilAssessingTMDLApproach2001; @schrammTotalMaximumDaily2022].
+This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge [@nationalresearchcouncilAssessingTMDLApproach2001; @benhamLessonsLearnedTMDL2008; @schrammTotalMaximumDaily2022].
 Increased pollutant loads and concentrations in runoff resulting from land use changes are a particular challenge.
@@
 We conducted a systematic review of recent (2000-2022) literature to compile field studies documenting the effectiveness of best management practices on fecal indicator bacteria and nutrients concentrations.
-Prior meta-analysis have utilized data reported in the International Stormwater Database, which consists of self-reported and quality checked BMP data [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014; @horvathEffectsRegionalClimate2023].
-The International Stormwater Database only recently added agricultural BMPs and has relatively sparse FIB data [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014].
+Prior meta-analysis have utilized data reported in the International Stormwater BMP Database (https://bmpdatabase.org/), which consists of self-reported and quality checked BMP data [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014; @horvathEffectsRegionalClimate2023].
+The International Stormwater BMP Database only recently added agricultural BMPs and has relatively sparse FIB data [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014].
 Since we had interest in both FIB performance and agricultural BMPs we chose to utilize a systematic review.
-We also assume that most published reports undergo some method of internal or external peer review prior to publication, which is not necessarily the case with all of the data published in the International Stormwater Database.
 
-The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines [@collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018; Figures S1-S2].
+The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines [@collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018].
 In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
@@
 Although results from Google Scholar are not always replicable, we utilized the service to maximize search results for studies not published in academic journals and presumably increase the chance of identifying studies with negative effects.
-Fecal indicator bacteria study seraches included the following query: “fecal indicator bacteria” OR “*E.
+Fecal indicator bacteria study searches included the following query: “fecal indicator bacteria” OR “*E.
 coli*” OR “*Escherichia coli*” OR “enterococci” OR “enterococcus” AND “best management practices” OR “BMPs” AND “effectiveness” OR “performance”.
@@
 Searches, review, and data extractions were conducted separately for FIB and nutrient related parameters.
+See Figure S1 and S2 for ROSES diagram.
 
@@
                 escape = FALSE, booktabs = TRUE,
-                caption = "Model selection AICc.") |> 
+                caption = "Sumary of AICc values used for model selection. Bolded values indicate the selected candidate model.") |> 
   add_header_above(c(" ", "AICc" = 6)) |> 
@@
 Changes in performance can also be described using a linear regression using date (transformed as a numeric variable) as an independent variable and log-transformed water quality as the dependent variable [@mitschCreatingWetlandsPrimary2012; @mitschValidationEcosystemServices2014; @pausAssessmentHydraulicToxic2014].
-Slopes are a valid effect size for use in meta-analysis but the set of covairates used between studies should be the same since the coeffecient of interest is adjusted to account for other terms in the the regression model [@beckerSynthesisRegressionSlopes2007].
+Slopes are a valid effect size for use in meta-analysis but the set of covariates used between studies should be the same since the coeffecient of interest is adjusted to account for other terms in the the regression model [@beckerSynthesisRegressionSlopes2007].
 It would be reasonable to assume that regressions equations vary between studies to adjust results for seasonality, flow rates, and other variables.
@@
 The results are in general agreement with previous reviews that found effective (but highly variable) removal efficiencies for nitrogen, phosphorus, and sediment [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014; @liuReviewEffectivenessBest2017].
-The FIB results are useful in particular because the general consensus has been that the effectiveness of BMPs at reducing FIBs and pathogens has been sparsely reviewed and understudied [@hagerPerformanceLowimpactDevelopment2019].
+The FIB results are useful in particular because FIBs performance by BMPs have been sparsely reviewed and generally understudied [@hagerPerformanceLowimpactDevelopment2019].
+
 The FIB reductions generally agreed with our hypothesis that BMP type, influent concentration, and aridity moderate the effectiveness.
@@
 Despite this, the results are promising considering the major limitations of using FIB as a water quality criteria.
-For example, FIB can originate from non-human source, naturalize in soils, and have considerably different fate and transport mechanisms than the human pathogens of underlying concern.
+It is important to note that FIB can originate from non-human source and naturalize in soils, and result in different underlying risk of illness [@ishiiEscherichiaColiEnvironment2008; @schoenAssessingPathogenRisk2010; @sollerEstimatedHumanHealth2010; @fujiokaRecreationalWaterQuality2015].
+Since the fate and transport of human pathogens within BMPs can potentially differ from FIB, BMP choices probably should not be based on FIB reduction alone as alternative indicators or even direct pathogen measurement becomes available [@pengIndicatorPathogenRemoval2016; @waltersPersistenceNucleicAcid2009].
+However, relatively few studies have compared human pathogen and FIB removal rates within BMPs [@rughHighlyVariableRemoval2022].
 
-The effect of higher FIB removal rates at high concentrations is expected as filtration and other removal processes associated with BMPs become saturated(?)...
-something about sediment associated fib...
+While we observed a strong relationship between FIB influent concentration and FIB removal across all BMPs, we anticipated this relationship to vary by BMP subcategory.
+The reliance on certain removal processes by BMP subcategories was expected to effect the ability of the BMP to retain FIB at higher or lower concentrations.
+We did have some evidence of differing BMP subcateogy removal under different ariditiy indecies.
 The impact of aridity might be due to differential fate and transport processes in arid versus humid environments. 
@@
 Conversly, these conditions are also less hospitable to the protoza, bacteriophages, and micro-zooplankton that can play a strong role in predating on and controlling FIB concentrations within BMP media  [@zhangCaptureDestructionEscherichia2010; @burtchettMicrozooplanktonGrazingMeans2017; @deanMetaanalysisAddressingImplications2022].
-Site specific conditions (such as retained soil moisture, turbidity, vegetation, and other factors) play an important role in these fate and transport mechanisms. 
-While our models capture some of the variance due to these differences as *between study* effects, including these as moderators in a meta-regression model would be valuable but these details are under reported in BMP studies.
+Site specific conditions (such as retained soil moisture, turbidity, vegetation, and other factors) play an important role in bacteria survival as well as for influencing the filtration and attachment processes that retain FIB within BMP media.
+For example, the presence or absence of a submerged zone within a bioretention BMP has a strong effect on FIB removal [@rippyMeetingCriteriaLinking2015; @pengIndicatorPathogenRemoval2016].
+While our models capture some of the variance due to these differences as between study effects, including these variables as fixed effect moderators in a meta-regression model would be valuable but these details are under reported in BMP studies.
 
@@
 Increased flow rates, which can reduce residence time and increase BMP flushing, lowers nitrogen retention [@wollheimRetentionUrbanizingHeadwater2005; @craigStreamRestorationStrategies2008].
-High nitrogen influent concentrations might be associated with higher flows and decresed BMP retention times in the included studies.
+High nitrogen influent concentrations might be associated with higher flows and decreased BMP retention times in the included studies.
 However, we did not collect associated flow data or discern between flow-weighted and mean concentration data within this study.
@@
 We also did not find evidence that BMP type or aridity moderated nitrogen removal.
-This is result is largely consistent with findings in reviews by @kochNitrogenRemovalStormwater2014, @hagerPerformanceLowimpactDevelopment2019 and @horvathEffectsRegionalClimate2023.
+This result is largely consistent with findings in reviews by @kochNitrogenRemovalStormwater2014, @hagerPerformanceLowimpactDevelopment2019 and @horvathEffectsRegionalClimate2023.
 There are a large number of abiotic and biotic processes that control nitrogen retention and removal in BMPs and these processes are moderated by both site specific climate and design factors [@lefevreReviewDissolvedPollutants2015; @valencaNitrateRemovalUncertainty2021].
 It is likely that these site specific factors (retained soil moisture, submerged anoxic zones, vegetation, media composition) are not captured by our broad categorization of BMP types and aridity index values.
-For example, @valencaNitrateRemovalUncertainty2021, using data from the International Stormwater Database, showed that the relative importance of climate and design variables for moderating nitrogen removal varied by BMP type.
+For example, @valencaNitrateRemovalUncertainty2021, using data from the International Stormwater BMP Database, showed that the relative importance of climate and design variables for moderating nitrogen removal varied by BMP type.
 
-*compare nutrient effects to koch, barrett, clary*
+Although influent phosphorus concentration was included in the selected TP and PO~4~ models, they provided relatively low explanatory ability.
+By comparison, @horvathEffectsRegionalClimate2023 found three types of BMPs resulted in differing TP and dissolved inorganic phosphorus removal rates with influent concentrations explaining a small proportion of removal rate variance.
+Again, site-specific factors not captured in our broad categorizations of BMP type and aridity index play a role in differential phosphorus removal rates.
+Soil physical characteristics and media amendments (iron for example) can play an critical role in sorption capacity and are dependent on covariates such as contact time and pH [@hoganBestManagementPractices2007; @lefevreReviewDissolvedPollutants2015].
 
-*Linking BMP performance to watershed scale improvement remains a challenge (Tomer and Locke, Melland et al, Meals).*
-*The succesful implementation and maintence of BMPs at watershed scales remains an area for more investigation (Liu et al., Lintern).*
-*Long-term trans-disciplinary monitoring projects not just on BMP performance but maintance, management, and social apsects may elucidate unknown factors influencing the ability to scale BMP reductions to watershed scale improvements (Lintern et al.)*
+A few reviews have noted a common trend of insufficient methodological and site specific data among peer-reviewed BMP performance studies [@grudzinskiDoesRiparianFencing2020; @eagleMetaAnalysisConstrained2017; @liuReviewEffectivenessBest2017].
+We confirm that inconsistent reporting among studies complicates data extraction, effect size calculations, and attributing important sources of variance.
+The International Stormwater BMP Database addresses some of these concerns through a standardized reporting format.
+Our future efforts will incorporate data from the International Stormwater BMP Database with data retrieved through a systematic review.
 
-*Runoff volume was not incorporated into our models, largely attributed to lack of available data.
-However, BMP size is typically scaled to accommodate runoff volumes of a particular return period volume (1-yr or 2-yr flows for example) (Hoss et al.).*
-*data exploration did not suggest this was the case, but it is reasonable to suggest the the underlying relationship between ROM and any of the moderators are a nonlinear function.
-Semi-parametric approaches might provide some additional  insight*
+In summary, we used multi-level random effects meta-regression models to estimate overall BMP effectiveness from systematically reviewed studies. 
+Although there was relatively high variance between studies, we found strong evidence that BMPs reduce overall mean FIB, TN, TP, and TSS concentrations.
+Influent concentrations moderated BMP efficiency for both FIB and PO~4~, with larger removal rates at high influent concentrations.
+We found that aridity and BMP subcategory moderated BMP performance for only FIB.
+Most likely, site specific design and climate variables not captured in our review or by our choice in BMP classification approach play a more important role in explaining BMP performance variability.
+Future efforts should seek to retrieve more detailed study information.
+Furthermore, our systematic review highlights the poor spatial coverage of BMP studies. 
+The reviewed studies therefore fail to incorporate the range of soil, climate, and runoff conditions needed to adequately link BMP performance to local predictors.
+
+Scaling BMP pollutant reductions to basin wide water quality improvement remains a substantial challenge [@tomerChallengeDocumentingWater2011; @mellandEffectsAgriculturalLand2018; @mealsLagTimeWater2010].
+The major identified challenges include the lack of long term studies, inadequate data collected on BMP management, and incomplete understanding of BMP function [@linternBestManagementPractices2020; @liuReviewEffectivenessBest2017].
+There is a clear need to fill knowledge gaps through additional long-term spatially relevant BMP studies.
+However, we emphasize the need for convergent research approaches that better align study design and reporting that produces data aligned with data synthesis and modeling approaches.
 
@@
 ```
-
-<!-- #  -->
-<!-- # ```{r marginalmeans} -->
-<!-- # #| fig.cap: "Marginal effects of influent pollutant concentrations on the log ratio of means (\\textit{ROM}).
-Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.
-The solid lines are the marginal mean effect, dashed lines are the 95% confidence interval around the mean, and the dotted lines are 95% prediction intervals.
-Here, *k* represents the number of effects with the number of studies in parenthesis." -->
-<!-- # #| fig.pos: "H" -->
-<!-- # #| out.extra: "" -->
-<!-- # #| out.width: "100%" -->
-<!-- #  -->
-<!-- # knitr::include_graphics("../figures/marginal_means.pdf") -->
-<!-- #  -->
-<!-- #  -->
-<!-- #``` -->

### 2024-02-07 frontiers_submission/frontiers_submission.tex
wrap up discussion

@@
 \begin{abstract}
-Abstract length and content varies depending on article type. Refer to
-\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract requirement
-and length according to article type.
+We conducted a systematic review of published nonpoint source water quality best management practice (BMP) studies and used a meta-analysis approach to describe variance in pollutant removal performance.
+Despite high variance, we found the BMPs on average were effective at reducing fecal indicator bacteria (FIB), total nitrogen (TN), total phosphorus (TP), and total suspended sediment (TSS) concentrations.
+We explored how much BMP pollutant removal process, influent pollutant concentration, and aridity effected BMP performance.
+We found that influent concentration and interaction effect between the BMP pollutant removal process and aridity explained a substantial amount of variance in BMP performance in FIB removal.
+Influent concentration explained a small amount of variability in BMP removal of TP and orthophosphate (PO\textsubscript{4}).
+We did not find evidence that any of our chosen variables moderated BMP performance in nitrogen or TSS removal.
+Through our systematic review, we found inadequate spatial representation of BMP studies to capture the underlying variability in climate, soil, and other conditions that could impact BMP performance.
 
@@
 In the United States (U.S.), major improvement in water quality has been achieved under the Clean Water Act of 1972.
-This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge \citep{benhamLessonsLearnedTMDL2008, nationalresearchcouncilAssessingTMDLApproach2001, schrammTotalMaximumDaily2022}.
+This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge \citep{nationalresearchcouncilAssessingTMDLApproach2001, benhamLessonsLearnedTMDL2008, schrammTotalMaximumDaily2022}.
 Increased pollutant loads and concentrations in runoff resulting from land use changes are a particular challenge.
@@
 We conducted a systematic review of recent (2000-2022) literature to compile field studies documenting the effectiveness of best management practices on fecal indicator bacteria and nutrients concentrations.
-Prior meta-analysis have utilized data reported in the International Stormwater Database, which consists of self-reported and quality checked BMP data \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, horvathEffectsRegionalClimate2023}.
-The International Stormwater Database only recently added agricultural BMPs and has relatively sparse FIB data \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014}.
+Prior meta-analysis have utilized data reported in the International Stormwater BMP Database (\url{https://bmpdatabase.org/}), which consists of self-reported and quality checked BMP data \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, horvathEffectsRegionalClimate2023}.
+The International Stormwater BMP Database only recently added agricultural BMPs and has relatively sparse FIB data \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014}.
 Since we had interest in both FIB performance and agricultural BMPs we chose to utilize a systematic review.
-We also assume that most published reports undergo some method of internal or external peer review prior to publication, which is not necessarily the case with all of the data published in the International Stormwater Database.
 
-The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines {[}\citet{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}; Figures S1-S2{]}.
+The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines \citep{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}.
 In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
@@
 Although results from Google Scholar are not always replicable, we utilized the service to maximize search results for studies not published in academic journals and presumably increase the chance of identifying studies with negative effects.
-Fecal indicator bacteria study seraches included the following query: ``fecal indicator bacteria'' OR ``\emph{E.
+Fecal indicator bacteria study searches included the following query: ``fecal indicator bacteria'' OR ``\emph{E.
 coli}'' OR ``\emph{Escherichia coli}'' OR ``enterococci'' OR ``enterococcus'' AND ``best management practices'' OR ``BMPs'' AND ``effectiveness'' OR ``performance''.
@@
 
-\caption{\label{tab:modelaic}Model selection AICc.}
+\caption{\label{tab:modelaic}Sumary of AICc values used for model selection. Bolded values indicate the selected candidate model.}
 \centering
@@
 Changes in performance can also be described using a linear regression using date (transformed as a numeric variable) as an independent variable and log-transformed water quality as the dependent variable \citep{mitschCreatingWetlandsPrimary2012, mitschValidationEcosystemServices2014, pausAssessmentHydraulicToxic2014}.
-Slopes are a valid effect size for use in meta-analysis but the set of covairates used between studies should be the same since the coeffecient of interest is adjusted to account for other terms in the the regression model \citep{beckerSynthesisRegressionSlopes2007}.
+Slopes are a valid effect size for use in meta-analysis but the set of covariates used between studies should be the same since the coeffecient of interest is adjusted to account for other terms in the the regression model \citep{beckerSynthesisRegressionSlopes2007}.
 It would be reasonable to assume that regressions equations vary between studies to adjust results for seasonality, flow rates, and other variables.
@@
 The results are in general agreement with previous reviews that found effective (but highly variable) removal efficiencies for nitrogen, phosphorus, and sediment \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, liuReviewEffectivenessBest2017}.
-The FIB results are useful in particular because the general consensus has been that the effectiveness of BMPs at reducing FIBs and pathogens has been sparsely reviewed and understudied \citep{hagerPerformanceLowimpactDevelopment2019}.
+The FIB results are useful in particular because FIBs performance by BMPs have been sparsely reviewed and generally understudied \citep{hagerPerformanceLowimpactDevelopment2019}.
+
 The FIB reductions generally agreed with our hypothesis that BMP type, influent concentration, and aridity moderate the effectiveness.
@@
 Despite this, the results are promising considering the major limitations of using FIB as a water quality criteria.
-For example, FIB can originate from non-human source, naturalize in soils, and have considerably different fate and transport mechanisms than the human pathogens of underlying concern.
+It is important to note that FIB can originate from non-human source and naturalize in soils, and result in different underlying risk of illness \citep{ishiiEscherichiaColiEnvironment2008, schoenAssessingPathogenRisk2010, sollerEstimatedHumanHealth2010, fujiokaRecreationalWaterQuality2015}.
+Since the fate and transport of human pathogens within BMPs can potentially differ from FIB, BMP choices probably should not be based on FIB reduction alone as alternative indicators or even direct pathogen measurement becomes available \citep{pengIndicatorPathogenRemoval2016, waltersPersistenceNucleicAcid2009}.
+However, relatively few studies have compared human pathogen and FIB removal rates within BMPs \citep{rughHighlyVariableRemoval2022}.
 
-The effect of higher FIB removal rates at high concentrations is expected as filtration and other removal processes associated with BMPs become saturated(?)\ldots{} something about sediment associated fib\ldots{}
+While we observed a strong relationship between FIB influent concentration and FIB removal across all BMPs, we anticipated this relationship to vary by BMP subcategory.
+The reliance on certain removal processes by BMP subcategories was expected to effect the ability of the BMP to retain FIB at higher or lower concentrations.
+We did have some evidence of differing BMP subcateogy removal under different ariditiy indecies.
 The impact of aridity might be due to differential fate and transport processes in arid versus humid environments.
@@
 Conversly, these conditions are also less hospitable to the protoza, bacteriophages, and micro-zooplankton that can play a strong role in predating on and controlling FIB concentrations within BMP media \citep{zhangCaptureDestructionEscherichia2010, burtchettMicrozooplanktonGrazingMeans2017, deanMetaanalysisAddressingImplications2022}.
-Site specific conditions (such as retained soil moisture, turbidity, vegetation, and other factors) play an important role in these fate and transport mechanisms.
-While our models capture some of the variance due to these differences as \emph{between study} effects, including these as moderators in a meta-regression model would be valuable but these details are under reported in BMP studies.
+Site specific conditions (such as retained soil moisture, turbidity, vegetation, and other factors) play an important role in bacteria survival as well as for influencing the filtration and attachment processes that retain FIB within BMP media.
+For example, the presence or absence of a submerged zone within a bioretention BMP has a strong effect on FIB removal \citep{rippyMeetingCriteriaLinking2015, pengIndicatorPathogenRemoval2016}.
+While our models capture some of the variance due to these differences as between study effects, including these variables as fixed effect moderators in a meta-regression model would be valuable but these details are under reported in BMP studies.
 
@@
 Increased flow rates, which can reduce residence time and increase BMP flushing, lowers nitrogen retention \citep{wollheimRetentionUrbanizingHeadwater2005, craigStreamRestorationStrategies2008}.
-High nitrogen influent concentrations might be associated with higher flows and decresed BMP retention times in the included studies.
+High nitrogen influent concentrations might be associated with higher flows and decreased BMP retention times in the included studies.
 However, we did not collect associated flow data or discern between flow-weighted and mean concentration data within this study.
@@
 We also did not find evidence that BMP type or aridity moderated nitrogen removal.
-This is result is largely consistent with findings in reviews by \citet{kochNitrogenRemovalStormwater2014}, \citet{hagerPerformanceLowimpactDevelopment2019} and \citet{horvathEffectsRegionalClimate2023}.
+This result is largely consistent with findings in reviews by \citet{kochNitrogenRemovalStormwater2014}, \citet{hagerPerformanceLowimpactDevelopment2019} and \citet{horvathEffectsRegionalClimate2023}.
 There are a large number of abiotic and biotic processes that control nitrogen retention and removal in BMPs and these processes are moderated by both site specific climate and design factors \citep{lefevreReviewDissolvedPollutants2015, valencaNitrateRemovalUncertainty2021}.
 It is likely that these site specific factors (retained soil moisture, submerged anoxic zones, vegetation, media composition) are not captured by our broad categorization of BMP types and aridity index values.
-For example, \citet{valencaNitrateRemovalUncertainty2021}, using data from the International Stormwater Database, showed that the relative importance of climate and design variables for moderating nitrogen removal varied by BMP type.
+For example, \citet{valencaNitrateRemovalUncertainty2021}, using data from the International Stormwater BMP Database, showed that the relative importance of climate and design variables for moderating nitrogen removal varied by BMP type.
 
-\emph{compare nutrient effects to koch, barrett, clary}
+Although influent phosphorus concentration was included in the selected TP and PO\textsubscript{4} models, they provided relatively low explanatory ability.
+By comparison, \citet{horvathEffectsRegionalClimate2023} found three types of BMPs resulted in differing TP and dissolved inorganic phosphorus removal rates with influent concentrations explaining a small proportion of removal rate variance.
+Again, site-specific factors not captured in our broad categorizations of BMP type and aridity index play a role in differential phosphorus removal rates.
+Soil physical characteristics and media amendments (iron for example) can play an critical role in sorption capacity and are dependent on covariates such as contact time and pH \citep{hoganBestManagementPractices2007, lefevreReviewDissolvedPollutants2015}.
 
-\emph{Linking BMP performance to watershed scale improvement remains a challenge (Tomer and Locke, Melland et al, Meals).}
-\emph{The succesful implementation and maintence of BMPs at watershed scales remains an area for more investigation (Liu et al., Lintern).}
-\emph{Long-term trans-disciplinary monitoring projects not just on BMP performance but maintance, management, and social apsects may elucidate unknown factors influencing the ability to scale BMP reductions to watershed scale improvements (Lintern et al.)}
+A few reviews have noted a common trend of insufficient methodological and site specific data among peer-reviewed BMP performance studies \citep{grudzinskiDoesRiparianFencing2020, eagleMetaAnalysisConstrained2017, liuReviewEffectivenessBest2017}.
+We confirm that inconsistent reporting among studies complicates data extraction, effect size calculations, and attributing important sources of variance.
+The International Stormwater BMP Database addresses some of these concerns through a standardized reporting format.
+Our future efforts will incorporate data from the International Stormwater BMP Database with data retrieved through a systematic review.
 
-\emph{Runoff volume was not incorporated into our models, largely attributed to lack of available data.
-However, BMP size is typically scaled to accommodate runoff volumes of a particular return period volume (1-yr or 2-yr flows for example) (Hoss et al.).}
-\emph{data exploration did not suggest this was the case, but it is reasonable to suggest the the underlying relationship between ROM and any of the moderators are a nonlinear function.
-Semi-parametric approaches might provide some additional insight}
+In summary, we used multi-level random effects meta-regression models to estimate overall BMP effectiveness from systematically reviewed studies.
+Although there was relatively high variance between studies, we found strong evidence that BMPs reduce overall mean FIB, TN, TP, and TSS concentrations.
+Influent concentrations moderated BMP efficiency for both FIB and PO\textsubscript{4}, with larger removal rates at high influent concentrations.
+We found that aridity and BMP subcategory moderated BMP performance for only FIB.
+Most likely, site specific design and climate variables not captured in our review or by our choice in BMP classification approach play a more important role in explaining BMP performance variability.
+Future efforts should seek to retrieve more detailed study information.
+Furthermore, our systematic review highlights the poor spatial coverage of BMP studies.
+The reviewed studies therefore fail to incorporate the range of soil, climate, and runoff conditions needed to adequately link BMP performance to local predictors.
+
+Scaling BMP pollutant reductions to basin wide water quality improvement remains a substantial challenge \citep{tomerChallengeDocumentingWater2011, mellandEffectsAgriculturalLand2018, mealsLagTimeWater2010}.
+The major identified challenges include the lack of long term studies, inadequate data collected on BMP management, and incomplete understanding of BMP function \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017}.
+There is a clear need to fill knowledge gaps through additional long-term spatially relevant BMP studies.
+However, we emphasize the need for convergent research approaches that better align study design and reporting that produces data aligned with data synthesis and modeling approaches.
 

### 2024-02-19 frontiers_submission/frontiers_submission.Rmd
incorporate internal comments.

@@
 
-In the United States (U.S.), major improvement in water quality has been achieved under the Clean Water Act of 1972.
-This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge [@nationalresearchcouncilAssessingTMDLApproach2001; @benhamLessonsLearnedTMDL2008; @schrammTotalMaximumDaily2022].
+In the United States (U.S.), major improvements in water quality have been achieved under the Clean Water Act of 1972.
+This progress has been largely attributed to investments and reductions in point source discharges while reduction in nonpoint source pollutants remains a substantial challenge [@nationalresearchcouncilAssessingTMDLApproach2001; @benhamLessonsLearnedTMDL2008; @schrammTotalMaximumDaily2022].
 Increased pollutant loads and concentrations in runoff resulting from land use changes are a particular challenge.
@@
 BMPs control the delivery of pollutants through a few possible mechanisms. 
-Structural BMPs reduce and retard total volume of runoff, thus reducing both the volume of water and pollutant load. 
+Structural BMPs (detention pond or vegetated filter strips as examples) reduce and retard total volume of runoff, thus reducing both the volume of water and pollutant load. 
 Structural BMPs may also provide a mechanism for physical, chemical, or biological removal of pollutant constituents suspended or dissolved in runoff.
@@
 The statistical properties of *ROM* (normal distribution around zero and additive properties) are preferable to using BMP efficiency [@osenbergEffectSizeEcological1997; @hedgesMetaanalysisResponseRatios1999].
-*ROM* \textgreater 0 indicates higher percent reductions and *ROM* \textless indicates pollutant leaching.
+*ROM* \textgreater 0 indicates higher percent reductions and *ROM* \textless 0 indicates pollutant leaching.
 One advantage of *ROM* is that the statistical results calculated using *ROM* are easily transformed to BMP efficiency for interpretation:
@@
 
-We conducted a systematic review of recent (2000-2022) literature to compile field studies documenting the effectiveness of best management practices on fecal indicator bacteria and nutrients concentrations.
+We conducted a systematic review of recent (2000-2022) literature to compile U.S.
+field studies documenting the effectiveness of best management practices on fecal indicator bacteria, nutrient, and TSS concentrations.
 Prior meta-analysis have utilized data reported in the International Stormwater BMP Database (https://bmpdatabase.org/), which consists of self-reported and quality checked BMP data [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014; @horvathEffectsRegionalClimate2023].
@@
 Searches, review, and data extractions were conducted separately for FIB and nutrient related parameters.
-See Figure S1 and S2 for ROSES diagram.
+See Figure S1 and S2 for ROSES diagram. BMP data from the systematic review is available in @kikoyoTWRINPSBMPDB2024.
 
@@
                            "Studies published from 2000 through 2022."),
-  "Excluision Criteria" = c("Synopsis or review studies, peports with reductions based on modeled or other estimated reductions (e.g. TMDLs, watershed plans, or modelling studies.",
+  "Excluision Criteria" = c("Synopsis or review studies, reports with reductions based on modeled or other estimated reductions (e.g. TMDLs, watershed plans, or modelling studies.",
                             "Studies not explicitly linking reductions to a specific BMP or insufficient information to quantify reductions.",
@@
   Description = c("Year the study was published",
-                  "The specific pollutant measured.",
-                  "Dominant source of runoff (crop fields, livestock pasture, commerical, residential",
+                  "The specific pollutant measured",
+                  "Dominant source of runoff (crop fields, livestock pasture, commerical, residential)",
                   "Major categorization of runoff source: agricultural or urban",
@@
                   "BMP subcategorization based on pollutant removal processes",
-                  'Spatial scale of the study area ("lot/field", "community", "watershed")',
+                  'Spatial scale of the study area (lot/field, community, watershed)',
                   "Location name used in the study description",
@@
 where *v* represents the sampling variance, $CV_{control,i}$ and $CV^2_{experiment,i}$ are the coefficients of variation from the *i*th study for studies 1, 2, ..., *K*.
+*n~control~* and *n~experiment~* are the number of samples in the control (pre-treatment) trial or experimental (post-treatment) trial respectively.
 
@@
 
-Our systematic review identified a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effects for PO~4~, and 33 studies with 125 effect sizes for TSS.
+Our systematic review identified a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effect sizes for PO~4~, and 33 studies with 125 effect sizes for TSS.
 The majority of studies were identified as smaller scaled lot or field studies (Figure \@ref(fig:bmpsummary) A).
-FIB studies had a roughly equal proportion large watershed/catchment studies and studies conducted at the commuity/farm scale or smaller.
+FIB studies had a roughly equal proportion of large watershed/catchment studies and studies conducted at the community/farm scale or smaller.
 We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \@ref(fig:bmpsummary) B).
 We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \@ref(fig:bmpsummary) C).
-Our review was resticted to studies published after 1999. 
+Our review was restricted to studies published after between 1999 and 2023.
 The number of studies published for each parameter were roughly uniformly distributed over time (Figure \@ref(fig:bmptemporal) A) and are not indicative of increases or decreases in the number of published studies.
-Study length was strongly skewed for all parameters (Figure \@ref(fig:bmptemporal) B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO~4), 2.5 (TN), 2.5 (TP), and 2 (TSS). 
+Study length was strongly skewed for all parameters (Figure \@ref(fig:bmptemporal) B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO~4), 2.5 (TN), 2.5 (TP), and 2 (TSS) years. 
 There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \@ref(fig:studymap)).
@@
 Moderator terms and interactions explained a high proportion of effect size variance (`r fib_r2`) in the FIB model.
-Increased influent concentrations ($\beta$ = 0.25, 95% CI [0.14, 0.37]) resulted in significantly larger ROM effect for FIB (Figure \@ref(fig:fibeffect).
+Increased influent concentrations ($\beta$ = 0.25, 95% CI [0.14, 0.37]) resulted in significantly larger ROM effect for FIB (Figure \@ref(fig:fibeffect)).
 Compared to the baseline aridity index\texttimes detention BMP subcategory interaction, infiltration ($\beta$ = -29.90, 95% CI [-50.34, -9.47]), livestock management ($\beta$ = -30.37, 95% CI [-50.93, -9.81]), and treatment ($\beta$ = -30.33, 95% CI [-49.62, -11.03]) interactions had significantly smaller slopes.
@@
 
-**Not complete**
-
-Our systematic review revealed strong spatial disparities in published BMP studies (Figure \@ref(fig:studymap).
+Our systematic review revealed strong spatial disparities in published BMP studies (Figure \@ref(fig:studymap)).
 Similar spatial disparities have been identified and discussed in @kochNitrogenRemovalStormwater2014, @grudzinskiDoesRiparianFencing2020 and can be problematic for extrapolating results to other regions of interest.
@@
 @horvathEffectsRegionalClimate2023 found overlapping BMP type and climate groups within their dataset that reduce the ability to distinguish effects due to either BMP type or climate.
-Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \@ref(fig:fibeffect).
+Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \@ref(fig:fibeffect)).
 Not only were there spatially disparities, but we observed that the relative distribution of aridity index values does not resemble the distribution of aridity values across the U.S.
-(Figure \@ref(fig:aiindex).
+(Figure \@ref(fig:aiindex)).
 Our review indicates that BMP studies are over represented in the generally humid regions of the country and underrepresented the more arid regions.
@@
 Study design prevented us from properly assessing BMP effectiveness as a function of age.
-Conducting a meta-analysis of BMP effectivness over time is hampered both by the lack of long-term studies and lack of standardized reporting mechanisms.
-Some studies simply describe the change in pollutant concentrations or loads at the beginning and end of the study as a percent change [@haileHydraulicPerformancePollutant2016] which presents statistical problems, espcially when sampling variance is not reported.
+Conducting a meta-analysis of BMP effectiveness over time is hampered both by the lack of long-term studies and lack of standardized reporting mechanisms.
+Some studies simply describe the change in pollutant concentrations or loads at the beginning and end of the study as a percent change [@haileHydraulicPerformancePollutant2016] which presents statistical problems, especially when sampling variance is not reported.
 Changes in performance can also be described using a linear regression using date (transformed as a numeric variable) as an independent variable and log-transformed water quality as the dependent variable [@mitschCreatingWetlandsPrimary2012; @mitschValidationEcosystemServices2014; @pausAssessmentHydraulicToxic2014].
-Slopes are a valid effect size for use in meta-analysis but the set of covariates used between studies should be the same since the coeffecient of interest is adjusted to account for other terms in the the regression model [@beckerSynthesisRegressionSlopes2007].
+Slopes are a valid effect size for use in meta-analysis but the set of covariates used between studies should be the same since the coefficient of interest is adjusted to account for other terms in the the regression model [@beckerSynthesisRegressionSlopes2007].
 It would be reasonable to assume that regressions equations vary between studies to adjust results for seasonality, flow rates, and other variables.
@@
 The reliance on certain removal processes by BMP subcategories was expected to effect the ability of the BMP to retain FIB at higher or lower concentrations.
-We did have some evidence of differing BMP subcateogy removal under different ariditiy indecies.
+We did have some evidence of differing BMP subcategory removal under different aridity.
 The impact of aridity might be due to differential fate and transport processes in arid versus humid environments. 
 On one hand, we assume that arid conditions might be less hospitable to FIBs due to increased UV exposure and osmotic stress. 
-Conversly, these conditions are also less hospitable to the protoza, bacteriophages, and micro-zooplankton that can play a strong role in predating on and controlling FIB concentrations within BMP media  [@zhangCaptureDestructionEscherichia2010; @burtchettMicrozooplanktonGrazingMeans2017; @deanMetaanalysisAddressingImplications2022].
+Conversely, these conditions are also less hospitable to the protozoa, bacteriophages, and micro-zooplankton that can play a strong role in predating on and controlling FIB concentrations within BMP media  [@zhangCaptureDestructionEscherichia2010; @burtchettMicrozooplanktonGrazingMeans2017; @deanMetaanalysisAddressingImplications2022].
 Site specific conditions (such as retained soil moisture, turbidity, vegetation, and other factors) play an important role in bacteria survival as well as for influencing the filtration and attachment processes that retain FIB within BMP media.
@@
 
-The data and R code used in this study are deposited in Zenodo: TBD.
+Raw data from the systematic review is deposited on Zenodo: https://zenodo.org/doi/10.5281/zenodo.8302472.
+R code used in the analysis is also deposited on Zenodo: TBD.
 

### 2024-02-19 frontiers_submission/frontiers_submission.tex
incorporate internal comments.

@@
 
-In the United States (U.S.), major improvement in water quality has been achieved under the Clean Water Act of 1972.
-This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge \citep{nationalresearchcouncilAssessingTMDLApproach2001, benhamLessonsLearnedTMDL2008, schrammTotalMaximumDaily2022}.
+In the United States (U.S.), major improvements in water quality have been achieved under the Clean Water Act of 1972.
+This progress has been largely attributed to investments and reductions in point source discharges while reduction in nonpoint source pollutants remains a substantial challenge \citep{nationalresearchcouncilAssessingTMDLApproach2001, benhamLessonsLearnedTMDL2008, schrammTotalMaximumDaily2022}.
 Increased pollutant loads and concentrations in runoff resulting from land use changes are a particular challenge.
@@
 BMPs control the delivery of pollutants through a few possible mechanisms.
-Structural BMPs reduce and retard total volume of runoff, thus reducing both the volume of water and pollutant load.
+Structural BMPs (detention pond or vegetated filter strips as examples) reduce and retard total volume of runoff, thus reducing both the volume of water and pollutant load.
 Structural BMPs may also provide a mechanism for physical, chemical, or biological removal of pollutant constituents suspended or dissolved in runoff.
@@
 The statistical properties of \emph{ROM} (normal distribution around zero and additive properties) are preferable to using BMP efficiency \citep{osenbergEffectSizeEcological1997, hedgesMetaanalysisResponseRatios1999}.
-\emph{ROM} \textgreater 0 indicates higher percent reductions and \emph{ROM} \textless indicates pollutant leaching.
+\emph{ROM} \textgreater 0 indicates higher percent reductions and \emph{ROM} \textless 0 indicates pollutant leaching.
 One advantage of \emph{ROM} is that the statistical results calculated using \emph{ROM} are easily transformed to BMP efficiency for interpretation:
@@
 
-We conducted a systematic review of recent (2000-2022) literature to compile field studies documenting the effectiveness of best management practices on fecal indicator bacteria and nutrients concentrations.
+We conducted a systematic review of recent (2000-2022) literature to compile U.S.
+field studies documenting the effectiveness of best management practices on fecal indicator bacteria, nutrient, and TSS concentrations.
 Prior meta-analysis have utilized data reported in the International Stormwater BMP Database (\url{https://bmpdatabase.org/}), which consists of self-reported and quality checked BMP data \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, horvathEffectsRegionalClimate2023}.
@@
 Searches, review, and data extractions were conducted separately for FIB and nutrient related parameters.
-See Figure S1 and S2 for ROSES diagram.
+See Figure S1 and S2 for ROSES diagram. BMP data from the systematic review is available in \citet{kikoyoTWRINPSBMPDB2024}.
 
@@
 Study type & Journal articles, book chapters, conference papers, unpublished research reports, thesis and dissertations, organizational and agency white papers.
-& Synopsis or review studies, peports with reductions based on modeled or other estimated reductions (e.g.
+& Synopsis or review studies, reports with reductions based on modeled or other estimated reductions (e.g.
 TMDLs, watershed plans, or modelling studies.\\
@@
 Publication Year & Year the study was published\\
-Parameter & The specific pollutant measured.\\
-Runoff source & Dominant source of runoff (crop fields, livestock pasture, commerical, residential\\
+Parameter & The specific pollutant measured\\
+Runoff source & Dominant source of runoff (crop fields, livestock pasture, commerical, residential)\\
 Source type & Major categorization of runoff source: agricultural or urban\\
@@
 BMP Subcategory & BMP subcategorization based on pollutant removal processes\\
-Study scale & Spatial scale of the study area ("lot/field", "community", "watershed")\\
+Study scale & Spatial scale of the study area (lot/field, community, watershed)\\
 Location & Location name used in the study description\\
@@
 where \emph{v} represents the sampling variance, \(CV_{control,i}\) and \(CV^2_{experiment,i}\) are the coefficients of variation from the \emph{i}th study for studies 1, 2, \ldots, \emph{K}.
+\emph{n\textsubscript{control}} and \emph{n\textsubscript{experiment}} are the number of samples in the control (pre-treatment) trial or experimental (post-treatment) trial respectively.
 
@@
 
-Our systematic review identified a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effects for PO\textsubscript{4}, and 33 studies with 125 effect sizes for TSS.
+Our systematic review identified a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effect sizes for PO\textsubscript{4}, and 33 studies with 125 effect sizes for TSS.
 The majority of studies were identified as smaller scaled lot or field studies (Figure \ref{fig:bmpsummary} A).
-FIB studies had a roughly equal proportion large watershed/catchment studies and studies conducted at the commuity/farm scale or smaller.
+FIB studies had a roughly equal proportion of large watershed/catchment studies and studies conducted at the community/farm scale or smaller.
 We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \ref{fig:bmpsummary} B).
 We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \ref{fig:bmpsummary} C).
-Our review was resticted to studies published after 1999.
+Our review was restricted to studies published after between 1999 and 2023.
 The number of studies published for each parameter were roughly uniformly distributed over time (Figure \ref{fig:bmptemporal} A) and are not indicative of increases or decreases in the number of published studies.
-Study length was strongly skewed for all parameters (Figure \ref{fig:bmptemporal} B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO\textasciitilde4), 2.5 (TN), 2.5 (TP), and 2 (TSS).
+Study length was strongly skewed for all parameters (Figure \ref{fig:bmptemporal} B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO\textasciitilde4), 2.5 (TN), 2.5 (TP), and 2 (TSS) years.
 There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \ref{fig:studymap}).
@@
 Moderator terms and interactions explained a high proportion of effect size variance (\textit{R\textsuperscript{2}\textsubscript{marginal}} =0.89) in the FIB model.
-Increased influent concentrations (\(\beta\) = 0.25, 95\% CI {[}0.14, 0.37{]}) resulted in significantly larger ROM effect for FIB (Figure \ref{fig:fibeffect}.
+Increased influent concentrations (\(\beta\) = 0.25, 95\% CI {[}0.14, 0.37{]}) resulted in significantly larger ROM effect for FIB (Figure \ref{fig:fibeffect}).
 Compared to the baseline aridity index\texttimes detention BMP subcategory interaction, infiltration (\(\beta\) = -29.90, 95\% CI {[}-50.34, -9.47{]}), livestock management (\(\beta\) = -30.37, 95\% CI {[}-50.93, -9.81{]}), and treatment (\(\beta\) = -30.33, 95\% CI {[}-49.62, -11.03{]}) interactions had significantly smaller slopes.
@@
 
-\textbf{Not complete}
-
-Our systematic review revealed strong spatial disparities in published BMP studies (Figure \ref{fig:studymap}.
+Our systematic review revealed strong spatial disparities in published BMP studies (Figure \ref{fig:studymap}).
 Similar spatial disparities have been identified and discussed in \citet{kochNitrogenRemovalStormwater2014}, \citet{grudzinskiDoesRiparianFencing2020} and can be problematic for extrapolating results to other regions of interest.
@@
 \citet{horvathEffectsRegionalClimate2023} found overlapping BMP type and climate groups within their dataset that reduce the ability to distinguish effects due to either BMP type or climate.
-Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \ref{fig:fibeffect}.
+Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \ref{fig:fibeffect}).
 Not only were there spatially disparities, but we observed that the relative distribution of aridity index values does not resemble the distribution of aridity values across the U.S.
-(Figure \ref{fig:aiindex}.
+(Figure \ref{fig:aiindex}).
 Our review indicates that BMP studies are over represented in the generally humid regions of the country and underrepresented the more arid regions.
@@
 Study design prevented us from properly assessing BMP effectiveness as a function of age.
-Conducting a meta-analysis of BMP effectivness over time is hampered both by the lack of long-term studies and lack of standardized reporting mechanisms.
-Some studies simply describe the change in pollutant concentrations or loads at the beginning and end of the study as a percent change \citep{haileHydraulicPerformancePollutant2016} which presents statistical problems, espcially when sampling variance is not reported.
+Conducting a meta-analysis of BMP effectiveness over time is hampered both by the lack of long-term studies and lack of standardized reporting mechanisms.
+Some studies simply describe the change in pollutant concentrations or loads at the beginning and end of the study as a percent change \citep{haileHydraulicPerformancePollutant2016} which presents statistical problems, especially when sampling variance is not reported.
 Changes in performance can also be described using a linear regression using date (transformed as a numeric variable) as an independent variable and log-transformed water quality as the dependent variable \citep{mitschCreatingWetlandsPrimary2012, mitschValidationEcosystemServices2014, pausAssessmentHydraulicToxic2014}.
-Slopes are a valid effect size for use in meta-analysis but the set of covariates used between studies should be the same since the coeffecient of interest is adjusted to account for other terms in the the regression model \citep{beckerSynthesisRegressionSlopes2007}.
+Slopes are a valid effect size for use in meta-analysis but the set of covariates used between studies should be the same since the coefficient of interest is adjusted to account for other terms in the the regression model \citep{beckerSynthesisRegressionSlopes2007}.
 It would be reasonable to assume that regressions equations vary between studies to adjust results for seasonality, flow rates, and other variables.
@@
 The reliance on certain removal processes by BMP subcategories was expected to effect the ability of the BMP to retain FIB at higher or lower concentrations.
-We did have some evidence of differing BMP subcateogy removal under different ariditiy indecies.
+We did have some evidence of differing BMP subcategory removal under different aridity.
 The impact of aridity might be due to differential fate and transport processes in arid versus humid environments.
 On one hand, we assume that arid conditions might be less hospitable to FIBs due to increased UV exposure and osmotic stress.
-Conversly, these conditions are also less hospitable to the protoza, bacteriophages, and micro-zooplankton that can play a strong role in predating on and controlling FIB concentrations within BMP media \citep{zhangCaptureDestructionEscherichia2010, burtchettMicrozooplanktonGrazingMeans2017, deanMetaanalysisAddressingImplications2022}.
+Conversely, these conditions are also less hospitable to the protozoa, bacteriophages, and micro-zooplankton that can play a strong role in predating on and controlling FIB concentrations within BMP media \citep{zhangCaptureDestructionEscherichia2010, burtchettMicrozooplanktonGrazingMeans2017, deanMetaanalysisAddressingImplications2022}.
 Site specific conditions (such as retained soil moisture, turbidity, vegetation, and other factors) play an important role in bacteria survival as well as for influencing the filtration and attachment processes that retain FIB within BMP media.
@@
 
-The data and R code used in this study are deposited in Zenodo: TBD.
+Raw data from the systematic review is deposited on Zenodo: \url{https://zenodo.org/doi/10.5281/zenodo.8302472}.
+R code used in the analysis is also deposited on Zenodo: TBD.
 

### 2024-03-07 README.Rmd
prep for submission

---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This repository contains the data and analysis used in the draft manuscript *"A  meta-analysis  of  the  impacts  of  best management  practices  on  nonpoint  source pollutant  concentration,"* by Schramm et al.

This project uses `{targets}` as the main organizing workflow. The `_targets.R` file defines the steps of the project workflow amd functions for those setps are stored in the `R/` directory. `renv` is used to manage package versions. To reproduce locally, clone or download this project to your computer. Open the project with RStudio or other IDE.

In the console:

```r
renv::restore()
```

then use `targets::tar_make()` to execute the project pipeline. Each step in the pipeline can be accessed with the `targets::tar_read()` function. For example, `targets::tar_read(tn_model)` will access the model object we fit to the total nitrogen data.

## Funding Statement

This project was funded by a nonpoint source grant from the Texas State Soil and Water Conservation Board.

## Working Environment Info

```{r}
sessioninfo::session_info()
```

### 2024-03-07 frontiers_submission/frontiers_submission.Rmd
prep for submission

@@
 
+Scaling BMP pollutant reductions to basin wide water quality improvement remains a substantial challenge [@tomerChallengeDocumentingWater2011; @mellandEffectsAgriculturalLand2018; @mealsLagTimeWater2010].
+The major identified challenges include the lack of long term studies, inadequate data collected on BMP management, and incomplete understanding of BMP function [@linternBestManagementPractices2020; @liuReviewEffectivenessBest2017].
+There is a clear need to fill knowledge gaps through additional long-term spatially relevant BMP studies.
+However, we emphasize the need for convergent research approaches that better align study design and reporting that produces data aligned with data synthesis and modeling approaches.
+
 In summary, we used multi-level random effects meta-regression models to estimate overall BMP effectiveness from systematically reviewed studies. 
@@
 The reviewed studies therefore fail to incorporate the range of soil, climate, and runoff conditions needed to adequately link BMP performance to local predictors.
-
-Scaling BMP pollutant reductions to basin wide water quality improvement remains a substantial challenge [@tomerChallengeDocumentingWater2011; @mellandEffectsAgriculturalLand2018; @mealsLagTimeWater2010].
-The major identified challenges include the lack of long term studies, inadequate data collected on BMP management, and incomplete understanding of BMP function [@linternBestManagementPractices2020; @liuReviewEffectivenessBest2017].
-There is a clear need to fill knowledge gaps through additional long-term spatially relevant BMP studies.
-However, we emphasize the need for convergent research approaches that better align study design and reporting that produces data aligned with data synthesis and modeling approaches.
+To adequately estimate the effects of moderating variable on BMP performance we suggest that there is a need for additional aligned BMP studies across regions and conditions.
 
@@
 
-Supplementary Material should be uploaded separately on submission, if there are
-Supplementary Figures, please include the caption in the same file as the
-figure. LaTeX Supplementary Material templates can be found in the Frontiers
-LaTeX folder
+Supplementary Material for this article has been uploaded with the submission.
 

### 2024-03-07 frontiers_submission/frontiers_submission.tex
prep for submission

@@
 
+Scaling BMP pollutant reductions to basin wide water quality improvement remains a substantial challenge \citep{tomerChallengeDocumentingWater2011, mellandEffectsAgriculturalLand2018, mealsLagTimeWater2010}.
+The major identified challenges include the lack of long term studies, inadequate data collected on BMP management, and incomplete understanding of BMP function \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017}.
+There is a clear need to fill knowledge gaps through additional long-term spatially relevant BMP studies.
+However, we emphasize the need for convergent research approaches that better align study design and reporting that produces data aligned with data synthesis and modeling approaches.
+
 In summary, we used multi-level random effects meta-regression models to estimate overall BMP effectiveness from systematically reviewed studies.
@@
 The reviewed studies therefore fail to incorporate the range of soil, climate, and runoff conditions needed to adequately link BMP performance to local predictors.
-
-Scaling BMP pollutant reductions to basin wide water quality improvement remains a substantial challenge \citep{tomerChallengeDocumentingWater2011, mellandEffectsAgriculturalLand2018, mealsLagTimeWater2010}.
-The major identified challenges include the lack of long term studies, inadequate data collected on BMP management, and incomplete understanding of BMP function \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017}.
-There is a clear need to fill knowledge gaps through additional long-term spatially relevant BMP studies.
-However, we emphasize the need for convergent research approaches that better align study design and reporting that produces data aligned with data synthesis and modeling approaches.
+To adequately estimate the effects of moderating variable on BMP performance we suggest that there is a need for additional aligned BMP studies across regions and conditions.
 

### 2024-05-23 frontiers_submission/reviewer_response/reviewer_response.qmd
working on responses

---
title: "Response to reviewer comments"
format:
  hikmah-response-typst: default
---

Thank you for giving us the opportunity to revise our manuscript. Reviewer comments were especially constructive and have improved the manuscript. The Reviewers' comments are cited or summarized in [red italicized text]{.memo-reviewer-inline}, and our responses are in black Roman text. We include excerpts from our revised manuscript in [blue]{.memo-excerpt-inline}.

# Reviewer 2

::: {.memo-reviewer}
***Reviewer 2***: The reported work only used BMP category and large-scale 
spatial indicators as predictors of their performance, but the parameters of 
BMPs (e.g., the size of infiltration) are missing. This may be why the model 
has a high variance.
:::

We generally found that specific BMP parameters such as infiltration size, 
drainage area, infiltration media/soil types, and other factors relevant to BMP 
performance were not well reported across studies. 
Our characterization of BMPs is further complicated by inclusion of BMPs that
rely on completely different pollutant removal or prevention mechanisms. 
For example, nutrient management BMPs do not rely on a infiltration area
and we would not be able to include these types of management BMPs if 
infiltration size were a regression moderator.
This is certainly a limitation of the study, but the purpose of the project
is a broad scale look at BMPs. Further studies investigating specific BMP types
with common removal mechanisms are certainly appropriate. 
To address the Reviewer's concerns, we have added the following to the Discussion:

::: {.memo-excerpt}
Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur? Quis autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur, vel illum qui dolorem eum fugiat quo voluptas nulla pariatur?
:::

---

::: {.memo-reviewer}
***Reviewer 2***: The heterogeneity of FIB, TN, TP, TSS are high, the study 
require subgroup analysis or sensitivity analysis on this issue.
:::

I appreciate this comment, but want to caution against extensive subgroup 
analysis/sensitivity analysis. While heterogeneity was indeed high, 
the primary purpose of the meta-regression models are to investigate how much
of the variance that a set of pre-determined moderators explain. Meta-regression
were conducted in lieu of traditional subgroup analysis. While additional 
subgroup analysis could be conducted based on additional moderators, we risk 
conducting a "data-dredging" study.

::: {.memo-excerpt}
Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt. Neque porro quisquam est, qui dolorem ipsum quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat voluptatem.
:::

# Reviewer 3

::: {.memo-reviewer}
***Reviewer 3***: Is there any moderating effects of study area size or scale (i.e., field, farm, and watershed) and urban versus agricultural source on the performance of BMPs?
:::

---

::: {.memo-reviewer}
***Reviewer 3***:  What is the temporal coverage of the aridity index? Does it reflect long-term climate patterns or only a single year? Using a snapshot of the climate variable would be less optimal.
:::

---

::: {.memo-reviewer}
***Reviewer 3***:  Since the locations of the studies are not precisely defined in many cases, please provide more details on how to spatially link studies to the aridity index map to obtain aridity values.
:::

---

::: {.memo-reviewer}
***Reviewer 3***:  Have you examined the effects of soil biophysical properties in addition to aridity? Can you obtain more detailed spatial locations of the various studies so to link these studies to local soil conditions in addition to climate?
:::

---

::: {.memo-reviewer}
***Reviewer 3***:  Could you please elaborate more on the topic “scaling BMP pollutant reductions to basin wide water quality improvement remains a substantial challenge”. This is a very interesting point, what modeling or mixed approaches exist to achieve this goal?
:::

---

::: {.memo-reviewer}
***Reviewer 3***:  Please expand the abbreviation “ROSES” in the main text.
:::

---

::: {.memo-reviewer}
***Reviewer 3***:  Change “Standard deviation of control measreuments” to “… measurements” in Table 2.
:::

---

::: {.memo-reviewer}
***Reviewer 3***:  Change “and outlier” to “an” in Figure S9 to S14.
:::

---

::: {.memo-reviewer}
***Reviewer 3***:  In the sentence "Any study values outside of the full model confidence intervals would be considered an outlier.", by “study values”, do you mean the intercept estimate of a study?
:::

---

::: {.memo-reviewer}
***Reviewer 3***:  Explain the abbreviated parameters in Figure 1 – 5 (e.g. TS, TP, TN, etc.)
:::

Added definitions to all the figure captions.

---

::: {.memo-reviewer}
***Reviewer 3***:   Delete one of the “the” at line 277.
:::

Fixed.

# Reviewer 5

::: {.memo-reviewer}
***Reviewer 5***:   Add spatial extent covered in abstract
:::

---

::: {.memo-reviewer}
***Reviewer 5***:   Add sediments or suspended sediments to keywords if there is space
:::

---

::: {.memo-reviewer}
***Reviewer 5***:   1st Paragraph of introduction (Line 28), add a sentence about sediment impairments to waters similar to the ones discussing FIBs and nutrients impairments
:::

---

::: {.memo-reviewer}
***Reviewer 5***:   Include small paragraph in discussion about how search terms of “BMPs” or “best management practices” and inclusion criteria (concentrations) may have not captured some studies. For example, if the study used the name of the practice instead of BMP. Also, there are a suite of agricultural drainage practices but studies generally focus on load reduction (from flow reduction) compared to concentration reduction (see Frankenberger et al., 2023 https://elibrary.asabe.org/abstract.asp?aid=54314 who’s review has a cluster of those studies in the Midwest, which is blank on this map). 
:::

---

::: {.memo-reviewer}
***Reviewer 5***:   Lines 61 – 64 consider adding an example BMP for instances where effluent is and is not related to influent concentration
:::

---

::: {.memo-reviewer}
***Reviewer 5***:   Line 136 – specify version of R used
:::

Added.

---

::: {.memo-reviewer}
***Reviewer 5***:   Line 149 – check for consistency in defining equation terms (CV v. CV2) 
:::

Fixed. Now reads as below:

::: {.memo-excerpt}
$$
v(ROM) = \frac{\sum_{i=1}^{K}{(CV^2_{control,i})/K}}{n_{control}} + \frac{\sum_{i=1}^{K}{(CV^2_{experiment,i})/K}}{n_{experiment}},
$$

where *v* represents the sampling variance, $CV^2_{control,i}$ and $CV^2_{experiment,i}$ are the squared coefficients of variation from the *i*th study for studies 1, 2, ..., *K*.
:::

---

::: {.memo-reviewer}
***Reviewer 5***:   Line 160 – mean annual potential (or reference) evapotranspiration
:::

Fixed.

---

::: {.memo-reviewer}
***Reviewer 5***:   Line 191 – “published after between 1999 and 2023” confusing wording, 2000-2022?
:::

Fixed.

---

::: {.memo-reviewer}
***Reviewer 5***:   Line 194 – PO~4
:::

Fixed.

---

::: {.memo-reviewer}
***Reviewer 5***:   Lines 213 – 214 – clarify “may not be reliable” in general, or just for lower aridity?
:::

---

::: {.memo-reviewer}
***Reviewer 5***:   Lines 61 – 64 consider adding an example BMP for instances where effluent is and is not related to influent concentration
:::

---

::: {.memo-reviewer}
***Reviewer 5***:   Lines 242-254: Nice discussion on spatial coverage with the caveat that we are not aware of the actual distribution of BMPs. However, I am curious, are BMPs over-represented or just predominately located in humid regions? Could it be related to larger population concentrations along the humid east coast than the arid west? Are detention practices common in locations with little rainfall and high potential loss of water to evaporation? See my recommendation for Figure 7 to potentially add a 3rd distribution of aridity index of urban & agricultural land covers across the US. 
:::

---

::: {.memo-reviewer}
***Reviewer 5***:  Lines 349 – 351: Based on your experience trying to synthesis information across studies in this review process, do you have recommendations for minimum reporting requirements or standards for future studies on BMP concentration reductions? If so, please list in text or perhaps in a box. 
:::

---

::: {.memo-reviewer}
***Reviewer 5***:   Consider defining abbreviated terms in the figure caption for clarity (e.g, DIN, FIB)
:::

We have added terms to all the figure captions.

---

::: {.memo-reviewer}
***Reviewer 5***:  Table 1
- Excluision Criteria → Exclusion Criteria
- Under Inclusion Criteria/Outcomes include sediments in addition for FIB and nutrients
- Under Exclusion criteria/Outcomes “Studies not explicitly linking reductions to a specific BMP…” it may be more clear if you say linking “concentration reductions”
:::

---

::: {.memo-reviewer}
***Reviewer 5***:  Figure 1
- While the color palette may be appropriate for normal, trichromatic, and dichromatic conditions there is very little contrast in the monochromatic or grey-scale view. If article is laser printed or photocopied, it is difficult to distinguish between the stacked elements of the bar graphs. Maybe white or black outline separating stacked sections could help.
- Please increase the length of the y-axis for panel C to for larger font size/less crowding
:::

---

::: {.memo-reviewer}
***Reviewer 5***:  Figure 3
- Consider adjusting the background grey value for no-data in maps. In grey-scale/monochromatic view the mid-tone greens are indistinguishable from the no-value grey. Labeling the states with the number of studies does help.
:::

We changed to a transparent background for states with no data.

---

::: {.memo-reviewer}
***Reviewer 5***:  Figure 4, 5, 6
- Consider adding including the range in sampling variance in the caption or provide a visual key to demonstrate the values represented by differently sized points. 
:::

---

::: {.memo-reviewer}
***Reviewer 5***:  Figure 5
- Add panel letters to figure caption to improve clarity.
- Consider adding note for aridity index (annual precipitation/potential evapotranspiration) in the x-label or in the figure caption that higher values indicate more humid conditions.
:::

Thank you for the suggestions. We have incorporated all the suggestions to the 
figure.

---

::: {.memo-reviewer}
***Reviewer 5***:  Figure 7
- It is difficult to see the tail of the US. Aridity Index values, consider having a slight offset from zero above the x-axis/y-axis intercept.
- Consider having a 3rd distribution shown – U.S. Aridity Index Values for urban and agricultural lands, with other land-covers filtered out, if these are the land-cover types that are associated with the placement of BMPs you analyzed. You could use a product with the same resolution as your Aridity Index values such as the 30m National Land Cover Database (https://www.usgs.gov/centers/eros/science/national-land-cover-database#overview )
- Consider adding note for aridity index (annual precipitation/potential evapotranspiration) in the x-label or in the figure caption that higher values indicate more humid conditions.
:::

Thank you for the suggestions. We have incorporated all the suggestions to the 
figure. 
We switched from overlapping densities to a ridge plot with each category labeled
on the y-axis since it gets difficult to discern densities with three overlapping
plots.

### 2024-05-23 frontiers_submission/frontiers_submission_revision_01.Rmd
working on responses

--- 2024-03-07 frontiers_submission/frontiers_submission.Rmd
@@
 Searches, review, and data extractions were conducted separately for FIB and nutrient related parameters.
-See Figure S1 and S2 for ROSES diagram. BMP data from the systematic review is available in @kikoyoTWRINPSBMPDB2024.
+See Figure S1 and S2 for RepOrting standards for Systematic Evidence Syntheses 
+(ROSES) diagram. 
+BMP data from the systematic review is available in @kikoyoTWRINPSBMPDB2024.
 
@@
                   "Maximum experiment measurement",
-                  "Standard deviation of control measreuments",
+                  "Standard deviation of control measurements",
                   "Standard deviation of experimental measurements",
@@
 
-We used the "rma.mv" function in the *metafor* R package to fit multilevel random effects regression models with *ROM* as the effect variable [@viechtbauerConductingMetaanalysesMetafor2010; @rcoreteamLanguageEnvironmentStatistical2023].
+We used the "rma.mv" function in the *metafor* R package (R version 4.3.1) to fit multilevel random effects regression models with *ROM* as the effect variable [@viechtbauerConductingMetaanalysesMetafor2010; @rcoreteamLanguageEnvironmentStatistical2023].
 We fit separate models for FIB, total nitrogen (TN), dissolved inorganic nitrogen (DIN), total phosphorus (TP), orthophosphate (PO~4~), and total suspended sediment (TSS).
@@
 
-where *v* represents the sampling variance, $CV_{control,i}$ and $CV^2_{experiment,i}$ are the coefficients of variation from the *i*th study for studies 1, 2, ..., *K*.
+where *v* represents the sampling variance, $CV^2_{control,i}$ and $CV^2_{experiment,i}$ are the squared coefficients of variation from the *i*th study for studies 1, 2, ..., *K*.
 *n~control~* and *n~experiment~* are the number of samples in the control (pre-treatment) trial or experimental (post-treatment) trial respectively.
@@
 We mapped study location coordinates to aridity index values published in the "Global Aridity Index and Potential Evapotranspiration Database - Version 3" (Global-AI_PET_v3) which provides gridded 30 arc-second annual average precipitation and potential evapotranspiration estimates [@zomerVersionGlobalAridity2022].
-The aridity index is calculated as the ratio of mean annual precipitation to mean annual evapotraspiration with values between 0 and 0.5 considered hyper to semi-arid, and values above 0.65 as humid.
+The aridity index is calculated as the ratio of mean annual precipitation to mean annual potential (or reference) evapotranspiration with values between 0 and 0.5 considered hyper to semi-arid, and values above 0.65 as humid.
 
@@
 We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \@ref(fig:bmpsummary) C).
-Our review was restricted to studies published after between 1999 and 2023.
+Our review was restricted to studies published from 2000 through 2022.
 The number of studies published for each parameter were roughly uniformly distributed over time (Figure \@ref(fig:bmptemporal) A) and are not indicative of increases or decreases in the number of published studies.
-Study length was strongly skewed for all parameters (Figure \@ref(fig:bmptemporal) B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO~4), 2.5 (TN), 2.5 (TP), and 2 (TSS) years. 
+Study length was strongly skewed for all parameters (Figure \@ref(fig:bmptemporal) B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO~4~), 2.5 (TN), 2.5 (TP), and 2 (TSS) years. 
 There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \@ref(fig:studymap)).
@@
 Changes in performance can also be described using a linear regression using date (transformed as a numeric variable) as an independent variable and log-transformed water quality as the dependent variable [@mitschCreatingWetlandsPrimary2012; @mitschValidationEcosystemServices2014; @pausAssessmentHydraulicToxic2014].
-Slopes are a valid effect size for use in meta-analysis but the set of covariates used between studies should be the same since the coefficient of interest is adjusted to account for other terms in the the regression model [@beckerSynthesisRegressionSlopes2007].
+Slopes are a valid effect size for use in meta-analysis but the set of covariates used between studies should be the same since the coefficient of interest is adjusted to account for other terms in the regression model [@beckerSynthesisRegressionSlopes2007].
 It would be reasonable to assume that regressions equations vary between studies to adjust results for seasonality, flow rates, and other variables.
@@
 Raw data from the systematic review is deposited on Zenodo: https://zenodo.org/doi/10.5281/zenodo.8302472.
-R code used in the analysis is also deposited on Zenodo: TBD.
+R code used in the analysis is also deposited on Zenodo: https://zenodo.org/records/10795231.
 
@@
 ```{r bmpsummary}
-#| fig.cap: "Summary of (A) study scale, (B) dominant runoff source, and (C) BMPs identified in the systematic review."
+#| fig.cap: "Summary of (A) study scale and (B) dominant runoff source.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN)."
 #| #| fig.pos: "p"
@@
 
+```{r bmpsummaryparam}
+#| fig.cap: "Summary BMPs identified in the systematic review by tested parameter.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN)."
+#| #| fig.pos: "p"
+#| out.extra: ""
+#| out.width: "100%"
+
+knitr::include_graphics("../figures/bmp_summary_param.pdf")
+
+```
+
 ```{r bmptemporal}
-#| fig.cap: "Number of studies identified in the systematic review summarized by (A) publication date and (B) study length."
+#| fig.cap: "Number of studies identified in the systematic review summarized by (A) publication date and (B) study length.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN)."
 #| #| fig.pos: "p"
@@
 ```{r studymap}
-#| fig.cap: "Distribution of studies identified in the systematic review by state and parameter."
+#| fig.cap: "Distribution of studies identified in the systematic review by state and parameter.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN)."
 #| fig.pos: "p"
@@
 The point estimate with uncertaintity bars indicate the estimated overall effect, 95% confidence intervals, and 95% prediction intervals.
-Here, \\textit{k} indicates the number of overall effects with the number of unique studies in parenthesis."
+Here, \\textit{k} indicates the number of overall effects with the number of unique studies in parenthesis.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN)."
 #| fig.pos: "p"
@@
 ```{r fibeffect}
-#| fig.cap: "Predicted marignal effect of influent FIB and aridity index (conditioned on BMP subcategory).
+#| fig.cap: "Predicted marginal effect of influent fecal indicator bacteria (FIB; Panel A) and aridity index (conditioned on BMP subcategory; Panel B).
 Solid lines are the predicted mean effect, dashed lines are the 95% confidence intervals, and the dotted lines are the 95% prediction intervals.
-Individual dots represent each effect size identified in the literature with the size scaled by sampling variance."
+Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.
+Higher aridity index (mean annual precipition/evapotranspiration) indicates more humid conditions."
 #| fig.pos: "p"
@@
 ```{r phoseffect}
-#| fig.cap: "Predicted marignal effect of influent pollutant concentration on TP (Panel A) and PO\\textsubscript{4} (Panel B) reductions.
+#| fig.cap: "Predicted marignal effect of influent pollutant concentration on total phosphorus (TP; Panel A) and orthophosphate (PO\\textsubscript{4}; Panel B) reductions.
 Solid lines are the predicted mean effect, dashed lines are the 95% confidence intervals, and the dotted lines are the 95% prediction intervals.
@@
 ```{r aiindex}
-#| fig.cap: "Comparison of the relative distributions (density) of aridity index values across the U.S. against aridity index values for studies in the systematic review."
+#| fig.cap: "Comparison of the relative distributions (density) of aridity index values (mean annual precipition/evapotranspiration) across the U.S., across U.S.
+agricultural and developed land uses, and for BMP study locations.
+Higher aridity index values indicate more humid conditions."
 #| fig.pos: "p"

### 2024-05-23 frontiers_submission/frontiers_submission_revision_01.tex
working on responses

--- 2024-03-07 frontiers_submission/frontiers_submission.tex
@@
 Searches, review, and data extractions were conducted separately for FIB and nutrient related parameters.
-See Figure S1 and S2 for ROSES diagram. BMP data from the systematic review is available in \citet{kikoyoTWRINPSBMPDB2024}.
+See Figure S1 and S2 for RepOrting standards for Systematic Evidence Syntheses
+(ROSES) diagram.
+BMP data from the systematic review is available in \citet{kikoyoTWRINPSBMPDB2024}.
 
@@
 \addlinespace
-SD control & Standard deviation of control measreuments\\
+SD control & Standard deviation of control measurements\\
 SD experiment & Standard deviation of experimental measurements\\
@@
 
-We used the ``rma.mv'' function in the \emph{metafor} R package to fit multilevel random effects regression models with \emph{ROM} as the effect variable \citep{viechtbauerConductingMetaanalysesMetafor2010, rcoreteamLanguageEnvironmentStatistical2023}.
+We used the ``rma.mv'' function in the \emph{metafor} R package (R version 4.3.1) to fit multilevel random effects regression models with \emph{ROM} as the effect variable \citep{viechtbauerConductingMetaanalysesMetafor2010, rcoreteamLanguageEnvironmentStatistical2023}.
 We fit separate models for FIB, total nitrogen (TN), dissolved inorganic nitrogen (DIN), total phosphorus (TP), orthophosphate (PO\textsubscript{4}), and total suspended sediment (TSS).
@@
 
-where \emph{v} represents the sampling variance, \(CV_{control,i}\) and \(CV^2_{experiment,i}\) are the coefficients of variation from the \emph{i}th study for studies 1, 2, \ldots, \emph{K}.
+where \emph{v} represents the sampling variance, \(CV^2_{control,i}\) and \(CV^2_{experiment,i}\) are the squared coefficients of variation from the \emph{i}th study for studies 1, 2, \ldots, \emph{K}.
 \emph{n\textsubscript{control}} and \emph{n\textsubscript{experiment}} are the number of samples in the control (pre-treatment) trial or experimental (post-treatment) trial respectively.
@@
 We mapped study location coordinates to aridity index values published in the ``Global Aridity Index and Potential Evapotranspiration Database - Version 3'' (Global-AI\_PET\_v3) which provides gridded 30 arc-second annual average precipitation and potential evapotranspiration estimates \citep{zomerVersionGlobalAridity2022}.
-The aridity index is calculated as the ratio of mean annual precipitation to mean annual evapotraspiration with values between 0 and 0.5 considered hyper to semi-arid, and values above 0.65 as humid.
+The aridity index is calculated as the ratio of mean annual precipitation to mean annual potential (or reference) evapotranspiration with values between 0 and 0.5 considered hyper to semi-arid, and values above 0.65 as humid.
 
@@
 We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \ref{fig:bmpsummary} C).
-Our review was restricted to studies published after between 1999 and 2023.
+Our review was restricted to studies published from 2000 through 2022.
 The number of studies published for each parameter were roughly uniformly distributed over time (Figure \ref{fig:bmptemporal} A) and are not indicative of increases or decreases in the number of published studies.
-Study length was strongly skewed for all parameters (Figure \ref{fig:bmptemporal} B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO\textasciitilde4), 2.5 (TN), 2.5 (TP), and 2 (TSS) years.
+Study length was strongly skewed for all parameters (Figure \ref{fig:bmptemporal} B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO\textsubscript{4}), 2.5 (TN), 2.5 (TP), and 2 (TSS) years.
 There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \ref{fig:studymap}).
@@
 Changes in performance can also be described using a linear regression using date (transformed as a numeric variable) as an independent variable and log-transformed water quality as the dependent variable \citep{mitschCreatingWetlandsPrimary2012, mitschValidationEcosystemServices2014, pausAssessmentHydraulicToxic2014}.
-Slopes are a valid effect size for use in meta-analysis but the set of covariates used between studies should be the same since the coefficient of interest is adjusted to account for other terms in the the regression model \citep{beckerSynthesisRegressionSlopes2007}.
+Slopes are a valid effect size for use in meta-analysis but the set of covariates used between studies should be the same since the coefficient of interest is adjusted to account for other terms in the regression model \citep{beckerSynthesisRegressionSlopes2007}.
 It would be reasonable to assume that regressions equations vary between studies to adjust results for seasonality, flow rates, and other variables.
@@
 Raw data from the systematic review is deposited on Zenodo: \url{https://zenodo.org/doi/10.5281/zenodo.8302472}.
-R code used in the analysis is also deposited on Zenodo: TBD.
+R code used in the analysis is also deposited on Zenodo: \url{https://zenodo.org/records/10795231}.
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,]{../figures/bmp_summary} \caption{Summary of (A) study scale, (B) dominant runoff source, and (C) BMPs identified in the systematic review.}\label{fig:bmpsummary}
+\includegraphics[width=1\linewidth,]{../figures/bmp_summary} \caption{Summary of (A) study scale and (B) dominant runoff source.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:bmpsummary}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,]{../figures/bmp_temporal_summary} \caption{Number of studies identified in the systematic review summarized by (A) publication date and (B) study length.}\label{fig:bmptemporal}
+\includegraphics[width=1\linewidth,]{../figures/bmp_summary_param} \caption{Summary BMPs identified in the systematic review by tested parameter.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:bmpsummaryparam}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth,]{../figures/bmp_temporal_summary} \caption{Number of studies identified in the systematic review summarized by (A) publication date and (B) study length.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:bmptemporal}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=1\linewidth,]{../figures/study_map} \caption{Distribution of studies identified in the systematic review by state and parameter.}\label{fig:studymap}
+\includegraphics[width=1\linewidth,]{../figures/study_map} \caption{Distribution of studies identified in the systematic review by state and parameter.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:studymap}
 \end{figure}
@@
 The point estimate with uncertaintity bars indicate the estimated overall effect, 95\% confidence intervals, and 95\% prediction intervals.
-Here, \textit{k} indicates the number of overall effects with the number of unique studies in parenthesis.}\label{fig:overalleffect}
+Here, \textit{k} indicates the number of overall effects with the number of unique studies in parenthesis.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:overalleffect}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=1\linewidth,]{../figures/bac_predicted} \caption{Predicted marignal effect of influent FIB and aridity index (conditioned on BMP subcategory).
+\includegraphics[width=1\linewidth,]{../figures/bac_predicted} \caption{Predicted marginal effect of influent fecal indicator bacteria (FIB; Panel A) and aridity index (conditioned on BMP subcategory; Panel B).
 Solid lines are the predicted mean effect, dashed lines are the 95\% confidence intervals, and the dotted lines are the 95\% prediction intervals.
-Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.}\label{fig:fibeffect}
+Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.
+Higher aridity index (mean annual precipition/evapotranspiration) indicates more humid conditions.}\label{fig:fibeffect}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=1\linewidth,]{../figures/phos_predicted} \caption{Predicted marignal effect of influent pollutant concentration on TP (Panel A) and PO\textsubscript{4} (Panel B) reductions.
+\includegraphics[width=1\linewidth,]{../figures/phos_predicted} \caption{Predicted marignal effect of influent pollutant concentration on total phosphorus (TP; Panel A) and orthophosphate (PO\textsubscript{4}; Panel B) reductions.
 Solid lines are the predicted mean effect, dashed lines are the 95\% confidence intervals, and the dotted lines are the 95\% prediction intervals.
@@
 \begin{figure}[p]
-\includegraphics[width=0.5\linewidth,]{../figures/aridity_density} \caption{Comparison of the relative distributions (density) of aridity index values across the U.S.
-against aridity index values for studies in the systematic review.}\label{fig:aiindex}
+\includegraphics[width=0.5\linewidth,]{../figures/aridity_density} \caption{Comparison of the relative distributions (density) of aridity index values (mean annual precipition/evapotranspiration) across the U.S., across U.S.
+agricultural and developed land uses, and for BMP study locations.
+Higher aridity index values indicate more humid conditions.}\label{fig:aiindex}
 \end{figure}

### 2024-05-23 R/01_read_data.R

@@ -60,6 +60,19 @@ download_aridity <- function(url) {
 }
 
 
+# download 2021 NLCD data https://s3-us-west-2.amazonaws.com/mrlc/nlcd_2021_land_cover_l48_20230630.zip
+download_nlcd <- function(url) {
+  options(timeout = max(3600, getOption("timeout")))
+  dir <- "data/nlcd"
+  temp_path <- tempfile()
+  if (file.exists(temp_path))  'file alredy exists' else download.file(url = url, 
+                                                                       destfile = temp_path,
+                                                                       method = "libcurl",
+                                                                       mode = "wb") 
+  unzip(zipfile = temp_path, exdir = dir)
+}
+
+
 
 add_aridity <- function(df) {
   ai <- rast("data/Global-AI_ET0_v3_annual/ai_v3_yr.tif")
@@ -72,6 +85,7 @@ add_aridity <- function(df) {
   
   ai_centered <- scale(df$ai, scale = FALSE)
   df <- df |> 
+    mutate(ai_uncentered = ai) |> 
     mutate(ai = ai_centered)
   
   return(list(model_df = df,

### 2024-05-23 R/05_literature_statistics.R

@@ -110,7 +110,7 @@ plot_map <- function(bac_df,
                     stat = "sf_coordinates", bg.color = "white", bg.r = 0.15,
                     family = "Manrope SemiBold", fontface = "plain",
                     force = 0) +
-    scale_fill_distiller(palette = "Greens", direction = 1, na.value = "grey80",
+    scale_fill_distiller(palette = "Greens", direction = 1, na.value = "transparent",
                          trans = "log10") +
     facet_wrap(~Parameter, ncol = 2) +
     labs(x = "", y = "") +
@@ -128,6 +128,80 @@ plot_map <- function(bac_df,
                                     color = "black",
                                     face = "plain",
                                     size = rel(1), hjust = 0))
+  
+}
+
+plot_bmp_param_summary <- function(bac_df, nut_df) {
+
+
+  ## BMPs
+  bmp_bac <- bac_df |> 
+    group_by(StudyID, BMP_Class) |> 
+    summarize(n = 1) |> 
+    ungroup() |> 
+    group_by(BMP_Class) |> 
+    summarise(n = sum(n)) |> 
+    mutate(Parameter = "FIB")
+
+
+  bmp_df <- nut_df |>
+    mutate(Parameter = case_when(
+      Parameter %in% c("TN") ~ "TN",
+      Parameter %in% c("NO3", "NOX", "NH3") ~ "DIN",
+      Parameter %in% c("TP") ~ "TP",
+      Parameter %in% c("PO4") ~ "PO4",
+      Parameter == "TSS" ~ "TSS",
+      .default = NA
+    )) |>
+    filter(!is.na(Parameter)) |>
+    group_by(StudyID, Parameter, BMP_Class) |>
+    summarize(n = 1) |>
+    ungroup() |>
+    group_by(Parameter, BMP_Class) |>
+    summarise(n = sum(n)) |>
+    bind_rows(bmp_bac) |>
+    mutate(BMP_Class = stringr::str_to_sentence(BMP_Class)) |>
+    mutate(BMP_Class = case_when(
+      BMP_Class == "Swale" ~ "Swales",
+      BMP_Class == "Proprietary" ~ "Proprietary devices",
+      BMP_Class == "Drainage systems" ~ "Drainage water management",
+      .default = BMP_Class
+    )) |>
+    ungroup() |>
+    mutate(BMP_Class = reorder(BMP_Class,
+                               n,
+                               FUN = sum))
+  
[…]

### 2024-05-23 _targets.R

@@ -8,6 +8,7 @@ tar_option_set(packages = c("broom",
                             "ggforce",
                             "ggtext",
                             "ggrepel",
+                            "ggridges",
                             "grid",
                             "gridtext",
                             "httr",
@@ -69,6 +70,18 @@ list(
       ai_file_names
     },
     pattern = map(ai_file_names),
+    format = "file"
+  ),
+  
+  tar_target(nlcd, download_nlcd("https://s3-us-west-2.amazonaws.com/mrlc/nlcd_2021_land_cover_l48_20230630.zip"),
+             format = "file"),
+  tar_target(nlcd_file_names, nlcd), # not a format = "file" target
+  tar_target(
+    nlcd_files, {
+      nlcd # Re-hash all the files separately if any file changes.
+      nlcd_file_names
+    },
+    pattern = map(nlcd_file_names),
     format = "file"
   ),
   
@@ -271,10 +284,19 @@ list(
                                     plot = bmp_summary,
                                     device = cairo_pdf,
                                     width = 6,
-                                    height = 6.5,
+                                    height = 3,
                                     units = "in"),
              format = "file"),
   
+  tar_target(bmp_summary_param, plot_bmp_param_summary(bacteria_df, nutrient_df)),
+  tar_target(bmp_summary_param_file, ggsave("figures/bmp_summary_param.pdf",
+                                      plot = bmp_summary_param,
+                                      device = cairo_pdf,
+                                      width = 6,
+                                      height = 6,
+                                      units = "in"),
+             format = "file"),
+
   tar_target(bmp_temporal, plot_bmp_temporal(bacteria_df, nutrient_df)),
   tar_target(bmp_temporal_file, ggsave("figures/bmp_temporal_summary.pdf",
                                       plot = bmp_temporal$p1,
@@ -322,8 +344,11 @@ list(
                                              units = "in"),
              format = "file"),
   
+  tar_target(aridity_landuse, crop_aridity_to_land_use()),
+  
   tar_target(aridity_density, plot_ai_summary(bac_model_df,
-                                              nut_model_df)),
+                                              nut_model_df,
+                                              aridity_landuse)),
   
   tar_target(aridity_file, ggsave("figures/aridity_density.pdf",
                                   plot = aridity_density,

### 2024-05-23 _targets_packages.R

@@ -1,0 +1,41 @@
+# Generated by targets::tar_renv(): do not edit by hand
+library(MuMIn)
+library(RColorBrewer)
+library(broom)
+library(bs4Dash)
+library(clustermq)
+library(dplyr)
+library(forcats)
+library(future)
+library(ggforce)
+library(ggplot2)
+library(ggrepel)
+library(ggtext)
+library(grid)
+library(gridtext)
+library(gt)
+library(httr)
+library(kableExtra)
+library(latex2exp)
+library(markdown)
+library(metafor)
+library(metagear)
+library(modelr)
+library(mpsTemplates)
+library(orchaRd)
+library(patchwork)
+library(pingr)
+library(purrr)
+library(readxl)
+library(rstudioapi)
+library(scales)
+library(sf)
+library(shiny)
+library(shinyWidgets)
+library(shinybusy)
+library(terra)
+library(tibble)
+library(tidyr)
+library(tigris)
+library(units)
+library(visNetwork)

### 2024-05-23 _temp_nlcd.R

@@ -1,0 +1,220 @@
+library(targets)
+library(MuMIn)
+library(RColorBrewer)
+library(broom)
+library(dplyr)
+library(forcats)
+library(ggforce)
+library(ggplot2)
+library(ggrepel)
+library(ggtext)
+library(grid)
+library(gridtext)
+library(gt)
+library(httr)
+library(kableExtra)
+library(latex2exp)
+library(markdown)
+library(metafor)
+library(metagear)
+library(modelr)
+library(mpsTemplates)
+library(orchaRd)
+library(patchwork)
+library(purrr)
+library(readxl)
+library(rstudioapi)
+library(scales)
+library(sf)
+library(shiny)
+library(terra)
+library(tibble)
+library(tidyr)
+library(tigris)
+library(units)
+library(visNetwork)
+
+
+
+run_sensitvity <- function(model) {
+  ## run leave one out analysis
+  
+  mods <- model$sel_model |> 
+    formula()
+  
+  df <- model$data |> 
+    group_by(Study) |> 
+    mutate(study_num = row_number(Study)) |> 
+    ungroup() |> 
+    mutate(Study = paste0(Study, ".", study_num)) |> 
+    select(-study_num) |> 
+    mutate(Study = as_factor(Study)) |> 
+    mutate(Source_Type = as_factor(Source_Type))
+  
+  VCV <- vcalc(vi = df$vi,
+               cluster = df$StudyID,
+               obs = df$ValueID,
+               rho = 0.5,
+               nearpd = TRUE)
+
+
+  ## leave one out
+  loo <- list()
+  study_levels <- levels(df$Source_Type)
+  print(study_levels)
+
+  for(i in 1:length(study_levels)) {
+    loo_df <- df |>
+      filter(Source_Type != Source_Type[i]) #|>
+      #mutate(Source_Type = fct_drop(Source_Type))
+    # VCV_leave1out <- vcalc(vi = loo_df$vi,
+    #                        cluster = loo_df$StudyID,
+    #                        obs = loo_df$ValueID,
+    #                        rho = 0.5,
+    #                        nearpd = TRUE)
+    # loo[[i]] <- rma.mv(yi = yi,
+    #                    V = VCV_leave1out,
+    #                    random = ~1 | StudyID / ValueID,
+    #                    mods = mods,
+    #                    data = loo_df,
+    #                    method = "REML",
+    #                    test = "t",
+    #                    dfs = "contain",
[…]

### 2024-05-24 frontiers_submission/reviewer_response/reviewer_response.qmd
add variance size legends

@@
 
+We moved panel c to a new figure (now figure 2). The parameters are not plotted on 
+individual facets with the overall distribution in light grey in the background
+to reduce reliance on color for parameter identification. We also improved
+legibility of the y-axis.
+
 ---
@@
 :::
+
+Legends indicating sampling variance have been added to figures.
 

### 2024-05-24 R/03_plot_models.R

@@ -137,7 +137,10 @@ plot_marginal_effect_lnRR <- function(model,
                                            name = "Efficiency (%)",
                                            breaks = sec_y_breaks,
                                            labels = sec_y_labels)) +
-    scale_radius(trans = "sqrt") +
+    scale_radius(name = "Precision (1/SE):",
+                 trans = "sqrt") +
+    
+    guides(size = guide_legend(title.position = "top")) +
     labs(x = xlab,
          y = ylab) +
     theme_mps_noto(base_family = "Manrope Regular") +
@@ -148,7 +151,8 @@ plot_marginal_effect_lnRR <- function(model,
           axis.text = element_text(family = "Manrope Light", face = "plain", size = rel(0.8)),
           axis.text.y.left = element_text(hjust = 1),
           axis.text.y.right = element_text(hjust = 0),
-          legend.position = "none",
+          legend.text = element_text(size = rel(0.8)),
+          legend.title = element_text(family = "Manrope SemiBold", face = "plain", size = rel(0.8)),
           panel.grid.major.y = element_blank(),
           strip.background = element_rect(fill = "grey90",
                                           color = "grey90"),
@@ -166,9 +170,14 @@ plot_marginal_effect_lnRR <- function(model,
     if(by == "BMP_SubCat") {
       p1 <- p1 +
         scale_fill_manual(values = pal_vals) +
-        scale_color_manual(values = pal_vals)
-    }    
-  }
+        scale_color_manual(values = pal_vals) +
+        guides(color = "none", fill = "none") +
+        theme(legend.position = c(0.2,0.06))
+    }
+  } else {
+    p1 <- p1 +
+      theme(legend.position = c(0.75,0.15))
+  }   
   if(!is.null(ylim)) {
     p1 <- p1 +
       coord_cartesian(ylim = ylim)
@@ -581,13 +590,16 @@ mult_orch_plot <- function(bac, tn, tin, tp, po4, tss) {
                                            labels = sec_y_labels)) +
     scale_color_manual(values = "#66C2A5") +
     scale_fill_manual(values = "#66C2A5") +
+    guides(size = guide_legend(title.position = "top")) +
     theme_mps_noto(base_family = "Manrope Regular") +
     theme(axis.title = element_text(family = "Manrope SemiBold", face = "plain", size = rel(1)),
           axis.title.x = element_text(hjust = 0),
           axis.title.y = element_text(hjust = 1),
           axis.text = element_text(family = "Manrope Light", face = "plain", size = rel(0.7)),
[…]

### 2024-05-24 R/05_literature_statistics.R

@@ -191,7 +191,6 @@ plot_bmp_param_summary <- function(bac_df, nut_df) {
           axis.title.x.top = element_blank(),
           axis.title.y = element_text(hjust = 1),
           axis.text = element_text(family = "Manrope Light", face = "plain", size = rel(0.7)),
-          axis.text.x.bottom = element_blank(),
           axis.text.x.top = element_blank(),
           axis.text.y.left = element_text(hjust = 1),
           legend.position = "none",

### 2024-05-28 frontiers_submission/reviewer_response/reviewer_response.qmd
post-hoc sensitivty

@@
 with common removal mechanisms are certainly appropriate. 
-To address the Reviewer's concerns, we have added the following to the Discussion:
+To address the Reviewer's concerns, we have incorporated a Study Limitation section to the Discussion:
 
 ::: {.memo-excerpt}
-Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur?
-Quis autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur, vel illum qui dolorem eum fugiat quo voluptas nulla pariatur?
+A few reviews have noted a common trend of insufficient methodological and site specific data among peer-reviewed BMP performance studies (Grudzinski et al., 2020; Eagle et al., 2017; Liu et al., 2017).
+We confirm that inconsistent reporting among studies complicates data extraction, effect size calculations, and attributing important sources of variance.
+Our desire to evaluate the effects of specific BMP parameters was hinderd by the overall lack of reporting of relevant parameters such as drainage area, infiltration media/soil, infiltration volume, riparian/buffer width and area, and other relevant factors.
+The lack of BMP specific parameters certainly contributes to our relatively high model variance.
+The International Stormwater BMP Database addresses some of these concerns through a standardized reporting format.
+Our future efforts will incorporate data from the International Stormwater BMP Database with data retrieved through a systematic review.
+Additionally, this study was a broad scale look across BMP types which limits factors that cannot be compared across different types of BMPs.
+For example, livestock management BMPs may not rely on or report parameters such as vegetative buffer width or infiltration area.
+Inclusion of such factors in our meta-regression approach would necessarily exclude certain types of BMPs.
+Future BMP-specific meta-analysis might be more useful and informative for practitioners by providing effect sizes of parameters that were necessarily excluded from our study.
 :::
@@
 
+It is annually averaged from 1970-2000 which should suffice as representative for climate patterns in this study. We have added clarification in the text which now reads:
+
+::: {.memo-excerpt}
+We mapped study location coordinates to aridity index values published in the "Global Aridity Index and Potential Evapotranspiration Database - Version 3" (Global-AI_PET_v3) which provides gridded 30 arc-second annual average (annual averaged from 1970-2000) precipitation and potential evapotranspiration estimates.
+:::
+
 ---
@@
 ***Reviewer 3***:  Since the locations of the studies are not precisely defined in many cases, please provide more details on how to spatially link studies to the aridity index map to obtain aridity values.
+:::
+
+We added the following text to describe how spatial locations were derived if they were not provided in the study:
+
+::: {.memo-excerpt}
+Study locations were recrorded as latitude and longitude coordinates as described in studies.
+If coordinates were not provided, the study team used Google Maps to locate approximate study location using site descriptions (county, city, municipal buildings, etc.) from the study and recorded the coordinates.
 :::
@@
 
+Added.
+
 ---
@@
 :::
+
+Fixed.
 
@@
 
+Fixed.
+
 ---
@@
 ***Reviewer 3***:  In the sentence "Any study values outside of the full model confidence intervals would be considered an outlier.", by “study values”, do you mean the intercept estimate of a study?
+:::
+
+I worded this poorly and have updated the figure captions with the following:
+
+::: {.memo-excerpt}
+Any intercept estimates outside of the full model confidence intervals implies the excluded study is a potential outlier.
 :::
@@
 
+# Reviewer 4
+
+::: {.memo-reviewer}
+The statistical analysis was rigorous and thorough.
+That, alone, warrants a recommendation for acceptance of this manuscript.
+My primary concern is that the models were fitted using only one of the four BMP variables – BMP Subcategory.
+Were models with the other three BMP variables evaluated?
+In the mid-Atlantic US region, critically important BMPs for rural and agricultural lands are vegetative buffer strips, and for urban areas constructed wetlands are quite important.
+For example, much work has been done by researchers in North Carolina on metrics of vegetative buffer strips related to their effectiveness.
+Evaluation of models with specific practice characterization would have been informative.
+:::
+
+---
+
+::: {.memo-reviewer}
+That leads me to my second concern.
+Only the last paragraph of the manuscript is presented as a summary.
+I recommend that this paragraph be reconstituted as a separate “Conclusion” section and that it be more thorough than this simple summary paragraph.
+For example, it would be useful to know what the authors expected to find in their modeling studies versus what actually was found.
+:::
+
 # Reviewer 5
@@
 :::
+
+Added the spatial extent.
 
@@
 
+The table has been updated with the suggested langauge.
+
 ---

### 2024-05-28 frontiers_submission/frontiers_submission_revision_01.Rmd
post-hoc sensitivty

@@
 abstract: |
-  We conducted a systematic review of published nonpoint source water quality best management practice (BMP) studies and used a meta-analysis approach to describe variance in pollutant removal performance.
+  We conducted a systematic review of published nonpoint source water quality best management practice (BMP) studies conducted in the United States and used a meta-analysis approach to describe variance in pollutant removal performance.
 Despite high variance, we found the BMPs on average were effective at reducing fecal indicator bacteria (FIB), total nitrogen (TN), total phosphorus (TP), and total suspended sediment (TSS) concentrations.
@@
 The remaining studies were split among team members for data extraction (Table \@ref(tab:datvars)), again with at least two team members reviewing each study.
+Study locations were recrorded as latitude and longitude coordinates as described in studies.
+If coordinates were not provided, the study team used Google Maps to locate approximate study location using site descriptions (county, city, municipal buildings, etc.) from the study and recorded the coordinates.
 If data was provided in figures, the data was extracted with the WebPlotDigitizer tool [@rohatgiWebPlotDigitizer2022]. 
@@
   "Excluision Criteria" = c("Synopsis or review studies, reports with reductions based on modeled or other estimated reductions (e.g. TMDLs, watershed plans, or modelling studies.",
-                            "Studies not explicitly linking reductions to a specific BMP or insufficient information to quantify reductions.",
+                            "Studies not explicitly linking concentration reductions to a specific BMP or insufficient information to quantify reductions.",
                             "Studies outside of the United States.",
@@
 Aridity index was the only moderator not obtained directly in the systematic review (Table \@ref(tab:datvars)).
-We mapped study location coordinates to aridity index values published in the "Global Aridity Index and Potential Evapotranspiration Database - Version 3" (Global-AI_PET_v3) which provides gridded 30 arc-second annual average precipitation and potential evapotranspiration estimates [@zomerVersionGlobalAridity2022].
+We mapped study location coordinates to aridity index values published in the "Global Aridity Index and Potential Evapotranspiration Database - Version 3" (Global-AI_PET_v3) which provides gridded 30 arc-second annual average (annual averaged from 1970-2000) precipitation and potential evapotranspiration estimates [@zomerVersionGlobalAridity2022].
 The aridity index is calculated as the ratio of mean annual precipitation to mean annual potential (or reference) evapotranspiration with values between 0 and 0.5 considered hyper to semi-arid, and values above 0.65 as humid.
@@
 Not only were there spatially disparities, but we observed that the relative distribution of aridity index values does not resemble the distribution of aridity values across the U.S.
+or the distribution of aridity values for agricultural and developed land uses across the U.S.
 (Figure \@ref(fig:aiindex)).
@@
 
+## Study Limitations
+
 A few reviews have noted a common trend of insufficient methodological and site specific data among peer-reviewed BMP performance studies [@grudzinskiDoesRiparianFencing2020; @eagleMetaAnalysisConstrained2017; @liuReviewEffectivenessBest2017].
 We confirm that inconsistent reporting among studies complicates data extraction, effect size calculations, and attributing important sources of variance.
+Our desire to evaluate the effects of specific BMP parameters was hinderd by the overall lack of reporting of relevant parameters such as drainage area, infiltration media/soil, infiltration volume, riparian/buffer width and area, and other relevant factors.
+The lack of BMP specific parameters certainly contributes to our relatively high model variance.
 The International Stormwater BMP Database addresses some of these concerns through a standardized reporting format.
 Our future efforts will incorporate data from the International Stormwater BMP Database with data retrieved through a systematic review.
+Additionally, this study was a broad scale look across BMP types which limits factors that cannot be compared across different types of BMPs.
+For example, livestock management BMPs may not rely on or report parameters such as vegetative buffer width or infiltration area. 
+Inclusion of such factors in our meta-regression approach would necessarily exclude certain types of BMPs.
+Future BMP-specific meta-analysis might be more useful and informative for practitioners by providing effect sizes of parameters that were necessarily excluded from our study.
+
+# Conclusion
 

### 2024-05-28 frontiers_submission/frontiers_submission_revision_01.tex
post-hoc sensitivty

@@
 \begin{abstract}
-We conducted a systematic review of published nonpoint source water quality best management practice (BMP) studies and used a meta-analysis approach to describe variance in pollutant removal performance.
+We conducted a systematic review of published nonpoint source water quality best management practice (BMP) studies conducted in the United States and used a meta-analysis approach to describe variance in pollutant removal performance.
 Despite high variance, we found the BMPs on average were effective at reducing fecal indicator bacteria (FIB), total nitrogen (TN), total phosphorus (TP), and total suspended sediment (TSS) concentrations.
@@
 The remaining studies were split among team members for data extraction (Table \ref{tab:datvars}), again with at least two team members reviewing each study.
+Study locations were recrorded as latitude and longitude coordinates as described in studies.
+If coordinates were not provided, the study team used Google Maps to locate approximate study location using site descriptions (county, city, municipal buildings, etc.) from the study and recorded the coordinates.
 If data was provided in figures, the data was extracted with the WebPlotDigitizer tool \citep{rohatgiWebPlotDigitizer2022}.
@@
 Outcomes & Field studies with measured effects on fecal indicator bacteria or nutrient concentrations.
-& Studies not explicitly linking reductions to a specific BMP or insufficient information to quantify reductions.\\
+& Studies not explicitly linking concentration reductions to a specific BMP or insufficient information to quantify reductions.\\
 Geographical context & Studies conducted within the United States. & Studies outside of the United States.\\
@@
 Aridity index was the only moderator not obtained directly in the systematic review (Table \ref{tab:datvars}).
-We mapped study location coordinates to aridity index values published in the ``Global Aridity Index and Potential Evapotranspiration Database - Version 3'' (Global-AI\_PET\_v3) which provides gridded 30 arc-second annual average precipitation and potential evapotranspiration estimates \citep{zomerVersionGlobalAridity2022}.
+We mapped study location coordinates to aridity index values published in the ``Global Aridity Index and Potential Evapotranspiration Database - Version 3'' (Global-AI\_PET\_v3) which provides gridded 30 arc-second annual average (annual averaged from 1970-2000) precipitation and potential evapotranspiration estimates \citep{zomerVersionGlobalAridity2022}.
 The aridity index is calculated as the ratio of mean annual precipitation to mean annual potential (or reference) evapotranspiration with values between 0 and 0.5 considered hyper to semi-arid, and values above 0.65 as humid.
@@
 Not only were there spatially disparities, but we observed that the relative distribution of aridity index values does not resemble the distribution of aridity values across the U.S.
+or the distribution of aridity values for agricultural and developed land uses across the U.S.
 (Figure \ref{fig:aiindex}).
@@
 
+\hypertarget{study-limitations}{%
+\subsection{Study Limitations}\label{study-limitations}}
+
 A few reviews have noted a common trend of insufficient methodological and site specific data among peer-reviewed BMP performance studies \citep{grudzinskiDoesRiparianFencing2020, eagleMetaAnalysisConstrained2017, liuReviewEffectivenessBest2017}.
 We confirm that inconsistent reporting among studies complicates data extraction, effect size calculations, and attributing important sources of variance.
+Our desire to evaluate the effects of specific BMP parameters was hinderd by the overall lack of reporting of relevant parameters such as drainage area, infiltration media/soil, infiltration volume, riparian/buffer width and area, and other relevant factors.
+The lack of BMP specific parameters certainly contributes to our relatively high model variance.
 The International Stormwater BMP Database addresses some of these concerns through a standardized reporting format.
 Our future efforts will incorporate data from the International Stormwater BMP Database with data retrieved through a systematic review.
+Additionally, this study was a broad scale look across BMP types which limits factors that cannot be compared across different types of BMPs.
+For example, livestock management BMPs may not rely on or report parameters such as vegetative buffer width or infiltration area.
+Inclusion of such factors in our meta-regression approach would necessarily exclude certain types of BMPs.
+Future BMP-specific meta-analysis might be more useful and informative for practitioners by providing effect sizes of parameters that were necessarily excluded from our study.
+
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 

### 2024-05-28 supplementary_materials/supplementary_materials.Rmd
post-hoc sensitivty

@@
 The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
-Any study values outside of the full model confidence intervals would be considered and outlier."
+Any intercept estimates outside of the full model confidence intervals implies the excluded study is a potential outlier."
 #| fig.height: 8
@@
 The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
-Any study values outside of the full model confidence intervals would be considered and outlier."
+Any intercept estimates outside of the full model confidence intervals implies the excluded study is a potential outlier."
 #| fig.height: 5.5
@@
 The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
-Any study values outside of the full model confidence intervals would be considered and outlier."
+Any intercept estimates outside of the full model confidence intervals implies the excluded study is a potential outlier."
 #| fig.height: 7
@@
 The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
-Any study values outside of the full model confidence intervals would be considered and outlier."
+Any intercept estimates outside of the full model confidence intervals implies the excluded study is a potential outlier."
 #| fig.height: 4
@@
 The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
-Any study values outside of the full model confidence intervals would be considered and outlier."
+Any intercept estimates outside of the full model confidence intervals implies the excluded study is a potential outlier."
 #| fig.height: 4.5
@@
 ```
+
+```{r bacsource}
+#| fig.cap: "Post-hoc sensitivty analysis of fecal indicator bacteria models to exclusion of agriculture or urban runoff sources.
+The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
+Any intercept estimates outside of the full model confidence intervals implies the excluded sources are overly influential."
+#| fig.height: 3
+#| fig.width: 8 
+#| out.width: "100%"
+ tar_read(bac_source, store = store)$p1
+```
+
+```{r tnsource}
+#| fig.cap: "Post-hoc sensitivty analysis of total nitrogen models to exclusion of agriculture or urban runoff sources.
+The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
+Any intercept estimates outside of the full model confidence intervals implies the excluded sources are overly influential."
+#| fig.height: 3
+#| fig.width: 8
+#| out.width: "100%"
+ tar_read(tn_source, store = store)$p1
+```
+
+```{r dinsource}
+#| fig.cap: "Post-hoc sensitivty analysis of dissolved inorganic nitrogen models to exclusion of agriculture or urban runoff sources.
+The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
+Any intercept estimates outside of the full model confidence intervals implies the excluded sources are overly influential."
+#| fig.height: 3
+#| fig.width: 8
+#| out.width: "100%"
+ tar_read(tin_source, store = store)$p1
+```
+
+```{r tpsource}
+#| fig.cap: "Post-hoc sensitivty analysis of total phosphorus models to exclusion of agriculture or urban runoff sources.
+The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
+Any intercept estimates outside of the full model confidence intervals implies the excluded sources are overly influential."
+#| fig.height: 3
+#| fig.width: 8 
+#| out.width: "100%"
+ tar_read(tp_source, store = store)$p1
+```
+
+```{r po4source}
+#| fig.cap: "Post-hoc sensitivty analysis of orthophosphate models to exclusion of agriculture or urban runoff sources.
+The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
+Any intercept estimates outside of the full model confidence intervals implies the excluded sources are overly influential."
+#| fig.height: 3
+#| fig.width: 8
+#| out.width: "100%"
+ tar_read(po4_source, store = store)$p1
+```
+
+```{r tsssource}
+#| fig.cap: "Post-hoc sensitivty analysis of total suspended sediment models to exclusion of agriculture or urban runoff sources.
+The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
+Any intercept estimates outside of the full model confidence intervals implies the excluded sources are overly influential."
+#| fig.height: 3
+#| fig.width: 8
+#| out.width: "100%"
+ tar_read(tss_source, store = store)$p1
+```
+
+```{r bacscale}
+#| fig.cap: "Post-hoc sensitivty analysis of fecal indicator bacteria models to study scale.
+The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
+Any intercept estimates outside of the full model confidence intervals implies the excluded sources are overly influential."
+#| fig.height: 3
+#| fig.width: 8 
+#| out.width: "100%"
+ tar_read(bac_scale, store = store)$p1
+```
+
+```{r tnscale}
+#| fig.cap: "Post-hoc sensitivty analysis of total nitrogen models to study scale.
+The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
+Any intercept estimates outside of the full model confidence intervals implies the excluded sources are overly influential."
+#| fig.height: 3
+#| fig.width: 8 
+#| out.width: "100%"
+ tar_read(tn_scale, store = store)$p1
+```
+
+```{r dinscale}
+#| fig.cap: "Post-hoc sensitivty analysis of dissolved inorganic nitrogen models to study scale.
+The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
+Any intercept estimates outside of the full model confidence intervals implies the excluded sources are overly influential."
+#| fig.height: 3
+#| fig.width: 8 
+#| out.width: "100%"
+ tar_read(tin_scale, store = store)$p1
+```
+
+```{r tpscale}
+#| fig.cap: "Post-hoc sensitivty analysis of total phosphorus models to study scale.
+The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
+Any intercept estimates outside of the full model confidence intervals implies the excluded sources are overly influential."
+#| fig.height: 3
+#| fig.width: 8 
+#| out.width: "100%"
+ tar_read(tp_scale, store = store)$p1
+```
+
+```{r po4scale}
+#| fig.cap: "Post-hoc sensitivty analysis of orthophosphate models to study scale.
+The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
+Any intercept estimates outside of the full model confidence intervals implies the excluded sources are overly influential."
+#| fig.height: 3
+#| fig.width: 8 
+#| out.width: "100%"
+ tar_read(po4_scale, store = store)$p1
+```
+
+```{r tssscale}
+#| fig.cap: "Post-hoc sensitivty analysis of total suspended sediment models to study scale.
+The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values.
+Any intercept estimates outside of the full model confidence intervals implies the excluded sources are overly influential."
+#| fig.height: 3
+#| fig.width: 8 
+#| out.width: "100%"
+ tar_read(tss_scale, store = store)$p1
+```

### 2024-05-28 R/post-hoc.R

@@ -1,0 +1,211 @@
+source_sensitvity <- function(model) {
+  ## run sensitivity on source type
+  
+  mods <- model$sel_model |> 
+    formula()
+  
+  
+  df <- model$data |> 
+    group_by(Study) |> 
+    mutate(study_num = row_number(Study)) |> 
+    ungroup() |> 
+    mutate(Study = paste0(Study, ".", study_num)) |> 
+    select(-study_num) |> 
+    mutate(Study = as_factor(Study)) |> 
+    mutate(Source_Type = as_factor(Source_Type)) |> 
+    mutate(Scale = as_factor(Scale))
+  
+  
+  
+  VCV <- vcalc(vi = df$vi,
+               cluster = df$StudyID,
+               obs = df$ValueID,
+               rho = 0.5,
+               nearpd = TRUE)
+  
+  
+  ## leave one out
+  loo <- list()
+  study_levels <- levels(df$Source_Type)
+  
+  
+  for(i in 1:length(study_levels)) {
+    loo_df <- df |>
+      filter(Source_Type != study_levels[i]) |>
+      mutate(Source_Type = fct_drop(Source_Type))
+    VCV_leave1out <- vcalc(vi = loo_df$vi,
+                           cluster = loo_df$StudyID,
+                           obs = loo_df$ValueID,
+                           rho = 0.5,
+                           nearpd = TRUE)
+    loo[[i]] <- rma.mv(yi = yi,
+                       V = VCV_leave1out,
+                       random = ~1 | StudyID / ValueID,
+                       mods = mods,
+                       data = loo_df,
+                       method = "REML",
+                       test = "t",
+                       dfs = "contain",
+                       sparse = FALSE,
+                       verbose = FALSE,
+                       control=list(rel.tol=1e-8))
+    
+  }
+  
+  loo <- loo |>
+    purrr::map(summary) |>
+    purrr::map(coef) |>
+    purrr::map(\(x) slice(x, 1)) |>
+    list_rbind() |>
+    tibble::rownames_to_column() |>
+    as_tibble() |>
+    mutate(studies = study_levels)
+  
+  
+  
+  m2_VCV <-  rma.mv(yi = yi,
+                    V = VCV,
+                    random = ~1 | StudyID / ValueID,
+                    mods = mods,
+                    data = df,
+                    method = "REML",
+                    test = "t",
+                    dfs = "contain",
+                    sparse = FALSE,
[…]

### 2024-05-28 _targets.R

@@ -43,6 +43,7 @@ source("R/02_fit_models.R")
 source("R/03_plot_models.R")
 source("R/04_tables.R")
 source("R/05_literature_statistics.R")
+source("R/post-hoc.R")
 
 ## prep fancy figure font
 font_hoist("Manrope")
@@ -268,6 +269,26 @@ list(
                                     i2 = tss_i2,
                                     r2 = tss_r2,
                                     caption = "Summary table of multilevel random effect model for effect of BMPs on TSS removal.")),
+  
+  
+  # Some post-hoc analysis for reviewers ---------------------------------------
+  ## Sensitivity analysis to the runoff type (ag or urban)
+  tar_target(bac_source, source_sensitvity(bac_model)),
+  tar_target(tn_source, source_sensitvity(tn_model)),
+  tar_target(tin_source, source_sensitvity(tin_model)),
+  tar_target(tp_source, source_sensitvity(tp_model)),
+  tar_target(po4_source, source_sensitvity(po4_model)),
+  tar_target(tss_source, source_sensitvity(tss_model)),
+  
+  ## Sensitivity to scale
+  tar_target(bac_scale, scale_sensitvity(bac_model)),
+  tar_target(tn_scale, scale_sensitvity(tn_model)),
+  tar_target(tin_scale, scale_sensitvity(tin_model)),
+  tar_target(tp_scale, scale_sensitvity(tp_model)),
+  tar_target(po4_scale, scale_sensitvity(po4_model)),
+  tar_target(tss_scale, scale_sensitvity(tss_model)),
+  
+  
 
   # Figures for Manuscript -----------------------------------------------------
   tar_target(study_map, plot_map(bacteria_df, nutrient_df)),

### 2024-05-28 _temp_nlcd.R

@@ -1,220 +1,0 @@
-library(targets)
-library(MuMIn)
-library(RColorBrewer)
-library(broom)
-library(dplyr)
-library(forcats)
-library(ggforce)
-library(ggplot2)
-library(ggrepel)
-library(ggtext)
-library(grid)
-library(gridtext)
-library(gt)
-library(httr)
-library(kableExtra)
-library(latex2exp)
-library(markdown)
-library(metafor)
-library(metagear)
-library(modelr)
-library(mpsTemplates)
-library(orchaRd)
-library(patchwork)
-library(purrr)
-library(readxl)
-library(rstudioapi)
-library(scales)
-library(sf)
-library(shiny)
-library(terra)
-library(tibble)
-library(tidyr)
-library(tigris)
-library(units)
-library(visNetwork)
-
-
-
-run_sensitvity <- function(model) {
-  ## run leave one out analysis
-  
-  mods <- model$sel_model |> 
-    formula()
-  
-  df <- model$data |> 
-    group_by(Study) |> 
-    mutate(study_num = row_number(Study)) |> 
-    ungroup() |> 
-    mutate(Study = paste0(Study, ".", study_num)) |> 
-    select(-study_num) |> 
-    mutate(Study = as_factor(Study)) |> 
-    mutate(Source_Type = as_factor(Source_Type))
-  
-  VCV <- vcalc(vi = df$vi,
-               cluster = df$StudyID,
-               obs = df$ValueID,
-               rho = 0.5,
-               nearpd = TRUE)
-
-
-  ## leave one out
-  loo <- list()
-  study_levels <- levels(df$Source_Type)
-  print(study_levels)
-
-  for(i in 1:length(study_levels)) {
-    loo_df <- df |>
-      filter(Source_Type != Source_Type[i]) #|>
-      #mutate(Source_Type = fct_drop(Source_Type))
-    # VCV_leave1out <- vcalc(vi = loo_df$vi,
-    #                        cluster = loo_df$StudyID,
-    #                        obs = loo_df$ValueID,
-    #                        rho = 0.5,
-    #                        nearpd = TRUE)
-    # loo[[i]] <- rma.mv(yi = yi,
-    #                    V = VCV_leave1out,
-    #                    random = ~1 | StudyID / ValueID,
-    #                    mods = mods,
-    #                    data = loo_df,
-    #                    method = "REML",
-    #                    test = "t",
-    #                    dfs = "contain",
[…]

### 2024-05-30 frontiers_submission/reviewer_response/reviewer_response.qmd
prep for resubmission

@@
   hikmah-response-typst: default
+
+bibliography: biblio.bib
 ---
@@
 subgroup analysis could be conducted based on additional moderators, we risk 
-conducting a "data-dredging" study.
-
-::: {.memo-excerpt}
-Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.
-Neque porro quisquam est, qui dolorem ipsum quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat voluptatem.
-:::
+conducting a "data-dredging" study [@thompsonHowShouldMeta2002]. However, we did add post-hoc sensitivity analysis for runoff source and catchment scale to the paper based on this suggestion. 
+We did not find strong evidence indicating these factors are highly
+influential to the results we found. This is likely, as you correctly point out,
+do to the lack of BMP specific parameters that better explain performance than our chosen moderators.
 
@@
 :::
+
+We added an additional post-hoc sensitivity analysis for both these factors to the study.
+While this is not the same as fully incorporating the moderators into the original model, the sensitivity analysis indicates that we don't have strong evidence that these terms are highly influential to our results.
+This is confounded in part by the fact some of our BMP types are only applied to urban or agricultural BMPs and highlights a need for future meta-analysis to be applied for specific BMP types instead of across multiple BMP types.
 
@@
 
+We did not include soil properties with the assumption that many infiltration based BMP modify soil as part of BMP construction.
+It is certainly possible that native soil conditions provide some additional moderating effect, but we did not obtain this data based on our assumptions.
+
 ---
@@
 :::
+
+We have expanded on this slightly in the conclusion.
+Essentially it is difficult to empirically link water quality improvements to BMP implementation in many studies due to short study lengths, underlying measurement and analytic uncertainty, and difficulty establishing adequate controls.
+Many watershed scale studies therefore rely on numeric watershed models to demonstrate BMP performance (see the USDA CEAP program for example).
 
@@
 
+We agree with the overall comment here.
+This paper was our first cut analysis across a broad range of BMPs.
+Our approach requires somewhat broad variables used in the regression models so we do not eliminate studies that do not include specific BMP data, such as buffer strip width.
+Further meta-analysis should focus on specific practices where relevant specific BMP parameters can be properly evaluated.
+We've added additional context in the Study Limitations section.
+
 ---
@@
 :::
+
+We have incorporated a conclusions section and expanded a bit more on our expectations as well as incorporated suggestions from other reviewers.
 
@@
 ***Reviewer 5***:   1st Paragraph of introduction (Line 28), add a sentence about sediment impairments to waters similar to the ones discussing FIBs and nutrients impairments
+:::
+
+Added the following:
+
+::: {.memo-excerpt}
+Approximately 15% of river and streams have excessive sedimentation, which leads to twice the likelihood of a stream to have poor biological condition (EPA, 2017).
 :::
@@
 
+This escaped our notice, thanks for raising this relevant shortcoming. We added the following to the Study Limitations section within the discussion:
+
+::: {.memo-excerpt}
+One reviewer also noted some examples of missing studies, attributable to our selection of search terms and inclusion criteria (Table 1).
+In particular, our search query did not include specifc practice names, instead using “BMP” or “best management practice” as a keyword search.
+This choice may have led us to miss studies that only included the practice name.
+Or inclusion criteria of concentration based studies may have also lead to the exclusion of studies that focused on load reductions (through flow reductions) but may have included data on relevant concentration effects.
+:::
+
 ---
@@
 ***Reviewer 5***:   Lines 61 – 64 consider adding an example BMP for instances where effluent is and is not related to influent concentration
+:::
+
+We modified the lines to include the following:
+
+::: {.memo-excerpt}
+Barrett (2005) demonstrated
+that percent pollutant reduction is often a function of influent quality. Specifically, for certain types of BMPs
+percent removal is low at low influent concentrations, and increases with increasing influent concentrations.
+Horvath et al. (2023) found that influent phosphorus concentrations had some explanatory ability for
+BMP performance among grass strips, bioretnetion, and grass swale BMPs. However, for some types
+of BMPs, such as media filters and permanently pooled retention basins (Barrett, 2005) and pollutant
+parameters, effluent concentration is unrelated to influent concentration. 
 :::
@@
 
+It would not be reliable when extrapolated to lower aridity regions due to lack of model coverage. Clarified this in text.
+
 ---
@@
 :::
+
+Added.
 
@@
 
+We have added a third distribution, based on your suggestion below, looking at just urban and agricultural land and still find some discrepancies in the distribution of aridity values.
+
 ---
@@
 
+We added the following text to the conclusions:
+
+::: {.memo-excerpt}
+To improve future data synthesis efforts, we highly reccomend future BMP studies
+follow the reporting guidelines provided in Eagle et al. (2017). In particular provide clearly defined water
+quality parameters, tabular data (either in the manuscript, as supplemental material, or in an open data
+repository), and error estimation at minimum. Control and treatment means should be made available in
+reports, not just the transformed efficiency or percent change values. Furthermore clearly defined controls
+and treatments are fundamental for comparing across studies. Finally, a major shortcoming in our synthesis
+was driven by the lack of reported covariate data. Although data such as soil type may not be an within
+study experimental covariate, documenting of these types of study variables is useful for data synthesis
+efforts.
+:::
 ---

### 2024-05-30 frontiers_submission/frontiers_submission.Rmd
prep for resubmission

@@
                       fig.width = 6,
-                      dpi = 300)
+                      dpi = 300,
+                      fig.process = 
+                        function(filename) {
+                          new_filename <- stringr::str_remove(string = filename,
+                                                              pattern = "-1")
+                          fs::file_move(path = filename, new_path = new_filename)
+                          ifelse(fs::file_exists(new_filename), new_filename, filename)
+                        })
 
@@
 Our systematic review identified a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effect sizes for PO~4~, and 33 studies with 125 effect sizes for TSS.
-The majority of studies were identified as smaller scaled lot or field studies (Figure \@ref(fig:bmpsummary) A).
+The majority of studies were identified as smaller scaled lot or field studies (Figure \@ref(fig:fig1) A).
 FIB studies had a roughly equal proportion of large watershed/catchment studies and studies conducted at the community/farm scale or smaller.
-We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \@ref(fig:bmpsummary) B).
-We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \@ref(fig:bmpsummary) C).
+We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \@ref(fig:fig1) B).
+We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \@ref(fig:fig1) C).
 Our review was restricted to studies published after between 1999 and 2023.
-The number of studies published for each parameter were roughly uniformly distributed over time (Figure \@ref(fig:bmptemporal) A) and are not indicative of increases or decreases in the number of published studies.
-Study length was strongly skewed for all parameters (Figure \@ref(fig:bmptemporal) B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO~4), 2.5 (TN), 2.5 (TP), and 2 (TSS) years. 
-There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \@ref(fig:studymap)).
+The number of studies published for each parameter were roughly uniformly distributed over time (Figure \@ref(fig:fig2) A) and are not indicative of increases or decreases in the number of published studies.
+Study length was strongly skewed for all parameters (Figure \@ref(fig:fig2) B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO~4), 2.5 (TN), 2.5 (TP), and 2 (TSS) years. 
+There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \@ref(fig:fig3)).
 
@@
 There were only 19 studies and 63 FIB effect sizes available to model after removal of studies and effects that only reported BMP~eff~.
-The overall mean effect (estimated with the intercept only multilevel random effects model) showed significant mean reductions in FIB (ROM = 0.85, 95% CI [0.36, 1.34]; BMP~eff~ = 57.4%, 95% CI [30.4%, 73.9%]; Figure \@ref(fig:overalleffect)) resulting from BMPs.
+The overall mean effect (estimated with the intercept only multilevel random effects model) showed significant mean reductions in FIB (ROM = 0.85, 95% CI [0.36, 1.34]; BMP~eff~ = 57.4%, 95% CI [30.4%, 73.9%]; Figure \@ref(fig:fig4)) resulting from BMPs.
 Total heterogeneity was moderate with a relatively large amount of heterogeneity observed due to differences within studies (`r fib_i2_null`).
@@
 Moderator terms and interactions explained a high proportion of effect size variance (`r fib_r2`) in the FIB model.
-Increased influent concentrations ($\beta$ = 0.25, 95% CI [0.14, 0.37]) resulted in significantly larger ROM effect for FIB (Figure \@ref(fig:fibeffect)).
+Increased influent concentrations ($\beta$ = 0.25, 95% CI [0.14, 0.37]) resulted in significantly larger ROM effect for FIB (Figure \@ref(fig:fig5)).
 Compared to the baseline aridity index\texttimes detention BMP subcategory interaction, infiltration ($\beta$ = -29.90, 95% CI [-50.34, -9.47]), livestock management ($\beta$ = -30.37, 95% CI [-50.93, -9.81]), and treatment ($\beta$ = -30.33, 95% CI [-49.62, -11.03]) interactions had significantly smaller slopes.
@@
 We identified 13 eligible TN studies and 14 DIN studies and 31 and 44 effect sizes respectively that could be included in the regression model.
-Overall effects showed that BMPs resulted in significant mean reductions in TN (ROM = 0.42, 95% CI [0.21, 0.62]; BMP~eff~ = 34.0%, 95% CI [18.7%, 46.4%]; Figure \@ref(fig:overalleffect)) but not in DIN (ROM = 0.64, 95% CI [-0.08, 1.35]; BMP~eff~ = 47.1%, 95% CI [-8.1%, 74.1%]; Figure \@ref(fig:overalleffect)).
+Overall effects showed that BMPs resulted in significant mean reductions in TN (ROM = 0.42, 95% CI [0.21, 0.62]; BMP~eff~ = 34.0%, 95% CI [18.7%, 46.4%]; Figure \@ref(fig:fig4)) but not in DIN (ROM = 0.64, 95% CI [-0.08, 1.35]; BMP~eff~ = 47.1%, 95% CI [-8.1%, 74.1%]; Figure \@ref(fig:fig4)).
 Heterogeneity was high for TN with a large proportion of heterogeneity attributed to within study effect (`r tn_i2_null`).
@@
 Moderators explained a relatively small amount of variance for both models (TP: `r tp_r2`, PO~4~: `r po4_r2`).
-Influent concentration ($\beta$ = 0.23, 95% CI [-0.035, 0.49]) was not significant at the 95% confidence level for the TP model (Figure \@ref(fig:phoseffect); Table S4).
-Influent concentration ($\beta$ = 0.27, 95% CI [0.085, 0.44]) was significant for the PO~4~ model (Figure \@ref(fig:phoseffect); Table S5).
+Influent concentration ($\beta$ = 0.23, 95% CI [-0.035, 0.49]) was not significant at the 95% confidence level for the TP model (Figure \@ref(fig:fig6); Table S4).
+Influent concentration ($\beta$ = 0.27, 95% CI [0.085, 0.44]) was significant for the PO~4~ model (Figure \@ref(fig:fig6); Table S5).
 
@@
 
-Our systematic review revealed strong spatial disparities in published BMP studies (Figure \@ref(fig:studymap)).
+Our systematic review revealed strong spatial disparities in published BMP studies (Figure \@ref(fig:fig3)).
 Similar spatial disparities have been identified and discussed in @kochNitrogenRemovalStormwater2014, @grudzinskiDoesRiparianFencing2020 and can be problematic for extrapolating results to other regions of interest.
@@
 @horvathEffectsRegionalClimate2023 found overlapping BMP type and climate groups within their dataset that reduce the ability to distinguish effects due to either BMP type or climate.
-Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \@ref(fig:fibeffect)).
+Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \@ref(fig:fig5)).
 Not only were there spatially disparities, but we observed that the relative distribution of aridity index values does not resemble the distribution of aridity values across the U.S.
-(Figure \@ref(fig:aiindex)).
+(Figure \@ref(fig:fig7)).
 Our review indicates that BMP studies are over represented in the generally humid regions of the country and underrepresented the more arid regions.
@@
 
-```{r bmpsummary}
+```{r fig1}
 #| fig.cap: "Summary of (A) study scale, (B) dominant runoff source, and (C) BMPs identified in the systematic review."
@@
 
-knitr::include_graphics("../figures/bmp_summary.pdf")
+knitr::include_graphics("frontiers_submission_files/figure-latex/fig1.pdf")
 
@@
 
-```{r bmptemporal}
+```{r fig2}
 #| fig.cap: "Number of studies identified in the systematic review summarized by (A) publication date and (B) study length."
@@
 
-knitr::include_graphics("../figures/bmp_temporal_summary.pdf")
+knitr::include_graphics("frontiers_submission_files/figure-latex/fig2.pdf")
 
@@
 
-```{r studymap}
+```{r fig3}
 #| fig.cap: "Distribution of studies identified in the systematic review by state and parameter."
@@
 
-knitr::include_graphics("../figures/study_map.pdf")
+knitr::include_graphics("frontiers_submission_files/figure-latex/fig3.pdf")
 
@@
 
-```{r overalleffect}
+```{r fig4}
 #| fig.cap: "Estimated effect sizes and intervals from the intercept only multilevel random effects model.
@@
 
-knitr::include_graphics("../figures/overall_effect.pdf")
+knitr::include_graphics("frontiers_submission_files/figure-latex/fig4.pdf")
 ```
 
-```{r fibeffect}
+```{r fig5}
 #| fig.cap: "Predicted marignal effect of influent FIB and aridity index (conditioned on BMP subcategory).
@@
 
-knitr::include_graphics("../figures/bac_predicted.pdf")
+knitr::include_graphics("frontiers_submission_files/figure-latex/fig5.pdf")
 ```
 
-```{r phoseffect}
+```{r fig6}
 #| fig.cap: "Predicted marignal effect of influent pollutant concentration on TP (Panel A) and PO\\textsubscript{4} (Panel B) reductions.
@@
 
-knitr::include_graphics("../figures/phos_predicted.pdf")
+knitr::include_graphics("frontiers_submission_files/figure-latex/fig6.pdf")
 ```
 
-```{r aiindex}
+```{r fig7}
 #| fig.cap: "Comparison of the relative distributions (density) of aridity index values across the U.S. against aridity index values for studies in the systematic review."
@@
 
-knitr::include_graphics("../figures/aridity_density.pdf")
+knitr::include_graphics("frontiers_submission_files/figure-latex/fig7.pdf")
 ```

### 2024-05-30 frontiers_submission/frontiers_submission.tex
prep for resubmission

@@
 Our systematic review identified a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effect sizes for PO\textsubscript{4}, and 33 studies with 125 effect sizes for TSS.
-The majority of studies were identified as smaller scaled lot or field studies (Figure \ref{fig:bmpsummary} A).
+The majority of studies were identified as smaller scaled lot or field studies (Figure \ref{fig:fig1} A).
 FIB studies had a roughly equal proportion of large watershed/catchment studies and studies conducted at the community/farm scale or smaller.
-We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \ref{fig:bmpsummary} B).
-We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \ref{fig:bmpsummary} C).
+We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \ref{fig:fig1} B).
+We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \ref{fig:fig1} C).
 Our review was restricted to studies published after between 1999 and 2023.
-The number of studies published for each parameter were roughly uniformly distributed over time (Figure \ref{fig:bmptemporal} A) and are not indicative of increases or decreases in the number of published studies.
-Study length was strongly skewed for all parameters (Figure \ref{fig:bmptemporal} B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO\textasciitilde4), 2.5 (TN), 2.5 (TP), and 2 (TSS) years.
-There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \ref{fig:studymap}).
+The number of studies published for each parameter were roughly uniformly distributed over time (Figure \ref{fig:fig2} A) and are not indicative of increases or decreases in the number of published studies.
+Study length was strongly skewed for all parameters (Figure \ref{fig:fig2} B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO\textasciitilde4), 2.5 (TN), 2.5 (TP), and 2 (TSS) years.
+There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \ref{fig:fig3}).
 
@@
 There were only 19 studies and 63 FIB effect sizes available to model after removal of studies and effects that only reported BMP\textsubscript{eff}.
-The overall mean effect (estimated with the intercept only multilevel random effects model) showed significant mean reductions in FIB (ROM = 0.85, 95\% CI {[}0.36, 1.34{]}; BMP\textsubscript{eff} = 57.4\%, 95\% CI {[}30.4\%, 73.9\%{]}; Figure \ref{fig:overalleffect}) resulting from BMPs.
+The overall mean effect (estimated with the intercept only multilevel random effects model) showed significant mean reductions in FIB (ROM = 0.85, 95\% CI {[}0.36, 1.34{]}; BMP\textsubscript{eff} = 57.4\%, 95\% CI {[}30.4\%, 73.9\%{]}; Figure \ref{fig:fig4}) resulting from BMPs.
 Total heterogeneity was moderate with a relatively large amount of heterogeneity observed due to differences within studies (\textit{I\textsuperscript{2}\textsubscript{total}} = 53.54, \textit{I\textsuperscript{2}\textsubscript{study}} = 10.03, \textit{I\textsuperscript{2}\textsubscript{effect}} = 43.51).
@@
 Moderator terms and interactions explained a high proportion of effect size variance (\textit{R\textsuperscript{2}\textsubscript{marginal}} =0.89) in the FIB model.
-Increased influent concentrations (\(\beta\) = 0.25, 95\% CI {[}0.14, 0.37{]}) resulted in significantly larger ROM effect for FIB (Figure \ref{fig:fibeffect}).
+Increased influent concentrations (\(\beta\) = 0.25, 95\% CI {[}0.14, 0.37{]}) resulted in significantly larger ROM effect for FIB (Figure \ref{fig:fig5}).
 Compared to the baseline aridity index\texttimes detention BMP subcategory interaction, infiltration (\(\beta\) = -29.90, 95\% CI {[}-50.34, -9.47{]}), livestock management (\(\beta\) = -30.37, 95\% CI {[}-50.93, -9.81{]}), and treatment (\(\beta\) = -30.33, 95\% CI {[}-49.62, -11.03{]}) interactions had significantly smaller slopes.
@@
 We identified 13 eligible TN studies and 14 DIN studies and 31 and 44 effect sizes respectively that could be included in the regression model.
-Overall effects showed that BMPs resulted in significant mean reductions in TN (ROM = 0.42, 95\% CI {[}0.21, 0.62{]}; BMP\textsubscript{eff} = 34.0\%, 95\% CI {[}18.7\%, 46.4\%{]}; Figure \ref{fig:overalleffect}) but not in DIN (ROM = 0.64, 95\% CI {[}-0.08, 1.35{]}; BMP\textsubscript{eff} = 47.1\%, 95\% CI {[}-8.1\%, 74.1\%{]}; Figure \ref{fig:overalleffect}).
+Overall effects showed that BMPs resulted in significant mean reductions in TN (ROM = 0.42, 95\% CI {[}0.21, 0.62{]}; BMP\textsubscript{eff} = 34.0\%, 95\% CI {[}18.7\%, 46.4\%{]}; Figure \ref{fig:fig4}) but not in DIN (ROM = 0.64, 95\% CI {[}-0.08, 1.35{]}; BMP\textsubscript{eff} = 47.1\%, 95\% CI {[}-8.1\%, 74.1\%{]}; Figure \ref{fig:fig4}).
 Heterogeneity was high for TN with a large proportion of heterogeneity attributed to within study effect (\textit{I\textsuperscript{2}\textsubscript{total}} = 77.12, \textit{I\textsuperscript{2}\textsubscript{study}} = 23.2, \textit{I\textsuperscript{2}\textsubscript{effect}} = 53.92).
@@
 Moderators explained a relatively small amount of variance for both models (TP: \textit{R\textsuperscript{2}\textsubscript{marginal}} =0.12, PO\textsubscript{4}: \textit{R\textsuperscript{2}\textsubscript{marginal}} =0.35).
-Influent concentration (\(\beta\) = 0.23, 95\% CI {[}-0.035, 0.49{]}) was not significant at the 95\% confidence level for the TP model (Figure \ref{fig:phoseffect}; Table S4).
-Influent concentration (\(\beta\) = 0.27, 95\% CI {[}0.085, 0.44{]}) was significant for the PO\textsubscript{4} model (Figure \ref{fig:phoseffect}; Table S5).
+Influent concentration (\(\beta\) = 0.23, 95\% CI {[}-0.035, 0.49{]}) was not significant at the 95\% confidence level for the TP model (Figure \ref{fig:fig6}; Table S4).
+Influent concentration (\(\beta\) = 0.27, 95\% CI {[}0.085, 0.44{]}) was significant for the PO\textsubscript{4} model (Figure \ref{fig:fig6}; Table S5).
 
@@
 
-Our systematic review revealed strong spatial disparities in published BMP studies (Figure \ref{fig:studymap}).
+Our systematic review revealed strong spatial disparities in published BMP studies (Figure \ref{fig:fig3}).
 Similar spatial disparities have been identified and discussed in \citet{kochNitrogenRemovalStormwater2014}, \citet{grudzinskiDoesRiparianFencing2020} and can be problematic for extrapolating results to other regions of interest.
@@
 \citet{horvathEffectsRegionalClimate2023} found overlapping BMP type and climate groups within their dataset that reduce the ability to distinguish effects due to either BMP type or climate.
-Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \ref{fig:fibeffect}).
+Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \ref{fig:fig5}).
 Not only were there spatially disparities, but we observed that the relative distribution of aridity index values does not resemble the distribution of aridity values across the U.S.
-(Figure \ref{fig:aiindex}).
+(Figure \ref{fig:fig7}).
 Our review indicates that BMP studies are over represented in the generally humid regions of the country and underrepresented the more arid regions.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,]{../figures/bmp_summary} \caption{Summary of (A) study scale, (B) dominant runoff source, and (C) BMPs identified in the systematic review.}\label{fig:bmpsummary}
+\includegraphics[width=1\linewidth,]{frontiers_submission_files/figure-latex/fig1} \caption{Summary of (A) study scale, (B) dominant runoff source, and (C) BMPs identified in the systematic review.}\label{fig:fig1}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,]{../figures/bmp_temporal_summary} \caption{Number of studies identified in the systematic review summarized by (A) publication date and (B) study length.}\label{fig:bmptemporal}
+\includegraphics[width=1\linewidth,]{frontiers_submission_files/figure-latex/fig2} \caption{Number of studies identified in the systematic review summarized by (A) publication date and (B) study length.}\label{fig:fig2}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=1\linewidth,]{../figures/study_map} \caption{Distribution of studies identified in the systematic review by state and parameter.}\label{fig:studymap}
+\includegraphics[width=1\linewidth,]{frontiers_submission_files/figure-latex/fig3} \caption{Distribution of studies identified in the systematic review by state and parameter.}\label{fig:fig3}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=1\linewidth,]{../figures/overall_effect} \caption{Estimated effect sizes and intervals from the intercept only multilevel random effects model.
+\includegraphics[width=1\linewidth,]{frontiers_submission_files/figure-latex/fig4} \caption{Estimated effect sizes and intervals from the intercept only multilevel random effects model.
 Individual points represent studies, with size scaled by sampling variance.
 The point estimate with uncertaintity bars indicate the estimated overall effect, 95\% confidence intervals, and 95\% prediction intervals.
-Here, \textit{k} indicates the number of overall effects with the number of unique studies in parenthesis.}\label{fig:overalleffect}
+Here, \textit{k} indicates the number of overall effects with the number of unique studies in parenthesis.}\label{fig:fig4}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=1\linewidth,]{../figures/bac_predicted} \caption{Predicted marignal effect of influent FIB and aridity index (conditioned on BMP subcategory).
+\includegraphics[width=1\linewidth,]{frontiers_submission_files/figure-latex/fig5} \caption{Predicted marignal effect of influent FIB and aridity index (conditioned on BMP subcategory).
 Solid lines are the predicted mean effect, dashed lines are the 95\% confidence intervals, and the dotted lines are the 95\% prediction intervals.
-Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.}\label{fig:fibeffect}
+Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.}\label{fig:fig5}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=1\linewidth,]{../figures/phos_predicted} \caption{Predicted marignal effect of influent pollutant concentration on TP (Panel A) and PO\textsubscript{4} (Panel B) reductions.
+\includegraphics[width=1\linewidth,]{frontiers_submission_files/figure-latex/fig6} \caption{Predicted marignal effect of influent pollutant concentration on TP (Panel A) and PO\textsubscript{4} (Panel B) reductions.
 Solid lines are the predicted mean effect, dashed lines are the 95\% confidence intervals, and the dotted lines are the 95\% prediction intervals.
-Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.}\label{fig:phoseffect}
+Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.}\label{fig:fig6}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=0.5\linewidth,]{../figures/aridity_density} \caption{Comparison of the relative distributions (density) of aridity index values across the U.S.
-against aridity index values for studies in the systematic review.}\label{fig:aiindex}
+\includegraphics[width=0.5\linewidth,]{frontiers_submission_files/figure-latex/fig7} \caption{Comparison of the relative distributions (density) of aridity index values across the U.S.
+against aridity index values for studies in the systematic review.}\label{fig:fig7}
 \end{figure}

### 2024-05-30 frontiers_submission/frontiers_submission_revision_01.Rmd
prep for resubmission

@@
   - nutrients
+  - suspended sediment
 output: 
@@
 FIB remains the leading cause of water body impairment on the Clean Water Act 303(d) list in the United States [@epaNationalWaterQuality2017].
+Approximately 15% of river and streams have excessive sedimentation, which leads to twice the likelihood of a stream to have poor biological condition [@epaNationalWaterQuality2017].
 
@@
 Specifically, for certain types of BMPs percent removal is low at low influent concentrations, and increases with increasing influent concentrations. 
-However, for some types of BMPs and pollutant parameters, effluent concentration is unrelated to influent concentration.
+@horvathEffectsRegionalClimate2023 found that influent phosphorus concentrations had some explanatory ability for BMP performance among grass strips, bioretnetion, and grass swale BMPs.
+However, for some types of BMPs, such as media filters and permanently pooled retention basins [@barrettPerformanceComparisonStructural2005], effluent concentration is unrelated to influent concentration.
 Second, local climatic conditions can be expected to influence BMP performance.
@@
 The remaining studies were split among team members for data extraction (Table \@ref(tab:datvars)), again with at least two team members reviewing each study.
-Study locations were recrorded as latitude and longitude coordinates as described in studies.
+Study locations were recorded as latitude and longitude coordinates as described in studies.
 If coordinates were not provided, the study team used Google Maps to locate approximate study location using site descriptions (county, city, municipal buildings, etc.) from the study and recorded the coordinates.
@@
 Aridity index was the only moderator not obtained directly in the systematic review (Table \@ref(tab:datvars)).
-We mapped study location coordinates to aridity index values published in the "Global Aridity Index and Potential Evapotranspiration Database - Version 3" (Global-AI_PET_v3) which provides gridded 30 arc-second annual average (annual averaged from 1970-2000) precipitation and potential evapotranspiration estimates [@zomerVersionGlobalAridity2022].
+We mapped study location coordinates to aridity index values published in the "Global Aridity Index and Potential Evapotranspiration Database - Version 3" (Global-AI_PET_v3) which provides gridded 30 arc-second annual average (annually averaged from 1970-2000) precipitation and potential evapotranspiration estimates [@zomerVersionGlobalAridity2022].
 The aridity index is calculated as the ratio of mean annual precipitation to mean annual potential (or reference) evapotranspiration with values between 0 and 0.5 considered hyper to semi-arid, and values above 0.65 as humid.
@@
 We did not identify evidence of outliers or overly influential studies for any of our models (Figure S9-S14).
+We also applied a sensitivity analysis to runoff source (agriculture and urban) and study catchment scale (community, lot, and watershed) and did not find evidence that these factors were especially influential to our results (Figure S21-S32).
 
@@
 Our systematic review identified a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effect sizes for PO~4~, and 33 studies with 125 effect sizes for TSS.
-The majority of studies were identified as smaller scaled lot or field studies (Figure \@ref(fig:bmpsummary) A).
+The majority of studies were identified as smaller scaled lot or field studies (Figure \@ref(fig:fig1) A).
 FIB studies had a roughly equal proportion of large watershed/catchment studies and studies conducted at the community/farm scale or smaller.
-We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \@ref(fig:bmpsummary) B).
-We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \@ref(fig:bmpsummary) C).
+We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \@ref(fig:fig1) B).
+We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \@ref(fig:fig2)).
 Our review was restricted to studies published from 2000 through 2022.
-The number of studies published for each parameter were roughly uniformly distributed over time (Figure \@ref(fig:bmptemporal) A) and are not indicative of increases or decreases in the number of published studies.
-Study length was strongly skewed for all parameters (Figure \@ref(fig:bmptemporal) B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO~4~), 2.5 (TN), 2.5 (TP), and 2 (TSS) years. 
-There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \@ref(fig:studymap)).
+The number of studies published for each parameter were roughly uniformly distributed over time (Figure \@ref(fig:fig3) A) and are not indicative of increases or decreases in the number of published studies.
+Study length was strongly skewed for all parameters (Figure \@ref(fig:fig3) B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO~4~), 2.5 (TN), 2.5 (TP), and 2 (TSS) years. 
+There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \@ref(fig:fig4)).
 
@@
 There were only 19 studies and 63 FIB effect sizes available to model after removal of studies and effects that only reported BMP~eff~.
-The overall mean effect (estimated with the intercept only multilevel random effects model) showed significant mean reductions in FIB (ROM = 0.85, 95% CI [0.36, 1.34]; BMP~eff~ = 57.4%, 95% CI [30.4%, 73.9%]; Figure \@ref(fig:overalleffect)) resulting from BMPs.
+The overall mean effect (estimated with the intercept only multilevel random effects model) showed significant mean reductions in FIB (ROM = 0.85, 95% CI [0.36, 1.34]; BMP~eff~ = 57.4%, 95% CI [30.4%, 73.9%]; Figure \@ref(fig:fig5)) resulting from BMPs.
 Total heterogeneity was moderate with a relatively large amount of heterogeneity observed due to differences within studies (`r fib_i2_null`).
@@
 Moderator terms and interactions explained a high proportion of effect size variance (`r fib_r2`) in the FIB model.
-Increased influent concentrations ($\beta$ = 0.25, 95% CI [0.14, 0.37]) resulted in significantly larger ROM effect for FIB (Figure \@ref(fig:fibeffect)).
+Increased influent concentrations ($\beta$ = 0.25, 95% CI [0.14, 0.37]) resulted in significantly larger ROM effect for FIB (Figure \@ref(fig:fig6)).
 Compared to the baseline aridity index\texttimes detention BMP subcategory interaction, infiltration ($\beta$ = -29.90, 95% CI [-50.34, -9.47]), livestock management ($\beta$ = -30.37, 95% CI [-50.93, -9.81]), and treatment ($\beta$ = -30.33, 95% CI [-49.62, -11.03]) interactions had significantly smaller slopes.
 However, the data had uneven coverage of BMP subcategories across the aridity index. 
-Effects for detention BMPs were clustered in humid climates (aridity index > 0.65) and the resulting estimate for the baseline interaction ($\beta$ = 32.63, 95% CI [12.57, 52.69]) may not be reliable.
+Effects for detention BMPs were clustered in humid climates (aridity index > 0.65) and the resulting estimate for the baseline interaction ($\beta$ = 32.63, 95% CI [12.57, 52.69]) may not be reliable when extrapolated to low-humidity regions.
 
@@
 We identified 13 eligible TN studies and 14 DIN studies and 31 and 44 effect sizes respectively that could be included in the regression model.
-Overall effects showed that BMPs resulted in significant mean reductions in TN (ROM = 0.42, 95% CI [0.21, 0.62]; BMP~eff~ = 34.0%, 95% CI [18.7%, 46.4%]; Figure \@ref(fig:overalleffect)) but not in DIN (ROM = 0.64, 95% CI [-0.08, 1.35]; BMP~eff~ = 47.1%, 95% CI [-8.1%, 74.1%]; Figure \@ref(fig:overalleffect)).
+Overall effects showed that BMPs resulted in significant mean reductions in TN (ROM = 0.42, 95% CI [0.21, 0.62]; BMP~eff~ = 34.0%, 95% CI [18.7%, 46.4%]; Figure \@ref(fig:fig5)) but not in DIN (ROM = 0.64, 95% CI [-0.08, 1.35]; BMP~eff~ = 47.1%, 95% CI [-8.1%, 74.1%]; Figure \@ref(fig:fig5)).
 Heterogeneity was high for TN with a large proportion of heterogeneity attributed to within study effect (`r tn_i2_null`).
@@
 Moderators explained a relatively small amount of variance for both models (TP: `r tp_r2`, PO~4~: `r po4_r2`).
-Influent concentration ($\beta$ = 0.23, 95% CI [-0.035, 0.49]) was not significant at the 95% confidence level for the TP model (Figure \@ref(fig:phoseffect); Table S4).
-Influent concentration ($\beta$ = 0.27, 95% CI [0.085, 0.44]) was significant for the PO~4~ model (Figure \@ref(fig:phoseffect); Table S5).
+Influent concentration ($\beta$ = 0.23, 95% CI [-0.035, 0.49]) was not significant at the 95% confidence level for the TP model (Figure \@ref(fig:fig7); Table S4).
+Influent concentration ($\beta$ = 0.27, 95% CI [0.085, 0.44]) was significant for the PO~4~ model (Figure \@ref(fig:fig7); Table S5).
 
@@
 
-Our systematic review revealed strong spatial disparities in published BMP studies (Figure \@ref(fig:studymap)).
+Our systematic review revealed strong spatial disparities in published BMP studies (Figure \@ref(fig:fig4)).
 Similar spatial disparities have been identified and discussed in @kochNitrogenRemovalStormwater2014, @grudzinskiDoesRiparianFencing2020 and can be problematic for extrapolating results to other regions of interest.
@@
 @horvathEffectsRegionalClimate2023 found overlapping BMP type and climate groups within their dataset that reduce the ability to distinguish effects due to either BMP type or climate.
-Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \@ref(fig:fibeffect)).
+Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \@ref(fig:fig6)).
 Not only were there spatially disparities, but we observed that the relative distribution of aridity index values does not resemble the distribution of aridity values across the U.S.
 or the distribution of aridity values for agricultural and developed land uses across the U.S.
-(Figure \@ref(fig:aiindex)).
+(Figure \@ref(fig:fig8)).
 Our review indicates that BMP studies are over represented in the generally humid regions of the country and underrepresented the more arid regions.
@@
 We confirm that inconsistent reporting among studies complicates data extraction, effect size calculations, and attributing important sources of variance.
-Our desire to evaluate the effects of specific BMP parameters was hinderd by the overall lack of reporting of relevant parameters such as drainage area, infiltration media/soil, infiltration volume, riparian/buffer width and area, and other relevant factors.
+Our desire to evaluate the effects of specific BMP parameters was hindered by the overall lack of reporting of relevant parameters such as drainage area, infiltration media/soil, infiltration volume, riparian/buffer width and area, and other relevant factors.
 The lack of BMP specific parameters certainly contributes to our relatively high model variance.
@@
 
+One reviewer also noted some examples of missing studies, attributable to our selection of search terms and inclusion criteria (Table \@ref(tab:criteria)).
+In particular, our search query did not include specifc practice names, instead using "BMP" or "best management practice" as a keyword search. 
+This choice may have led us to miss studies that only included the practice name. 
+Our inclusion criteria of concentration based studies may have also lead to the exclusion of studies that focused on load reductions (through flow reductions) but may have included data on relevant concentration effects.
+
 # Conclusion
 
-Scaling BMP pollutant reductions to basin wide water quality improvement remains a substantial challenge [@tomerChallengeDocumentingWater2011; @mellandEffectsAgriculturalLand2018; @mealsLagTimeWater2010].
+Scaling BMP pollutant reductions to basin wide water quality improvement remains a substantial challenge.
+While there are numerous studies of field scale practices, linking BMPs with large scale watershed improvements is hindered by lack of adequate controls, scaling of measurement and analytic uncertainty, and substantial lags in downstream water quality improvement [@tomerChallengeDocumentingWater2011; @mellandEffectsAgriculturalLand2018; @mealsLagTimeWater2010].
 The major identified challenges include the lack of long term studies, inadequate data collected on BMP management, and incomplete understanding of BMP function [@linternBestManagementPractices2020; @liuReviewEffectivenessBest2017].
+As a result there is strong reliance on numeric watershed models to assess performance of BMPs at watershed scales, but @liuReviewEffectivenessBest2017 notes there is discrepancy between rates of water quality improvements found in modelling studies and empirical studies.
 There is a clear need to fill knowledge gaps through additional long-term spatially relevant BMP studies.
@@
 
+Improved reporting and data availability provides opportunity to improve decision-making through applied numeric modeling approaches that better represent BMP systems, or to advance new modelling methods for decision-making with statistical based approaches such as machine-learning.
+To improve future data synthesis efforts, we highly recommend future BMP studies follow the reporting guidelines provided in @eagleMetaAnalysisConstrained2017.
+In particular, authors should provide clearly defined water quality parameters, tabular data (either in the manuscript, as supplemental material, or in an open data repository), and error estimation at minimum.
+Control and treatment means should be made available in reports, not just the transformed efficiency or percent change values.
+Furthermore clearly defined controls and treatments are fundamental for comparing across studies.
+Finally, a major shortcoming in our synthesis was driven by the lack of reported covariate data.
+Although data such as soil type may not be an within study experimental covariate, documenting of these types of study variables is useful for data synthesis efforts.
+
 In summary, we used multi-level random effects meta-regression models to estimate overall BMP effectiveness from systematically reviewed studies.
 Although there was relatively high variance between studies, we found strong evidence that BMPs reduce overall mean FIB, TN, TP, and TSS concentrations.
+These results are generally consistent with results from prior reviews that used different approaches to synthesize results.
 Influent concentrations moderated BMP efficiency for both FIB and PO~4~, with larger removal rates at high influent concentrations.
 We found that aridity and BMP subcategory moderated BMP performance for only FIB.
+We anticipated stronger interaction effects between inflow concentrations and BMP subcategory due to differences in influent based performance demonstrated in prior studies.
 Most likely, site specific design and climate variables not captured in our review or by our choice in BMP classification approach play a more important role in explaining BMP performance variability.
@@
 
-```{r bmpsummary}
+```{r fig1}
 #| fig.cap: "Summary of (A) study scale and (B) dominant runoff source.
@@
 
-knitr::include_graphics("../figures/bmp_summary.pdf")
+knitr::include_graphics("frontiers_submission_revision_01_files/figure-latex/fig1.pdf")
 
@@
 
-```{r bmpsummaryparam}
-#| fig.cap: "Summary BMPs identified in the systematic review by tested parameter.
+```{r fig2}
+#| fig.cap: "Summary of BMPs identified in the systematic review by tested parameter.
 Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN)."
@@
 
-knitr::include_graphics("../figures/bmp_summary_param.pdf")
+knitr::include_graphics("frontiers_submission_revision_01_files/figure-latex/fig2.pdf")
 
@@
 
-```{r bmptemporal}
+```{r fig3}
 #| fig.cap: "Number of studies identified in the systematic review summarized by (A) publication date and (B) study length.
@@
 
-knitr::include_graphics("../figures/bmp_temporal_summary.pdf")
+knitr::include_graphics("frontiers_submission_revision_01_files/figure-latex/fig3.pdf")
 
@@
 
-```{r studymap}
+```{r fig4}
 #| fig.cap: "Distribution of studies identified in the systematic review by state and parameter.
@@
 
-knitr::include_graphics("../figures/study_map.pdf")
+knitr::include_graphics("frontiers_submission_revision_01_files/figure-latex/fig4.pdf")
 
@@
 
-```{r overalleffect}
+```{r fig5}
 #| fig.cap: "Estimated effect sizes and intervals from the intercept only multilevel random effects model.
@@
 
-knitr::include_graphics("../figures/overall_effect.pdf")
+knitr::include_graphics("frontiers_submission_revision_01_files/figure-latex/fig5.pdf")
 ```
 
-```{r fibeffect}
+```{r fig6}
 #| fig.cap: "Predicted marginal effect of influent fecal indicator bacteria (FIB; Panel A) and aridity index (conditioned on BMP subcategory; Panel B).
@@
 
-knitr::include_graphics("../figures/bac_predicted.pdf")
+knitr::include_graphics("frontiers_submission_revision_01_files/figure-latex/fig6.pdf")
 ```
 
-```{r phoseffect}
+```{r fig7}
 #| fig.cap: "Predicted marignal effect of influent pollutant concentration on total phosphorus (TP; Panel A) and orthophosphate (PO\\textsubscript{4}; Panel B) reductions.
@@
 
-knitr::include_graphics("../figures/phos_predicted.pdf")
+knitr::include_graphics("frontiers_submission_revision_01_files/figure-latex/fig7.pdf")
 ```
 
-```{r aiindex}
+```{r fig8}
 #| fig.cap: "Comparison of the relative distributions (density) of aridity index values (mean annual precipition/evapotranspiration) across the U.S., across U.S.
@@
 
-knitr::include_graphics("../figures/aridity_density.pdf")
+knitr::include_graphics("frontiers_submission_revision_01_files/figure-latex/fig8.pdf")
 ```

### 2024-05-30 frontiers_submission/frontiers_submission_revision_01.tex
prep for resubmission

@@
 \tiny
- \keyFont{ \section{Keywords:} best management practice, water quality, nonpoint source pollution, fecal indicator bacteria, nutrients} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
+ \keyFont{ \section{Keywords:} best management practice, water quality, nonpoint source pollution, fecal indicator bacteria, nutrients, suspended sediment} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
 \end{abstract}
@@
 FIB remains the leading cause of water body impairment on the Clean Water Act 303(d) list in the United States \citep{epaNationalWaterQuality2017}.
+Approximately 15\% of river and streams have excessive sedimentation, which leads to twice the likelihood of a stream to have poor biological condition \citep{epaNationalWaterQuality2017}.
 
@@
 Specifically, for certain types of BMPs percent removal is low at low influent concentrations, and increases with increasing influent concentrations.
-However, for some types of BMPs and pollutant parameters, effluent concentration is unrelated to influent concentration.
+\citet{horvathEffectsRegionalClimate2023} found that influent phosphorus concentrations had some explanatory ability for BMP performance among grass strips, bioretnetion, and grass swale BMPs.
+However, for some types of BMPs, such as media filters and permanently pooled retention basins \citep{barrettPerformanceComparisonStructural2005}, effluent concentration is unrelated to influent concentration.
 Second, local climatic conditions can be expected to influence BMP performance.
@@
 The remaining studies were split among team members for data extraction (Table \ref{tab:datvars}), again with at least two team members reviewing each study.
-Study locations were recrorded as latitude and longitude coordinates as described in studies.
+Study locations were recorded as latitude and longitude coordinates as described in studies.
 If coordinates were not provided, the study team used Google Maps to locate approximate study location using site descriptions (county, city, municipal buildings, etc.) from the study and recorded the coordinates.
@@
 Aridity index was the only moderator not obtained directly in the systematic review (Table \ref{tab:datvars}).
-We mapped study location coordinates to aridity index values published in the ``Global Aridity Index and Potential Evapotranspiration Database - Version 3'' (Global-AI\_PET\_v3) which provides gridded 30 arc-second annual average (annual averaged from 1970-2000) precipitation and potential evapotranspiration estimates \citep{zomerVersionGlobalAridity2022}.
+We mapped study location coordinates to aridity index values published in the ``Global Aridity Index and Potential Evapotranspiration Database - Version 3'' (Global-AI\_PET\_v3) which provides gridded 30 arc-second annual average (annually averaged from 1970-2000) precipitation and potential evapotranspiration estimates \citep{zomerVersionGlobalAridity2022}.
 The aridity index is calculated as the ratio of mean annual precipitation to mean annual potential (or reference) evapotranspiration with values between 0 and 0.5 considered hyper to semi-arid, and values above 0.65 as humid.
@@
 Our systematic review identified a total of 33 studies and 125 effect sizes on FIB, 24 studies and 50 effect sizes on TN, 31 studies and 88 effect sizes for DIN, 31 studies and 61 effect sizes for TP, 17 studies and 36 effect sizes for PO\textsubscript{4}, and 33 studies with 125 effect sizes for TSS.
-The majority of studies were identified as smaller scaled lot or field studies (Figure \ref{fig:bmpsummary} A).
+The majority of studies were identified as smaller scaled lot or field studies (Figure \ref{fig:fig1} A).
 FIB studies had a roughly equal proportion of large watershed/catchment studies and studies conducted at the community/farm scale or smaller.
-We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \ref{fig:bmpsummary} B).
-We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \ref{fig:bmpsummary} C).
+We also identified that the majority of studies (all parameters) were conducted on urban or non-agricultural runoff (Figure \ref{fig:fig1} B).
+We did identify a wide variety of BMPs in the review, but it did not appear that any particular type of BMP was responsible for the majority of studies for any given parameter (Figure \ref{fig:fig2}).
 Our review was restricted to studies published from 2000 through 2022.
-The number of studies published for each parameter were roughly uniformly distributed over time (Figure \ref{fig:bmptemporal} A) and are not indicative of increases or decreases in the number of published studies.
-Study length was strongly skewed for all parameters (Figure \ref{fig:bmptemporal} B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO\textsubscript{4}), 2.5 (TN), 2.5 (TP), and 2 (TSS) years.
-There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \ref{fig:studymap}).
+The number of studies published for each parameter were roughly uniformly distributed over time (Figure \ref{fig:fig3} A) and are not indicative of increases or decreases in the number of published studies.
+Study length was strongly skewed for all parameters (Figure \ref{fig:fig3} B). Median study lengths were 3 (DIN), 2 (FIB), 2.5 (PO\textsubscript{4}), 2.5 (TN), 2.5 (TP), and 2 (TSS) years.
+There appears to be a strong clustering of BMP studies in the mid-Atlantic region (North Carolina, Virginia, Maryland) with other states sparsely represented or completely absent from the review (Figure \ref{fig:fig4}).
 
@@
 There were only 19 studies and 63 FIB effect sizes available to model after removal of studies and effects that only reported BMP\textsubscript{eff}.
-The overall mean effect (estimated with the intercept only multilevel random effects model) showed significant mean reductions in FIB (ROM = 0.85, 95\% CI {[}0.36, 1.34{]}; BMP\textsubscript{eff} = 57.4\%, 95\% CI {[}30.4\%, 73.9\%{]}; Figure \ref{fig:overalleffect}) resulting from BMPs.
+The overall mean effect (estimated with the intercept only multilevel random effects model) showed significant mean reductions in FIB (ROM = 0.85, 95\% CI {[}0.36, 1.34{]}; BMP\textsubscript{eff} = 57.4\%, 95\% CI {[}30.4\%, 73.9\%{]}; Figure \ref{fig:fig5}) resulting from BMPs.
 Total heterogeneity was moderate with a relatively large amount of heterogeneity observed due to differences within studies (\textit{I\textsuperscript{2}\textsubscript{total}} = 53.54, \textit{I\textsuperscript{2}\textsubscript{study}} = 10.03, \textit{I\textsuperscript{2}\textsubscript{effect}} = 43.51).
@@
 Moderator terms and interactions explained a high proportion of effect size variance (\textit{R\textsuperscript{2}\textsubscript{marginal}} =0.89) in the FIB model.
-Increased influent concentrations (\(\beta\) = 0.25, 95\% CI {[}0.14, 0.37{]}) resulted in significantly larger ROM effect for FIB (Figure \ref{fig:fibeffect}).
+Increased influent concentrations (\(\beta\) = 0.25, 95\% CI {[}0.14, 0.37{]}) resulted in significantly larger ROM effect for FIB (Figure \ref{fig:fig6}).
 Compared to the baseline aridity index\texttimes detention BMP subcategory interaction, infiltration (\(\beta\) = -29.90, 95\% CI {[}-50.34, -9.47{]}), livestock management (\(\beta\) = -30.37, 95\% CI {[}-50.93, -9.81{]}), and treatment (\(\beta\) = -30.33, 95\% CI {[}-49.62, -11.03{]}) interactions had significantly smaller slopes.
 However, the data had uneven coverage of BMP subcategories across the aridity index.
-Effects for detention BMPs were clustered in humid climates (aridity index \textgreater{} 0.65) and the resulting estimate for the baseline interaction (\(\beta\) = 32.63, 95\% CI {[}12.57, 52.69{]}) may not be reliable.
+Effects for detention BMPs were clustered in humid climates (aridity index \textgreater{} 0.65) and the resulting estimate for the baseline interaction (\(\beta\) = 32.63, 95\% CI {[}12.57, 52.69{]}) may not be reliable when extrapolated to low-humidity regions.
 
@@
 We identified 13 eligible TN studies and 14 DIN studies and 31 and 44 effect sizes respectively that could be included in the regression model.
-Overall effects showed that BMPs resulted in significant mean reductions in TN (ROM = 0.42, 95\% CI {[}0.21, 0.62{]}; BMP\textsubscript{eff} = 34.0\%, 95\% CI {[}18.7\%, 46.4\%{]}; Figure \ref{fig:overalleffect}) but not in DIN (ROM = 0.64, 95\% CI {[}-0.08, 1.35{]}; BMP\textsubscript{eff} = 47.1\%, 95\% CI {[}-8.1\%, 74.1\%{]}; Figure \ref{fig:overalleffect}).
+Overall effects showed that BMPs resulted in significant mean reductions in TN (ROM = 0.42, 95\% CI {[}0.21, 0.62{]}; BMP\textsubscript{eff} = 34.0\%, 95\% CI {[}18.7\%, 46.4\%{]}; Figure \ref{fig:fig5}) but not in DIN (ROM = 0.64, 95\% CI {[}-0.08, 1.35{]}; BMP\textsubscript{eff} = 47.1\%, 95\% CI {[}-8.1\%, 74.1\%{]}; Figure \ref{fig:fig5}).
 Heterogeneity was high for TN with a large proportion of heterogeneity attributed to within study effect (\textit{I\textsuperscript{2}\textsubscript{total}} = 77.12, \textit{I\textsuperscript{2}\textsubscript{study}} = 23.2, \textit{I\textsuperscript{2}\textsubscript{effect}} = 53.92).
@@
 Moderators explained a relatively small amount of variance for both models (TP: \textit{R\textsuperscript{2}\textsubscript{marginal}} =0.12, PO\textsubscript{4}: \textit{R\textsuperscript{2}\textsubscript{marginal}} =0.35).
-Influent concentration (\(\beta\) = 0.23, 95\% CI {[}-0.035, 0.49{]}) was not significant at the 95\% confidence level for the TP model (Figure \ref{fig:phoseffect}; Table S4).
-Influent concentration (\(\beta\) = 0.27, 95\% CI {[}0.085, 0.44{]}) was significant for the PO\textsubscript{4} model (Figure \ref{fig:phoseffect}; Table S5).
+Influent concentration (\(\beta\) = 0.23, 95\% CI {[}-0.035, 0.49{]}) was not significant at the 95\% confidence level for the TP model (Figure \ref{fig:fig7}; Table S4).
+Influent concentration (\(\beta\) = 0.27, 95\% CI {[}0.085, 0.44{]}) was significant for the PO\textsubscript{4} model (Figure \ref{fig:fig7}; Table S5).
 
@@
 
-Our systematic review revealed strong spatial disparities in published BMP studies (Figure \ref{fig:studymap}).
+Our systematic review revealed strong spatial disparities in published BMP studies (Figure \ref{fig:fig4}).
 Similar spatial disparities have been identified and discussed in \citet{kochNitrogenRemovalStormwater2014}, \citet{grudzinskiDoesRiparianFencing2020} and can be problematic for extrapolating results to other regions of interest.
@@
 \citet{horvathEffectsRegionalClimate2023} found overlapping BMP type and climate groups within their dataset that reduce the ability to distinguish effects due to either BMP type or climate.
-Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \ref{fig:fibeffect}).
+Similarly, we found detention type BMPs clustered only in humid climates (high aridity index) which reduces our confidence in extrapolating the interaction between BMP types and aridity index for FIB (Figure \ref{fig:fig6}).
 Not only were there spatially disparities, but we observed that the relative distribution of aridity index values does not resemble the distribution of aridity values across the U.S.
 or the distribution of aridity values for agricultural and developed land uses across the U.S.
-(Figure \ref{fig:aiindex}).
+(Figure \ref{fig:fig8}).
 Our review indicates that BMP studies are over represented in the generally humid regions of the country and underrepresented the more arid regions.
@@
 
+One reviewer also noted some examples of missing studies, attributable to our selection of search terms and inclusion criteria (Table \ref{tab:criteria}).
+In particular, our search query did not include specifc practice names, instead using ``BMP'' or ``best management practice'' as a keyword search.
+This choice may have led us to miss studies that only included the practice name.
+Our inclusion criteria of concentration based studies may have also lead to the exclusion of studies that focused on load reductions (through flow reductions) but may have included data on relevant concentration effects.
+
 \hypertarget{conclusion}{%
@@
 
-Scaling BMP pollutant reductions to basin wide water quality improvement remains a substantial challenge \citep{tomerChallengeDocumentingWater2011, mellandEffectsAgriculturalLand2018, mealsLagTimeWater2010}.
+Scaling BMP pollutant reductions to basin wide water quality improvement remains a substantial challenge.
+While there are numerous studies of field scale practices, linking BMPs with large scale watershed improvements is hindered by lack of adequate controls, scaling of measurement and analytic uncertainty, and substantial lags in downstream water quality improvement \citep{tomerChallengeDocumentingWater2011, mellandEffectsAgriculturalLand2018, mealsLagTimeWater2010}.
 The major identified challenges include the lack of long term studies, inadequate data collected on BMP management, and incomplete understanding of BMP function \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017}.
+As a result there is strong reliance on numeric watershed models to assess performance of BMPs at watershed scales, but \citet{liuReviewEffectivenessBest2017} notes there is discrepancy between rates of water quality improvements found in modelling studies and empirical studies.
 There is a clear need to fill knowledge gaps through additional long-term spatially relevant BMP studies.
@@
 
+Improved reporting and data availability provides opportunity to improve decision-making through applied numeric modeling approaches that better represent BMP systems, or to advance new modelling methods for decision-making with statistical based approaches such as machine-learning.
+To improve future data synthesis efforts, we highly recommend future BMP studies follow the reporting guidelines provided in \citet{eagleMetaAnalysisConstrained2017}.
+In particular, authors should provide clearly defined water quality parameters, tabular data (either in the manuscript, as supplemental material, or in an open data repository), and error estimation at minimum.
+Control and treatment means should be made available in reports, not just the transformed efficiency or percent change values.
+Furthermore clearly defined controls and treatments are fundamental for comparing across studies.
+Finally, a major shortcoming in our synthesis was driven by the lack of reported covariate data.
+Although data such as soil type may not be an within study experimental covariate, documenting of these types of study variables is useful for data synthesis efforts.
+
 In summary, we used multi-level random effects meta-regression models to estimate overall BMP effectiveness from systematically reviewed studies.
 Although there was relatively high variance between studies, we found strong evidence that BMPs reduce overall mean FIB, TN, TP, and TSS concentrations.
+These results are generally consistent with results from prior reviews that used different approaches to synthesize results.
 Influent concentrations moderated BMP efficiency for both FIB and PO\textsubscript{4}, with larger removal rates at high influent concentrations.
 We found that aridity and BMP subcategory moderated BMP performance for only FIB.
+We anticipated stronger interaction effects between inflow concentrations and BMP subcategory due to differences in influent based performance demonstrated in prior studies.
 Most likely, site specific design and climate variables not captured in our review or by our choice in BMP classification approach play a more important role in explaining BMP performance variability.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,]{../figures/bmp_summary} \caption{Summary of (A) study scale and (B) dominant runoff source.
-Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:bmpsummary}
+\includegraphics[width=1\linewidth,]{frontiers_submission_revision_01_files/figure-latex/fig1} \caption{Summary of (A) study scale and (B) dominant runoff source.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:fig1}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,]{../figures/bmp_summary_param} \caption{Summary BMPs identified in the systematic review by tested parameter.
-Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:bmpsummaryparam}
+\includegraphics[width=1\linewidth,]{frontiers_submission_revision_01_files/figure-latex/fig2} \caption{Summary of BMPs identified in the systematic review by tested parameter.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:fig2}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,]{../figures/bmp_temporal_summary} \caption{Number of studies identified in the systematic review summarized by (A) publication date and (B) study length.
-Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:bmptemporal}
+\includegraphics[width=1\linewidth,]{frontiers_submission_revision_01_files/figure-latex/fig3} \caption{Number of studies identified in the systematic review summarized by (A) publication date and (B) study length.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:fig3}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=1\linewidth,]{../figures/study_map} \caption{Distribution of studies identified in the systematic review by state and parameter.
-Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:studymap}
+\includegraphics[width=1\linewidth,]{frontiers_submission_revision_01_files/figure-latex/fig4} \caption{Distribution of studies identified in the systematic review by state and parameter.
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:fig4}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=1\linewidth,]{../figures/overall_effect} \caption{Estimated effect sizes and intervals from the intercept only multilevel random effects model.
+\includegraphics[width=1\linewidth,]{frontiers_submission_revision_01_files/figure-latex/fig5} \caption{Estimated effect sizes and intervals from the intercept only multilevel random effects model.
 Individual points represent studies, with size scaled by sampling variance.
@@
 Here, \textit{k} indicates the number of overall effects with the number of unique studies in parenthesis.
-Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:overalleffect}
+Water quality parameters include total suspended sediment (TSS), total phosphorus (TP), total nitrogen (TN), orthophosphate (PO\textsubscript{4}), fecal indicator bacteria (FIB), and dissolved inorganic nitrogen (DIN).}\label{fig:fig5}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=1\linewidth,]{../figures/bac_predicted} \caption{Predicted marginal effect of influent fecal indicator bacteria (FIB; Panel A) and aridity index (conditioned on BMP subcategory; Panel B).
+\includegraphics[width=1\linewidth,]{frontiers_submission_revision_01_files/figure-latex/fig6} \caption{Predicted marginal effect of influent fecal indicator bacteria (FIB; Panel A) and aridity index (conditioned on BMP subcategory; Panel B).
 Solid lines are the predicted mean effect, dashed lines are the 95\% confidence intervals, and the dotted lines are the 95\% prediction intervals.
 Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.
-Higher aridity index (mean annual precipition/evapotranspiration) indicates more humid conditions.}\label{fig:fibeffect}
+Higher aridity index (mean annual precipition/evapotranspiration) indicates more humid conditions.}\label{fig:fig6}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=1\linewidth,]{../figures/phos_predicted} \caption{Predicted marignal effect of influent pollutant concentration on total phosphorus (TP; Panel A) and orthophosphate (PO\textsubscript{4}; Panel B) reductions.
+\includegraphics[width=1\linewidth,]{frontiers_submission_revision_01_files/figure-latex/fig7} \caption{Predicted marignal effect of influent pollutant concentration on total phosphorus (TP; Panel A) and orthophosphate (PO\textsubscript{4}; Panel B) reductions.
 Solid lines are the predicted mean effect, dashed lines are the 95\% confidence intervals, and the dotted lines are the 95\% prediction intervals.
-Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.}\label{fig:phoseffect}
+Individual dots represent each effect size identified in the literature with the size scaled by sampling variance.}\label{fig:fig7}
 \end{figure}
@@
 \begin{figure}[p]
-\includegraphics[width=0.5\linewidth,]{../figures/aridity_density} \caption{Comparison of the relative distributions (density) of aridity index values (mean annual precipition/evapotranspiration) across the U.S., across U.S.
+\includegraphics[width=0.5\linewidth,]{frontiers_submission_revision_01_files/figure-latex/fig8} \caption{Comparison of the relative distributions (density) of aridity index values (mean annual precipition/evapotranspiration) across the U.S., across U.S.
 agricultural and developed land uses, and for BMP study locations.
-Higher aridity index values indicate more humid conditions.}\label{fig:aiindex}
+Higher aridity index values indicate more humid conditions.}\label{fig:fig8}
 \end{figure}

### 2024-05-30 R/05_literature_statistics.R

@@ -242,7 +242,7 @@ plot_bmp_summary <- function(bac_df, nut_df) {
     mutate(Scale = forcats::fct_relevel(Scale, "Lot/Field", "Community/Farm", "Catchment/Watershed"))
   
   p1 <- ggplot(scale_df) +
-    geom_col(aes(n, Parameter, fill = Scale)) +
+    geom_col(aes(n, Parameter, fill = Scale), color = "black", linewidth = 0.2) +
     geom_text_repel(aes(n, Parameter, label = n, group = Scale),
                     position = position_stack(vjust = 0.5, reverse = FALSE),
                     bg.color = "#FFFFFF33", bg.r = 0.15,
@@ -250,7 +250,11 @@ plot_bmp_summary <- function(bac_df, nut_df) {
                     force = 0, size = 3) +
     scale_x_continuous("Studies (n)", expand = expansion(mult = c(0, 0.65))) +
     scale_y_discrete("") +
-    scale_fill_manual("Study Scale:", values = c("#1B9E77","#D95F02","#7570B3"),
+    # scale_fill_manual("Study Scale:", values = c("#1B9E77","#D95F02","#7570B3"),
+    #                   guide = guide_legend(reverse = TRUE)) +
+    scale_fill_brewer("Study Scale:",
+                      type = "div",
+                      palette = "PuOr",
                       guide = guide_legend(reverse = TRUE)) +
     theme_mps_noto(base_family = "Manrope Regular") +
     theme(axis.title = element_text(family = "Manrope SemiBold", face = "plain", size = rel(1)),
@@ -303,7 +307,7 @@ plot_bmp_summary <- function(bac_df, nut_df) {
     bind_rows(source_bac)
   
   p2 <- ggplot(source_df) +
-    geom_col(aes(n, Parameter, fill = Source_Type)) +
+    geom_col(aes(n, Parameter, fill = Source_Type), color = "black", linewidth = 0.2) +
     geom_text_repel(aes(n, Parameter, label = n, group = Source_Type),
                     position = position_stack(vjust = 0.5, reverse = FALSE),
                     bg.color = "#FFFFFF33", bg.r = 0.15,
@@ -311,7 +315,11 @@ plot_bmp_summary <- function(bac_df, nut_df) {
                     force = 0, size = 3) +
     scale_x_continuous("Studies (n)", expand = expansion(mult = c(0, 0.6))) +
     scale_y_discrete("") +
[…]

### 2024-05-30 diffs/diff.R

@@ -1,0 +1,17 @@
+library(latexdiffr)
+Sys.setenv("LATEXDIFF_PATH" = "c:/Users/michael.schramm/AppData/Roaming/TinyTeX/texmf-dist/scripts/latexdiff/latexdiff.pl")
+tinytex::tlmgr_path("add")
+latexdiff("frontiers_submission/frontiers_submission.tex", 
+          "frontiers_submission/frontiers_submission_revision_01.tex",
+          output = "frontiers_submission/diff",
+          quiet = FALSE,
+          compile = FALSE)
+
+# 
+# tinytex::latexmk("diff.tex",
+#                  engine = "xelatex",
+#                  bib_engine = "biber",
+#                  pdf_file = "frontiers_submission/diff.pdf")
+
+system2(command = "latexmk",
+        args = c("-pdf", "diff.tex", "-f"))

### 2024-06-10 frontiers_submission/frontiers_submission_revision_02.Rmd
fix typo in table 1

--- 2024-05-30 frontiers_submission/frontiers_submission_revision_01.Rmd
@@
                            "Studies published from 2000 through 2022."),
-  "Excluision Criteria" = c("Synopsis or review studies, reports with reductions based on modeled or other estimated reductions (e.g. TMDLs, watershed plans, or modelling studies.",
+  "Exclusion Criteria" = c("Synopsis or review studies, reports with reductions based on modeled or other estimated reductions (e.g. TMDLs, watershed plans, or modelling studies.",
                             "Studies not explicitly linking concentration reductions to a specific BMP or insufficient information to quantify reductions.",

### 2024-07-12 README.Rmd
clean up targets and update readme with renv diagnostics.
add doi badge
Final published manuscript

@@
 ---
+
+# A  meta-analysis  of  the  impacts  of  best management  practices  on  nonpoint  source pollutant  concentration
+
+[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.10795231.svg)](https://doi.org/10.5281/zenodo.10795231)
 
@@
 
-This repository contains the data and analysis used in the draft manuscript *"A  meta-analysis  of  the  impacts  of  best management  practices  on  nonpoint  source pollutant  concentration,"* by Schramm et al.
+This repository contains the data and analysis used in:
+
+Schramm, M., Kikoyo, D., Wright, J., Jain, S.
+(2024) [*"A  meta-analysis  of  the  impacts  of  best management  practices  on  nonpoint  source pollutant  concentration,"*](https://doi.org/10.3389/frwa.2024.1397615).
+Frontiers in Water.
+doi:10.3389/frwa.2024.1397615.
 
@@
 ```{r}
-sessioninfo::session_info()
+renv::diagnostics()
 ```

### 2024-09-24 ucowr-2024/Schramm-UCOWR-2024.qmd
add UCOWR presentation

---
title: "The Meta-Analysis of Nonpoint Source BMPs: Can We Improve the Empirical Knowledge Associated with BMP Water Quality Improvement?"
author: "Michael Schramm, Duncan Kikoyo, Janelle Wright, Shubham Jain"
institute: "Texas Water Resources Institute, Texas A&M AgriLife Research.\\hfill \\break Project funding provided by the Texas State Soil and Water Conservation Board. \\hfill \\break 2024 AWRA, UCOWR, & NIWR Joint Water Resources Conference"
date: 8/30/2024
format: 
  beamer:
    theme: moloch
    themeoptions: progressbar=frametitle
    aspectratio: 169
    keep-tex: true
    include-in-header: preamble.tex
    bibliography: bibliography.bib
    cite-method: biblatex
    biblio-style: apa
    include-before-body: before-body.tex
    template-partials: 
      - biblio.tex

---

```{r setup}
library(targets)
library(ggtext)
library(kableExtra)
library(patchwork)
library(ragg)
library(targets)
library(tidyverse)

## flowchart
library(ggdag)
library(rsvg)

## this sets our default code chunk options
knitr::opts_chunk$set(dev = "ragg_png",
                      echo = FALSE,
                      ## figure chunk options
                      fig.width = 6,
                      dpi = 300)

## make fonts available
font_hoist <- function(family, silent = FALSE) {
  font_specs <- systemfonts::system_fonts() |> 
    filter(family == {{family}}) |> 
    mutate(family = paste(family, style)) |> 
    select(plain = path, name = family)

  purrr::pwalk(as.list(font_specs), systemfonts::register_font)

  if (!silent)  message(paste0("Hoisted ", nrow(font_specs), " variants:\n",
                               paste(font_specs$name, collapse = "\n")))
}

font_hoist("Manrope")

## grab output from targets to use in manuscript
store <- "../_targets"
```

## Introduction

### Purpose

 - Assess scope and scale of published water quality BMP field studies.
 - Assess mean reductions in pollutant *concentrations* in BMP studies.
 - Methods and results fully detailed in Frontiers article \footfullcite{schrammMetaanalysisImpactsBest2024}

## Methods

### Systematic Review
 - Followed systematic review guidance in Collaboration for Environmental Evidence \footfullcite{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}
  - Review identified **90 studies** and **642 effect sizes**
 - Collected data included: *Publication date, measured parameters, runoff source, BMP type, study scale, location, study length, measurement units, control and treatment sample size, standard error, standard deviation, mean concentration, etc.*

 - Raw data available on Zenodo \footfullcite{kikoyoTWRINonpointSource2023}

## Methods

```{r}
#| id: bmpstruct
#| fig-width: 6
#| fig-height: 3
dag <- dagify(
  manure_management ~ livestock,
  prescribed_grazing ~ livestock,
  livestock_exclusion ~ livestock,
  alternate_water ~ livestock,
  reduced_till ~ crop,
  no_till ~ crop,
  nutrient_management ~ crop,
  cover_crop ~ crop,
  crop ~ management,
  livestock ~ management,
  sand_filter ~ filtration,
  swale ~ filtration,
  filter_strip ~ filtration,
  infil_trench ~ infiltration,
  infil_chamber ~ infiltration,
  infil_basin ~ infiltration,
  retention_basin ~ detention,
  sediment_trap ~ detention,
  detention_basin ~ detention,
  tree_filter ~ treatment,
  bioreactor ~ treatment,
  wetland ~ treatment,
  bioret ~ treatment,
  dtr ~ drainage,
  proprietary ~ drainage,
  green_roof ~ drainage,
  gss ~ drainage,
  controlled_drainage ~ drainage,
  filtration ~ structural,
  infiltration ~ structural,
  detention ~ structural,
  treatment ~ structural,
  drainage ~ structural,
  labels = c(
    "manure_management" = "Manure Management",
    "prescribed_grazing" = "Prescribed Grazing",
    "livestock_exclusion" = "Livestock Exclusion",
    "alternate_water" = "Alternate Water",
    "reduced_till" = "Reduced Till",
    "no_till" = "No Till",
    "nutrient_management" = "Nutrient Management",
    "cover_crop" = "Cover Crop",
    "crop" = "Crop",
    "livestock" = "Livestock",
    "sand_filter" = "Sand Filter",
    "swale" = "Swale",
    "filter_strip" = "Filter Strip",
    "infil_trench" = "Infiltration Trench",
    "infil_chamber" = "Infiltration Chamber",
    "infil_basin" = "Infiltration Basin",
    "retention_basin" = "Retention Basin",
    "sediment_trap" = "Sediment Trap",
    "detention_basin" = "Detention Basin",
    "tree_filter" = "Tree Filter",
    "bioret" = "Bioretention",
    "bioreactor" = "Bioreactor",
    "wetland" = "Wetland",
    "dtr" = "Drainage Tailwater Recovery",
    "proprietary" = "Proprietary",
    "green_roof" = "Green Roof",
    "gss" = "Grade Stabilization Structure",
    "controlled_drainage" = "Controlled Drainage",
    "filtration" = "Filtration",
    "infiltration" = "Infiltration",
    "detention" = "Detention",
    "treatment" = "Treatment",
    "drainage" = "Drainage",
    "management" = "Management\nBMPs",
    "structural" = "Structural\nBMPs"
  )
) |> 
  tidy_dagitty(layout = "tree", circular = FALSE) |>
  mutate(rotate = case_when(
    !(name %in% c("structural", "drainage", "treatment", "detention", "infiltration", "management", "crop", "livestock", "filtration")) ~ 90,
    .default = 0
  ),
  node_size = case_when(
    !(name %in% c("structural", "drainage", "treatment", "detention", "infiltration", "management", "crop", "livestock", "filtration")) ~ NA,
    .default = 20
  ),
  hjust = case_when(
    !(name %in% c("structural", "drainage", "treatment", "detention", "infiltration", "management", "crop", "livestock", "filtration")) ~ 0,
    .default = 0.5
  ),
  nudge_y = case_when(
    !(name %in% c("structural", "drainage", "treatment", "detention", "infiltration", "management", "crop", "livestock", "filtration")) ~ 1,
    .default = 0
  ))

ggplot(dag, aes(x = -y, y = x, xend = -yend, yend = xend)) +
  geom_dag_node(aes(size = node_size), color = "wheat") +
  geom_dag_edges_link() +
  geom_dag_text(data = dag |> 
                  filter(nudge_y == 0),
                aes(angle = rotate, hjust = hjust, label = label), size = 2, color = "black") +
  geom_dag_text(data = dag |>
                  filter(nudge_y == 1),
                aes(#angle = rotate, 
                    hjust = hjust, label = label), size = 2, color = "black",
                nudge_x = -0.1) +
  scale_size_area(max_size = 20) +
  scale_y_continuous(limits = c(-11,17),
                     expand = expansion(0,0)) +
  scale_x_continuous(limits = c(-2.5, 1),
                     expand = expansion(0,0)) +
  theme_dag() +
  theme(legend.position = "none")

```

## Methods

### Meta-Analysis

 - Describe overall effect and heterogeneity
 - Weighted multilevel random effects models
    - Weights individual effect sizes based on sampling variance
 - Model terms:
    - Influent pollutant concentration
    - BMP category
    - Aridity index: $\frac{\bar{P}}{\bar{PET}}$ \footfullcite{zomerVersionGlobalAridity2022}
    - Fixed effect interactions
    - Nested random effects $(Effect|Study)$

## Methods

### Meta-Analysis

 - Dependent variable: log ratio of means

$$
ROM_i = ln(\frac{x_{i,control}}{x_{i,trmt}}) = ln(x_{i,control}) - ln(x_{i,trmt})
$$
 - Separate models for each parameter of interest: **fecal indicator bacteria** (FIB), **total nitrogen** (TN), **total phosphorus** (TP), **total suspended sediment** (TSS), **orthophosphate** (PO~4~), **dissolved inorganic nitrogen** (DIN).

## Study Scale & Source

```{r}
#| label: p1
#| fig-height: 3
#| fig-width: 8
#| out-width: "100%"
#| fig-dpi: 300

tar_read(bmp_summary, store = store) &
  theme(legend.text = element_text(size = rel(0.65)),
        legend.title = element_text(size = rel(0.65)),
        legend.position = c(0.8, 0.65),)
```

- Most studies were lot/field scale and investigated urbanized runoff.
- FIB studies tended toward larger catchments.

## Publication Date and Study Length

```{r}
#| label: p2
#| fig-dpi: 300
#| fig-width: 6
#| fig-height: 6
#| out-width: "50%"
#| 
tar_read(bmp_temporal, store = store)$p1
```

## Spatial Distribution

```{r}
#| label: p3
#| fig-dpi: 300
#| fig-width: 6.5
#| fig-height: 6.5
#| out-width: "50%"
#| 
tar_read(study_map, store = store)
```

## Results

::::: {.columns}

:::: {.column width=60%}

```{r}
#| label: p4
#| fig-height: 5.75
#| fig-width: 6.5
#| out-width: "100%"
#| fig-dpi: 300
tar_read(overall_effect, store = store)
```
::::

:::: {.column width=40%}

### Overall Effect

* Strong evidence across studies for mean reductions in:
  + **FIB (57%)** 
  + **TN (34%)** 
  + **TP (32%)** 
  + **TSS (81%)**

* Prediction intervals provide low certainty about the performance of future *individual* studies.

::::

:::::

## Attempts to Explain Variation in BMP Performance

::::: {.columns}

:::: {.column width=40%}

```{r}
tibble::tibble("Parameter" = c("FIB", "TP", "PO\\textsubscript{4}"),
               "Moderators" = kableExtra::linebreak(c("Influent Concentration,\nAridity × BMP Subcategory",
                                "Influent Concentration",
                                "Influent Concentration"))) |> 
  kableExtra::kbl(escape = FALSE)
```

Influent concentration + Aridity×BMP Subcategory explained ~89% of effect size variance in FIB models.

::::

:::: {.column width=60%}

```{r}
#| label: p5
#| fig-width: 6
#| fig-height: 4
#| out-width: "100%"
#| fig-dpi: 300
tar_read(bac_marginal_means, store = store) &
  theme(legend.position = "none")
```

::::

:::::

## Attempts to Explain Variation in BMP Performance

::::: {.columns}

:::: {.column width=50%}

```{r}
tibble::tibble("Parameter" = c("FIB", "TP", "PO\\textsubscript{4}"),
               "Moderators" = kableExtra::linebreak(c("Influent Concentration,\nAridity × BMP Subcategory",
                                "Influent Concentration",
                                "Influent Concentration"))) |> 
  kableExtra::kbl(escape = FALSE)
```

Influent concentration explained 12% and 35% of effect size variance in TP and PO~4~ models.

::::

:::: {.column width=50%}

```{r}
#| label: p6
#| fig-width: 6
#| fig-height: 4
#| out-width: "100%"
#| fig-dpi: 300
tar_read(phos_marginal_means, store = store) &
  theme(legend.position = "none")
```

::::

:::::

## Overview

* Few long-term studies (>4 year).
* Studies clustered in the mid-Atlantic, low spatial variability.
* Studies demonstrated mean reductions in FIB, TN, TP, and TSS.
* Did not find evidence for mean reductions in DIN, PO~4~.
* Influent concentration has strong influence on FIB, TP, and PO~4~ performance.
* Failed to explain majority of variance in nutrient and TSS models. **(Why???)**

## Lessons Learned

**Issues in Published BMP Studies**

* Very inconsistent reporting of non-manipulated study conditions.
* Earlier studies tended to report BMP efficiency only.
* Many studies fail to report sample size, variance, or uncertainty measures (59% of identified effect sizes).
* There is an opportunity to develop reporting protocols for BMP studies. See Eagle et al. \footfullcite{eagleMetaAnalysisConstrained2017}
  - Environmental variables, reported metrics, regression model reporting...

## Lessons Learned

**Study/Systematic Review Issues**

* Broad study design (focus on specific BMP types).
* Keyword comprehensiveness.
* *metafor* R package facilitates meta-analysis.\footfullcite{viechtbauerConductingMetaanalysesMetafor2010} 
* Developing and training students/staff on data extraction is difficult.
* Data entry interfaces that reinforce data consistency.

## Thank You

**Contact:** michael.schramm@ag.tamu.edu

**Acknowledgements:** Texas State Soil and Water Conservation Board for funding.

**Paper:** Schramm M, Kikoyo D, Wright J and Jain S (2024) A meta-analysis of the impacts of best management practices on nonpoint source pollutant concentration. Front. Water 6:1397615. doi: 10.3389/frwa.2024.1397615

**Data:** Kikoyo D, Jain S, Wright J, and Schramm M. (2024). TWRI NPS-BMP-DB (v1.05). Zenodo. doi: 10.5281/zenodo.10451303

### 2024-09-24 ucowr-2024/Schramm-UCOWR-2024.tex
add UCOWR presentation

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
  aspectratio=169,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\usetheme[progressbar=frametitle]{moloch}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
\usepackage{caption}
% Make caption package work with longtable
\makeatletter
\def\fnum@table{\tablename~\thetable}
\makeatother
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\setbeamertemplate{caption}[default]
%\setbeamertemplate{caption name}{\raggedright\insertcaption\par}
\setbeamertemplate{caption label seperator}[none]
%\setbeamerfont{footnote}{size=\tiny}
\definecolor{twriBlue}{HTML}{0054a4}
\definecolor{spotMaroon}{HTML}{5d0025}
\setbeamercolor{alerted text}{fg=spotMaroon}
\setbeamercolor{frametitle}{bg=twriBlue}
\usepackage{fontspec}

\setsansfont{Open Sans}

% title logos
\titlegraphic{%
    \includegraphics[width=.2\textwidth]{logo}\hfill
  \includegraphics[width=.2\textwidth]{logo-AgriLife}\hfill
  \includegraphics[width=.2\textwidth]{logo-TSSWCB}
}

\makeatletter
\setbeamertemplate{title page}{
  \begin{minipage}[b][\paperheight]{\textwidth}
    \vfill%
    \ifx\inserttitle\@empty\else\usebeamertemplate*{title}\fi
    \ifx\insertsubtitle\@empty\else\usebeamertemplate*{subtitle}\fi
    \usebeamertemplate*{title separator}
    \ifx\beamer@shortauthor\@empty\else\usebeamertemplate*{author}\fi
    \ifx\insertdate\@empty\else\usebeamertemplate*{date}\fi
    \ifx\insertinstitute\@empty\else\usebeamertemplate*{institute}\fi
    \vfill
    \ifx\inserttitlegraphic\@empty\else\inserttitlegraphic\fi
    \vspace*{1cm}
  \end{minipage}
}
\makeatother
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[style=apa,]{biblatex}
\addbibresource{bibliography.bib}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Meta-Analysis of Nonpoint Source BMPs: Can We Improve the Empirical Knowledge Associated with BMP Water Quality Improvement?},
  pdfauthor={Michael Schramm, Duncan Kikoyo, Janelle Wright, Shubham Jain},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{The Meta-Analysis of Nonpoint Source BMPs: Can We Improve the
Empirical Knowledge Associated with BMP Water Quality Improvement?}
\author{Michael Schramm, Duncan Kikoyo, Janelle Wright, Shubham Jain}
\date{2024-08-30}
\institute{Texas Water Resources Institute, Texas A\&M AgriLife
Research.\hfill \break Project funding provided by the Texas State Soil
and Water Conservation Board. \hfill \break 2024 AWRA, UCOWR, \& NIWR
Joint Water Resources Conference}

\begin{document}
\frame{\titlepage}

\begin{frame}{Introduction}
\phantomsection\label{introduction}
\begin{block}{Purpose}
\phantomsection\label{purpose}
\begin{itemize}
\tightlist
\item
  Assess scope and scale of published water quality BMP field studies.
\item
  Assess mean reductions in pollutant \emph{concentrations} in BMP
  studies.
\item
  Methods and results fully detailed in Frontiers article
  \footfullcite{schrammMetaanalysisImpactsBest2024}
\end{itemize}
\end{block}
\end{frame}

\begin{frame}{Methods}
\phantomsection\label{methods}
\begin{block}{Systematic Review}
\phantomsection\label{systematic-review}
\begin{itemize}
\item
  Followed systematic review guidance in Collaboration for Environmental
  Evidence
  \footfullcite{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}
\item
  Review identified \textbf{90 studies} and \textbf{642 effect sizes}
\item
  Collected data included: \emph{Publication date, measured parameters,
  runoff source, BMP type, study scale, location, study length,
  measurement units, control and treatment sample size, standard error,
  standard deviation, mean concentration, etc.}
\item
  Raw data available on Zenodo
  \footfullcite{kikoyoTWRINonpointSource2023}
\end{itemize}
\end{block}
\end{frame}

\begin{frame}{Methods}
\phantomsection\label{methods-1}
\includegraphics{Schramm-UCOWR-2024_files/figure-beamer/bmpstruct-1.png}
\end{frame}

\begin{frame}{Methods}
\phantomsection\label{methods-2}
\begin{block}{Meta-Analysis}
\phantomsection\label{meta-analysis}
\begin{itemize}
\tightlist
\item
  Describe overall effect and heterogeneity
\item
  Weighted multilevel random effects models

  \begin{itemize}
  \tightlist
  \item
    Weights individual effect sizes based on sampling variance
  \end{itemize}
\item
  Model terms:

  \begin{itemize}
  \tightlist
  \item
    Influent pollutant concentration
  \item
    BMP category
  \item
    Aridity index: \(\frac{\bar{P}}{\bar{PET}}\)
    \footfullcite{zomerVersionGlobalAridity2022}
  \item
    Fixed effect interactions
  \item
    Nested random effects \((Effect|Study)\)
  \end{itemize}
\end{itemize}
\end{block}
\end{frame}

\begin{frame}{Methods}
\phantomsection\label{methods-3}
\begin{block}{Meta-Analysis}
\phantomsection\label{meta-analysis-1}
\begin{itemize}
\tightlist
\item
  Dependent variable: log ratio of means
\end{itemize}

\[
ROM_i = ln(\frac{x_{i,control}}{x_{i,trmt}}) = ln(x_{i,control}) - ln(x_{i,trmt})
\] - Separate models for each parameter of interest: \textbf{fecal
indicator bacteria} (FIB), \textbf{total nitrogen} (TN), \textbf{total
phosphorus} (TP), \textbf{total suspended sediment} (TSS),
\textbf{orthophosphate} (PO\textsubscript{4}), \textbf{dissolved
inorganic nitrogen} (DIN).
\end{block}
\end{frame}

\begin{frame}{Study Scale \& Source}
\phantomsection\label{study-scale-source}
\includegraphics[width=1\textwidth,height=\textheight]{Schramm-UCOWR-2024_files/figure-beamer/p1-1.png}

\begin{itemize}
\tightlist
\item
  Most studies were lot/field scale and investigated urbanized runoff.
\item
  FIB studies tended toward larger catchments.
\end{itemize}
\end{frame}

\begin{frame}{Publication Date and Study Length}
\phantomsection\label{publication-date-and-study-length}
\includegraphics[width=0.5\textwidth,height=\textheight]{Schramm-UCOWR-2024_files/figure-beamer/p2-1.png}
\end{frame}

\begin{frame}{Spatial Distribution}
\phantomsection\label{spatial-distribution}
\includegraphics[width=0.5\textwidth,height=\textheight]{Schramm-UCOWR-2024_files/figure-beamer/p3-1.png}
\end{frame}

\begin{frame}{Results}
\phantomsection\label{results}
\begin{columns}[T]
\begin{column}{0.6\textwidth}
\includegraphics[width=1\textwidth,height=\textheight]{Schramm-UCOWR-2024_files/figure-beamer/p4-1.png}
\end{column}

\begin{column}{0.4\textwidth}
\begin{block}{Overall Effect}
\phantomsection\label{overall-effect}
\begin{itemize}
\tightlist
\item
  Strong evidence across studies for mean reductions in:

  \begin{itemize}
  \tightlist
  \item
    \textbf{FIB (57\%)}
  \item
    \textbf{TN (34\%)}
  \item
    \textbf{TP (32\%)}
  \item
    \textbf{TSS (81\%)}
  \end{itemize}
\item
  Prediction intervals provide low certainty about the performance of
  future \emph{individual} studies.
\end{itemize}
\end{block}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Attempts to Explain Variation in BMP Performance}
\phantomsection\label{attempts-to-explain-variation-in-bmp-performance}
\begin{columns}[T]
\begin{column}{0.4\textwidth}
\begin{tabular}[t]{l|l}
\hline
Parameter & Moderators\\
\hline
FIB & \makecell[l]{Influent Concentration,\\Aridity × BMP Subcategory}\\
\hline
TP & Influent Concentration\\
\hline
PO\textsubscript{4} & Influent Concentration\\
\hline
\end{tabular}

Influent concentration + Aridity×BMP Subcategory explained
\textasciitilde89\% of effect size variance in FIB models.
\end{column}

\begin{column}{0.6\textwidth}
\includegraphics[width=1\textwidth,height=\textheight]{Schramm-UCOWR-2024_files/figure-beamer/p5-1.png}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Attempts to Explain Variation in BMP Performance}
\phantomsection\label{attempts-to-explain-variation-in-bmp-performance-1}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{tabular}[t]{l|l}
\hline
Parameter & Moderators\\
\hline
FIB & \makecell[l]{Influent Concentration,\\Aridity × BMP Subcategory}\\
\hline
TP & Influent Concentration\\
\hline
PO\textsubscript{4} & Influent Concentration\\
\hline
\end{tabular}

Influent concentration explained 12\% and 35\% of effect size variance
in TP and PO\textsubscript{4} models.
\end{column}

\begin{column}{0.5\textwidth}
\includegraphics[width=1\textwidth,height=\textheight]{Schramm-UCOWR-2024_files/figure-beamer/p6-1.png}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Overview}
\phantomsection\label{overview}
\begin{itemize}
\tightlist
\item
  Few long-term studies (\textgreater4 year).
\item
  Studies clustered in the mid-Atlantic, low spatial variability.
\item
  Studies demonstrated mean reductions in FIB, TN, TP, and TSS.
\item
  Did not find evidence for mean reductions in DIN, PO\textsubscript{4}.
\item
  Influent concentration has strong influence on FIB, TP, and
  PO\textsubscript{4} performance.
\item
  Failed to explain majority of variance in nutrient and TSS models.
  \textbf{(Why???)}
\end{itemize}
\end{frame}

\begin{frame}{Lessons Learned}
\phantomsection\label{lessons-learned}
\textbf{Issues in Published BMP Studies}

\begin{itemize}
\tightlist
\item
  Very inconsistent reporting of non-manipulated study conditions.
\item
  Earlier studies tended to report BMP efficiency only.
\item
  Many studies fail to report sample size, variance, or uncertainty
  measures (59\% of identified effect sizes).
\item
  There is an opportunity to develop reporting protocols for BMP
  studies. See Eagle et
  al.~\footfullcite{eagleMetaAnalysisConstrained2017}

  \begin{itemize}
  \tightlist
  \item
    Environmental variables, reported metrics, regression model
    reporting\ldots{}
  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Lessons Learned}
\phantomsection\label{lessons-learned-1}
\textbf{Study/Systematic Review Issues}

\begin{itemize}
\tightlist
\item
  Broad study design (focus on specific BMP types).
\item
  Keyword comprehensiveness.
\item
  \emph{metafor} R package facilitates
  meta-analysis.\footfullcite{viechtbauerConductingMetaanalysesMetafor2010}
\item
  Developing and training students/staff on data extraction is
  difficult.
\item
  Data entry interfaces that reinforce data consistency.
\end{itemize}
\end{frame}

\begin{frame}{Thank You}
\phantomsection\label{thank-you}
\textbf{Contact:} michael.schramm@ag.tamu.edu

\textbf{Acknowledgements:} Texas State Soil and Water Conservation Board
for funding.

\textbf{Paper:} Schramm M, Kikoyo D, Wright J and Jain S (2024) A
meta-analysis of the impacts of best management practices on nonpoint
source pollutant concentration. Front. Water 6:1397615. doi:
10.3389/frwa.2024.1397615

\textbf{Data:} Kikoyo D, Jain S, Wright J, and Schramm M. (2024). TWRI
NPS-BMP-DB (v1.05). Zenodo. doi: 10.5281/zenodo.10451303
\end{frame}

%\begin{frame}[allowframebreaks]{}
%  \bibliographytrue
%  %\printbibliography[heading=none]
%\end{frame}

\end{document}
